### 2024-01-18 meetings/18-01-2024_slides.qmd
slides meeting 18-01-2024

---
title: "Meeting --- Summary LLM"
author: "Bas Doners, Dennis Fok, Finn-Ole Höner"
date: 2024-01-18
date-format: long
format: beamer
aspectratio: 169
incremental: false
include-in-header: ../beamer_preamble.tex
bibliography: ../references.bib
fig-align: center
fig-cap-location: bottom
fig-format: pdf
links-as-notes: true
number-depth: 1
---

# Terms

target string
: The string that we want to "re-create"

input embedding
: $n_{tokens} \times n_{vocab}$ input embedding

flat input embedding
: $1 \times n_{vocab}$ input embedding

# Can we speed things up? {.allowframebreaks}

- flat input embedding of target string helps
- GPU helps a lot
- RMSProp instead of ADAM helps
  - Adjust learning rate based on input embedding and generation technique
- $n$-Beam-search is slightly slower than greedy search, but not by factor $n$ (finding better steps that take longer to compute?)

- Stop optimization based on hitting the target sequence instead? About half the iterations
- Is there a way to move to "conditional likelihoods", once we found one correct token?
- How stable is the optimization?

---

"Does it make sense to compare apples and oranges?", flat input embedding, RMSProp(lr=1e-1), T4GPU

:::{layout-ncol="2"}

![Convergence](image002.png)

![NLL](image003.png)

:::

# Generation technique?

- We run into problems if the generation technique samples: Then we cannot track the likelihood anymore (inplace gradient change)
- Likely beam search with many beams is the most balanced option?
- Contrastive search, which penalizes repeated tokens, does not work as it relies on sampling behind the scenes
- Maybe we can penalize the likelihood itself for repetitions? Or solve it through restrictions on the optimization?

# Number of parameters?

- With flat input embedding, we have $n_{vocab}$ ($768$ for GPT-2) parameters. 
- Many hyperparameters: Starting values, optimizer, generation settings, LLM
- Maybe we can reduce the number of parameters by having a smaller layer in front of the flat input-embedding. This smaller layer plus the weights connecting it to the input embedding could be less than $n_{vocab}$. (Backwards convolution?)

# Other

- Need to read up on the Google Colab resources, whether it is suitable for computing larger datasets
- GPU on Snellius (would that be feasible)?
- GitHub
- Can we compare the likelihood across different settings? (Likely not?)
- Difference to "Reverse Prompt-Engineering"

### 2024-01-18 meetings/meeting_notes.qmd
slides meeting 18-01-2024

---
title: "Meeting Notes --- Summary LLM"
author: "Bas Doners, Dennis Fok, Finn-Ole Höner"
date: last-modified
date-format: long
format:
    pdf:
        documentclass: scrartcl
        toc: false
        number-sections: false
        number-depth: 2
        colorlinks: true
        fig-pos: 'H'
        fig-align: center
        geometry:
            - top = 30mm
            - bottom = 30mm
            - left = 20mm
            - right = 20mm
        papersize: dina4
aspectratio: 169
incremental: false
bibliography: ../references.bib
fig-align: center
fig-cap-location: bottom
fig-format: pdf
links-as-notes: true
number-depth: 1
---

# 18-01-2024

> First steps towards a summary LLM, basic computations seem to work.

## Can we speed things up?

## Generation technique?

## Number of parameters?

## Other

## Next steps

### 2024-01-31 meetings/meeting_notes.qmd
meeting notes

@@
 ---
-title: "Meeting Notes --- Summary LLM"
-author: "Bas Doners, Dennis Fok, Finn-Ole Höner"
+title: "Meeting Notes---Summary LLM"
+author: "Bas Donkers, Dennis Fok, Finn-Ole Höner"
 date: last-modified
@@
         papersize: dina4
-aspectratio: 169
-incremental: false
 bibliography: ../references.bib
@@
 fig-format: pdf
-links-as-notes: true
+links-as-notes: false
 number-depth: 1
@@
 
-> First steps towards a summary LLM, basic computations seem to work.
+- Status: First steps towards a summary LLM, basic computations seem to work.
+- Admin: Setup of MS Teams group
+- Technical Note: The `output` object is a tuple of length "positions", and each row of the contained matrix is one beam.
 
-## Can we speed things up?
+## Dutch LLMs
 
-## Generation technique?
+> I found some dutch language models on Huggingface, e.g.
+a [GPT-2](https://huggingface.co/GroNLP/gpt2-small-dutch) version which I use for the illustration below (also available with more parameters, see [corresponding paper](https://arxiv.org/abs/2012.05628)).
+There are also other “dutch” versions of other language models, see this [collection](https://huggingface.co/collections/BramVanroy/dutch-llms-64f99dddbb86c0fa80846f89).
+One caveat might be that all these models are in some way fine-tuned/transfer learned from the English version of the model, so not “natively” trained on only dutch corpora.
+However, it seems to work in our current “brute force” optimization.See attachements for the targetHeeft het zin om appels en sinaasappelen te vergelijken??
+(sic)
 
-## Number of parameters?
+## Speeding up computations
 
-## Other
+> Currently, we are performing too many computations.
+We can make use of the fact that we know the target string.
+Essentially, we can compute each "next word" in parallel, instead of sequentially.
+This also circumvents the problem that we need to think about generation strategies, as we can essentially just make a forward pass through the LLM.
 
-## Next steps
+## TO-DOs
+
+- Finn speeds computations up
+- Finn tries to write draft for grant proposal
+- Reaching out to SKIM, slide deck, NDA
+
+<!-- TODO: Finn speeds computations up -->
+<!-- TODO: Finn tries to write draft for grant proposal -->
+<!-- TODO: Reaching out to SKIM, slide deck, NDA -->

### 2024-02-21 src/llm/llm.py
first three versions of optimization

@@ -1,127 +1,0 @@
-# SETUP --------------------------
-import torch
-from transformers import AutoTokenizer, GPT2LMHeadModel
-import pandas as pd
-import numpy as np
-
-# MAGICKS --------------------------
-s_model = 'gpt2'
-s_string = "Finn writes code"
-
-# FUNCTIONS --------------------------
-def calc_joint_prob(transition_scores: torch.Tensor) -> float:
-    # calculate the joint probability of a generated text
-    return np.exp(np.sum(transition_scores.numpy()))
-
-def calc_trans_prob(transition_scores: torch.Tensor) -> float:
-    return np.exp(transition_scores.numpy())
-
-def table_trans_prob(generated_tokens: torch.Tensor , transition_scores: torch.Tensor) -> pd.DataFrame : 
-    # for beam search, loops over all generated texts
-    l_df = list()
-    for i in range(len(generated_tokens)):
-        # create the dataframe and append to list
-        l_df.append(pd.DataFrame({'token': generated_tokens[i].numpy(),
-                        'trans_scores': transition_scores[i].numpy()}))
-        # create other columns
-        l_df[i]['token_str'] = l_df[i]['token'].apply(lambda x: tokenizer.decode(x))
-        l_df[i]['trans_prob'] = l_df[i]['trans_scores'].apply(lambda x: np.exp(x))
-        l_df[i]['generation'] = i+1
-        l_df[i]['joint'] = calc_joint_prob(transition_scores[i])
-    
-    return pd.concat(l_df)
-
-def passthrough(input_embedding, fn_model, model):
-    outputs = fn_model(input_embedding)
-    
-    try:
-        transition_scores = model.compute_transition_scores(
-            outputs.sequences, outputs.scores, outputs.beam_indices, normalize_logits=True
-        )
-    except AttributeError:
-        #print("\nNo beam search!\n")
-        transition_scores = model.compute_transition_scores(
-            outputs.sequences, outputs.scores, normalize_logits=True
-        )
-    
-    return calc_trans_prob(transition_scores[0]), outputs
-        
-# MAIN --------------------------
-if __name__ == "__main__":
-    
-
-    # MAGICKS
-    i_seed_value = 42
-    
-    s_string = "Chicken and the"
-    s_model = 'gpt2'
-    
-    i_num_beams = 1
-    i_new_tokens = 5
-    
-    # BUG: Does this seed work?
[…]

### 2024-02-21 src/llm/training.py

@@ -1,185 +1,0 @@
-# SETUP --------------------------
-import pandas as pd
-import numpy as np
-import matplotlib.pyplot as plt
-from tqdm import tqdm
-
-import torch
-import torch.autograd as autograd
-from torchviz import make_dot
-from transformers import AutoTokenizer, GPT2LMHeadModel
-
-from undecorated import undecorated
-from types import MethodType
-
-# FUNCTIONS --------------------------
-
-def llikelihood(output: tuple, target_sequence: torch.Tensor) -> float:
-    """Calculate the log-likelihood of the target tokens in the output sequence.
-    NOTE: This only works for greedy search, not beam search (i.e. set num_beams=1).
-
-    Args:
-        output (tuple): Output from a HF LLM text generation model. Only pass one sequence.
-        target_sequence (torch.Tensor): The target tokens (the sequence) to calculate the log-likelihood of.
-
-    Returns:
-        float: log-llikelihood for the sequence
-    """
-    
-    # get scores and the sequences
-    #scores = autograd.Variable(torch.stack(output.scores).reshape(len(output.scores), -1).transpose(0, 1), requires_grad=True)
-    scores = torch.stack(output.scores).reshape(len(output.scores), -1).transpose(0, 1)
-    
-    # get the scores into a matrix, use log_softmax
-    #m_prob = autograd.Variable(torch.nn.functional.log_softmax(scores, dim=0), requires_grad=True)
-    m_prob = torch.nn.functional.log_softmax(scores, dim=0)
-    #m_prob_sub = autograd.Variable(m_prob[target_sequence.squeeze()], requires_grad=True)
-    m_prob_sub = m_prob[target_sequence.squeeze()]
-    
-    # get the diagonal of the matrix 
-    #v_trans_prob = autograd.Variable(torch.diag(m_prob_sub), requires_grad=True)
-    v_trans_prob = torch.diag(m_prob_sub)
-    #d_llikelihood = autograd.Variable(torch.sum(v_trans_prob) * (-1), requires_grad=True)
-    # NOTE: we use a minimizer
-    d_llikelihood = torch.sum(v_trans_prob) * (-1)
-    
-    # return the log-likelihood of the target tokens and the diagonal
-    #return m_prob, v_trans_prob, d_llikelihood
-    return d_llikelihood
-
-
-# TODO: Criterion
-
-# MAGICKS --------------------------
-# add other "special" backends
-b_use_backend = False
-i_seed_value = 42
-s_model = 'gpt2'
-
[…]

### 2024-02-21 beamer_preamble.tex
another commit

\usepackage{appendixnumberbeamer}
\usepackage{graphicx}
\usepackage{url}
\usetheme[progressbar=frametitle, block=fill, numbering=fraction, sectionpage=progressbar]{metropolis}
\newcommand{\themename}{\textbf{\textsc{metropolis}}\xspace}
\usepackage{xcolor}

\definecolor{UBCblue}{RGB}{106, 52, 107} % UBC Blue (primary)
\definecolor{UBCgrey}{rgb}{0.3686, 0.5255, 0.6235} % UBC Grey (secondary)

\setbeamercolor{palette primary}{bg=UBCblue,fg=white}
\setbeamercolor{palette secondary}{bg=UBCblue,fg=white}
\setbeamercolor{palette tertiary}{bg=UBCblue,fg=white}
\setbeamercolor{palette quaternary}{bg=UBCblue,fg=white}
\setbeamercolor{structure}{fg=UBCblue} % itemize, enumerate, etc
\setbeamercolor{section in toc}{fg=UBCblue} % TOC sections

\setbeamercolor{progress bar}{bg=gray,fg=UBCblue} 
\setbeamercolor{title separator}{bg=gray,fg=UBCblue}
\setbeamercolor{progress bar in head/foot}{bg=gray,fg=UBCblue} 
\setbeamercolor{progress bar in section page}{bg=gray,fg=orange}
\setbeamercolor{background canvas}{bg=white}

% Override palette coloring with secondary
\setbeamercolor{subsection in head/foot}{bg=UBCgrey,fg=white}

### 2024-02-21 meetings/slides_SKIM.qmd
another commit

---
title: "Learning about consumer preferences for media through Generative AI"
author: "Bas Donkers, Dennis Fok, and Finn-Ole Höner"
format: 
    beamer: default
    revealjs: default
    pptx:
        reference-doc: template.pptx
        slide-level: 1

aspectratio: 169
---

# The marketing problem

Serving media that resonates with people through advertising, branding, or in social media posts.

Our running example here are slogans, but these could be other types of media as well.

For this, we need to learn different consumers' preferences for these slogans

When we try to learn about product preferences, there are some aspects which marketers can easily tweak, e.g. the price, brand, or quantity of a product.

For media or brand elements, this is more difficult, as it is often unclear how we can alter e.g. a slogan to make it more appealing to a specific consumer.

Currently, marketers either rely on theoretical motivations, e.g. heuristics of a "brand personality", or they can perform a conjoint analysis on a set of specific material to learn about consumers' preferences for these slogans.

# Shortcoming of the current methods

Heuristic approaches are highly subjective. One drawback of this is, that the company strongly depends on its marketer. If the marketer leaves, the company loses its expertise. Heuristics can also be difficult to scale up.

Choice-study based approaches are more objective and better replicable. However, in the case of media, they are limited to the material that is part of the study and it is unclear how these findings generalize. Essentially, each variation in a campaign requires a new study. And what if the ideal slogan is not part of the study set?

## Idea: Can we learn about *how* to create these slogans instead?

This would allow us to create new slogans, where we already have an idea on which consumers might like it.

Our idea is, to get a Large Language Model to re-create a specific Slogan on which we already know the consumer's preferences. We are trying to "reverse-engineer" what creates a slogan that is appealing to a specific consumer.

# What this brings to SKIM

Currently, SKIM offers different forms of Conjoint Analysis, e.g. Choice-Based, Menu-Based, or Adaptive Choice-Based Conjoint Analysis. All of these can provide insights on consumers' preferences for a specific slogan, but they provide little help in creating new, "good", slogans.

This could be a first-step towards a new form of consumer preference studies, that directly helps you to create new taglines for your customers, or to transfer learnings from one brand to another.

## What we ask from you

For a test-run on whether this idea works, we would need a dataset which contains a choice based conjoint analysis on (short) text-media, e.g. slogans, taglines, or tweets. This would allow us to learn about a relationship between consumer's preferences and the "generation space" of these text-media.

<!-- 
# Generative AI and Preferences

- Generative personalization based on consumer-preferences
- Provide a continuous representation of consumer preferences, that allows e.g. personalized communication
- Performing a conjoint study on slogans, and mapping these slogans into a representation that an LLM can use to generate new slogans in accordance with consumers preferences
- Deterministic summary for a given text based on an LLM

*First progress on this idea, working on summary embedding for LLMs in MPhil*

Can make these space's dimensions "interpretable", about established branding concepts [@sisodiaAutomaticallyDiscoveringVisual2022]

Maybe, this "reverse engineering" of a generative model yields interesting insights in other fields as well. E.g. in ideation, maybe we can use these "deterministic" summaries to identify new ideas or to validate ideas.
 -->

### 2024-02-21 src/llm/llm.py

@@ -1,0 +1,127 @@
+# SETUP --------------------------
+import torch
+from transformers import AutoTokenizer, GPT2LMHeadModel
+import pandas as pd
+import numpy as np
+
+# MAGICKS --------------------------
+s_model = 'gpt2'
+s_string = "Finn writes code"
+
+# FUNCTIONS --------------------------
+def calc_joint_prob(transition_scores: torch.Tensor) -> float:
+    # calculate the joint probability of a generated text
+    return np.exp(np.sum(transition_scores.numpy()))
+
+def calc_trans_prob(transition_scores: torch.Tensor) -> float:
+    return np.exp(transition_scores.numpy())
+
+def table_trans_prob(generated_tokens: torch.Tensor , transition_scores: torch.Tensor) -> pd.DataFrame : 
+    # for beam search, loops over all generated texts
+    l_df = list()
+    for i in range(len(generated_tokens)):
+        # create the dataframe and append to list
+        l_df.append(pd.DataFrame({'token': generated_tokens[i].numpy(),
+                        'trans_scores': transition_scores[i].numpy()}))
+        # create other columns
+        l_df[i]['token_str'] = l_df[i]['token'].apply(lambda x: tokenizer.decode(x))
+        l_df[i]['trans_prob'] = l_df[i]['trans_scores'].apply(lambda x: np.exp(x))
+        l_df[i]['generation'] = i+1
+        l_df[i]['joint'] = calc_joint_prob(transition_scores[i])
+    
+    return pd.concat(l_df)
+
+def passthrough(input_embedding, fn_model, model):
+    outputs = fn_model(input_embedding)
+    
+    try:
+        transition_scores = model.compute_transition_scores(
+            outputs.sequences, outputs.scores, outputs.beam_indices, normalize_logits=True
+        )
+    except AttributeError:
+        #print("\nNo beam search!\n")
+        transition_scores = model.compute_transition_scores(
+            outputs.sequences, outputs.scores, normalize_logits=True
+        )
+    
+    return calc_trans_prob(transition_scores[0]), outputs
+        
+# MAIN --------------------------
+if __name__ == "__main__":
+    
+
+    # MAGICKS
+    i_seed_value = 42
+    
+    s_string = "Chicken and the"
+    s_model = 'gpt2'
+    
+    i_num_beams = 1
+    i_new_tokens = 5
+    
+    # BUG: Does this seed work?
[…]

### 2024-02-21 src/llm/sketch.py

@@ -1,0 +1,64 @@
+# SETUP --------------------------
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+from tqdm import tqdm
+
+import torch
+import torch.autograd as autograd
+#from torchviz import make_dot
+from transformers import AutoTokenizer, GPT2LMHeadModel
+
+from undecorated import undecorated
+from types import MethodType
+
+# FUNCTIONS --------------------------
+import torch
+
+b_use_backend = True
+i_seed_value = 42
+s_model = 'gpt2'
+#s_model = 'GroNLP/gpt2-small-dutch'
+
+# INIT
+if b_use_backend:
+    if torch.backends.mps.is_available():
+        mps_device = torch.device("mps")
+        x = torch.ones(1, device=mps_device)
+        torch.set_default_device('mps')
+        print (f"Using MPS backend {x}")
+    elif torch.cuda.is_available():
+        print ("Using CUDA backend")
+        torch.set_default_device('cuda')
+    else:
+        print ("Using CPU.")
+
+#s_prompt_string = "A hot"
+#s_target_string = "Potato"
+# s_target_string = "easy example." # text generated by this prompt for this seed!
+# s_target_string = "does it make sense to compare apples and oranges?" # text generated by this prompt for this seed!
+s_target_string = "one, two," # text generated by this prompt for this seed!
+# s_target_string = "Transform Your Hair's Health with our Revolutionary Haircare System: Experience Unparalleled Shine, Strength, and Vitality!" # text generated by this prompt for this seed!
+# #s_target_string = "Heeft het zin om appels en sinaasappelen te vergelijken?" # text generated by this prompt for this seed!
+
+torch.manual_seed(i_seed_value)
+model = GPT2LMHeadModel.from_pretrained(s_model)
+tokenizer = AutoTokenizer.from_pretrained(s_model)
+tokenizer.pad_token_id = tokenizer.eos_token_id
+
+# BUG: Issue that greedy search does not find max likelihood sequence? Does this matter?
+i_num_beams = 1 # greedy search for i_num_beams = 1
+i_no_repeat_ngram_size = 1 # TODO: This creates inplace gradient error
+i_num_return = 1
+i_max_new_tokens = tokenizer.encode(s_target_string, return_tensors="pt").shape[1]
+
+inputs = tokenizer(s_target_string, return_tensors="pt")
[…]

### 2024-02-21 src/llm/training.py

@@ -1,0 +1,343 @@
+# SETUP --------------------------
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+from tqdm import tqdm
+
+import torch
+import torch.autograd as autograd
+#from torchviz import make_dot
+from transformers import AutoTokenizer, GPT2LMHeadModel
+
+from undecorated import undecorated
+from types import MethodType
+import time
+from concurrent.futures import ThreadPoolExecutor, ProcessPoolExecutor
+
+# FUNCTIONS --------------------------
+
+def llikelihood(output: tuple, target_sequence: torch.Tensor) -> float:
+    """Calculate the log-likelihood of the target tokens in the output sequence.
+    NOTE: This only works for greedy search, not beam search (i.e. set num_beams=1).
+
+    Args:
+        output (tuple): Output from a HF LLM text generation model. Only pass one sequence.
+        target_sequence (torch.Tensor): The target tokens (the sequence) to calculate the log-likelihood of.
+
+    Returns:
+        float: log-llikelihood for the sequence
+    """
+
+    # get scores and the sequences
+    #print("raw scores comparison")
+    #print(output.scores[0])
+    #print(output.scores[1])
+    #print("len tuple")
+    #print(len(output.scores))
+
+    #scores = autograd.Variable(torch.stack(output.scores).reshape(len(output.scores), -1).transpose(0, 1), requires_grad=True)
+    scores = torch.stack(output.scores).reshape(len(output.scores), -1).transpose(0, 1)
+
+    #print("scores shape")
+    #print(scores.shape)
+
+    # get the scores into a matrix, use log_softmax
+    #m_prob = autograd.Variable(torch.nn.functional.log_softmax(scores, dim=0), requires_grad=True)
+    m_prob = torch.nn.functional.log_softmax(scores, dim=0)
+    #m_prob_sub = autograd.Variable(m_prob[target_sequence.squeeze()], requires_grad=True)
+    m_prob_sub = m_prob[target_sequence.squeeze()]
+
+    # get the diagonal of the matrix
+    #v_trans_prob = autograd.Variable(torch.diag(m_prob_sub), requires_grad=True)
+    v_trans_prob = torch.diag(m_prob_sub)
+    #d_llikelihood = autograd.Variable(torch.sum(v_trans_prob) * (-1), requires_grad=True)
+    # NOTE: we use a minimizer
[…]

### 2024-02-21 src/llm/training.py
todos

@@ -217,6 +217,9 @@ def training_parallel(target_sequence, summary_embedding, tokenizer, gd, fn_gene
             
         return summary_embedding
 
+# TODO: Implement reverse training, see forward expanding generation, but start from last token instead
+# TODO: Compare the times of threading inside the function vs multiprocessing on different observations
+
 # MAIN --------------------------
 if __name__ == "__main__":
     # MAGICKS --------------------------

### 2024-02-21 writing/draft_grant_proposal.qmd
writing research grant

---
format: docx
pdf-engine: pdflatex
header-includes:
  \usepackage{fancyhdr}
  \usepackage{fancyvrb}
  \usepackage{float}
  \usepackage{booktabs}
  \usepackage{lscape}
  \usepackage{amsmath}
  \usepackage{graphicx}
  \usepackage{algorithm}
  \usepackage{algpseudocode}
  \DeclareMathOperator*{\argmin}{arg\,min}
  \DeclareMathOperator*{\argmax}{arg\,max}
---

# Introduction

Difference of written summary and embedding summary.

Learning summaries of text is a fundamental problem in natural language processing, and of special interest in the context of Large Language Models (LLMs). While current approaches aim at summarizing text with a written summary that is shorter than the original text, we propose a novel approach, which reverse engineers these summaries. Rather than prompting a LLM to write a summary of a given text, we perform an optimization to find the input embedding to this LLM, which maximizes the likelihood for the LLM to generate the given text.

We aim to link these summary embeddings to consumer preferences, and plan to use these data to inform the generation of new claims for specific contexts.

One way how such a generation could work, is by averaging the embeddings of a set of claims, and then using this average embedding to generate a new claim.

<!-- TODO: Double check whether this is the right bandit -->
We could use Combinatorial Bandits to learn about how to create linear combinations of existing embeddings, such that the generated new claim is more appealing to the consumer.

A more sophisticated approach would lie in the use of Multi-Armed Bandits, where we set the arms to be certain elements of the embedding vector. By creating a stack of MAB and LLM, we could give the MAB the ability to generate new claims and to learn about the preferences of the consumer for these new claims. In this, our summary embeddings play a key role, as they allow us to generate new claims in a deterministic fashion, which in-turn enables us to learn about the preferences of the consumer for these new claims.

Another empirical applications could lie in learning about consumer preferences for headlines of news articles or posts, and to use these data to inform the generation of new headlines.

There are ample use cases and possible empirical investigations in a marketing context for these summary embeddings. In the following, we want to highlight three different applications, that might deliver novel insights due to the use of these summary embeddings.

<!-- TODO: double checks -->
First, we want to research how marketers design choices on the creation of product claims relates to consumer preferences. For example, a marketer might decide between making a more factual or a more emotional claim. While there are theoretical motivations for these types of decisions, it would allow us to quantitatively assess these styles of claims and to learn how consumer preferences relate to them. An extension of this research could lie in the relation of human marketers and the help of AI, namely LLMs, in the creation of these claims. Investigating how AI and human generated claims that are supposed to fulfill a certain characteristic, such as being more emotional, are actually perceived as such by consumers, could deliver novel insights.

Second, in a more field-based setting, we want to investigate how we can use these summary embeddings to adapt online descriptions to the preferences of the consumer. While there are techniques based on Reinforcement Learning, which is used to adapt the website to the preferences of the consumer, already in use, these types of algorithms cannot make adaptations "out-of-sample", i.e. they cannot generate new content that is not already predefined. We believe that the use of summary embeddings could allow us to generate adaptive text-based content, such as product descriptions or article headlines, that are not part of a predefined set.

Third,

<!-- TODO: True? Also add more examples -->
In preliminary tests, we have found that these summary embeddings appear to be able to differentiate different styles of claims in the same domain.

# Methods

- Novel, “reverse engineering” approach to finding text summaries (find a better word… in an information sense, not in a writing short summary sense). Given a target sequence and a given LLM, we want to find the input embedding to this LLM, which maximizes the likelihood for the LLM to generate the target sequence.

- Training can be done with simply forward passes through the LLM model

## Problem Formulation

For a given Large Language Model, $\mathcal{M}$, we want to find the input embedding, $\boldsymbol{s^{*}}$, that maximizes the likelihood of generating a given target sequence, $\lbrace t_i \rbrace_{i=1}^{L}$, where $L$ denotes the length of the target sequence and $t_i$ represents token $i$.

<!-- TODO: What does it really return? -->

To express that we generate tokens with the LLM, based on length $E$ input-embedding $\boldsymbol{e}$, we use the notation $\mathcal{M}(\boldsymbol{e})$. To express the number of tokens, $k$ that we generate, we use the subscript $\mathcal{M}_k(\boldsymbol{e})$. This function call returns an output vector of length $V$, where $V$ denotes the vocabulary size. We denote the output vector as $\boldsymbol{o}$. Each element $o_v$ of $\boldsymbol{o}$ represents the probability of generating token $v$ as the next token. We predict the next token, by selecting the largest element of $\boldsymbol{o}$. Explicitly, $t_{1} = \argmax \mathcal{M}_1(\boldsymbol{e})$.

When $k = 1$, we use the previous tokens to predict the next token. When $k > 1$, we append the predicted tokens autoregressively to the input embedding, such that we can predict more than one token. While there are many ways to design this procedure and select tokens along the way, we perform a standard greedy procedure, where we select the token with the highest probability at each step.

We denote the probability that $\mathcal{M}$ generates the target sequence, given the input embedding, as $p(t_1, \ldots, t_L | \boldsymbol{s})$.

We estimate these embeddings, by maximizing the conditional likelihood that this embedding generates the target sequence for a given LLM. We use a gradient-based optimization algorithm to maximize this likelihood.

These embeddings have the same dimension as the input embedding of $\mathcal{M}$, length $E$.

$$
\boldsymbol{s^{*}} = \argmax_{\boldsymbol{s}} \prod_{i=1}^{L} p(t_1, \ldots, t_L | \boldsymbol{s})
$$

$$
\boldsymbol{s^{*}} = \argmax_{\boldsymbol{s}} \sum_{i=1}^{L} \log p(t_1, \ldots, t_L | \boldsymbol{s})
$$

$$
\boldsymbol{s^{*}} = \argmax_{\boldsymbol{s}} \log p(t_1, | \boldsymbol{s}) + \log p(t_2, | \boldsymbol{s}, t_1) + \ldots + \log p(t_L, | \boldsymbol{s}, t_1, \ldots, t_{L-1}) 
$$

$$
\boldsymbol{s^{*}} = \argmax_{\boldsymbol{s}} \log \mathcal{M}_1(\boldsymbol{s}) + \log \mathcal{M}_1(\left [\boldsymbol{s}, \boldsymbol{e}_1 \right]) + \ldots + \log \mathcal{M}_1(\left [\boldsymbol{s}, \boldsymbol{e}_1, \ldots, \boldsymbol{e}_{L-1} \right])
$$

$$
\boldsymbol{s^{*}} = \argmax_{\boldsymbol{s}} \sum_{i=1}^{L} \log \mathcal{M}_L (\boldsymbol{s})
$$

We can make this expression more explicit, by using the LLM in our notation. For example $p(t_1, | \boldsymbol{s}) = \mathcal{M}(\boldsymbol{s})$

In the training process, we execute the following algorithm. First, we initialize the input embedding, $\boldsymbol{s}$. Second, we generate a sequence of length $L$, with the LLM and the current input embedding.

\begin{algorithm}
\begin{algorithmic}
\State Initialize $\boldsymbol{s}$
\State $i \gets 0$
\While{$Convergence == False$}
    \State $i \gets i + 1$
    \State $\boldsymbol{o}^{(i)} \gets \mathcal{M}(\boldsymbol{s})$
    \State $l^{(i)} \gets \operatorname{Likelihood}(\boldsymbol{o}^{(i)}, \lbrace t_i \rbrace_{1}^{L})$
    \State $\boldsymbol{s}^{(i + 1)} \gets \operatorname{GradientDecent}(\boldsymbol{s}^{(i)}, l^{(i)})$
    \State $Convergence \gets \operatorname{CheckConvergence}(\boldsymbol{s}^{(i + 1)}, \boldsymbol{s}^{(i)})$
\EndWhile
\end{algorithmic}
\end{algorithm}

### 2024-02-21 src/llm/sketch.py

@@ -21,44 +21,49 @@ s_model = 'gpt2'
 #s_model = 'GroNLP/gpt2-small-dutch'
 
 # INIT
-if b_use_backend:
-    if torch.backends.mps.is_available():
-        mps_device = torch.device("mps")
-        x = torch.ones(1, device=mps_device)
-        torch.set_default_device('mps')
-        print (f"Using MPS backend {x}")
-    elif torch.cuda.is_available():
-        print ("Using CUDA backend")
-        torch.set_default_device('cuda')
-    else:
-        print ("Using CPU.")
+# if b_use_backend:
+#     if torch.backends.mps.is_available():
+#         mps_device = torch.device("mps")
+#         x = torch.ones(1, device=mps_device)
+#         torch.set_default_device('mps')
+#         print (f"Using MPS backend {x}")
+#     elif torch.cuda.is_available():
+#         print ("Using CUDA backend")
+#         torch.set_default_device('cuda')
+#     else:
+#         print ("Using CPU.")
 
-#s_prompt_string = "A hot"
-#s_target_string = "Potato"
-# s_target_string = "easy example." # text generated by this prompt for this seed!
-# s_target_string = "does it make sense to compare apples and oranges?" # text generated by this prompt for this seed!
-s_target_string = "one, two," # text generated by this prompt for this seed!
-# s_target_string = "Transform Your Hair's Health with our Revolutionary Haircare System: Experience Unparalleled Shine, Strength, and Vitality!" # text generated by this prompt for this seed!
-# #s_target_string = "Heeft het zin om appels en sinaasappelen te vergelijken?" # text generated by this prompt for this seed!
+# #s_prompt_string = "A hot"
+# #s_target_string = "Potato"
+# # s_target_string = "easy example." # text generated by this prompt for this seed!
+# # s_target_string = "does it make sense to compare apples and oranges?" # text generated by this prompt for this seed!
+# s_target_string = "one, two," # text generated by this prompt for this seed!
+# # s_target_string = "Transform Your Hair's Health with our Revolutionary Haircare System: Experience Unparalleled Shine, Strength, and Vitality!" # text generated by this prompt for this seed!
[…]

### 2024-02-21 src/llm/training.py

@@ -216,6 +216,64 @@ def training_parallel(target_sequence, summary_embedding, tokenizer, gd, fn_gene
                 break
             
         return summary_embedding
+    
+
+def training_fwinduction(target_sequence, summary_embedding, tokenizer, gd, fn_generate_i, fn_generate_full, model):
+    print(summary_embedding.shape)
+    # triangle shape
+    l_target_sequence = list()
+    l_target_sequence_embeddings = list()
+    for i, target in enumerate(target_sequence[0]):
+        l_target_sequence.append(target_sequence[0][0:(i+1)].reshape(1, -1))
+        l_target_sequence_embeddings.append(model.get_input_embeddings()(l_target_sequence[i]))
+
+    summary_embedding_start = torch.clone(summary_embedding) # save it
+    output_start = fn_generate_full(summary_embedding_start)
+
+    print("START\n")
+    print(summary_embedding_start)
+    print(tokenizer.decode(output_start.sequences[0]))
+    print(output_start.sequences[0])
+
+    ll = llikelihood(output_start, target_sequence)
+    
+    for j in range(len(l_target_sequence)):
+        print(j)
+        print(ll)
+        history_gd = []
+
+        for i in tqdm(range(10000)):
+            if i+1 == 1:
+                print("\nFirst iteration (before any training)\n")
+                print(tokenizer.decode(fn_generate_full(summary_embedding).sequences[0]))
+
+            gd.zero_grad()
+
+            # generate output for the current input
+            objective = llikelihood(fn_generate_i(summary_embedding, j + 1), target_sequence)
+            objective.backward()
+            gd.step()
+            history_gd.append(objective.item())
+
+            if (i > 1) and (i % 10 == 0):
+                #print(summary_embedding)
+                print("\n")
+                print(tokenizer.decode(fn_generate_full(summary_embedding).sequences[0]))
+                print(f"neg. LL: {history_gd[-1]}")
+                print(f"Delta: {np.abs(history_gd[-1] - history_gd[-2])}")
+                print("Mean abs gradient: ", torch.abs(summary_embedding.grad).mean().item())
+
+            if (i>1) and ((torch.abs(summary_embedding.grad).mean().item()) < .001):
+                #if (i>1) and (np.abs(history_gd[-1] - history_gd[-2]) < .01):
+                print("\nConvergence achieved in ", i+1, " iterations", " for ", j+1, " tokens")
[…]

### 2024-02-25 writing/draft_grant_proposal.qmd
more writing on Grant

@@
 ---
-format: docx
+format: pdf
 pdf-engine: pdflatex
@@
   \DeclareMathOperator*{\argmax}{arg\,max}
+bibliography: references.bib
 ---
@@
 
-Difference of written summary and embedding summary.
+<!--
+Many firms adopt personalization strategies to remain competitive in digital marketing. It is desirable to
+be able to learn and execute the personalization policy in real-time. This is especially true in a dynamic
+environment where the payoffs of strategies can change over time, or where even the pool of available
+actions can change. Advertising is a typical example of a such a dynamic setting. Ad effectiveness often
+diminishes over time: repeated display of the same ads may incur negative responses from consumers,
+known as wear out of advertising. Ad wear out effects have been empirically examined since the 1980s
+[1, 2], but it remains difficult to develop an optimal personalized advertising that takes into account non-
+stationary ad returns. Ad wear out in turn forces companies to frequently refresh the pool of ads they
+serve to their customers. This brings in new options of which the rewards still need to be learned and
+further increases the need for real-time personalization. As the pandemic accelerates the
+transformation from brick-and-mortar to online business for many companies, real-time personalization
+in dynamic environments can be the fuel for reshaping interactions between firms and consumers. Firms
+have been aware of the value of real-time personalization, but still lack real-time engines that handle
+non-stationary reward functions and arm pools.
+-->
+
+#### Business -> Challenge
+
+<!-- TODO: References to examples -->
 
 Learning summaries of text is a fundamental problem in natural language processing, and of special interest in the context of Large Language Models (LLMs).
-While current approaches aim at summarizing text with a written summary that is shorter than the original text, we propose a novel approach, which reverse engineers these summaries.
-Rather than prompting a LLM to write a summary of a given text, we perform an optimization to find the input embedding to this LLM, which maximizes the likelihood for the LLM to generate the given text.
+
+#### Why a challenge? -> Idea
+
+<!-- TODO: REFERENCES, what about SBERT -->
+<!-- TODO: "Deterministic" right word? Needs better structure here -->
+Current approaches that summarize text work in two different ways.
+The first one is to make use of pretrained embeddings, such as BERT or Word2VEC [@devlinBERTPretrainingDeep2018; @mikolov2013distributed], to apply them to the tokens in the text, and then aggregate these embeddings to a single vector [@shenBaselineNeedsMore2018].
+The second, more recent, approach is to use generative Large Language Models (LLM) to generate a summary of a text through prompt engineering.
+Here, we see a gap between these two approaches: While the former can deliver a deterministic numeric summary of the text, it is not clear how to use these summaries for the generation of new text.
+The latter, on the other hand, can generate new text by prompting for similar writing, but the summary itself is not deterministic (i.e.
+when re-running the summary prompt, we are not guaranteed the same summary).
+In our research we aim to bridge this gap by finding a summary embedding that is both deterministic and can be used to generate new text.
+
+Prompt engineering can deliver astonishing results and illustrates the flexibility of LLMs.
+However, these strategies are difficult to replicate, as they are prone to uncertainty from the sampling process of an LLM, as well as to changes in the model, which are often managed by third parties.
+For a more robust use in business and academic applications, we need to develop a more systematic approach to the use of LLMs.
+One prominent application of LLMs is the engineering of features for other downstream tasks.
+
+Our approach, also hinges on LLMs.
+While these prompt engineering approaches aim at summarizing text with a written summary that is shorter than the original text, we propose a novel approach, which reverse engineers these summaries.
+Rather than prompting the LLM to write a summary of a given text (the "target sequence"), we perform an optimization to find the input embedding to this LLM, which maximizes the likelihood for the LLM to generate the target sequence.
+
+There are ample use cases and possible empirical investigations for these summary embeddings.
+In the following, we want to highlight applications both from the perspective of AI development and a domain-specific application in marketing.
+
+This novel training framework can be an important stepping stone towards making generative language models more interpretable and safe.
+If we can reverse engineer the generation of a specific output sequence, we might be able to perform a sensitivity analysis.
+Similar to adversarial attacks in computer vision, we can investigate how small changes in the input embedding affect the output sequence.
+Additionally, there is a potential to learn about the meaning behind different dimensions of the input embedding, as we can investigate how different elements of the input embedding affect the generated text.
+An interesting open-source implementation of this could be to build uppon the [PyTorch Captum](https://github.com/pytorch/captum) framework, as especially the gradient based methods of this toolbox seem adquate for our problem.
+Notice that our approach is not limited to a specific domain or Language Model.
+While we aim to research these summary embeddings in a marketing context, we believe that the use of these embeddings can be extended to other domains in text data, such as pieces of source code, analysis of human feedback, or medical records.
+It will also be interesting to research, how these summary embedding can be brought to recent developments in the field of multimodal learning, where we aim to learn from both text and images, e.g.
+to analyze the alignment of text and image for fake news detection [e.g.
+@uppadaImageTextbasedMultimodal2023; @liuMultiModalFakeNews2023].
+
+<!-- TODO: double checks -->
+In the domain of marketing, we envision two projects with these summary embeddings.
+First, we want to research how marketers design-choices of product claims relates to consumer preferences.
+For example, a marketer might decide between making a more factual or a more emotional claim.
+While there are theoretical motivations for these types of decisions, it would allow us to quantitatively assess these styles of claims and to learn how consumer preferences relate to them.
+An extension of this research could lie in the relation of human marketers and the help of AI, namely LLMs, in the creation of these claims.
+Investigating how AI and human generated claims that are supposed to fulfill a certain characteristic, such as being more emotional, are actually perceived as such by consumers, could deliver novel insights.
+Second, in a more field-based setting, we want to investigate how we can use these summary embeddings to adapt online descriptions to the preferences of the consumer.
+While there are techniques based on Reinforcement Learning, which is used to adapt the website to users' tastes, already in use, these types of algorithms cannot make adaptations "out-of-sample", i.e.
+they cannot generate new content that is not already pre-defined.
+We believe that the use of summary embeddings could allow us to generate adaptive text-based content, such as product descriptions or article headlines, that are not part of a predefined set, allowing for a true form of personalization.
+
+#### Building blocks -> Approach
 
@@
 
-<!-- TODO: Double check whether this is the right bandit -->
-We could use Combinatorial Bandits to learn about how to create linear combinations of existing embeddings, such that the generated new claim is more appealing to the consumer.
-
 A more sophisticated approach would lie in the use of Multi-Armed Bandits, where we set the arms to be certain elements of the embedding vector.
@@
 
+<!-- TODO: Double check whether this is the right bandit -->
+We could use Combinatorial Bandits to learn about how to create linear combinations of existing embeddings, such that the generated new claim is more appealing to the consumer.
+
 Another empirical applications could lie in learning about consumer preferences for headlines of news articles or posts, and to use these data to inform the generation of new headlines.
 
-There are ample use cases and possible empirical investigations in a marketing context for these summary embeddings.
-In the following, we want to highlight three different applications, that might deliver novel insights due to the use of these summary embeddings.
+We believe this project fits the mission of Amazon Science's Foundation Model Development call for proposals.
+As it investigates a novel training-framework, which aims to find embeddings that make the LLM generate a pre-determined text, it fits into Theme 2: 'Reducing the sensitivity to tweaks to the input prompt'.
 
-<!-- TODO: double checks -->
-First, we want to research how marketers design choices on the creation of product claims relates to consumer preferences.
-For example, a marketer might decide between making a more factual or a more emotional claim.
-While there are theoretical motivations for these types of decisions, it would allow us to quantitatively assess these styles of claims and to learn how consumer preferences relate to them.
-An extension of this research could lie in the relation of human marketers and the help of AI, namely LLMs, in the creation of these claims.
-Investigating how AI and human generated claims that are supposed to fulfill a certain characteristic, such as being more emotional, are actually perceived as such by consumers, could deliver novel insights.
+## Prior Work
 
-Second, in a more field-based setting, we want to investigate how we can use these summary embeddings to adapt online descriptions to the preferences of the consumer.
-While there are techniques based on Reinforcement Learning, which is used to adapt the website to the preferences of the consumer, already in use, these types of algorithms cannot make adaptations "out-of-sample", i.e.
-they cannot generate new content that is not already predefined.
-We believe that the use of summary embeddings could allow us to generate adaptive text-based content, such as product descriptions or article headlines, that are not part of a predefined set.
-
-Third,
-
-<!-- TODO: True? Also add more examples -->
-In preliminary tests, we have found that these summary embeddings appear to be able to differentiate different styles of claims in the same domain.
+<!-- TODO: References -->
 
 # Methods
+
+#### Overview of the Method
 
@@
 
-For a given Large Language Model, $\mathcal{M}$, we want to find the input embedding, $\boldsymbol{s^{*}}$, that maximizes the likelihood of generating a given target sequence, $\lbrace t_i \rbrace_{i=1}^{L}$, where $L$ denotes the length of the target sequence and $t_i$ represents token $i$.
-
 <!-- TODO: What does it really return? -->
 
-To express that we generate tokens with the LLM, based on length $E$ input-embedding $\boldsymbol{e}$, we use the notation $\mathcal{M}(\boldsymbol{e})$.
-To express the number of tokens, $k$ that we generate, we use the subscript $\mathcal{M}_k(\boldsymbol{e})$.
+<!-- TODO: Good idea or not? -->
+To express that we generate tokens with the LLM, $\mathcal{M}$, based on length $E$ input-embedding $\boldsymbol{e}$, we use the notation $\mathcal{M}(\boldsymbol{e})$.
+To express the number of tokens, $k$, that we generate, we use the subscript $\mathcal{M}_k(\boldsymbol{e})$.
 This function call returns an output vector of length $V$, where $V$ denotes the vocabulary size.
 We denote the output vector as $\boldsymbol{o}$.
-Each element $o_v$ of $\boldsymbol{o}$ represents the probability of generating token $v$ as the next token.
+Each element $o_v$ of this output vector represents the probability of generating token $v$ as the next token.
 We predict the next token, by selecting the largest element of $\boldsymbol{o}$.
-Explicitly, $t_{1} = \argmax \mathcal{M}_1(\boldsymbol{e})$.
-
 When $k = 1$, we use the previous tokens to predict the next token.
@@
 While there are many ways to design this procedure and select tokens along the way, we perform a standard greedy procedure, where we select the token with the highest probability at each step.
+We denote the probability that $\mathcal{M}$ generates the target sequence, given the input embedding, as $p(t_1, \ldots, t_L | \boldsymbol{e})$.
+For a given Large Language Model, we define $\boldsymbol{e^{*}}$,
 
-We denote the probability that $\mathcal{M}$ generates the target sequence, given the input embedding, as $p(t_1, \ldots, t_L | \boldsymbol{s})$.
+$$
+\boldsymbol{e^{*}} = \argmax_{\boldsymbol{e}} \sum_{i=1}^{L} \log p(t_1, \ldots, t_L | \boldsymbol{e}) \text{, }
+$$
 
-We estimate these embeddings, by maximizing the conditional likelihood that this embedding generates the target sequence for a given LLM.
+as the input-embedding that maximizes the likelihood of generating a given target sequence, $\lbrace t_i \rbrace_{i=1}^{L}$, where $L$ denotes the length of the target sequence and $t_i$ represents token $i$.
+We estimate these summary embeddings $\boldsymbol{e^{*}}$, by maximizing the conditional likelihood that this embedding generates the target sequence for a given LLM.
 We use a gradient-based optimization algorithm to maximize this likelihood.
 
-These embeddings have the same dimension as the input embedding of $\mathcal{M}$, length $E$.
+<!-- 
+There are options to make this optimization more efficient, by decomposing this likelihood in a product of conditional likelihoods,
+$$
+\boldsymbol{e^{*}} = \argmax_{\boldsymbol{e}} \log p(t_1, | \boldsymbol{e}) + \log p(t_2, | \boldsymbol{e}, t_1) + \ldots + \log p(t_L, | \boldsymbol{e}, t_1, \ldots, t_{L-1}) \text{,}
+$$
+
+and parallelizing the computation of these conditional likelihoods.
+-->
+
+<!--
 
 $$
-\boldsymbol{s^{*}} = \argmax_{\boldsymbol{s}} \prod_{i=1}^{L} p(t_1, \ldots, t_L | \boldsymbol{s})
+\boldsymbol{e^{*}} = \argmax_{\boldsymbol{e}} \prod_{i=1}^{L} p(t_1, \ldots, t_L | \boldsymbol{e})
 $$
@@
 $$
-\boldsymbol{s^{*}} = \argmax_{\boldsymbol{s}} \sum_{i=1}^{L} \log p(t_1, \ldots, t_L | \boldsymbol{s})
+\boldsymbol{e^{*}} = \argmax_{\boldsymbol{e}} \log p(t_1, | \boldsymbol{e}) + \log p(t_2, | \boldsymbol{e}, t_1) + \ldots + \log p(t_L, | \boldsymbol{e}, t_1, \ldots, t_{L-1}) 
 $$
@@
 $$
-\boldsymbol{s^{*}} = \argmax_{\boldsymbol{s}} \log p(t_1, | \boldsymbol{s}) + \log p(t_2, | \boldsymbol{s}, t_1) + \ldots + \log p(t_L, | \boldsymbol{s}, t_1, \ldots, t_{L-1}) 
+\boldsymbol{e^{*}} = \argmax_{\boldsymbol{e}} \log \mathcal{M}_1(\boldsymbol{e}) + \log \mathcal{M}_1(\left [\boldsymbol{e}, \boldsymbol{e}_1 \right]) + \ldots + \log \mathcal{M}_1(\left [\boldsymbol{e}, \boldsymbol{e}_1, \ldots, \boldsymbol{e}_{L-1} \right])
 $$
@@
 $$
-\boldsymbol{s^{*}} = \argmax_{\boldsymbol{s}} \log \mathcal{M}_1(\boldsymbol{s}) + \log \mathcal{M}_1(\left [\boldsymbol{s}, \boldsymbol{e}_1 \right]) + \ldots + \log \mathcal{M}_1(\left [\boldsymbol{s}, \boldsymbol{e}_1, \ldots, \boldsymbol{e}_{L-1} \right])
-$$
-
-$$
-\boldsymbol{s^{*}} = \argmax_{\boldsymbol{s}} \sum_{i=1}^{L} \log \mathcal{M}_L (\boldsymbol{s})
+\boldsymbol{e^{*}} = \argmax_{\boldsymbol{s}} \sum_{i=1}^{L} \log \mathcal{M}_L (\boldsymbol{s})
 $$
@@
 Second, we generate a sequence of length $L$, with the LLM and the current input embedding.
+-->
+
+Currently, we initialize the summary embedding as the element-wise average of the embedding of the target sequence.
+This speeds up the computation, compared to a randomly intialized embedding.
+We want to investigate further improvements to the implementation, such as computing the likelihood contributions of the tokens in a distributed fashion.
+In Algorithm 1, we summarize the training process.
 
@@
 \begin{algorithmic}
+\label{alg:training}
+\caption{Training of Summary Embeddings}
 \State Initialize $\boldsymbol{s}$
@@
 \end{algorithm}
+
+## Method in a Nutshell
+
+# Expected Results
+
+## Goal
+
+<!-- TODO: Right interpretation? -->
+
+Moving forward, we expect to find embeddings that can identify different writing styles across domains.
+We plan to use these embeddings to generate new text, for example by transfering a certain writing style from one domain to another, by investigating whether linear combinations of embeddings can be used to generate new, meaningful, text and how this new text relates to the weighted input embeddings.
+Since this is a novel framework for the training of summary embeddings, we plan to implement this framework as a open-source Python module which is adjusted to the API of the [HuggingFace Transformers library](https://huggingface.co/) library, to ensure ease of use and adoption by other researchers and practitioners, and to scale our research to a large set of LLMs.
+As mentioned above, we also want to investigate the use of this approach in LLM safety and interpretability, which could also lead to a second open-source implementation, connected to the [PyTorch Captum](https://github.com/pytorch/captum) module.
+
+## Status quo
+
+<!-- TODO: True? Also add more examples -->
+At the time of this application, we have implemented an intial version of the training process and have tested it on a small set of target sequences.
+We have found that the training process is able to find embeddings that make the LLM generate the target sequence.
+The speed of the optimization depends on the length of the target sequence, size of the LLM, and parameters of the optimization.
+It seems that even with small LLMs, like GPT-2, the found summary embeddings capture information about the target sequence.
+For this preliminary test we generated a selection of target sequences through GPT-3.5, prompting it to create advertising slogans for the same product, but with either tangible or more abstract claims.
+Upon investigating a small selection of the resulting embeddings, we find that they seem to capture inherent information of the target sequence, such as this specific writing style.
+When visualizing these embeddings, ignoring information on which embedding belong to the tangible and abstract groups, we find a separation between the two groups.
+On closer inspection, we also find that embeddings of similar target sequences are closer to each other in this embedding space.
+
+## Future work
+
+The next steps of this project have different angles.
+
+The first angle is to investigate and understand these summary embeddings better.
+How do they relate to each other?
+What types of concepts can we capture with these embeddings?
+To which degree are they transferable between different versions of the same LLM?
+How does the captured information compare to existing approaches, such as the use BERT embeddings?
+Can we learn concepts across domains, such as different writing techniques?
+A valuable extension here lies in the use of these embeddings together with context information.
+Can we incorporate information about the author of the target sequence, or the context in which the target sequence was created, into the training process?
+If so, how does that change the captured information in these embeddings?
+
+As a second angle, we want to investigate how we can use these embeddings to generate new text.
+A question here is whether linear combinations of these summary embeddings lead to the generation of new, meaningful, text.
+If this would be the case, then one possible avenue for the generation of new, targeted, text could lie in the use of Multi-Armed Bandits, where we learn how to linearly combine summary embeddings to maximize a reward function, e.g.
+generating product descriptions to maximize a key performance measure such as click-through rate.
+
+We plan the following timeline:
+
+# Funds needed
+
+- Current implementation needs around 3 minutes per claim, on Google Colab T4 GPU.
+- Training of embeddings in parallel
+- GPU is the limiting factor
+
+{{< pagebreak >}}
+
+# References
+
+<!-- TODO: good idea or bad idea? -->
+Since we are working within the real of LLMs, we aim to incorporate our training framework in the API of the HuggingFace Transformers library.
+This would ensure ease of use and adoption by other researchers and practitioners, and would allow us to scale our research to a larger set of LLMs.

### 2024-02-25 src/llm/training.py

@@ -372,7 +372,7 @@ if __name__ == "__main__":
 
     input_embedding1 = autograd.Variable(torch.mean(model.get_input_embeddings()(tokenizer.encode(s_target_string, return_tensors="pt")), 1, keepdim=True), requires_grad=True)
     gd = torch.optim.RMSprop([input_embedding1],
-                            lr=1e-1, alpha=0.99, eps=1e-08, weight_decay=0, momentum=0, centered=False, foreach=None, maximize=False, differentiable=False)
+                            lr=1, alpha=0.99, eps=1e-08, weight_decay=0, momentum=0, centered=False, foreach=None, maximize=False, differentiable=False)
 
     start_time = time.time()
     summary_embedding_1 = training_joint(target_sequence, input_embedding1, tokenizer, gd, fn_generate_full)
@@ -381,7 +381,7 @@ if __name__ == "__main__":
 
     input_embedding4 = autograd.Variable(torch.mean(model.get_input_embeddings()(tokenizer.encode(s_target_string, return_tensors="pt")), 1, keepdim=True), requires_grad=True)
     gd = torch.optim.RMSprop([input_embedding4],
-                            lr=1e-1, alpha=0.99, eps=1e-08, weight_decay=0, momentum=0, centered=False, foreach=None, maximize=False, differentiable=False)
+                            lr=1, alpha=0.99, eps=1e-08, weight_decay=0, momentum=0, centered=False, foreach=None, maximize=False, differentiable=False)
     
     start_time = time.time()
     summary_embedding_4, history_gd = training_fwinduction(target_sequence, input_embedding4, tokenizer, gd, fn_generate_i, fn_generate_full, model)
@@ -399,12 +399,12 @@ if __name__ == "__main__":
     end_time = time.time()
     execution_time_2 = end_time - start_time
     
-    input_embedding = autograd.Variable(torch.mean(model.get_input_embeddings()(tokenizer.encode(s_target_string, return_tensors="pt")), 1, keepdim=True), requires_grad=True)
+    input_embedding_3 = autograd.Variable(torch.mean(model.get_input_embeddings()(tokenizer.encode(s_target_string, return_tensors="pt")), 1, keepdim=True), requires_grad=True)
     gd = torch.optim.RMSprop([input_embedding],
                             lr=1e-1, alpha=0.99, eps=1e-08, weight_decay=0, momentum=0, centered=False, foreach=None, maximize=False, differentiable=False)
 
     start_time = time.time()
[…]

### 2024-03-12 meetings/meeting_notes.qmd
notes SURF workshop

@@
 <!-- TODO: Reaching out to SKIM, slide deck, NDA -->
+
+# 12-03-2024 ([Surf Research Cloud Workshop](https://www.surf.nl/en/agenda/introduction-to-surf-research-cloud-11))
+
+- [Available flavors](https://servicedesk.surf.nl/wiki/display/WIKI/SRC+Available+Flavours)
+- [Small Compute Applications (E-Infra grant)](https://www.surf.nl/en/small-compute-applications-nwo)
+  - [documentation](https://servicedesk.surf.nl/wiki/display/WIKI/Paying+for+usage)
+  - 1 year valid, can apply every *calendar* year
+  - up to 5,000 GPU hours
+  - can extend by 6 months if resources are left
+  - goes through SURF, can help with resources
+  - Estimate by iteration time, core hours + 20% overhead
+  - Need somebody who has permanent contract with Dutch institution
+  - hard *not* to get
+- NWO Groot grant is bigger, but slower and harder to get
+  - Can chain this together with the small compute applications
+- At EUR will create new policy on this next month
+- Sounds like there will be resources that your department can free up directly
+- RCCS contract
+- Likly much faster GPUs then e.g. Google Colab
+- Can also have budgets at other Cloud services and use research cloud for collaboration, e.g. Azure, AWS
+- Could we get resources for the NLP course? They are also dedicated to education, not only research
+
+## People
+
+- Pieter Meijndert, ESE, Head of Business Intelligence and Application Management
+- Yuliia Orlova, Liasion User-Developers
+- Carsten Schelp
+
+## Workshop
+
+- Pick collaboration and budget work a workspace, people from collaboration get access
+- e.g RStudio, Jupyter available
+- Disk budget will expire when grant expires, not an archive. Quite fast discs. Only upload data once and use from multiple instances.
+- Storage is persistent, data is not lost if VM crashes.
+- Attach the storage to the workspace
+- Set deadline for the machine, normal lifetime is 5 days
+- Shortlifed instances support reproduceability
+- RStudio "username": Profile tab of the portal; time-based password. For Jupyter its the portal credentials.
+  - Can also be SSH for some instances
+- Can probably connect through VSCode and SSH to instance directly
+- Whisper instance on cloud, useful for transcription?
+
+## Questions
+
+- What is a GPU core hour?

### 2024-03-14 docs/conf.py
bug fixes in visualize embeddings

@@ -1,244 +1,0 @@
-# -*- coding: utf-8 -*-
-#
-# mphil documentation build configuration file, created by
-# sphinx-quickstart.
-#
-# This file is execfile()d with the current directory set to its containing dir.
-#
-# Note that not all possible configuration values are present in this
-# autogenerated file.
-#
-# All configuration values have a default; values that are commented out
-# serve to show the default.
-
-import os
-import sys
-
-# If extensions (or modules to document with autodoc) are in another directory,
-# add these directories to sys.path here. If the directory is relative to the
-# documentation root, use os.path.abspath to make it absolute, like shown here.
-# sys.path.insert(0, os.path.abspath('.'))
-
-# -- General configuration -----------------------------------------------------
-
-# If your documentation needs a minimal Sphinx version, state it here.
-# needs_sphinx = '1.0'
-
-# Add any Sphinx extension module names here, as strings. They can be extensions
-# coming with Sphinx (named 'sphinx.ext.*') or your custom ones.
-extensions = []
-
-# Add any paths that contain templates here, relative to this directory.
-templates_path = ['_templates']
-
-# The suffix of source filenames.
-source_suffix = '.rst'
-
-# The encoding of source files.
-# source_encoding = 'utf-8-sig'
-
-# The master toctree document.
-master_doc = 'index'
-
-# General information about the project.
-project = u'mphil'
-
-# The version info for the project you're documenting, acts as replacement for
-# |version| and |release|, also used in various other places throughout the
-# built documents.
-#
-# The short X.Y version.
-version = '0.1'
-# The full version, including alpha/beta/rc tags.
-release = '0.1'
-
-# The language for content autogenerated by Sphinx. Refer to documentation
-# for a list of supported languages.
-# language = None
-
-# There are two options for replacing |today|: either, you set today to some
-# non-false value, then it is used:
-# today = ''
-# Else, today_fmt is used as the format for a strftime call.
-# today_fmt = '%B %d, %Y'
-
-# List of patterns, relative to source directory, that match files and
-# directories to ignore when looking for source files.
-exclude_patterns = ['_build']
-
-# The reST default role (used for this markup: `text`) to use for all documents.
-# default_role = None
-
-# If true, '()' will be appended to :func: etc. cross-reference text.
-# add_function_parentheses = True
-
-# If true, the current module name will be prepended to all description
[…]

### 2024-03-14 setup.py

@@ -1,10 +1,0 @@
-from setuptools import find_packages, setup
-
-setup(
-    name='src',
-    packages=find_packages(),
-    version='0.1.0',
-    description='MPhil thesis for the research master Business Data Science at the Tinbergen Institute',
-    author='Finn-Ole Hoener',
-    license='',
-)

### 2024-03-14 src/llm/sketch.py

@@ -1,69 +1,0 @@
-# SETUP --------------------------
-import pandas as pd
-import numpy as np
-import matplotlib.pyplot as plt
-from tqdm import tqdm
-
-import torch
-import torch.autograd as autograd
-#from torchviz import make_dot
-from transformers import AutoTokenizer, GPT2LMHeadModel
-
-from undecorated import undecorated
-from types import MethodType
-
-# FUNCTIONS --------------------------
-import torch
-
-b_use_backend = True
-i_seed_value = 42
-s_model = 'gpt2'
-#s_model = 'GroNLP/gpt2-small-dutch'
-
-# INIT
-# if b_use_backend:
-#     if torch.backends.mps.is_available():
-#         mps_device = torch.device("mps")
-#         x = torch.ones(1, device=mps_device)
-#         torch.set_default_device('mps')
-#         print (f"Using MPS backend {x}")
-#     elif torch.cuda.is_available():
-#         print ("Using CUDA backend")
-#         torch.set_default_device('cuda')
-#     else:
-#         print ("Using CPU.")
-
-# #s_prompt_string = "A hot"
-# #s_target_string = "Potato"
-# # s_target_string = "easy example." # text generated by this prompt for this seed!
-# # s_target_string = "does it make sense to compare apples and oranges?" # text generated by this prompt for this seed!
-# s_target_string = "one, two," # text generated by this prompt for this seed!
-# # s_target_string = "Transform Your Hair's Health with our Revolutionary Haircare System: Experience Unparalleled Shine, Strength, and Vitality!" # text generated by this prompt for this seed!
-# # #s_target_string = "Heeft het zin om appels en sinaasappelen te vergelijken?" # text generated by this prompt for this seed!
-
-# torch.manual_seed(i_seed_value)
-# model = GPT2LMHeadModel.from_pretrained(s_model)
-# tokenizer = AutoTokenizer.from_pretrained(s_model)
-# tokenizer.pad_token_id = tokenizer.eos_token_id
-
-# # BUG: Issue that greedy search does not find max likelihood sequence? Does this matter?
-# i_num_beams = 1 # greedy search for i_num_beams = 1
-# i_no_repeat_ngram_size = 1 # TODO: This creates inplace gradient error
-# i_num_return = 1
-# i_max_new_tokens = tokenizer.encode(s_target_string, return_tensors="pt").shape[1]
-
[…]

### 2024-03-14 src/llm/training.py

@@ -1,433 +1,0 @@
-# SETUP --------------------------
-import pandas as pd
-import numpy as np
-import matplotlib.pyplot as plt
-from tqdm import tqdm
-
-import torch
-import torch.autograd as autograd
-#from torchviz import make_dot
-from transformers import AutoTokenizer, GPT2LMHeadModel
-
-from undecorated import undecorated
-from types import MethodType
-import time
-from concurrent.futures import ThreadPoolExecutor, ProcessPoolExecutor
-
-# FUNCTIONS --------------------------
-
-def llikelihood(output: tuple, target_sequence: torch.Tensor) -> float:
-    """Calculate the log-likelihood of the target tokens in the output sequence.
-    NOTE: This only works for greedy search, not beam search (i.e. set num_beams=1).
-
-    Args:
-        output (tuple): Output from a HF LLM text generation model. Only pass one sequence.
-        target_sequence (torch.Tensor): The target tokens (the sequence) to calculate the log-likelihood of.
-
-    Returns:
-        float: log-llikelihood for the sequence
-    """
-
-    # get scores and the sequences
-    #print("raw scores comparison")
-    #print(output.scores[0])
-    #print(output.scores[1])
-    #print("len tuple")
-    #print(len(output.scores))
-
-    #scores = autograd.Variable(torch.stack(output.scores).reshape(len(output.scores), -1).transpose(0, 1), requires_grad=True)
-    scores = torch.stack(output.scores).reshape(len(output.scores), -1).transpose(0, 1)
-
-    #print("scores shape")
-    #print(scores.shape)
-
-    # get the scores into a matrix, use log_softmax
-    #m_prob = autograd.Variable(torch.nn.functional.log_softmax(scores, dim=0), requires_grad=True)
-    m_prob = torch.nn.functional.log_softmax(scores, dim=0)
-    #m_prob_sub = autograd.Variable(m_prob[target_sequence.squeeze()], requires_grad=True)
-    m_prob_sub = m_prob[target_sequence.squeeze()]
-
-    # get the diagonal of the matrix
-    #v_trans_prob = autograd.Variable(torch.diag(m_prob_sub), requires_grad=True)
-    v_trans_prob = torch.diag(m_prob_sub)
-    #d_llikelihood = autograd.Variable(torch.sum(v_trans_prob) * (-1), requires_grad=True)
-    # NOTE: we use a minimizer
[…]

### 2024-03-14 writing/draft_grant_proposal.qmd
sync

@@
 ---
-format: pdf
+format: docx
 pdf-engine: pdflatex
@@
 
-#### Business -> Challenge
-
 <!-- TODO: References to examples -->
@@
 Learning summaries of text is a fundamental problem in natural language processing, and of special interest in the context of Large Language Models (LLMs).
-
-#### Why a challenge? -> Idea
 
@@
 
-#### Building blocks -> Approach
-
 We aim to link these summary embeddings to consumer preferences, and plan to use these data to inform the generation of new claims for specific contexts.
@@
 
+To summarize the information contained in a piece of text, we want to highlight the work on word2vec [@mikolov2013distributed], GloVe [@penningtonGloVeGlobalVectors2014], BERT [@devlinBERTPretrainingDeep2018], and SBERT [@reimersSentenceBERTSentenceEmbeddings2019].
+The first two word representations are based on the idea that words derive the meaning from the company that they keep, while the BERT and SBERT embeddings are rooted in the use of Transformers [see e.g.
+@vaswani2017attention].
+Aggregating these word embeddings for sentences can serve as a way to capture the information content in a sentence [see @shenBaselineNeedsMore2018].
+With respect to the task of computing the similarity between sentences, @reimersSentenceBERTSentenceEmbeddings2019 show that their adaptation of the BERT algorithm, SBERT, outperforms other methods with respect to this task and does a good job at serving as a sentence embedding in classification tasks.
+
 <!-- TODO: References -->
@@
 # Methods
-
-#### Overview of the Method
-
-- Novel, “reverse engineering” approach to finding text summaries (find a better word… in an information sense, not in a writing short summary sense).
-Given a target sequence and a given LLM, we want to find the input embedding to this LLM, which maximizes the likelihood for the LLM to generate the target sequence.
-
-- Training can be done with simply forward passes through the LLM model
 
@@
 
-Currently, we initialize the summary embedding as the element-wise average of the embedding of the target sequence.
-This speeds up the computation, compared to a randomly intialized embedding.
+## Method in a Nutshell
+
+We aim to find a summary embedding, $\boldsymbol{e^{*}}$, that maximizes the likelihood of generating a given target sequence, $\lbrace t_i \rbrace_{i=1}^{L}$, with a given LLM, $\mathcal{M}$.
+We use a gradient-based optimization algorithm to maximize this likelihood.
+We initialize the summary embedding as the element-wise average of the embedding of the target sequence.
+We then generate a sequence of length $L$, with the LLM and the current input embedding.
+For the resulting output layer, we compute the likelihood that this layer will generate the target sequence.
+We backpropagate the gradients and adjust the input layer accordingly.
+We repeat this process until the optimization converges.
 We want to investigate further improvements to the implementation, such as computing the likelihood contributions of the tokens in a distributed fashion.
@@
 \end{algorithm}
-
-## Method in a Nutshell
 
@@
 
+<!-- TODO: TIMELINE -->
+
+# Novel hardware exploration
+
+Despite working on an efficient implementation of our algorithm and the ability to use pre-trained LLMs, we are currently constrained by the computational resources available to us, which limits the lengths of our feasible target sequences.
+While we can already train summary embeddings for target sequences of lengths that are useful for applications e.g.
+in marketing, summarizing full documents has prohibitively high computational costs.
+Being able to perform these calculations for full "target documents", could lead valuable insights for example when researching political polarization.
+If these summary embeddings truly pick up on nuances in the writing style, we could use them to investigate how different political groups and news outlets differ in the bias of their writing style.
+Since, we are interested in using these embeddings for the generation of new text, we might be able to use these deterministic summary to "de-bias" a piece of writing in a generative and deterministic fashion.
+These types of applications hinge on the availability of computational resources, and the use of the AWS Trainium2 chips could be a valuable asset in this regard.
+
+The higher computational efficiency of the Trainium2 chip also aligns with our view, that we need to be mindful of the environmental impact of our research.
+We try to keep this impact low by only making use of pre-trained LLMs and only having to perform forward-passes through these.
+However, we are aware that the repetitive
+
+For applications of these summary embeddings to massive datasets of millions of target tokens, a high degree of parallelization will be necessary to perform this in adequate time.
+While each separate optimization only requires a sequence of forward-passes through the LLM, running these optimizations in parallel will require a high degree of computational resources.
+We would be excited to investigate how such a training process could be tailored to the architecture of the Trainium2 hardware.
+
+Currently, we are working with small LLMs such as GPT-2, to keep our training times at bay and computational costs low.
+Naturally, we are interested in the potential of using state-of-the art LLMs instead, such as e.g.
+LLaMA2.
+Not only can these types of models generate more coherent and meaningful text, but their input embeddings are of higher dimensions than the current LLMs we are using.
+This could potentially lead to more expressive and more detailed summary embeddings.
+
+# Open-source intention
+
+This grant would enable us to contribute to the open-source community in two ways.
+The first way is by open-sourcing our code in a publicly hosted Python module.
+Ideally, this module would be adjusted to the API of the HuggingFace Transformers library.
+This would ensure ease of use and adoption by other researchers and practitioners, and would allow us to scale our research to a large set of LLMs.
+In the context of using these summary embeddings for researching LLM safety and interpretability, an extension to the PyTorch Captum module would prove to be valuable.
+The second way how we can contribute to the open-source community, is not through code, but by providing access to pre-trained summary embeddings.
+With sufficient computing resources, we would be able to train these summary embeddings for a large set of target sequences from a public dataset, and make these embeddings publicly available to the research community for benchmarking purposes.
+One example could be to provide summary embeddings for the [Microsoft News Dataset](https://msnews.github.io/#getting-start), which serves as a benchmark for news recommendation systems.
+
 # Funds needed

### 2024-03-14 docs/conf.py

@@ -1,0 +1,244 @@
+# -*- coding: utf-8 -*-
+#
+# mphil documentation build configuration file, created by
+# sphinx-quickstart.
+#
+# This file is execfile()d with the current directory set to its containing dir.
+#
+# Note that not all possible configuration values are present in this
+# autogenerated file.
+#
+# All configuration values have a default; values that are commented out
+# serve to show the default.
+
+import os
+import sys
+
+# If extensions (or modules to document with autodoc) are in another directory,
+# add these directories to sys.path here. If the directory is relative to the
+# documentation root, use os.path.abspath to make it absolute, like shown here.
+# sys.path.insert(0, os.path.abspath('.'))
+
+# -- General configuration -----------------------------------------------------
+
+# If your documentation needs a minimal Sphinx version, state it here.
+# needs_sphinx = '1.0'
+
+# Add any Sphinx extension module names here, as strings. They can be extensions
+# coming with Sphinx (named 'sphinx.ext.*') or your custom ones.
+extensions = []
+
+# Add any paths that contain templates here, relative to this directory.
+templates_path = ['_templates']
+
+# The suffix of source filenames.
+source_suffix = '.rst'
+
+# The encoding of source files.
+# source_encoding = 'utf-8-sig'
+
+# The master toctree document.
+master_doc = 'index'
+
+# General information about the project.
+project = u'mphil'
+
+# The version info for the project you're documenting, acts as replacement for
+# |version| and |release|, also used in various other places throughout the
+# built documents.
+#
+# The short X.Y version.
+version = '0.1'
+# The full version, including alpha/beta/rc tags.
+release = '0.1'
+
+# The language for content autogenerated by Sphinx. Refer to documentation
+# for a list of supported languages.
+# language = None
+
+# There are two options for replacing |today|: either, you set today to some
+# non-false value, then it is used:
+# today = ''
+# Else, today_fmt is used as the format for a strftime call.
+# today_fmt = '%B %d, %Y'
+
+# List of patterns, relative to source directory, that match files and
+# directories to ignore when looking for source files.
+exclude_patterns = ['_build']
+
+# The reST default role (used for this markup: `text`) to use for all documents.
+# default_role = None
+
+# If true, '()' will be appended to :func: etc. cross-reference text.
+# add_function_parentheses = True
+
+# If true, the current module name will be prepended to all description
[…]

### 2024-03-14 src/analysis/visualize_embeddings.py

@@ -1,0 +1,245 @@
+# SETUP --------------------------
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+import pickle
+import torch
+import io
+from sklearn.manifold import TSNE
+from sklearn.decomposition import PCA
+import statsmodels.api as sm
+from transformers import BertTokenizer, BertModel
+
+# MAGICKS --------------------------
+sPathEmbeddings = "data/processed/summary_embeddings"
+sNameEmbeddingA = "tangible_hair.pkl"
+sNameEmbeddingB = "intangible_hair.pkl"
+
+# FUNCTIONS --------------------------
+# https://stackoverflow.com/questions/56369030/runtimeerror-attempting-to-deserialize-object-on-a-cuda-device
+class CPU_Unpickler(pickle.Unpickler):
+    def find_class(self, module, name):
+        if module == 'torch.storage' and name == '_load_from_bytes':
+            return lambda b: torch.load(io.BytesIO(b), map_location='cpu')
+        else:
+            return super().find_class(module, name)
+
+def load_embeddings(sPathEmbeddings, sNameEmbeddingA, sNameEmbeddingB):
+    
+    # Load pickle file
+    with open(sPathEmbeddings + "/" + sNameEmbeddingA, 'rb') as f:
+        lEmbeddingsA = CPU_Unpickler(f).load()
+    
+    with open(sPathEmbeddings + "/" + sNameEmbeddingB, 'rb') as f:
+        lEmbeddingsB = CPU_Unpickler(f).load()
+        
+    tEmbeddingA = torch.concatenate(lEmbeddingsA, dim=1)
+    tEmbeddingB = torch.concatenate(lEmbeddingsB, dim=1)
+    
+    tEmbeddingA = tEmbeddingA.squeeze()
+    tEmbeddingB = tEmbeddingB.squeeze()
+        
+    return tEmbeddingA, tEmbeddingB
+
+class EmbeddingVisualizer:
+    def __init__(self, lEmbeddingA, lEmbeddingB):
+        self.lEmbeddingA = lEmbeddingA
+        self.lEmbeddingB = lEmbeddingB
+        
+        
+def visualize_correlations(similarity_matrix, sName, bEmbedding=False):
+
+    plt.imshow(similarity_matrix, cmap='coolwarm', vmin=-1, vmax=1)
+    plt.colorbar()
+    plt.title(f'Similarity Matrix ({sName})')
[…]

### 2024-03-14 src/llm/sketch.py

@@ -1,0 +1,69 @@
+# SETUP --------------------------
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+from tqdm import tqdm
+
+import torch
+import torch.autograd as autograd
+#from torchviz import make_dot
+from transformers import AutoTokenizer, GPT2LMHeadModel
+
+from undecorated import undecorated
+from types import MethodType
+
+# FUNCTIONS --------------------------
+import torch
+
+b_use_backend = True
+i_seed_value = 42
+s_model = 'gpt2'
+#s_model = 'GroNLP/gpt2-small-dutch'
+
+# INIT
+# if b_use_backend:
+#     if torch.backends.mps.is_available():
+#         mps_device = torch.device("mps")
+#         x = torch.ones(1, device=mps_device)
+#         torch.set_default_device('mps')
+#         print (f"Using MPS backend {x}")
+#     elif torch.cuda.is_available():
+#         print ("Using CUDA backend")
+#         torch.set_default_device('cuda')
+#     else:
+#         print ("Using CPU.")
+
+# #s_prompt_string = "A hot"
+# #s_target_string = "Potato"
+# # s_target_string = "easy example." # text generated by this prompt for this seed!
+# # s_target_string = "does it make sense to compare apples and oranges?" # text generated by this prompt for this seed!
+# s_target_string = "one, two," # text generated by this prompt for this seed!
+# # s_target_string = "Transform Your Hair's Health with our Revolutionary Haircare System: Experience Unparalleled Shine, Strength, and Vitality!" # text generated by this prompt for this seed!
+# # #s_target_string = "Heeft het zin om appels en sinaasappelen te vergelijken?" # text generated by this prompt for this seed!
+
+# torch.manual_seed(i_seed_value)
+# model = GPT2LMHeadModel.from_pretrained(s_model)
+# tokenizer = AutoTokenizer.from_pretrained(s_model)
+# tokenizer.pad_token_id = tokenizer.eos_token_id
+
+# # BUG: Issue that greedy search does not find max likelihood sequence? Does this matter?
+# i_num_beams = 1 # greedy search for i_num_beams = 1
+# i_no_repeat_ngram_size = 1 # TODO: This creates inplace gradient error
+# i_num_return = 1
+# i_max_new_tokens = tokenizer.encode(s_target_string, return_tensors="pt").shape[1]
+
[…]

### 2024-03-14 src/llm/training.py

@@ -1,0 +1,433 @@
+# SETUP --------------------------
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+from tqdm import tqdm
+
+import torch
+import torch.autograd as autograd
+#from torchviz import make_dot
+from transformers import AutoTokenizer, GPT2LMHeadModel
+
+from undecorated import undecorated
+from types import MethodType
+import time
+from concurrent.futures import ThreadPoolExecutor, ProcessPoolExecutor
+
+# FUNCTIONS --------------------------
+
+def llikelihood(output: tuple, target_sequence: torch.Tensor) -> float:
+    """Calculate the log-likelihood of the target tokens in the output sequence.
+    NOTE: This only works for greedy search, not beam search (i.e. set num_beams=1).
+
+    Args:
+        output (tuple): Output from a HF LLM text generation model. Only pass one sequence.
+        target_sequence (torch.Tensor): The target tokens (the sequence) to calculate the log-likelihood of.
+
+    Returns:
+        float: log-llikelihood for the sequence
+    """
+
+    # get scores and the sequences
+    #print("raw scores comparison")
+    #print(output.scores[0])
+    #print(output.scores[1])
+    #print("len tuple")
+    #print(len(output.scores))
+
+    #scores = autograd.Variable(torch.stack(output.scores).reshape(len(output.scores), -1).transpose(0, 1), requires_grad=True)
+    scores = torch.stack(output.scores).reshape(len(output.scores), -1).transpose(0, 1)
+
+    #print("scores shape")
+    #print(scores.shape)
+
+    # get the scores into a matrix, use log_softmax
+    #m_prob = autograd.Variable(torch.nn.functional.log_softmax(scores, dim=0), requires_grad=True)
+    m_prob = torch.nn.functional.log_softmax(scores, dim=0)
+    #m_prob_sub = autograd.Variable(m_prob[target_sequence.squeeze()], requires_grad=True)
+    m_prob_sub = m_prob[target_sequence.squeeze()]
+
+    # get the diagonal of the matrix
+    #v_trans_prob = autograd.Variable(torch.diag(m_prob_sub), requires_grad=True)
+    v_trans_prob = torch.diag(m_prob_sub)
+    #d_llikelihood = autograd.Variable(torch.sum(v_trans_prob) * (-1), requires_grad=True)
+    # NOTE: we use a minimizer
[…]

### 2024-03-14 meetings/14-03-2024_slides.qmd
typos

---
title: "Meeting --- Summary LLM"
author: "Bas Donkers, Dennis Fok, Finn-Ole Höner"
date: 2024-03-14
date-format: long
format: beamer
aspectratio: 169
incremental: false
include-in-header: ../beamer_preamble.tex
bibliography: ../references.bib
fig-align: center
fig-cap-location: bottom
fig-format: pdf
links-as-notes: true
number-depth: 1
---

# Updates

- Grant proposal handed in
- SURF Research Cloud Workshop
  - Pieter Meijndert, Yuliia Orlova
  - Small SURF Grant (quick, easy to get, uncomplicated); up to 5k GPU hours; once per calendar year
- (Start thinking about thesis comittee)
- Improving implementation of optimization still work in progress

---

{{< embed ../src/analysis/visualize_embeddings.ipynb#fig-corr1 >}}

---

{{< embed ../src/analysis/visualize_embeddings.ipynb#fig-corr2 >}}

---

{{< embed ../src/analysis/visualize_embeddings.ipynb#fig-corr3 >}}

---

{{< embed ../src/analysis/visualize_embeddings.ipynb#fig-pca >}}

---

{{< embed ../src/analysis/visualize_embeddings.ipynb#fig-tsne >}}

---

{{< embed ../src/analysis/visualize_embeddings.ipynb#fig-loo1 >}}

---

{{< embed ../src/analysis/visualize_embeddings.ipynb#fig-loo2 >}}

---

{{< embed ../src/analysis/visualize_embeddings.ipynb#fig-loo3 >}}

---

{{< embed ../src/analysis/visualize_embeddings.ipynb#fig-corr-all >}}

---

{{< embed ../src/analysis/visualize_embeddings.ipynb#fig-pca-all >}}

### 2024-03-14 meetings/18-01-2024_slides.qmd
typos

@@
 title: "Meeting --- Summary LLM"
-author: "Bas Doners, Dennis Fok, Finn-Ole Höner"
+author: "Bas Donkers, Dennis Fok, Finn-Ole Höner"
 date: 2024-01-18

### 2024-04-23 meetings/12-04-2024_slides.qmd
eos works

---
title: "Meeting --- Summary LLM"
author: "Bas Donkers, Dennis Fok, Finn-Ole Höner"
date: 2024-04-12
date-format: long
format: beamer
aspectratio: 169
incremental: false
include-in-header: ../beamer_preamble.tex
bibliography: ../references.bib
fig-align: center
fig-cap-location: bottom
fig-format: pdf
links-as-notes: true
number-depth: 1
eval: false
echo: true
fig-width: 12
fig-height: 8
---

# Updates

- Improved the training, now works with tensors, GPT batching (~5-10x faster)
- Monitoring steps made is more involved than expected
- Normalization layer
- EOS token (GPT-2 has no padding token)
- "Exploration" of the encoder space
- Scales linearly with `iMaxTokens`, by the cube with `iClaims` (i.e. $O(n^3)$)

---

:::: {.columns}
::: {.column}
![Hair 10](../models/training_monitor_autoencoder.png)
:::
::: {.column}
![Hair 20](../models/training_monitor_autoencoder20.png)
:::
::::

---

# Examples of "New" Claims

Issues at $(0, 0)$?

Exact reproduction of training data, dead space, and "new" claims; i.e. EOS token and not in training data.

---

Examples:

<|endoftext|>Locks in in moisture to enhance<|endoftext|>"

<|endoftext|>Clin by enhance by enhance by saturation.<|endoftext|>

<|endoftext|>Linininineineine<|endoftext|>

<|endoftext|>Lin by enhance by enhanceistedistedistedisted by<|endoftext|>

<|endoftext|>Locks in moisture to amplify amplify l to enhance line<|endoftext|>

<|endoftext|>Elevate your confidence with confidence.<|endoftext|>

<|endoftext|>Elevate your confidence with hair that gleams in any light.<|endoftext|>

<|endoftext|>Clinically proven to enhance shine by up to%.<|endoftext|>

---

![Cleaner](./figures/cleaner_vis_norm.png)

---

![New Hair, bad grid?](./figures/output_new.png)

---

![New Hair, bad grid?](./figures/hair_new2.png)

---

![New Hair, bad grid?](./figures/hair_new3.png)

---

Inspriration from molecule generation / drug discovery [e.g @MegaSynIntegratingGenerative]

- Balancing exploration: Searching for molecules that are not only enumerations of the training data, but that are also feasible
- Primed models, focussing on specific sub-structures and saving models every $n$ epochs. This yields a spectrum of models that generate more general to more specific solutions.
- Have model that scores the quality of a claim
- Hill-climb MLE (Feedback with top X% of claims)
- Fine-tune GPT to advertising claims
- Wrap MAB around this: Play arm for new generation, evaluate with scoring model, learn from feedback (*also online on websites? This model might be rather fast? Retrain the latent space in batches based on the feedback*)
- Do we need to retrain from scratch or can we simply keep training with the existing model as a starting point?

### 2024-04-23 meetings/27-03-2024_slides.qmd
eos works

---
title: "Meeting --- Summary LLM"
author: "Bas Donkers, Dennis Fok, Finn-Ole Höner"
date: 2024-03-27
date-format: long
format: beamer
aspectratio: 169
incremental: false
include-in-header: ../beamer_preamble.tex
bibliography: ../references.bib
fig-align: center
fig-cap-location: bottom
fig-format: pdf
links-as-notes: true
number-depth: 1
eval: false
echo: true
---

# Updates

- Bias terms in encoder and decoder
- More nodes in hidden layer speeds up convergence
- Also tried out 3 nodes, results are "more noisy". However, could be interesting when claims have more "dimensions" behind them.
- Threshold: Each log-likelihood below 1 (this is softer than)

# Relevant Code

```{python}
while not bStop:
    aStore = np.zeros((1, tOHETarget.shape[0]))
    lLoss = []
    for i in range(tOHETarget.shape[0]):
        ae_outputs = ae(tOHETarget[i, :])
        outputs = lGenerate[i](ae_outputs.reshape(1, 1, -1))
        loss = llikelihood(outputs, lTargetSequence[i])

        optimizer.zero_grad()
        loss.backward()
        optimizer.step()
    bStop = np.all(aLLHistory[epoch, :] < dEps)
```

---

::: {.columns}

::: {.column}

![4:15h MacBook, Hair, 10 Claims, ADAM lr=3]("./figures/ae_training_hair_2d.png")

:::

::: {.column}
![14h MacBook, Cleaner, 10 Claims, ADAM lr=3]("./figures/training_cleaner_2d.png")

:::

:::

---

::: {.columns}

::: {.column}

![Autoencoder values of Hair 2 dimensions]("./figures/ae_2d_hair.png")

:::

::: {.column}

![Autoencoder values of Cleaner 2 dimensions]("./figures/ae_sc.png")

:::

:::

---

::: {.columns}

::: {.column}

![Embeddings AE-2D]("./figures/ae_structure_hair_2d.png")

:::

::: {.column}

![Embeddings Previously]("./figures/structure_embeddings_before.png")

:::

:::

---

Hair
```{python}
    "Experience mirror-like shine with every wash.",
    "Get dazzling, glossy hair in just one use.",
    "Transform dull strands into radiant locks.",
    "Achieve salon-quality shine at home.",
    "Enhance natural brilliance with our advanced formula.",
    "Unlock the confidence of luminous hair that turns heads.",
    "Experience the joy of hair that gleams with vitality.",
    "Embrace the radiance of hair that reflects your inner beauty.",
    "Elevate your everyday with hair that shines with sophistication.",
    "Feel the allure of irresistibly shiny hair that captivates."
```

---

::: {.columns}

::: {.column}

![Corr. Mat. 3D-AE Hair]("./figures/ae_3d_corr.png")

:::

::: {.column}

![Corr. Mat. BERT Hair]("./figures/bert_corr_hair.png")

:::

:::

---

::: {.columns}

::: {.column}

![Corr. Mat. 2D-AE Hair]("./figures/ae_corr_hair_2d.png")

:::

::: {.column}

![Corr. Mat. BERT Hair]("./figures/bert_corr_hair.png")

:::

:::

---

Surface Cleaner
```{python}
    "Our surface cleaner delivers a mirror-like shine every time.",
    "Say goodbye to streaks and smudges with our powerful formula.",
    "Achieve a noticeable sparkle on all surfaces with just one application.",
    "Experience surfaces so clean, you can see your reflection in them.",
    "Our cleaner leaves behind a gleaming, polished finish that lasts.",
    "Elevate your surroundings with a touch of brilliance.",
    "Unleash the hidden beauty of your home with our surface cleaner.",
    "Experience the joy of pristine surfaces that brighten your space.",
    "Bring a sense of clarity and freshness to your environment.",
    "Enhance the ambiance of any room with surfaces that radiate."
```

---

::: {.columns}

::: {.column}

![Corr. Mat. 2D-AE Cleaner]("./figures/ae_sc_corr.png")

:::

::: {.column}

![Corr. Mat. BERT Cleaner]("./figures/bert_sc_corr.png")

:::

:::

---

::: {.columns}

::: {.column}

![PCA 2D-AE Hair]("./figures/ae_pca_hair_2d.png")

:::

::: {.column}

![PCA BERT Hair]("./figures/bert_pca_hair.png")

:::

:::

---

::: {.columns}

::: {.column}

![PCA 2D-AE Cleaner]("./figures/ae_sc_pca.png")

:::

::: {.column}

![PCA BERT Cleaner]("./figures/bert_sc_pca.png")

:::

:::

---

::: {.columns}

::: {.column}

![LOO 2D-AE Hair]("./figures/ae_loo_hair_2d.png")

:::

::: {.column}

![LOO 2D-AE Cleaner]("./figures/ae_sc_loo.png")

:::

:::

---

![Embeddings AE-3D]("./figures/ae_structure_both.png")

---

::: {.columns}

::: {.column}

![Corr. Mat. BERT Hair]("./figures/ae_both_3d_corr.png")

:::

::: {.column}

![Corr. Mat. BERT Hair]("./figures/bert_corr_both.png")

:::

:::

### 2024-04-23 src/analysis/toolbox.py

@@ -1,0 +1,156 @@
+# Read in the sentences from lClaims
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+
+import torch
+import torch.nn as nn
+import torch.optim as optim
+
+from transformers import AutoTokenizer, GPT2LMHeadModel
+from undecorated import undecorated
+from types import MethodType
+
+import time
+import pickle
+import io
+import os
+
+
+def TokenizeClaims(lClaims, tokenizer):
+    # Tokenize the sentences
+    tokenized_sentences = [tokenizer.encode(sentence, add_special_tokens=True) for sentence in lClaims]
+
+    # Find the maximum length of the tokenized sentences
+    lLengths = [len(sentence) for sentence in tokenized_sentences]
+    max_length = max(len(sentence) for sentence in tokenized_sentences) + 1 # add on to account for the eos token
+
+    # Pad the tokenized sentences on the left to have the same length
+    padded_sentences = [sentence + [tokenizer.eos_token_id] + [tokenizer.pad_token_id] * (max_length - len(sentence) - 1) for sentence in tokenized_sentences]
+    
+    # Convert the padded sentences to tensors
+    padded_sentences = torch.tensor(padded_sentences)
+
+    return padded_sentences, torch.tensor(lLengths)
+
+def create_input_data(iClaims):
+    return torch.eye(iClaims)
+
+# TODO: Consider removing this function
+def PadScores(tOut, tLengths, iClaims, iVocab: int=50257):
+
+    tOutAdj = tOut.clone()
+    for i in range(iClaims):
+        tOutAdj[i, tLengths[i]:, :-1] = float('-inf')
+        tOutAdj[i, tLengths[i]:, -1] = 0
+        
+    # tOutAdj[:, tLengths[:, None]:, :-1] = float('-inf')
+    # tOutAdj[:, tLengths[:, None]:, -1] = 0
+
+    # TODO: Fix and double-check this, tOut, get rid of things that are not needed
+    # TODO: Set to zero to get rid of it
+    # TODO: Do we get the gradient? Ignore the signal from excess tokens
+
+    
+    return tOutAdj
+
+def LLikelihood(lScores, tTargetStrings, iMaxTokens, iClaims, iVocab: int=50257):
+    
+    if len(lScores) < iMaxTokens:
+        print("Appending")
[…]

### 2024-04-23 src/analysis/visualize_embeddings.py

@@ -9,6 +9,11 @@ from sklearn.manifold import TSNE
 from sklearn.decomposition import PCA
 import statsmodels.api as sm
 from transformers import BertTokenizer, BertModel
+from sklearn.feature_extraction.text import CountVectorizer
+import torch
+from transformers import BertTokenizer, BertModel
+import numpy as np
+from gensim.models import Word2Vec
 
 # MAGICKS --------------------------
 sPathEmbeddings = "data/processed/summary_embeddings"
@@ -189,12 +194,6 @@ def loo_regression(tEmbeddingsA, tEmbeddingsB, tEmbeddings_demeaned, sName, vCol
     plt.show()
     
     return lResults
-
-from sklearn.feature_extraction.text import CountVectorizer
-import torch
-from transformers import BertTokenizer, BertModel
-import numpy as np
-from gensim.models import Word2Vec
 class SummaryEmbeddings:
     def __init__(self, lClaims):
         self.lClaims = lClaims
@@ -214,7 +213,7 @@ class SummaryEmbeddings:
         # Extract the embeddings for the [CLS] token, which represents the entire sentence
         # This index corresponds to the first token in each sequence
         embeddings = outputs.last_hidden_state[:, 0, :]
-        
+        print(embeddings.shape)
         return embeddings.detach().numpy()
     
     def word2vec(self):

### 2024-04-23 src/estimate_ae.py

@@ -1,0 +1,265 @@
+# SETUP --------------------------
+from analysis.toolbox import *
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+
+import torch
+import torch.nn as nn
+import torch.optim as optim
+
+from transformers import AutoTokenizer, GPT2LMHeadModel
+from undecorated import undecorated
+from types import MethodType
+
+import time
+import pickle
+import io
+import os
+
+device = (
+    "cuda"
+    if torch.cuda.is_available()
+    else "mps"
+    if torch.backends.mps.is_available()
+    else "cpu"
+)
+torch.set_default_device(device)
+print(f"Using {device} device")
+print(torch.tensor([1, 2, 3]).device)
+
+# MAGICKS --------------------------
+s_model = "gpt2"
+iSeed = 4523522
+sPadding = "left"
+sNameAEModel = "autoencoder_test"
+
+lTargetStrings = [
+    "Experience 50% more visible shine after just one use.",
+    "Formulated with light-reflecting technology for a glossy finish.",
+    "Transform dull strands into radiant, luminous locks.",
+    "Infused with nourishing oils that enhance natural shine.",
+    "See instant brilliance with our advanced shine-boosting formula.",
+    "Locks in moisture to amplify hair's natural luster.",
+    "Achieve salon-quality shine without leaving home.",
+    "Visible reduction in dullness, replaced with stunning shine.",
+    "Say goodbye to lackluster hair, hello to mirror-like shine.",
+    "Clinically proven to enhance shine by up to 70%.", # ^tangible
+    "Elevate your confidence with hair that gleams under any light.",
+    "Embrace the allure of luminous hair that turns heads.",
+    "Unleash the power of radiant hair that speaks volumes.",
+    "Transform your look with hair that exudes brilliance.",
+    "Feel the difference of hair that shines with vitality and health.",
+    "Rediscover the joy of hair that beams with inner vibrancy.",
+    "Indulge in the luxury of hair that shimmers with elegance.",
+    "Step into the spotlight with hair that radiates beauty.",
[…]

### 2024-04-23 src/llm/ae_summary.py

@@ -1,0 +1,215 @@
+import numpy as np
+import pandas as pd
+import torch
+import torch.nn as nn
+import torch.optim as optim
+import torch
+import torch.autograd as autograd
+import os
+import torch
+from torch import nn
+from torch.utils.data import DataLoader
+from torchvision import datasets, transforms
+#from torchviz import make_dot
+from transformers import AutoTokenizer, GPT2LMHeadModel
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+from tqdm import tqdm
+
+import torch
+import torch.autograd as autograd
+import os
+import torch
+from torch import nn
+from torch.utils.data import DataLoader
+from torchvision import datasets, transforms
+#from torchviz import make_dot
+from transformers import AutoTokenizer, GPT2LMHeadModel
+
+from undecorated import undecorated
+from types import MethodType
+import time
+from concurrent.futures import ThreadPoolExecutor, ProcessPoolExecutor
+
+import pickle
+import io
+import os
+
+print(os.getcwd())
+
+class CPU_Unpickler(pickle.Unpickler):
+    def find_class(self, module, name):
+        if module == 'torch.storage' and name == '_load_from_bytes':
+            return lambda b: torch.load(io.BytesIO(b), map_location='cpu')
+        else:
+            return super().find_class(module, name)
+
+def load_embeddings(sPathEmbeddings, sNameEmbeddingA, sNameEmbeddingB):
+    
+    # Load pickle file
+    with open(sPathEmbeddings + "/" + sNameEmbeddingA, 'rb') as f:
+        lEmbeddingsA = CPU_Unpickler(f).load()
+    
+    with open(sPathEmbeddings + "/" + sNameEmbeddingB, 'rb') as f:
+        lEmbeddingsB = CPU_Unpickler(f).load()
+        
+    tEmbeddingA = torch.concatenate(lEmbeddingsA, dim=1)
+    tEmbeddingB = torch.concatenate(lEmbeddingsB, dim=1)
+    
+    tEmbeddingA = tEmbeddingA.squeeze()
+    tEmbeddingB = tEmbeddingB.squeeze()
+        
+    return tEmbeddingA, tEmbeddingB
+
+sPathEmbeddings = "data/processed/summary_embeddings"
+sNameEmbeddingA = "tangible_hair_prec.pkl"
+sNameEmbeddingB = "intangible_hair_prec.pkl"
+
+device = (
+    "cuda"
+    if torch.cuda.is_available()
+    else "mps"
[…]

### 2024-04-23 src/llm/training.py

@@ -16,123 +16,87 @@ from concurrent.futures import ThreadPoolExecutor, ProcessPoolExecutor
 
 # FUNCTIONS --------------------------
 
-def llikelihood(output: tuple, target_sequence: torch.Tensor) -> float:
-    """Calculate the log-likelihood of the target tokens in the output sequence.
-    NOTE: This only works for greedy search, not beam search (i.e. set num_beams=1).
+class ReverseEngineer:
+    
+    def __init__(self, target_sequence: torch.Tensor):
+        self.target_sequence = target_sequence
 
-    Args:
-        output (tuple): Output from a HF LLM text generation model. Only pass one sequence.
-        target_sequence (torch.Tensor): The target tokens (the sequence) to calculate the log-likelihood of.
+    def llikelihood(self, output: tuple) -> float:
+        """Calculate the log-likelihood of the target tokens in the output sequence.
+        NOTE: This only works for greedy search, not beam search (i.e. set num_beams=1).
+        NOTE: We use a minimizer
 
-    Returns:
-        float: log-llikelihood for the sequence
-    """
+        Args:
+            output (tuple): Output from a HF LLM text generation model. Only pass one sequence.
+            target_sequence (torch.Tensor): The target tokens (the sequence) to calculate the log-likelihood of.
 
-    # get scores and the sequences
-    #print("raw scores comparison")
-    #print(output.scores[0])
-    #print(output.scores[1])
-    #print("len tuple")
-    #print(len(output.scores))
+        Returns:
+            float: log-llikelihood for the sequence
+        """
 
-    #scores = autograd.Variable(torch.stack(output.scores).reshape(len(output.scores), -1).transpose(0, 1), requires_grad=True)
-    scores = torch.stack(output.scores).reshape(len(output.scores), -1).transpose(0, 1)
+        # get scores and the sequences
+        scores = torch.stack(output.scores).reshape(len(output.scores), -1).transpose(0, 1)
 
-    #print("scores shape")
-    #print(scores.shape)
+        # get the scores into a matrix, use log_softmax
+        m_prob = torch.nn.functional.log_softmax(scores, dim=0)
+        m_prob_sub = m_prob[self.target_sequence.squeeze()]
 
-    # get the scores into a matrix, use log_softmax
-    #m_prob = autograd.Variable(torch.nn.functional.log_softmax(scores, dim=0), requires_grad=True)
-    m_prob = torch.nn.functional.log_softmax(scores, dim=0)
[…]

### 2024-04-24 writing/paper.qmd
pre-training, init works

--- 2024-02-21 writing/draft_grant_proposal.qmd
@@
----
-format: docx
-pdf-engine: pdflatex
-header-includes:
-  \usepackage{fancyhdr}
-  \usepackage{fancyvrb}
-  \usepackage{float}
-  \usepackage{booktabs}
-  \usepackage{lscape}
-  \usepackage{amsmath}
-  \usepackage{graphicx}
-  \usepackage{algorithm}
-  \usepackage{algpseudocode}
-  \DeclareMathOperator*{\argmin}{arg\,min}
-  \DeclareMathOperator*{\argmax}{arg\,max}
----
-
 # Introduction
 
-Difference of written summary and embedding summary.
+- Media products increase relevance of content
+  - Social media, short videos TikTok, Images, Netflix, Audio / Podcasts
 
-Learning summaries of text is a fundamental problem in natural language processing, and of special interest in the context of Large Language Models (LLMs).
-While current approaches aim at summarizing text with a written summary that is shorter than the original text, we propose a novel approach, which reverse engineers these summaries.
-Rather than prompting a LLM to write a summary of a given text, we perform an optimization to find the input embedding to this LLM, which maximizes the likelihood for the LLM to generate the given text.
+- Market research is expensive. How can we leverage these insights better to improve its benefits?
 
-We aim to link these summary embeddings to consumer preferences, and plan to use these data to inform the generation of new claims for specific contexts.
+- EU AI Act
 
-One way how such a generation could work, is by averaging the embeddings of a set of claims, and then using this average embedding to generate a new claim.
+- GDPR
+  - Challenges to cookies in website morphing
 
-<!-- TODO: Double check whether this is the right bandit -->
-We could use Combinatorial Bandits to learn about how to create linear combinations of existing embeddings, such that the generated new claim is more appealing to the consumer.
+- Generative AI for personalization
+  - Previously: Set pieces, e.g. recommendation systems
+  - True personalization: Generate solution for each individual consumer
 
-A more sophisticated approach would lie in the use of Multi-Armed Bandits, where we set the arms to be certain elements of the embedding vector.
-By creating a stack of MAB and LLM, we could give the MAB the ability to generate new claims and to learn about the preferences of the consumer for these new claims.
-In this, our summary embeddings play a key role, as they allow us to generate new claims in a deterministic fashion, which in-turn enables us to learn about the preferences of the consumer for these new claims.
+- Consumer preferences in latent generation space
+  - We can view products as attribute bundles and get utilities for these attributes <!-- TODO: What about interactions of attributes? -->
+  - How do we do this for products that cannot be "discretized"?
 
-Another empirical applications could lie in learning about consumer preferences for headlines of news articles or posts, and to use these data to inform the generation of new headlines.
+- Novel training framework
+  - Circumvent issues with the LLMs decoder
+  - Input embedding represents the information contained in the sequence, as it is the starting point which makes the LLM generate the target sequence
 
-There are ample use cases and possible empirical investigations in a marketing context for these summary embeddings.
-In the following, we want to highlight three different applications, that might deliver novel insights due to the use of these summary embeddings.
+<!-- TODO: Consider renaming this headline -->
+# Theory
 
-<!-- TODO: double checks -->
-First, we want to research how marketers design choices on the creation of product claims relates to consumer preferences.
-For example, a marketer might decide between making a more factual or a more emotional claim.
-While there are theoretical motivations for these types of decisions, it would allow us to quantitatively assess these styles of claims and to learn how consumer preferences relate to them.
-An extension of this research could lie in the relation of human marketers and the help of AI, namely LLMs, in the creation of these claims.
-Investigating how AI and human generated claims that are supposed to fulfill a certain characteristic, such as being more emotional, are actually perceived as such by consumers, could deliver novel insights.
+- Generative AI for text
+    - Attention
+    - Transformer
+    - GPT
 
-Second, in a more field-based setting, we want to investigate how we can use these summary embeddings to adapt online descriptions to the preferences of the consumer.
-While there are techniques based on Reinforcement Learning, which is used to adapt the website to the preferences of the consumer, already in use, these types of algorithms cannot make adaptations "out-of-sample", i.e.
-they cannot generate new content that is not already predefined.
-We believe that the use of summary embeddings could allow us to generate adaptive text-based content, such as product descriptions or article headlines, that are not part of a predefined set.
+- Autoencoder
 
-Third,
+- Maximize the likelihood to generate the data
 
-<!-- TODO: True? Also add more examples -->
-In preliminary tests, we have found that these summary embeddings appear to be able to differentiate different styles of claims in the same domain.
+The Autoencoder $\operatorname{AE}_{\Beta} \left(\cdot \right): \lbrace 0, 1 \rbrace^{C} \to \mathbb{R}^E$, with parameters $\Beta$
 
-# Methods
+The log-likelihood function $\operatorname{\mathcal{L}} \left ( \cdot \right): \mathbb{R}^E \to \left (- \infty, 0 \right]$
 
-- Novel, “reverse engineering” approach to finding text summaries (find a better word… in an information sense, not in a writing short summary sense).
-Given a target sequence and a given LLM, we want to find the input embedding to this LLM, which maximizes the likelihood for the LLM to generate the target sequence.
+The length $E$ summary-embedding $\boldsymbol{e}$.
 
-- Training can be done with simply forward passes through the LLM model
-
-## Problem Formulation
-
-For a given Large Language Model, $\mathcal{M}$, we want to find the input embedding, $\boldsymbol{s^{*}}$, that maximizes the likelihood of generating a given target sequence, $\lbrace t_i \rbrace_{i=1}^{L}$, where $L$ denotes the length of the target sequence and $t_i$ represents token $i$.
+The length $T$ vector of target tokens $\boldsymbol{t}$, where the element $t_i$ is a unique integer code representing a token, e.g. in GPT-2, $50267$ is the End-of-Text token `<|eos|>`.
 
 <!-- TODO: What does it really return? -->
-
-To express that we generate tokens with the LLM, based on length $E$ input-embedding $\boldsymbol{e}$, we use the notation $\mathcal{M}(\boldsymbol{e})$.
-To express the number of tokens, $k$ that we generate, we use the subscript $\mathcal{M}_k(\boldsymbol{e})$.
+<!-- TODO: Good idea or not? -->
+To express that we generate tokens with the LLM, $\mathcal{M}$, based on length $E$ input-embedding $\boldsymbol{e}$, we use the notation $\mathcal{M}(\boldsymbol{e})$.
+To express the number of tokens, $k$, that we generate, we use the subscript $\mathcal{M}_k(\boldsymbol{e})$.
 This function call returns an output vector of length $V$, where $V$ denotes the vocabulary size.
 We denote the output vector as $\boldsymbol{o}$.
-Each element $o_v$ of $\boldsymbol{o}$ represents the probability of generating token $v$ as the next token.
+Each element $o_v$ of this output vector represents the probability of generating token $v$ as the next token.
 We predict the next token, by selecting the largest element of $\boldsymbol{o}$.
-Explicitly, $t_{1} = \argmax \mathcal{M}_1(\boldsymbol{e})$.
-
 When $k = 1$, we use the previous tokens to predict the next token.
@@
 While there are many ways to design this procedure and select tokens along the way, we perform a standard greedy procedure, where we select the token with the highest probability at each step.
+We denote the probability that $\mathcal{M}$ generates the target sequence, given the input embedding, as $p(t_1, \ldots, t_L | \boldsymbol{e})$.
+For a given Large Language Model, we define $\boldsymbol{e^{*}}$,
 
-We denote the probability that $\mathcal{M}$ generates the target sequence, given the input embedding, as $p(t_1, \ldots, t_L | \boldsymbol{s})$.
+$$
+\boldsymbol{e^{*}} = \argmax_{\boldsymbol{e}} \sum_{i=1}^{L} \log p(t_1, \ldots, t_L | \boldsymbol{e}) \text{, }
+$$
 
-We estimate these embeddings, by maximizing the conditional likelihood that this embedding generates the target sequence for a given LLM.
+as the input-embedding that maximizes the likelihood of generating a given target sequence, $\lbrace t_i \rbrace_{i=1}^{L}$, where $L$ denotes the length of the target sequence and $t_i$ represents token $i$.
+We estimate these summary embeddings $\boldsymbol{e^{*}}$, by maximizing the conditional likelihood that this embedding generates the target sequence for a given LLM.
 We use a gradient-based optimization algorithm to maximize this likelihood.
 
-These embeddings have the same dimension as the input embedding of $\mathcal{M}$, length $E$.
+<!-- TODO: Consider renaming this headline -->
+# Implementation
 
-$$
-\boldsymbol{s^{*}} = \argmax_{\boldsymbol{s}} \prod_{i=1}^{L} p(t_1, \ldots, t_L | \boldsymbol{s})
-$$
+- Autoencoder
 
-$$
-\boldsymbol{s^{*}} = \argmax_{\boldsymbol{s}} \sum_{i=1}^{L} \log p(t_1, \ldots, t_L | \boldsymbol{s})
-$$
+- Configuration of LLM
 
-$$
-\boldsymbol{s^{*}} = \argmax_{\boldsymbol{s}} \log p(t_1, | \boldsymbol{s}) + \log p(t_2, | \boldsymbol{s}, t_1) + \ldots + \log p(t_L, | \boldsymbol{s}, t_1, \ldots, t_{L-1}) 
-$$
+- Optimization
 
-$$
-\boldsymbol{s^{*}} = \argmax_{\boldsymbol{s}} \log \mathcal{M}_1(\boldsymbol{s}) + \log \mathcal{M}_1(\left [\boldsymbol{s}, \boldsymbol{e}_1 \right]) + \ldots + \log \mathcal{M}_1(\left [\boldsymbol{s}, \boldsymbol{e}_1, \ldots, \boldsymbol{e}_{L-1} \right])
-$$
+- Connection to other measures, e.g. syntax or rating of consumers
 
-$$
-\boldsymbol{s^{*}} = \argmax_{\boldsymbol{s}} \sum_{i=1}^{L} \log \mathcal{M}_L (\boldsymbol{s})
-$$
+# Data
 
-We can make this expression more explicit, by using the LLM in our notation. For example $p(t_1, | \boldsymbol{s}) = \mathcal{M}(\boldsymbol{s})$
+- Descriptives of the data
 
-In the training process, we execute the following algorithm.
-First, we initialize the input embedding, $\boldsymbol{s}$.
-Second, we generate a sequence of length $L$, with the LLM and the current input embedding.
+- Prepocessing of the data
 
-\begin{algorithm}
-\begin{algorithmic}
-\State Initialize $\boldsymbol{s}$
-\State $i \gets 0$
-\While{$Convergence == False$}
-    \State $i \gets i + 1$
-    \State $\boldsymbol{o}^{(i)} \gets \mathcal{M}(\boldsymbol{s})$
-    \State $l^{(i)} \gets \operatorname{Likelihood}(\boldsymbol{o}^{(i)}, \lbrace t_i \rbrace_{1}^{L})$
-    \State $\boldsymbol{s}^{(i + 1)} \gets \operatorname{GradientDecent}(\boldsymbol{s}^{(i)}, l^{(i)})$
-    \State $Convergence \gets \operatorname{CheckConvergence}(\boldsymbol{s}^{(i + 1)}, \boldsymbol{s}^{(i)})$
-\EndWhile
-\end{algorithmic}
-\end{algorithm}
+# Results
+
+- Optimization
+  - Computation time, scaling
+- Encoding space
+  - example tangible, intangible
+  - generations from encoding space
+  - Wasteland, geometrical shape and where it comes from
+- Application to CBC data
+
+# Managerial Implications
+
+- Generative AI for personalization
+
+# Discussion
+
+- Problems
+  - Computation time
+
+- Expansion of this research
+  - MAB around the fitted AE, to incorporate consumer feedback in online learning: Which areas of the space are valuable for which consumers?
+  - Better LLM
+  - Image, video, website, audio data. Perhaps website a good application. Perhaps audio a rather simple way to make this multi-modal (content and sound)

### 2024-04-24 src/analysis/toolbox.py

@@ -7,7 +7,7 @@ import torch
 import torch.nn as nn
 import torch.optim as optim
 
-from transformers import AutoTokenizer, GPT2LMHeadModel
+from transformers import AutoTokenizer, GPT2LMHeadModel, PhrasalConstraint
 from undecorated import undecorated
 from types import MethodType
 
@@ -154,3 +154,49 @@ def plot_results(sNameAEModel, iEncodingDim, ae, lLoss, aLLHistory, lStepsWeight
             tGenNew = fn_generate_max(ae(tOHETarget).reshape(tOHETarget.shape[0], 1, -1)).sequences
             for i in range(tOHETarget.shape[0]):
                 print(f"Claim {i}: {tokenizer.decode(tGenNew[i])}")
+                
+                
+def fit_ae(fn_generate_max_constr, fn_generate_max, tokenizer, ae, optimizer, OHETarget, TargetStrings, iMaxTokens, iClaims, iVocab, dEps, i):
+    iEpoch = 0
+    lLoss = list()
+    bStop = False
+    tic = time.time()
+    while not bStop:
+
+        # Forward pass, Generate and compute loss
+        tAEOutputs = ae(OHETarget).reshape(OHETarget.shape[0], 1, -1)
+
+        # BUG: Generated less that iMaxTokens elements for all claims, then crashed
+        outputs = fn_generate_max_constr(tAEOutputs)
+        
+        # TODO revert to outputs.scores
+        loss, loss_monitor = LLikelihood(outputs.scores, TargetStrings, iMaxTokens, iClaims, iVocab)
+        lLoss.append(loss.item())
+
+        # Backward pass and optimization
+        optimizer.zero_grad()
+        loss.backward()
+
+        # Check stopping criterion; logging
+        bStop = loss < dEps        
+        optimizer.step()
+
+        # Monitoring
+        if (iEpoch % 10 == 0) or bStop:
+            toc = time.time()
+            plt.plot(np.log1p(lLoss))
+            plt.axhline(y=np.log1p(dEps), linestyle='--', color='black')
+            plt.legend()
+            print(f"time: {np.round(toc - tic, 2)} s")
+            plt.savefig(f"./models/training_monitor_monitor.png")
+            tic = time.time()
+            tGenNew = fn_generate_max(ae(OHETarget).reshape(OHETarget.shape[0], 1, -1)).sequences
[…]

### 2024-04-24 src/estimate_ae.py

@@ -8,7 +8,7 @@ import torch
 import torch.nn as nn
 import torch.optim as optim
 
-from transformers import AutoTokenizer, GPT2LMHeadModel
+from transformers import AutoTokenizer, GPT2LMHeadModel, PhrasalConstraint
 from undecorated import undecorated
 from types import MethodType
 
@@ -16,6 +16,7 @@ import time
 import pickle
 import io
 import os
+from tqdm import tqdm
 
 device = (
     "cuda"
@@ -24,6 +25,7 @@ device = (
     if torch.backends.mps.is_available()
     else "cpu"
 )
+device = "cpu"
 torch.set_default_device(device)
 print(f"Using {device} device")
 print(torch.tensor([1, 2, 3]).device)
@@ -32,7 +34,7 @@ print(torch.tensor([1, 2, 3]).device)
 s_model = "gpt2"
 iSeed = 4523522
 sPadding = "left"
-sNameAEModel = "autoencoder_test"
+sNameAEModel = "autoencoder_hair20_eos"
 
 lTargetStrings = [
     "Experience 50% more visible shine after just one use.",
@@ -56,28 +58,22 @@ lTargetStrings = [
     "Experience the magic of hair that dazzles with every movement.",
     "Unlock the secret to hair that shines from within, reflecting your inner glow."
 ]
-lTargetStrings = [
-    "Locks in moisture to amplify hair's natural luster.",
-    "Achieve salon-quality shine without leaving home.",
-    "Visible reduction in dullness, replaced with stunning shine.",
-    "Say goodbye to lackluster hair, hello to mirror-like shine.",
-    "Clinically proven to enhance shine by up to 70%.", # ^tangible
-    "Elevate your confidence with hair that gleams under any light.",
-    "Embrace the allure of luminous hair that turns heads.",
-    "Unleash the power of radiant hair that speaks volumes.",
-    "Transform your look with hair that exudes brilliance.",
-    "Feel the difference of hair that shines with vitality and health."]
-lPromptStrings = [
-    "This new haircare product will",
-    "This new haircare product will",
-    "This new haircare product will",
-    "This new haircare product will",
-    "This new haircare product will",
[…]

### 2024-05-03 src/analysis/toolbox.py
explore likelihood start

@@ -83,12 +83,12 @@ class Autoencoder(nn.Module):
        super(Autoencoder, self).__init__()
        self.encoder = nn.Sequential(
            # TODO: Consider killing this bias term
-           nn.Linear(input_size, encoding_dim, bias=False)
+           nn.Linear(input_size, encoding_dim, bias=True)
        )
        self.decoder = nn.Sequential(
-           nn.Linear(encoding_dim, output_size, bias=True),
+           nn.Linear(encoding_dim, output_size, bias=True)#,
            #nn.Dropout(p = 10/768),
-           nn.LayerNorm(output_size)
+           #nn.LayerNorm(output_size)
        )
 
    def forward(self, x):
@@ -97,11 +97,11 @@ class Autoencoder(nn.Module):
        return x
    
    
-def plot_results(sNameAEModel, iEncodingDim, ae, lLoss, aLLHistory, lStepsWeightsEncoder, lStepsBiasEncoder, lStepsWeightsDecoder, lStepsBiasDecoder, lGradWeightsEncoder, lGradBiasEncoder, lGradWeightsDecoder, lGradBiasDecoder, iEpoch, dEps, tOHETarget, fn_generate_max, tokenizer):
+def plot_results(dTime, sNameAEModel, iEncodingDim, ae, lLoss, aLLHistory, lStepsWeightsEncoder, lStepsBiasEncoder, lStepsWeightsDecoder, lStepsBiasDecoder, lGradWeightsEncoder, lGradBiasEncoder, lGradWeightsDecoder, lGradBiasDecoder, iEpoch, dEps, tOHETarget, fn_generate_max, tokenizer):
             
             #print(f"Progress: {aLLHistory}\n")
             
-            fig, axs = plt.subplots(2, 3, figsize=(16, 16))
+            fig, axs = plt.subplots(2, 4, figsize=(32, 16))
             
             # Plot 1
             for i in range(tOHETarget.shape[0]):
@@ -110,7 +110,7 @@ def plot_results(sNameAEModel, iEncodingDim, ae, lLoss, aLLHistory, lStepsWeight
             axs[0, 0].legend()
             axs[0, 0].set_xlabel("Epoch")
             axs[0, 0].set_ylabel("log1p (-LL)")
-            axs[0, 0].set_title(f"N-LL {iEpoch}")
+            axs[0, 0].set_title(f"N-LL {iEpoch}, time per 10 epochs: {dTime} s")
 
[…]

### 2024-05-03 src/estimate_ae.py

@@ -8,7 +8,7 @@ import torch
 import torch.nn as nn
 import torch.optim as optim
 
-from transformers import AutoTokenizer, GPT2LMHeadModel, PhrasalConstraint
+from transformers import AutoTokenizer, GPT2LMHeadModel, PhrasalConstraint, QuantoConfig
 from undecorated import undecorated
 from types import MethodType
 
@@ -25,7 +25,7 @@ device = (
     if torch.backends.mps.is_available()
     else "cpu"
 )
-device = "cpu"
+
 torch.set_default_device(device)
 print(f"Using {device} device")
 print(torch.tensor([1, 2, 3]).device)
@@ -95,6 +95,7 @@ ae_burnin = Autoencoder(iClaims, iEncodingDim, iOutputSize)
 dLearningRatePrefit = 1e1
 dLearningRate = 1
 dLearningRate = 5
+dLearningRate = 5e-1
 optimizer = optim.Adam(ae.parameters(), lr=dLearningRate)
 iEpoch = 0
 iEpochBurnin = 0
@@ -104,7 +105,10 @@ dEps = 1
 # INIT --------------------------
 torch.manual_seed(iSeed)
 
-model = GPT2LMHeadModel.from_pretrained(s_model, resume_download=True)
+quantization_config = QuantoConfig(weights="int8")
+# TODO: Try out flash-attention
+#model = GPT2LMHeadModel.from_pretrained(s_model, device_map=device, resume_download=True, quantization_config=quantization_config)
+model = GPT2LMHeadModel.from_pretrained(s_model, device_map=device, resume_download=True)
 # TODO: Link to this https://github.com/huggingface/transformers/issues/15552 (last comment)
 generate_with_grad = model.generate.__closure__[1].cell_contents
 model.generate_with_grad = MethodType(generate_with_grad, model)
@@ -118,17 +122,8 @@ tOHETarget = create_input_data(iClaims)
 
 lConstraints = [PhrasalConstraint(tokenizer(sentence, add_special_tokens=False).input_ids) for sentence in lTargetStrings]
 lForceWords = [tokenizer(sentence, add_special_tokens=False).input_ids for sentence in lTargetStrings]
-print("Force Words")
-print(lForceWords)
 
 tTargetStrings, lLengths = TokenizeClaims(lTargetStrings, tokenizer)
[…]

### 2024-05-03 meetings/07-05-2024_slides.qmd
working implementation

---
title: "Meeting --- Summary LLM"
author: "Bas Donkers, Dennis Fok, Finn-Ole Höner"
date: 2024-05-07
date-format: long
format: beamer
aspectratio: 169
incremental: false
include-in-header: ../beamer_preamble.tex
bibliography: ../references.bib
fig-align: center
fig-cap-location: bottom
fig-format: pdf
links-as-notes: true
number-depth: 1
eval: false
echo: true
fig-width: 12
fig-height: 8
---

# Updates

- Possible source for the spikes: The way the likelihoods are generated by the LLM.
  - We always select the respective token for each position in the target sequence.
  - However, if we generate a wrong token earlier in the sequence, then we select the probability of this token, given the wrong token.
  - Instead, we should still select the "chain" of probabilities for the target sequence
  - A possible solution: Apply a logits processor, that forces the model to always generate the correct sequence. Summary embedding that we train this way, still generate the target sequence with an unconstrained model, but should be more smooth to train!

"Transition scores for each vocabulary token at each generation step. Beam transition scores consisting of log probabilities of tokens conditioned on log softmax of previously generated tokens in this beam. Tuple of torch.FloatTensor with up to max_new_tokens elements (one element for each generated token), with each tensor of shape (batch_size*num_beams, config.vocab_size)." ([link](https://huggingface.co/docs/transformers/en/main_classes/text_generation#transformers.GenerationMixin.compute_transition_scores))

`.compute_transition_scores()` very similar to our functions; Computes the conditional probabilities for the generated sequence. Taking the sum, yields the likelihood of the generated sequence.

Does calling this function with the target sequence, yield the correct target likelihood?

### 2024-05-03 src/test.py

@@ -1,0 +1,220 @@
+# SETUP --------------------------
+from analysis.toolbox import *
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+
+import torch
+import torch.nn as nn
+import torch.optim as optim
+
+from transformers import AutoTokenizer, GPT2LMHeadModel, PhrasalConstraint, QuantoConfig
+from undecorated import undecorated
+from types import MethodType
+
+import time
+import pickle
+import io
+import os
+from tqdm import tqdm
+
+device = (
+    "cuda"
+    if torch.cuda.is_available()
+    else "mps"
+    if torch.backends.mps.is_available()
+    else "cpu"
+)
+device = "cpu"
+torch.set_default_device(device)
+print(f"Using {device} device")
+print(torch.tensor([1, 2, 3]).device)
+
+# MAGICKS --------------------------
+s_model = "gpt2"
+iSeed = 4523522
+sPadding = "left"
+sNameAEModel = "autoencoder_hair20_eos"
+
+lTargetStrings = [
+    "Experience 50% more visible shine after just one use.",
+]
+
+i_num_beams = 1
+i_no_repeat_ngram_size = 1
+i_num_return = 1
+iClaims = len(lTargetStrings)
+print(f"Number of claims: {iClaims}")
+iEncodingDim = 2  # Desired number of outtput dimensions
+iOutputSize = 768  # Number of output features
+iVocab = 50257
+iBurnIn = 500
+ae = Autoencoder(iClaims, iEncodingDim, iOutputSize)
+ae_burnin = Autoencoder(iClaims, iEncodingDim, iOutputSize)
+dLearningRatePrefit = 1e1
+dLearningRate = 1
+dLearningRate = 5
+dLearningRate = 5e-1
+optimizer = optim.Adam(ae.parameters(), lr=dLearningRate)
+iEpoch = 0
+iEpochBurnin = 0
+dEps = 1
+bStop = False
+
+torch.manual_seed(iSeed)
+
+quantization_config = QuantoConfig(weights="int8")
+# TODO: Try out flash-attention
+#model = GPT2LMHeadModel.from_pretrained(s_model, device_map=device, resume_download=True, quantization_config=quantization_config)
+model = GPT2LMHeadModel.from_pretrained(s_model, device_map=device, resume_download=True)
[…]

### 2024-05-05 meetings/07-05-2024_slides.qmd
batching works
runs and slides

@@
 
-- Possible source for the spikes: The way the likelihoods are generated by the LLM.
-  - We always select the respective token for each position in the target sequence.
-  - However, if we generate a wrong token earlier in the sequence, then we select the probability of this token, given the wrong token.
-  - Instead, we should still select the "chain" of probabilities for the target sequence
-  - A possible solution: Apply a logits processor, that forces the model to always generate the correct sequence.
-Summary embedding that we train this way, still generate the target sequence with an unconstrained model, but should be more smooth to train!
+Found the source for the spikes and fixed it: `.generate()` is the wrong method, need `.forward()` $\to$ wrong conditionals! Now maximizes the likelihood and faster!
 
-"Transition scores for each vocabulary token at each generation step.
-Beam transition scores consisting of log probabilities of tokens conditioned on log softmax of previously generated tokens in this beam.
-Tuple of torch.FloatTensor with up to max_new_tokens elements (one element for each generated token), with each tensor of shape (batch_size*num_beams, config.vocab_size)." ([link](https://huggingface.co/docs/transformers/en/main_classes/text_generation#transformers.GenerationMixin.compute_transition_scores))
+- New generated slogans seem better, space looks more interesting (have yet to do the "box" normalization)
+- Maybe even further performance gains by using `past_key_values`
 
-`.compute_transition_scores()` very similar to our functions; Computes the conditional probabilities for the generated sequence. Taking the sum, yields the likelihood of the generated sequence.
+- Idea: LLM judge to assess grammar / quality of slogans?
+- (MLFlow)
 
-Does calling this function with the target sequence, yield the correct target likelihood?
+\textbf{Admin}:
+
+- [ ] SLURM application
+- [ ] Set appointment for pre-defense
+- [ ] Find third assessor 
+- [ ] Complete data management survey

### 2024-05-13 meetings/07-05-2024_slides.qmd
improve monitoring

@@
 
-Found the source for the spikes and fixed it: `.generate()` is the wrong method, need `.forward()` $\to$ wrong conditionals! Now maximizes the likelihood and faster!
+Improved the optimization: `.generate()` is the wrong method, need `.forward()` $\to$ wrong conditionals! Now maximizes the likelihood and is faster!
 
-- New generated slogans seem better, space looks more interesting (have yet to do the "box" normalization)
-- Maybe even further performance gains by using `past_key_values`
+- Space looks different now; Does maximizing the likelihood collapse the encoding space?
+- Maybe even further performance gains by using `past_key_values`?
 
@@
 
+MPhil: Need to start working towards the paper now
+
+---
+
 \textbf{Admin}:
 
+- [ ] There seem to be issues with the access to the data for me now
 - [ ] SLURM application
@@
 - [ ] Complete data management survey
+
+---
+
+![Maxmimum Likelihood (nLL)](./figures/07-05-24-04.png)
+
+---
+
+![Encoded data points](./figures/07-05-24-03.png)
+
+---
+
+![Exploring the encoding space with generation, x are candidates](./figures/07-05-24-01.png)
+
+---
+
+![Cosine similarity between the candidate points](./figures/07-05-24-02.png)

### 2024-05-13 meetings/summary.qmd
improve monitoring

---
title: "BDS MPhil: Reverse Engineering GPT"
author: "Finn-Ole Höner; Supervisors: Bas Donkers, Dennis Fok"
date: last-modified
date-format: long
format: beamer
aspectratio: 169
incremental: false
pdf-engine: pdflatex
include-in-header: ../beamer_preamble.tex
bibliography: ../references.bib
fig-align: center
fig-cap-location: bottom
fig-format: pdf
links-as-notes: true
number-depth: 1
eval: false
echo: true
fig-width: 12
fig-height: 8
---

# Setup

**Can consumer preferences inform LLM text generation?**

- **Application:** Generating advertising claims with an LLM, based on consumer preferences.

- **Methodology:** Find input-embeddings for GPT, that generate a specific advertising claim (we call these "summary embeddings"). We model these "summary embeddings" with an Autoencoder, which gives us a "generation space" for advertising claims.

- **Toy-Example:** Tangible (e.g. "50% more visible shine after one use") and intangible (e.g. "Rediscover vibrant, joyful hair") advertising claims.

# Idea

We try to find an input-embedding that makes an LLM generate a specific output-sequence: Find the summary embedding $\boldsymbol{e^{*}}$ that maximizes the likelihood of the target-sequence $t_1, \ldots, t_L$ given the summary embedding $\boldsymbol{e}$:

$$
\boldsymbol{e^{*}} = \argmax_{\boldsymbol{e}} \sum_{i=1}^{L} \log p(t_1, \ldots, t_L | \boldsymbol{e}) \text{. }
$$

We impose restrictions on $\boldsymbol{e^{*}}$ by using an Autoencoder (AE) to model its elements. The hidden layer of this AE is our "generation space".

# Current State

- Summary embeddings generate target sequences.
- Separation of tangible and intangible claims in generation space (Figure 1).
- Grid-search exploration reveals "candidates" for new claims (red crosses) and "islands" regenerating training data claims (color-coded circles) (Figure 2).

**Next Steps:**

- Optimize optimization process.
- Enhance generation mode for clearer new claims.
- Integrate consumer preferences/ratings for generated claims.
---

::: {.columns}
::: {.column}
![Tangible and intangible claims in 2D-generation space.](./figures/07-05-24-03.png)
:::
::: {.column}
![Grid-search across 2D-generation space.](./figures/slides_pitch.png)
:::
:::

### 2024-05-13 writing/application_SLURM_small24.qmd
improve monitoring

Our research is relevant to the Creative Industry, as our algorithm can be used to personalize text media to specific consumer's of these media. Also, the novel approach that we propose of estimating a generative summary embedding, might find applications in this area to represent the information captured in a piece of writing.

#	Description

This project proposes an new method to estimate summary embeddings for Large Language Models (LLMs), that enable them to generate specific output sequences. We apply this novel method to tailor advertising claims, harnessing these "summary embeddings." Essentially, our technique reverse-engineers the generation of advertising claims into unique vector representations, which we correlate with market preferences for these claims. Additionally, we delve into the interaction between brands and these summary embeddings within the generation space. Our project aims to refine advertising strategies and identify market gaps, akin to perceptual maps. The potential marketing applications for these summary embeddings are extensive. Beyond personalizing advertising claims, they can potentially enhance recommendation engines and help with website morphing. Unlike existing personalization methods, our approach doesn't rely on a predefined set of offerings but can generate new offerings that weren't part of the training data.

# Scientific project description

The crux of our proposed method revolves around understanding how Language Models (LLMs) generate text—namely, by predicting the next token in a sequence of tokens. Our approach involves estimating summary embeddings, which serve as the input embeddings to the LLM. These embeddings are designed to maximize the likelihood of the LLM generating a specific output sequence. Through an autoencoder, we can model these input embeddings, capturing them in a shared, compact "generation space." Within this space, we can explore various points to generate new claims, evaluate their quality, and gauge the suitability of a summary embedding for a particular claim.

Following validation on a small dataset, our next step involves scaling up to a larger dataset sourced from a market research firm. These data consist of aggregated ratings provided by respondents regarding advertising claims, with no individual personal information included. Additionally, we possess metadata pertaining to these advertising claims, such as the associated brand. Our intention is to integrate this metadata, including brand information, into our autoencoder model. However, such computational tasks demand more power than a personal laptop can offer, making GPU resources necessary. The support of a NWO small compute grant, would enable us to pursue this project further.

# Technical project requirements

A typical application run on a CPU takes 36 hours and uses less than 16GB of memory, with a small LLM (GPT-2). Since, we will be computing on a GPU, we expect it to take around 12h on a GPU, as we observed a speed-up by a factor of 3 on Google Colab's T4 GPU. The input data is less than 1 GB in size. The application uses around 10 software packages in Python to execute the runs, among them the pytorch and transformers packages. The application runs are independent of each other. Since, we are developing and exploring a novel algorithm, we cannot put a fixed amount of runs as an upper bound. However, for our current application, we expect to need about 150 runs. We plan to employ a larger LLM (e.g. LLAMA 3) in our research, which is more computationally demanding than GPT-2. LLAMA 3 8B requires around 16GB of disk space and 20GB of VRAM, this is about 4 times more than GPT-2. For this reason, and to account for further developments of our algorithm, we add another 50% on top of this amount (12h x 150runs = 1,800 GPU hours) in our budget request (i.e. 2,700 GPU hours).

345,600 SBUs, 2,700 GPU hours

### 2024-05-13 writing/paper.qmd
improve monitoring

@@
+---
+author: Finn-Ole Höner (657110)
+title: "MPhil"
+date: last-modified
+date-format: long
+code-color-bg: lightgray
+header-includes:
+  \usepackage{fancyhdr}
+  \usepackage{fancyvrb}
+  \usepackage{float}
+  \usepackage{booktabs}
+  \usepackage{lscape}
+  \usepackage{amsmath}
+  \usepackage{graphicx}
+  \usepackage{algorithm}
+  \usepackage{algpseudocode}
+  \usepackage{kpfonts}
+  \DeclareMathOperator*{\argmin}{arg\,min}
+  \DeclareMathOperator*{\argmax}{arg\,max}
+pdf-engine: pdflatex
+format:
+    pdf:
+        documentclass: article
+        toc: false
+        number-sections: false
+        number-depth: 2
+        colorlinks: true
+        fig-pos: 'H'
+        fig-align: center
+        geometry:
+            - top = 20mm
+            - bottom = 20mm
+            - left = 20mm
+            - right = 20mm
+        papersize: dina4
+bibliography: references.bib
+echo: false
+eval: true
+messages: false
+error: false
+warning: false
+cache: false
+---
+
 # Introduction
@@
   - Social media, short videos TikTok, Images, Netflix, Audio / Podcasts
+
+- What is missing from preference learning? How is this different?
+  - Media products / things are different from products as attribute bundles
+    - We can directly represent the product itself?
+  - Brands and their role in advertising
 
@@
 <!-- TODO: Consider renaming this headline -->
+
 # Theory
@@
 
-The Autoencoder $\operatorname{AE}_{\Beta} \left(\cdot \right): \lbrace 0, 1 \rbrace^{C} \to \mathbb{R}^E$, with parameters $\Beta$
+The Autoencoder $\operatorname{AE}_{\mathcal{B}} \left(\cdot \right): \lbrace 0, 1 \rbrace^{C} \to \mathbb{R}^E$, with parameters $\mathcal{B}$
 
@@
 
+- Adding covariates to the AE
+
+- LLM judge to rate claims
+
 <!-- TODO: Consider renaming this headline -->
@@
 - Connection to other measures, e.g. syntax or rating of consumers
+
+- Different generation strategies, use beam search to get closer to max LL
 
@@
   - Wasteland, geometrical shape and where it comes from
-- Application to CBC data
+- Measures of fit
+  - areas in the encoding space (count out the pixels and compare to curvature of LL?)
+  - analysis of the trees
+  - Gap between most likely and second most likely token
+- Application to MR data
+
+- Comparison to BERT embeddings
 
@@
 - Generative AI for personalization
+- Brands
+- Wording
 
@@
 
+- Comparison to BERT embeddings
+
 - Expansion of this research

### 2024-05-13 writing/thesis_update.qmd
improve monitoring

---
bibliography: references.bib
title: "Decoding Consumer Preferences: Reverse-Engineering LLMs for Personalization"
subtitle: MPhil Thesis Update, Tentative Pre-Defense Date and Committee
format: pdf
---

# Update

- **Application:** Generating advertising claims with an LLM, based on consumer preferences.

- **Method:** Development of a new method. Find input-embeddings for LLM, that generate a specific advertising claim (we call these "summary embeddings"). We model these "summary embeddings" with an autoencoder, which gives us a "generation space" for advertising claims. These summary embeddings maximize the likelihood to generate a specific advertising claim.

- **Data:** Advertising claims from market research company, different brands, different products. Ratings by respondents on different dimensions, e.g. fit with brand, and preference over other claims.

- **Intended contribution:**
  - Introduction of reverse-engineered "summary embeddings"
  - Use of generative models for personalization in marketing
  - Explore the role of brands in the generation space
  - Learning about consumer preferences and linking them to a "generation-space", from which we can generate new advertising claims. Possibly, relate this to the idea of perceptual maps, i.e. mapping the different offerings on the market and identifying the "gaps" in-between.

- **Key references:**
    - @radford2018improving
    - @devlinBERTPretrainingDeep2018
    - @mullainathanPredictiveAlgorithmsAutomatic2023
    - @pangDeepGenerativeModels2023
    - @morozovWhereDoesAdvertising2024
    - @burnapProductAestheticDesign2023
    - @liFrontiersDeterminingValidity2024a
    - @schmalenseePerceptualMapsOptimal1988

- **Progress:** Data acquisition, development of algorithm, validation on toy-data, exploration of generation space.
  - Summary embeddings generate target sequences.
  - Separation of tangible and intangible claims in generation space (Figure 1).
  - Grid-search exploration reveals "candidates" for new claims (red crosses) and "islands" regenerating training data claims (color-coded circles) (Figure 2).

::: {layout-ncol=2 #fig-fig}
![Tangible and intangible claims in 2D-generation space.](../meetings/figures/07-05-24-03.png){#fig-fig1}

![Grid-search across 2D-generation space.](../meetings/figures/slides_pitch.png){#fig-fig2}

Two preliminary results on the toy-data. The toy-data are a collection of ChatGPT generated advertising claims for a hair-shampoo product, where the first half of claims is tangible and the second half is intangible in their wording.
:::

- **Next steps:** Improve training, introduce LLM judge for newly generated claims, incorporate e.g. brands in training process, perform analysis on market research data, explore relation to perceptual maps.

# Pre-Defense Admin

- **Expected defence date:** Fri, 28th June 2024

- **Thesis Committee**: Prof. Meike Morren (VU), Prof. Jonne Guyt (UvA); Supervisors: Prof. Fok (EUR), Prof. Donkers (EUR)

# References

### 2024-05-13 src/analysis/toolbox.py

@@ -16,6 +16,7 @@ import pickle
 import io
 import os
 
+import mlflow
 
 def TokenizeClaims(lClaims, tokenizer):
     # Tokenize the sentences
@@ -101,16 +102,17 @@ def plot_results(dTime, sNameAEModel, iEncodingDim, ae, lLoss, aLLHistory, lStep
             
             #print(f"Progress: {aLLHistory}\n")
             
-            fig, axs = plt.subplots(2, 4, figsize=(32, 16))
+            fig, axs = plt.subplots(2, 4, figsize=(14, 7))
             
-            # Plot 1
-            for i in range(tOHETarget.shape[0]):
-                axs[0, 0].plot(np.log1p(aLLHistory[0:iEpoch, i]), label=f"Claim {i + 1}")
-            axs[0, 0].axhline(y=np.log1p(dEps), linestyle='--', color='black', label="Threshold")
-            axs[0, 0].legend()
-            axs[0, 0].set_xlabel("Epoch")
-            axs[0, 0].set_ylabel("log1p (-LL)")
-            axs[0, 0].set_title(f"N-LL {iEpoch}, time per 10 epochs: {dTime} s")
+            if aLLHistory is not None:
+                # Plot 1
+                for i in range(tOHETarget.shape[0]):
+                    axs[0, 0].plot(np.log1p(aLLHistory[0:iEpoch, i]), label=f"Claim {i + 1}")
+                axs[0, 0].axhline(y=np.log1p(dEps), linestyle='--', color='black', label="Threshold")
+                axs[0, 0].legend()
+                axs[0, 0].set_xlabel("Epoch")
+                axs[0, 0].set_ylabel("log1p (-LL)")
+                axs[0, 0].set_title(f"N-LL {iEpoch}, time per 10 epochs: {dTime} s")
 
             # Plot 1
             axs[0, 1].plot(np.log1p(np.asarray(lLoss)))
@@ -175,15 +177,17 @@ def plot_results(dTime, sNameAEModel, iEncodingDim, ae, lLoss, aLLHistory, lStep
             axs[0, 3].set_ylabel("Stepsize")
             axs[0, 3].set_title(f"Grad Weights{iEpoch}")
             axs[0, 3].legend()
+            
+            mlflow.log_figure(fig, 'dashboard.png')
 
-            try:
-                plt.savefig(f"./models/training_monitor_{sNameAEModel}.png")
-            except:
[…]

### 2024-05-13 writing/paper.qmd
last10 monitor, set up paper

@@
 author: Finn-Ole Höner (657110)
-title: "MPhil"
+title: "Decoding Consumer Preferences with Large Language Models: A Novel Training Framework"
 date: last-modified

### 2024-05-13 src/analysis/toolbox.py

@@ -128,27 +128,27 @@ def plot_results(dTime, sNameAEModel, iEncodingDim, ae, lLoss, aLLHistory, lStep
                 axs[0, 2].plot(np.mean(np.asarray(lStepsWeightsEncoder)[0:iEpoch, i, :], axis=1), linestyle='--', label=f"Weights {i + 1}")
             axs[0, 2].set_xlabel("Epoch")
             axs[0, 2].set_ylabel("Stepsize")
-            axs[0, 2].set_title(f"Step made Encoder {iEpoch}")
+            axs[0, 2].set_title(f"SM Encoder {iEpoch}")
             #axs0, [2].legend()
             
             for i in range(iEncodingDim):
                 axs[1, 0].plot(np.mean(np.asarray(lStepsWeightsDecoder)[0:iEpoch, :, i], axis=1), linestyle='--', label=f"Weights {i + 1}")
             axs[1, 0].set_xlabel("Epoch")
             axs[1, 0].set_ylabel("Stepsize")
-            axs[1, 0].set_title(f"Step made Decoder {iEpoch}")
+            axs[1, 0].set_title(f"SM Decoder {iEpoch}")
             #axs[3].legend()
             
             # # # Plot 3
             axs[1, 1].plot(np.asarray(lStepsBiasEncoder)[:, 0], linestyle='--', label=f"Bias Encoder")
             axs[1, 1].set_xlabel("Epoch")
             axs[1, 1].set_ylabel("Stepsize")
-            axs[1, 1].set_title(f"Step made Bias Encoder {iEpoch}")
+            axs[1, 1].set_title(f"SM Bias Encoder {iEpoch}")
             axs[1, 1].legend()
 
             axs[1, 2].plot(np.asarray(lStepsBiasDecoder)[:, 0], linestyle='--', label=f"Bias Decoder")
             axs[1, 2].set_xlabel("Epoch")
             axs[1, 2].set_ylabel("Stepsize")
-            axs[1, 2].set_title(f"Step made Bias Decoder {iEpoch}")
+            axs[1, 2].set_title(f"SM Bias Decoder {iEpoch}")
             axs[1, 2].legend()
             
             # # # Plot 4
@@ -156,29 +156,120 @@ def plot_results(dTime, sNameAEModel, iEncodingDim, ae, lLoss, aLLHistory, lStep
             axs[1, 3].plot(np.asarray(lGradBiasEncoder), linestyle='--', label=f"Grad Bias Encoder")
[…]

### 2024-05-13 meetings/xx-xx-2024.qmd
insights on optimization issues

---
title: "Meeting --- Summary LLM"
author: "Bas Donkers, Dennis Fok, Finn-Ole Höner"
date: last-modified
date-format: long
format: beamer
aspectratio: 169
incremental: false
include-in-header: ../beamer_preamble.tex
bibliography: ../references.bib
fig-align: center
fig-cap-location: bottom
fig-format: pdf
links-as-notes: true
number-depth: 1
eval: false
echo: true
fig-width: 12
fig-height: 8
---

# Updates

- Gradient spikes first, then large step in decoder / encoder
- These steps seem to affect one hidden unit more strongly than the other
- Increase of Decoder Weight gradients seems to precede the increase of the -LL
- The other gradients seem to react to this jump
- Gradient-Decoder $\to$ Step $\to$ Other gradients $\to$ Step ...

---

::: {#fig-elephants layout-ncol=2}

![Epoch 970-980](./figures/dashboard_last10_980.png)

![Epoch 980-990](./figures/dashboard_last10_990.png)

Spikes: Notice how the decoder gradient norm increases first, causing a large step and in-turn affecting the other gradients. The scales of the y-axis differ across all plots.
:::

---

![](./figures/dashboard_last10_540.png)

---

![](./figures/dashboard_last10_160.png)

---

![](./figures/dashboard_last10_350.png)

### 2024-05-13 src/analysis/toolbox.py

@@ -178,7 +178,7 @@ def plot_results(dTime, sNameAEModel, iEncodingDim, ae, lLoss, aLLHistory, lStep
             axs[0, 3].set_title(f"Grad Weights")
             axs[0, 3].legend()
             
-            mlflow.log_figure(fig, 'dashboard.png')
+            mlflow.log_figure(fig, './plots/dashboard.png')
 
             # try:
             #     plt.savefig(f"./models/training_monitor_{sNameAEModel}.png")
@@ -269,7 +269,7 @@ def plot_results_last10(dTime, sNameAEModel, iEncodingDim, ae, lLoss, aLLHistory
             axs[0, 3].set_title(f"Grad Weights")
             axs[0, 3].legend()
             
-            mlflow.log_figure(fig, f'dashboard_last10_{iEpoch}.png')
+            mlflow.log_figure(fig, f'./plots/dashboard_last10_{iEpoch}.png')
 
             # try:
             #     plt.savefig(f"./models/training_monitor_{sNameAEModel}.png")

### 2024-05-13 meetings/xx-xx-2024.qmd
runs

@@
 - Gradient-Decoder $\to$ Step $\to$ Other gradients $\to$ Step ...
+- [`torch.nn.utils.clip_grad_norm_(ae.parameters(), max_norm=dMaxGrad)`](https://pytorch.org/docs/stable/generated/torch.nn.utils.clip_grad_norm_.html#torch.nn.utils.clip_grad_norm_) seems to help.
+This clips the gradient if the norm exceeds `dMaxGrad`.
+[See here](https://www.cs.toronto.edu/~rgrosse/courses/csc321_2017/readings/L15%20Exploding%20and%20Vanishing%20Gradients.pdf#page=6.78).
 
@@
 ![](./figures/dashboard_last10_350.png)
+
+---
+
+![Fixed clipping at 0.3. Later in the optimization this is too large.](./figures/dashboard_clipping_fixed.png)
+
+---
+
+Likely, you only want to clip true "outliers" in the gradients and not be too conservative.
+
+![You can overdo it with the gradient clipping.](./figures/dashboard_clip_01.png)

### 2024-05-13 src/analysis/toolbox.py

@@ -107,16 +107,16 @@ def plot_results(dTime, sNameAEModel, iEncodingDim, ae, lLoss, aLLHistory, lStep
             if aLLHistory is not None:
                 # Plot 1
                 for i in range(tOHETarget.shape[0]):
-                    axs[0, 0].plot(np.log1p(aLLHistory[0:iEpoch, i]), label=f"Claim {i + 1}")
-                axs[0, 0].axhline(y=np.log1p(dEps), linestyle='--', color='black', label="Threshold")
+                    axs[0, 0].plot(aLLHistory[0:iEpoch, i], label=f"Claim {i + 1}")
+                axs[0, 0].axhline(y=dEps, linestyle='--', color='black', label="Threshold")
                 axs[0, 0].legend()
                 axs[0, 0].set_xlabel("Epoch")
                 axs[0, 0].set_ylabel("log1p (-LL)")
                 axs[0, 0].set_title(f"N-LL {iEpoch}, time per 10 epochs: {dTime} s")
 
             # Plot 1
-            axs[0, 1].plot(np.log1p(np.asarray(lLoss)))
-            axs[0, 1].axhline(y=np.log1p(dEps), linestyle='--', color='black', label="Threshold")
+            axs[0, 1].plot(np.asarray(lLoss))
+            axs[0, 1].axhline(y=dEps, linestyle='--', color='black', label="Threshold")
             axs[0, 1].legend()
             axs[0, 1].set_xlabel("Epoch")
             axs[0, 1].set_ylabel("log1p (-LL)")
@@ -198,16 +198,16 @@ def plot_results_last10(dTime, sNameAEModel, iEncodingDim, ae, lLoss, aLLHistory
             if aLLHistory is not None:
                 # Plot 1
                 for i in range(tOHETarget.shape[0]):
-                    axs[0, 0].plot(np.log1p(aLLHistory[0:iEpoch, i]), label=f"Claim {i + 1}")
-                axs[0, 0].axhline(y=np.log1p(dEps), linestyle='--', color='black', label="Threshold")
+                    axs[0, 0].plot(aLLHistory[0:iEpoch, i], label=f"Claim {i + 1}")
+                axs[0, 0].axhline(y=dEps, linestyle='--', color='black', label="Threshold")
                 axs[0, 0].legend()
                 axs[0, 0].set_xlabel("Epoch")
                 axs[0, 0].set_ylabel("log1p (-LL)")
[…]

### 2024-05-15 meetings/summary.qmd
1-slide pitch

@@
 ---
-title: "BDS MPhil: Reverse Engineering GPT"
+title: "BDS MPhil"
 author: "Finn-Ole Höner; Supervisors: Bas Donkers, Dennis Fok"
@@
 ---
+
+# Decoding Marketers' Intentions: Deterministic Generation of Ad-Claims
+
+::: {.columns}
+::: {.column}
+
+**Setting**
+
+- Advertising claims with meta information & ratings
+
+**Contribution**
+
+- Novel "summary embeddings", that make an LLM generate a specific ad-claim.
+
+**What's done?**
+
+- Algorithm, validation, data acquisition, Snellius application
+
+**What's next?**
+
+- Empirical application, first-draft
+
+<!--
+
+- Match and mismatch of intention and wording?
+
+- Comparing focal brand with competitors, country, consumer ratings, brand fit, different languages, ranking of claims within set, compare branded and unbranded, sustainability messages, tangible & intangible
+
+- Different sub-categories of hair-care e.g. intense damage, hydro, color
+
+- mixtures of certain phrasings
+
+Using model for prediction? Have training claims and test claims embedded in the same space. Then use the training claims to predict the test claims based on encoding space.
+
+male and female targeting
+
+short-term, long-term benefits
+
+fit with brand measure
+
+Pre-training based on GPT generated claims
+
+Dairy: Naturality, nutrition, ingredients, ferments
+-->
+
+:::
+::: {.column}
+![2D-generation space.](./figures/slides_pitch.png)
+:::
+:::
 
@@
 - Integrate consumer preferences/ratings for generated claims.
+
 ---

### 2024-05-15 writing/paper.qmd
1-slide pitch

@@
         geometry:
-            - top = 20mm
-            - bottom = 20mm
-            - left = 20mm
-            - right = 20mm
+            - top = 1in
+            - bottom = 1in
+            - left = 1in
+            - right = 1in
         papersize: dina4
@@
 
+- Comparison to prompt engineering
+
 <!-- TODO: Consider renaming this headline -->
 
-# Theory
+# Methodology
 
-- Generative AI for text
+- Overview
+  - LLM
+  - maximize likelihood of target sequence
+  - autoencoder on these summary embeddings (+ covariates?)
+  - also yields low-dim generation space
+  - generation
+  - rating of claims
+
+- LLM
   - Attention
@@
   - GPT
-
-- Autoencoder
+  - The length $E$ summary-embedding $\boldsymbol{e}$.
+  - The length $T$ vector of target tokens $\boldsymbol{t}$, where the element $t_i$ is a unique integer code representing a token, e.g. in GPT-2, $50267$ is the End-of-Text token `<|eos|>`.
 
 - Maximize the likelihood to generate the data
-
-The Autoencoder $\operatorname{AE}_{\mathcal{B}} \left(\cdot \right): \lbrace 0, 1 \rbrace^{C} \to \mathbb{R}^E$, with parameters $\mathcal{B}$
-
-The log-likelihood function $\operatorname{\mathcal{L}} \left ( \cdot \right): \mathbb{R}^E \to \left (- \infty, 0 \right]$
-
-The length $E$ summary-embedding $\boldsymbol{e}$.
-
-The length $T$ vector of target tokens $\boldsymbol{t}$, where the element $t_i$ is a unique integer code representing a token, e.g. in GPT-2, $50267$ is the End-of-Text token `<|eos|>`.
+  - The log-likelihood function $\operatorname{\mathcal{L}} \left ( \cdot \right): \mathbb{R}^E \to \left (- \infty, 0 \right]$
 
@@
 
+- Autoencoder
+  - The Autoencoder $\operatorname{AE}_{\mathcal{B}} \left(\cdot \right): \lbrace 0, 1 \rbrace^{C} \to \mathbb{R}^E$, with parameters $\mathcal{B}$
   - Adding covariates to the AE
+
+- Figure of generation tree
+  - beam-search for generation
 
@@
 <!-- TODO: Consider renaming this headline -->
-# Implementation
+
+## Implementation
 
@@
 - Generative AI for personalization
+
 - Brands
+
 - Wording
+
+- Interactive dashboard to explore the embedding space
 
@@
 
-- Comparison to BERT embeddings
-
 - Expansion of this research
   - MAB around the fitted AE, to incorporate consumer feedback in online learning: Which areas of the space are valuable for which consumers?
+  - For online learning, could wrap the decoder into an MAB and let the MAB explore the space by pulling arms <!-- TODO: Which kind of MAB? -->
+  - Could also learn preferences online with this setup
   - Better LLM

### 2024-05-21 meetings/summary.qmd
ignore pngs in artifacts, monitor training, constant lr optimizer

@@
 
-**Setting**
-
-- Advertising claims with meta information & ratings
-
 **Contribution**
 
-- Novel "summary embeddings", that make an LLM generate a specific ad-claim.
+- Novel Method: A Summary Embeddings, that make an LLM generate a specific ad-claim.
+
+**Data**
+
+- Advertising claims with meta information & ratings from choice-experiment
 
@@
 - Empirical application, first-draft
+
+:::
+::: {.column}
+![2D-generation space](./figures/slides_pitch.png)
+:::
+:::
 
@@
 -->
-
-:::
-::: {.column}
-![2D-generation space.](./figures/slides_pitch.png)
-:::
-:::
 

### 2024-05-21 src/analysis/toolbox.py

@@ -125,18 +125,17 @@ def plot_results(dTime, sNameAEModel, iEncodingDim, ae, lLoss, aLLHistory, lStep
             # # Plot 3
             # print(np.asarray(lStepsWeightsEncoder).shape)
             for i in range(iEncodingDim):
-                axs[0, 2].plot(np.mean(np.asarray(lStepsWeightsEncoder)[0:iEpoch, i, :], axis=1), linestyle='--', label=f"Weights {i + 1}")
+                axs[0, 2].plot(np.max(np.asarray(lStepsWeightsEncoder)[0:iEpoch, i, :], axis=1), linestyle='--', label=f"Weights {i + 1}")
             axs[0, 2].set_xlabel("Epoch")
             axs[0, 2].set_ylabel("Stepsize")
             axs[0, 2].set_title(f"SM Encoder {iEpoch}")
             #axs0, [2].legend()
             
             for i in range(iEncodingDim):
-                axs[1, 0].plot(np.mean(np.asarray(lStepsWeightsDecoder)[0:iEpoch, :, i], axis=1), linestyle='--', label=f"Weights {i + 1}")
+                axs[1, 0].plot(np.max(np.asarray(lStepsWeightsDecoder)[0:iEpoch, :, i], axis=1), linestyle='--', label=f"Weights {i + 1}")
             axs[1, 0].set_xlabel("Epoch")
             axs[1, 0].set_ylabel("Stepsize")
             axs[1, 0].set_title(f"SM Decoder {iEpoch}")
-            #axs[3].legend()
             
             # # # Plot 3
             axs[1, 1].plot(np.asarray(lStepsBiasEncoder)[:, 0], linestyle='--', label=f"Bias Encoder")
@@ -216,14 +215,14 @@ def plot_results_last10(dTime, sNameAEModel, iEncodingDim, ae, lLoss, aLLHistory
             # # Plot 3
             # print(np.asarray(lStepsWeightsEncoder).shape)
             for i in range(iEncodingDim):
-                axs[0, 2].plot(np.mean(np.asarray(lStepsWeightsEncoder)[-10:, i, :], axis=1), linestyle='--', label=f"Weights {i + 1}")
+                axs[0, 2].plot(np.max(np.asarray(lStepsWeightsEncoder)[-10:, i, :], axis=1), linestyle='--', label=f"Weights {i + 1}")
             axs[0, 2].set_xlabel("Epoch")
[…]

### 2024-05-28 meetings/xx-xx-2024.qmd
improvements on gradients

--- 2024-05-13 meetings/xx-xx-2024.qmd
@@
 - Gradient spikes first, then large step in decoder / encoder
-- These steps seem to affect one hidden unit more strongly than the other
-- Increase of Decoder Weight gradients seems to precede the increase of the -LL
 - The other gradients seem to react to this jump
-- Gradient-Decoder $\to$ Step $\to$ Other gradients $\to$ Step ...
+- Gradient $\to$ Step $\to$ Other gradients $\to$ Step $\to$ -LL jumps 
+- [`torch.nn.utils.clip_grad_norm_(ae.parameters(), max_norm=dMaxGrad)`](https://pytorch.org/docs/stable/generated/torch.nn.utils.clip_grad_norm_.html#torch.nn.utils.clip_grad_norm_) seems to help.
+This clips the gradient if the norm exceeds `dMaxGrad`.
+[See here](https://www.cs.toronto.edu/~rgrosse/courses/csc321_2017/readings/L15%20Exploding%20and%20Vanishing%20Gradients.pdf#page=6.78).
 
@@
 
-![](./figures/dashboard_last10_540.png)
+![More spikes I](./figures/dashboard_last10_540.png)
 
@@
 
-![](./figures/dashboard_last10_160.png)
+![More spikes II](./figures/dashboard_last10_160.png)
 
@@
 
-![](./figures/dashboard_last10_350.png)
+![More spikes III](./figures/dashboard_last10_350.png)
+
+---
+
+Optimization with gradient-clipping at $dMaxGrad = 0.3$.
+
+![](./figures/dashboard_clipping_fixed.png)
+
+---
+
+Optimization with gradient-clipping at $dMaxGrad = 0.1$. Deeper in the optimization those spikes seem to become a problem again, at a lower scale. Oscill. of gradients later in the optimization.
+
+![Clipping at 0.1](./figures/dashboard_clip_01.png)
+
+---
+
+Optimization with gradient-clipping at $dMaxGrad = 0.05$. This model also has a normalization layer after the hidden layer.
+
+![Clipping at 0.05 and normalization layer](../mlartifacts/437740661089244805/50054b1717e444eeb1d4acd408e1e77d/artifacts/plots/dashboard.png)

### 2024-06-01 writing/paper.qmd
analyze LL, draft of modifier

@@
 - What is missing from preference learning? How is this different?
-  - Media products / things are different from products as attribute bundles
+  - Media products/things are different from products as attribute bundles
     - We can directly represent the product itself?
@@
 - EU AI Act
+  - [Content generation is considered low risk](https://www.growthloop.com/university/article/generative-ai-in-marketing)
+
+- Main risk of Gen AI stem from unpredictable behaviors, as generation is based on random samples from a highly non-linear model
+  - lawsuits
+  - damages to brands
+  - missed opportunities
+  - tailoring of models based on prompt engineering and RL-HF
 
@@
 
+We reverse-engineer the input to a Large Language Model (LLM), such that it generates a pre-defined text.
+
+For a given target sequence, we find the input-embedding that maximizes the likelihood to generate this target sequence with a unidirectional LLM (e.g.
+GPT, @radford2018improving).
+We call this input embedding, "summary embedding" of the target sequence.
+
+We restrict the degrees of freedom in our search for these input embeddings, by making the elements of the input embedding linear combinations of a few hidden factors.
+
+In turn, these hidden factors yield a generation space, where we locate our target sequences and search for new sequences.
+
+@goodfellowDeepLearning2016
+
+- Autoencoder
+  - The Autoencoder $\operatorname{AE}_{\mathcal{B}} \left(\cdot \right): \lbrace 0, 1 \rbrace^{C} \to \mathbb{R}^E$, with parameters $\mathcal{B}$
+  - Adding covariates to the AE
+
 - Overview
@@
 
-- LLM
-  - Attention
-  - Transformer
-  - GPT
-  - The length $E$ summary-embedding $\boldsymbol{e}$.
-  - The length $T$ vector of target tokens $\boldsymbol{t}$, where the element $t_i$ is a unique integer code representing a token, e.g. in GPT-2, $50267$ is the End-of-Text token `<|eos|>`.
+Unidirectional LLMs, such as GPT or LLaMa [@radford2018improving; @metaMetaLlama], are built for text-generation.
+They predict the next token in a sequence of tokens autoregressively, therby, these models only work from left-to-right (or right-to-left).
+Bidirectional LLMs, such as BERT [@devlinBERTPretrainingDeep2018], are built for representing text.
+They predict a token given the context around this token, and can hence read the text both from left-to-right and right-to-left.
+Both types of models share two note-worthy components, that account for a large share of their power.
+One, developers pre-train both types of models on large bodies of text, to learn about the structure of language itself.
+Two, by using a transformer architecture [@vaswani2017attention], which enables these language models to learn dependencies between words, such as negations (short-range dependency), or document structures, e.g.
+salutations in letters (long-range dependency).
 
-- Maximize the likelihood to generate the data
-  - The log-likelihood function $\operatorname{\mathcal{L}} \left ( \cdot \right): \mathbb{R}^E \to \left (- \infty, 0 \right]$
+- Notation
+  - A large language model $\mathcal{M}$, with parameters $\Omega$.
+  - The length $E$ summary-embedding $\boldsymbol{e}^{*}$.
+  - The length $T$ vector of target tokens $\boldsymbol{t}$, where the element $t_i$ is a unique integer code representing a token, e.g.
+in GPT-2, the number $50267$ is the End-of-Text token: `<|eos|>`.
+  - The log-likelihood function $\operatorname{\mathcal{L}}_{\mathcal{M}} \left ( \boldsymbol{i}, \boldsymbol{o} \right): \mathbb{R}^E \to \left (- \infty, 0 \right]$, that returns the log-likelihood for LLM $\mathcal{M}$ to generate the output sequence $\boldsymbol{o}$, given the input embedding $\boldsymbol{i}$.
+
+$$
+\boldsymbol{e^{*}} = \argmax_{\boldsymbol{e}} \operatorname{\mathcal{L}}_{\mathcal{M}} \left ( \boldsymbol{e}, \boldsymbol{t} \right) = \argmax_{\boldsymbol{e}} \sum_{i=1}^{L} \log p(t_1, \ldots, t_L | \boldsymbol{e})
+$$
+
+We aim to find a summary embedding, $\boldsymbol{e^{*}}$, that maximizes the likelihood of generating a given target sequence, $\lbrace t_i \rbrace_{i=1}^{L}$, with a given LLM, $\mathcal{M}$.
+We use a gradient-based optimization algorithm to maximize this likelihood.
+We initialize the summary embedding as the element-wise average of the embedding of the target sequence.
+We then generate a sequence of length $L$, with the LLM and the current input embedding.
+For the resulting output layer, we compute the likelihood that this layer will generate the target sequence.
+We backpropagate the gradients and adjust the input layer accordingly.
+We repeat this process until the optimization converges.
+We want to investigate further improvements to the implementation, such as computing the likelihood contributions of the tokens in a distributed fashion.
+In Algorithm 1, we summarize the training process.
+
+<!-- TODO: Update this algorithm -->
+
+\begin{algorithm}
+\begin{algorithmic}
+\label{alg:training}
+\caption{Training of Summary Embeddings}
+\State Initialize $\boldsymbol{s}$
+\State $i \gets 0$
+\While{$Convergence == False$}
+    \State $i \gets i + 1$
+    \State $\boldsymbol{o}^{(i)} \gets \mathcal{M}(\boldsymbol{s})$
+    \State $l^{(i)} \gets \operatorname{Likelihood}(\boldsymbol{o}^{(i)}, \lbrace t_i \rbrace_{1}^{L})$
+    \State $\boldsymbol{s}^{(i + 1)} \gets \operatorname{GradientDecent}(\boldsymbol{s}^{(i)}, l^{(i)})$
+    \State $Convergence \gets \operatorname{CheckConvergence}(\boldsymbol{s}^{(i + 1)}, \boldsymbol{s}^{(i)})$
+\EndWhile
+\end{algorithmic}
+\end{algorithm}
 
@@
 <!-- TODO: Good idea or not? -->
-To express that we generate tokens with the LLM, $\mathcal{M}$, based on length $E$ input-embedding $\boldsymbol{e}$, we use the notation $\mathcal{M}(\boldsymbol{e})$.
-To express the number of tokens, $k$, that we generate, we use the subscript $\mathcal{M}_k(\boldsymbol{e})$.
-This function call returns an output vector of length $V$, where $V$ denotes the vocabulary size.
-We denote the output vector as $\boldsymbol{o}$.
-Each element $o_v$ of this output vector represents the probability of generating token $v$ as the next token.
-We predict the next token, by selecting the largest element of $\boldsymbol{o}$.
-When $k = 1$, we use the previous tokens to predict the next token.
-When $k > 1$, we append the predicted tokens autoregressively to the input embedding, such that we can predict more than one token.
-While there are many ways to design this procedure and select tokens along the way, we perform a standard greedy procedure, where we select the token with the highest probability at each step.
-We denote the probability that $\mathcal{M}$ generates the target sequence, given the input embedding, as $p(t_1, \ldots, t_L | \boldsymbol{e})$.
-For a given Large Language Model, we define $\boldsymbol{e^{*}}$,
-
-$$
-\boldsymbol{e^{*}} = \argmax_{\boldsymbol{e}} \sum_{i=1}^{L} \log p(t_1, \ldots, t_L | \boldsymbol{e}) \text{, }
-$$
 
@@
 
-- Autoencoder
-  - The Autoencoder $\operatorname{AE}_{\mathcal{B}} \left(\cdot \right): \lbrace 0, 1 \rbrace^{C} \to \mathbb{R}^E$, with parameters $\mathcal{B}$
-  - Adding covariates to the AE
-
 - Figure of generation tree
-  - beam-search for generation
+  - For generation of sequences, the LLM autoregressively predicts the next token, given the previous tokens.
+At each step in the generation, a selection of the next token needs to take place, for which there exist various procedures (CITE).
+As we are optimizing with respect to the joint likelihood of the target sequence, we also need to generate with a procedure that generates this most likely sequence.
+Finding this sequence from the generation tree is a combinatorical problem and computationally not feasible.
+Hence, we use the beam-search heuristic to make our generation.
+Rather than only tracking the most likely token at each generation step, beam search also tracks the $n$-most likely tokens.
+The more beams we search, the more likely we are to find a sequence with a higher likelihood, then we would have found otherwise [@huggingfaceGenerateText].
 
@@
 
-## Implementation
+# Data
+
+We obtain data from a market research company, that includes advertising claims for different FMCG.
+In our application, we focus on dairy and hair shampoo products.
+These data, stem from different variations of choice-based conjoint studies [@eggersChoiceBasedConjointAnalysis2022], which we cannot further disclose due to confidentiality agreements.
+
+The data includes choice-experiments aggregated at the level of the advertising claims, hence we do not have responses of individuals.
+These measurements of consumer-preferences are on multiple dimensions, such as relevance, brand fit, or an overarching rating of the claim.
+
+The advertising claims were designed for markets in different countries and are in english (US, UK) or translated into english (others). The date of the study is also included in the data.
+
+We also have information on the marketers motivation behind the design of an advertising claims, such as whether she designed the claims to pronounce the health benefits of a yoghurt, or the taste yoghurt.
+
+The data also include the brand of a product.
+
+- Descriptives of the data
+
+- Prepocessing of the data
+
+<!-- TODO: Provide examples that are created with GPT, i.e. are fake -->
+
+# Results
+
+### Implementation
 
@@
 
-# Data
-
-- Descriptives of the data
-
-- Prepocessing of the data
-
-# Results
-
 - Optimization
   - Computation time, scaling
+
 - Encoding space
@@
   - Wasteland, geometrical shape and where it comes from
+
 - Measures of fit
@@
   - Gap between most likely and second most likely token
-- Application to MR data
 
 - Comparison to BERT embeddings
+- Comparison to prompt engineering results
+
+# Empirical Application
+
+- Application to MR data
 
@@
   - Computation time
+  - Benefits over existing approaches in practice?
 
@@
   - Image, video, website, audio data. Perhaps website a good application. Perhaps audio a rather simple way to make this multi-modal (content and sound)
+
+{{< pagebreak >}}
+
+# References

### 2024-06-04 writing/factor_tikz.tex
graph NN

\def\layersep{2.5cm}
\begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=\layersep]
    \tikzstyle{every pin edge}=[<-,shorten <=1pt]
    \tikzstyle{neuron}=[circle,fill=black!25,minimum size=17pt,inner sep=0pt]
    \tikzstyle{LLM}=[draw,thick,minimum width=2cm, minimum height=2cm]
    \tikzstyle{NLL}=[draw,thick,minimum width=1cm, minimum height=1cm]
    \tikzstyle{bias neuron}=[neuron, fill=white];
    \tikzstyle{input neuron}=[neuron, fill=green!50];
    \tikzstyle{output neuron}=[neuron, fill=red!50];
    \tikzstyle{hidden neuron}=[neuron, fill=blue!50];
    \tikzstyle{annot} = [text width=7em, text centered]

    \node[input neuron] (I-1) at (0, -1) {$i_1$};
    \node[input neuron] (I-2) at (0, -2) {$i_2$};
    \node[input neuron] (I-3) at (0, -3) {$\vdots$};
    \node[input neuron] (I-4) at (0, -4) {$i_I$};
    \node[bias neuron] (I-5) at (0, -5) {$B^{E}$};

    \node[hidden neuron] (H-1) at (\layersep, -1.5) {$h_1$};
    \node[bias neuron] (H-2) at (\layersep, -2.5) {$\vdots$};
    \node[hidden neuron] (H-3) at (\layersep, -3.5) {$h_H$};
    \node[bias neuron] (H-4) at   (\layersep, -4.5) {$B^{D}$};

    \node[output neuron] (O-1) at (\layersep * 2, -0.5) {$e_1$};
    \node[output neuron] (O-2) at (\layersep * 2, -1.5) {$e_2$};
    \node[bias neuron] (O-3) at (\layersep * 2, -2.5) {$\vdots$};
    \node[output neuron] (O-4) at (\layersep * 2, -3.5) {$e_{E-1}$};
    \node[output neuron] (O-5) at (\layersep * 2, -4.5) {$e_E$};

    \node[LLM] (M-1) at (\layersep * 3, -2.5) {LLM};

    \node[NLL] (N-1) at (\layersep * 4, -2.5) {$- log \mathcal{L}$};

    \foreach \source in {1,...,5}
      \foreach \dest in {1,...,3}
        \path (I-\source) edge (H-\dest);

    \foreach \source in {1,...,4}
      \foreach \dest in {1,...,5}
        \path (H-\source) edge (O-\dest);

    \foreach \source in {1,...,5}
        \path (O-\source) edge (M-1);

    \draw (I-1) -- (H-1) node [midway, fill=white, minimum size=2pt] {$W_{1,1}^{E}$};
    \draw (H-1) -- (O-1) node [midway, fill=white, minimum size=2pt] {$W_{1,1}^{D}$};

    \path (M-1) edge (N-1);

    % Annotate the layers
    \node[annot,above of=H-1, node distance=2cm] (hl) {Hidden ($H$)};
    \node[annot,right of=hl] {Summary-Embedding ($E$)};
    \node[annot,left of=hl] {Input ($I$)};

\end{tikzpicture}

### 2024-06-04 writing/paper.qmd
graph NN

@@
   \usepackage{kpfonts}
+  \usepackage{tikz}
   \DeclareMathOperator*{\argmin}{arg\,min}
@@
 
+- Neural Text degeneration (@ariholtzmanCuriousCaseNeural2020)
+- BERT and RoBERTa @devlinBERTPretrainingDeep2018, @yinhanliuRoBERTaRobustlyOptimized2019, @jeffreypenningtonGloveGlobalVectors2014, @tomasmikolovDistributedRepresentationsWords2013 @tomasmikolovEfficientEstimationWord2013
+
 <!-- TODO: Consider renaming this headline -->
@@
 # Methodology
+
+- Distributional hypothesis of language @sahlgren2008distributional
 
@@
 GPT, @radford2018improving).
-We call this input embedding, "summary embedding" of the target sequence.
+We call this input embedding, the "summary embedding" of the target sequence.
 
 We restrict the degrees of freedom in our search for these input embeddings, by making the elements of the input embedding linear combinations of a few hidden factors.
+This amounts to a linear factor model.
+
+![Factor model in Neural Network form](factor_tikz.tex)
 
@@
 
-<!-- TODO: Consider renaming this headline -->
-
 # Data
@@
 
-- Optimization
-
 - Connection to other measures, e.g. syntax or rating of consumers
@@
 - Optimization
+  - @diederikp.kingmaAdamMethodStochastic2014
   - Computation time, scaling
+  - Issues with spikes; Figures where they come from
 
@@
 
-- Comparison to BERT embeddings
-- Comparison to prompt engineering results
+- Comparison to BERT embeddings?
+- Comparison to prompt engineering results?
 
@@
 
+- Costs of tailoring advertising claims
+  - When we take current prices for the [GPT-4o LLM; June 2024](https://openai.com/api/pricing/) (\$5 per 1 Mio tokens input, \$15 per 1 Mio tokens output), then training a dataset of 100 advertising claims, for 5,000 epochs will cost \$150.
+Freelancers on platforms such as [Upwork.com](https://www.upwork.com/) charge between \$20 and \$200 per hour for work on tasks such as SEO optimization, copy writing, and creation of advertising claims.
+    - Online services for conjoint studies, such as [Conjointly.com](https://conjointly.com/) and [Sawtooth](https://sawtoothsoftware.com/pricing) charge a few thousand dollars per year to use their services.
+Leveraging the resulting insights as much as possible, by using generative AI, could hence make market research more cost effective.
+
 # Discussion
+
+- Using embedding rather than identity matrix as the input turns this into an autoencoder, rather than a factor model.
+According to @goodfellowDeepLearning2016, generation from factor models leads to a combination of the learned features in these factors, but not to the generation of sensible data points.
+Existing, autoencoder based, approaches that have succeeded in generation tasks, might be more suitable for this task (e.g.
+CITE: Variational Autoencoder).
 
@@
   - Benefits over existing approaches in practice?
+  - Conceptual issues of using AI in decision making @cathyoneilWeaponsMathDestruction2016; Gender biases @emilyshengWomanWorkedBabysitter2019
+  - How to evaluate the quality of the generated ad claims @tatsunorihashimotoUnifyingHumanStatistical2019
 

### 2024-06-14 writing/paper.qmd
writing session

@@
 author: Finn-Ole Höner (657110)
-title: "Decoding Consumer Preferences with Large Language Models: A Novel Training Framework"
+title: "working title"
 date: last-modified
@@
   \usepackage{kpfonts}
+  \usepackage{tcolorbox}
   \usepackage{tikz}
@@
 
-- Media products increase relevance of content
-  - Social media, short videos TikTok, Images, Netflix, Audio / Podcasts
+> *"Prompt engineering is the art of communicating eloquently to an AI."----Greg Brockman, President OpenAI*
 
+<!-- TODO: Have more of an advertising claims example? -->
+<!-- TODO: Make this more about a summary, than about generation and prompts? Build the bridge of the two -->
+<!-- TODO: Talk about low-dim representation -->
+
+Generative Large Language Models (LLM) are powerful, but unreliable tools for generating text.
+[A study by the NGO "Algorithm Watch"](https://algorithmwatch.ch/en/wp-content/uploads/2023/12/AlgorithmWatch_Generative-AI-and-elections_An-analysis-of-Microsofts-Bing-Chat.pdf), which is partially funded by the European Union, finds that Microsoft's chatbot Bing Chat, delivers wrong information in a third of its replies to election related questions, posing a risk to exacerbate desinformation.
+
+Prompt engineering defines the task of finding the right prompt to generate a desired output.
+Typically, this is an interactive and iterative process between human and LLM.
+There are certain writing techniques that improve the quality of
+the LLM's response, such as providing examples for what the desired output should look like.
+There are various [blog posts](https://huggingface.co/docs/transformers/en/tasks/prompting#best-practices-of-llm-prompting) and books [e.g.
+@phoenix2024prompt] about how to write good prompts, in fact businesses are [hiring "Prompt Engineers"](https://indeed.com/q-prompt-engineer-jobs.html?vjk=98ac28acfd1328d7) even on dedicated [job boards](https://prompt-engineering-jobs.com/jobs/prompt-engineer-qwvmju/).
+Despite these, supposedly low-barrier, resources, end users without AI knowledge still struggle with prompt engineering [@zamfirescu2023johnny].
+Not only can prompt engineering pose a [security risk](https://www.promptingguide.ai/risks/adversarial), but the process is also fuzzy, hard to replicate, and might still lead to unpredictable behaviors of the model.
+These issues pose risks for organizations, as the obscurity of manual prompt engineering can lead to loss of know-how, and the unpredictiability of the model can lead to lawsuits, damages to brands, or missed opportunities.
+
+Problems of reproduceability, transparency, and documentation, become critical for organizations with the advent of the [EU AI Act](https://digital-strategy.ec.europa.eu/en/policies/regulatory-framework-ai).
+The act requires for "High risk applications", that the AI model features, e.g.
+"adequate mitigation systems", "traceability of results", and, "robustness".
+High risk applications are not common and important for society, they include applications e.g.
+in education, employment, and public services.
+
+We propose a new type of numeric document summary, which captures all the information contained in the document, as it can be used to recreate the document itself.
+This document summary comes in the form of a numeric vector, which we call the summary embedding.
+To obtain this summary embedding, we maximize the likelihood of generating the document with a Large Language Model (LLM).
+In this sense, these summary embeddings are optimal.
+We show, that these embeddings capture inherent information about the document, can be used for classification, and generation of new documents.
+There are only practical requirements for estimating these summary embeddings, namely that backpropagation of gradients is possible from the likelihood to the input of the LLM.
+
+Beyond the methodological contribution, we show a marketing application of these summary embeddings and make a step towards using Generative AI in market research and design of the marketing-mix.
+We find that we can use these embeddings to capture design motivations of marketers behind advertising claims, and establish a measure for linguistic uniqueness of an advertising claim.
+Suprisingly, we find that this linguistic uniqueness is negatively correlated with the perceived uniqueness of an advertising claim by consumers.
+We explore these linguistic features and find that claims with certain words and their synonyms, cluster together, e.g.
+claims that emphasize the taste of a yoghurt drink or present it as a breakfast.
+Furthermore, we show that claims that live in certain areas of a low dimensional representation of the summary embeddings, are more likely to be rated favorably by consumers, but that small differences in wording can lead to relatively large differences in evaluation.
+This suggests a two-step approach to designing a market research study for advertising claims, where the first stage should focus on a wide exploration of the design space, while the second stage should focus on a fine-grained evaluation of the most promising area in this design space.
+We end with a sketch on how these summary embeddings can be used in an AI augmented design process, similar to applications in image data by @burnap2023product and @ludwigMachineLearningTool2023.
+
+<!--
+
+- Consumer preferences in latent generation space & Ad-claims
+  - Market research is expensive. How can we leverage these insights better to improve its benefits?
   - What is missing from preference learning? How is this different?
@@
     - Brands and their role in advertising
-
-- Market research is expensive. How can we leverage these insights better to improve its benefits?
-
-- EU AI Act
-  - [Content generation is considered low risk](https://www.growthloop.com/university/article/generative-ai-in-marketing)
-
-- Main risk of Gen AI stem from unpredictable behaviors, as generation is based on random samples from a highly non-linear model
-  - lawsuits
-  - damages to brands
-  - missed opportunities
-  - tailoring of models based on prompt engineering and RL-HF
-
-- GDPR
-  - Challenges to cookies in website morphing
-
+  - We can view products as attribute bundles and get utilities for these attributes 
+  - How do we do this for products that cannot be "discretized"?
   - Generative AI for personalization
@@
     - True personalization: Generate solution for each individual consumer
+  - TODO: Add more product claim specific
+  - Advertisements can be effective, @daiFrontiersWhichFirms2023a observe an increase in purchase intentions for restaurants that adapt online advertising by 7%-19%.
+  - they are relevant @kongMadeUSAClaims2021
+  - costs
+  - need flexibility in online settings
+  - closely related problems, such as email targeting @sahniPersonalizationEmailMarketing2018
+  - advertising is one of the marketing-mix instruments which companies use to position themselves in the market and shapes the incentives of the system itself @moorthyTargetingAdvertisingSpending2023 @renAdvertisingContentCreation2023
+  - design question
+  - Ban of cookie-based advertising due to GDPR, adapting to observable behavior in live sessions or to context on the website itself becomes more important. GDPR
+  - Challenges to cookies in website morphing
+  - @beckerDoesItPay2019, point to the nuanced role that authenticity plays in advertising, especially with respect to its alignment with the brand and product. 
+  - @edelmanBrandingDigitalAge2010 Concept of the consumer decision journey, product claims could vary across multiple visits or encounters of the same customer
 
-- Consumer preferences in latent generation space
-  - We can view products as attribute bundles and get utilities for these attributes <!-- TODO: What about interactions of attributes? -->
-  - How do we do this for products that cannot be "discretized"?
+-->
 
-- Novel training framework
-  - Circumvent issues with the LLMs decoder
-  - Input embedding represents the information contained in the sequence, as it is the starting point which makes the LLM generate the target sequence
+## Relevant Literature
 
-- Comparison to prompt engineering
+In the following, we compare our summary embeddings with three established methods to summarize documents: The Bidirectional Encoder Representations from Transformers (BERT) model's `[CLS]` token [@devlinBERTPretrainingDeep2018], pooled word embeddings [@shenBaselineNeedsMore2018], and prompt engineering approaches [@@huangenhanced].
 
-- Neural Text degeneration (@ariholtzmanCuriousCaseNeural2020)
-- BERT and RoBERTa @devlinBERTPretrainingDeep2018, @yinhanliuRoBERTaRobustlyOptimized2019, @jeffreypenningtonGloveGlobalVectors2014, @tomasmikolovDistributedRepresentationsWords2013 @tomasmikolovEfficientEstimationWord2013
+The BERT model is a transformer [see @vaswani2017attention] based model to represent text.
+Its developers have pre-trained BERT on a large corpus of text, using two different self-supervised learning objectives.
+The first objective is the so called "Masked Language Model" task.
+Here, the researchers replace a random selection of tokens in the text by a mask token, and the goal of the BERT model is to predict which token is underlying this mask.
+In order to learn about the information captures in a sentence, the researchers also pose the "Next Sentence Prediction" task to the model.
+For this training objective, the BERT model receives a pair of two sentences and needs to predict whether the second sentence follows the first one in a text.
+This task trains the `[CLS]` token: As it is pre-pended before every such sentence pair, it learns a representation of the information contained in the focal sentence.
+In practice, the output for a `[CLS]` token serves as a powerful feature e.g.
+for sentence classification.
+BERT embeddings are versatile, but can also be fine-tuned to a specific task such as text summarization.
+However, such a fine-tuning requires the user to compile a dataset of text-summary and text pairs, that are representative of their application.
+These BERT representations are only descriptive and cannot be used for the generation of new text [see @devlinBERTPretrainingDeep2018].
+There exist expansions to BERT, such as RoBERTa by @yinhanliuRoBERTaRobustlyOptimized2019, which improve the optimization procedure.
 
-<!-- TODO: Consider renaming this headline -->
+@shenBaselineNeedsMore2018 points to the value of using pooled word embeddings to represent documents.
+The idea behind a pooled word embedding is, that we can represent a sequence of tokens by aggregating the word-embeddings of these tokens.
+There are different methods we can use for this aggregation, some examples are the use of taking the average across the tokens (mean-pooling) or taking the maximum value across the tokens (max-pooling).
+There are various types of word embeddings that we could use for the pooling, often, these are non-attention word embeddings, such as Word2Vec [@tomasmikolovDistributedRepresentationsWords2013; @tomasmikolovEfficientEstimationWord2013] or Global Vectors (GLoVe) [@jeffreypenningtonGloveGlobalVectors2014].
+These pooled word embeddings, cannot be used for text generation and perform worse than BERT embeddings in language tasks, as they do not employ the attention mechanism and loose information in the aggregation step [see e.g.
+@onanHierarchicalGraphbasedText2023].
+
+We can also summarize documents by passing these to an LLM and prompting the model to write a summary for us [e.g.
+@huangenhanced; @chakrabortyAbstractiveSummarizationEvaluation2024].
+However, these summaries are textual, not deterministic, and their quality depends on the formulation of the prompt.
+For example, @liuLostMiddleHow2023 find that answers of LLMs are better, when we place relevant information at the beginning or end of the prompt, while inserting emotions into prompts also improves the quality of the LLMs response, in some cases even doubling its performance [see @liLargeLanguageModels2023].
+Another issue is that these summaries can vary in length, and one needs to specify how detailed a summary should be.
+The difference in length between the summary and focal document, also makes it challenging to evaluate how much of the information of the focal document is captured by the summary [@chakrabortyAbstractiveSummarizationEvaluation2024].
+If these summaries are of a high quality, i.e.
+represent the information in the focal document well, then they can be useful when we want to share a written summary of a document.
+However, for many data science and automation tasks, we need a reliable numeric representation of the document, which is the focus of this paper.
+
+In this work, we propose a method that optimizes the input prompt to an LLM, such that we obtain a desired outcome.
+We call these input prompts "document summaries", as the perfectly replicate a focal document.
+These input embeddings capture all information that is contained in a document, because we can re-generate the focal document just based on this summary embedding.
+These document summaries are deterministic, because we obtain them through maximum likelihood estimation and live in continuous space.
+We can evaluate the fit of such a document summary, as the likelihood indicates its fit directly.
+
+<!-- TODO:  @khattabDSPyCompilingDeclarative2023 -->
+
+| Method                              | Origin                                          | Reference                              |  Deterministic? |  Generation?  |  Type?  |
+|:------------------------------------|:------------------------------------|:--------------------|:----------------------:|:----------------------:|:----------------------:|
+| BERT `[CLS]` token                       | Next-sentence prediction task                  | @devlinBERTPretrainingDeep2018         | $\checkmark$   | $\times$       | Numeric    |
+| Pooled Word Embeddings              | Aggregation of token information               | @shenBaselineNeedsMore2018             | $\checkmark$   | $\times$       | Numeric    |
+| Prompt engineering                  | Emergent capability of LLM                     | CITE                                   | $\times$       | $\times$       | Textual    |
+| Reverse engineered document summaries | Maximize the likelihood to re-generate the focal document | *Proposed method*                        | $\checkmark$   | $\checkmark$   | Numeric |
+: Overview of existing methods to summarize documents. {#tbl-docsum}
 
@@
 
-- Distributional hypothesis of language @sahlgren2008distributional
+<!-- TODO: Intuitive method explainer -->
 
-We reverse-engineer the input to a Large Language Model (LLM), such that it generates a pre-defined text.
+<!-- TODO: Talk about low-dim representation and regularization -->
 
-For a given target sequence, we find the input-embedding that maximizes the likelihood to generate this target sequence with a unidirectional LLM (e.g.
-GPT, @radford2018improving).
-We call this input embedding, the "summary embedding" of the target sequence.
+<!-- TODO: Normalization layer and its role; Does it work without it? Why / not? -->
 
-We restrict the degrees of freedom in our search for these input embeddings, by making the elements of the input embedding linear combinations of a few hidden factors.
-This amounts to a linear factor model.
-
-![Factor model in Neural Network form](factor_tikz.tex)
-
-In turn, these hidden factors yield a generation space, where we locate our target sequences and search for new sequences.
-
-@goodfellowDeepLearning2016
-
-- Autoencoder
-  - The Autoencoder $\operatorname{AE}_{\mathcal{B}} \left(\cdot \right): \lbrace 0, 1 \rbrace^{C} \to \mathbb{R}^E$, with parameters $\mathcal{B}$
-  - Adding covariates to the AE
-
-- Overview
-  - LLM
-  - maximize likelihood of target sequence
-  - autoencoder on these summary embeddings (+ covariates?)
-  - also yields low-dim generation space
-  - generation
-  - rating of claims
-
+Generative Large Language Models (LLM), such as [ChatGPT](https://chat.openai.com/) [@radford2018improving], generate text based on textual inputs, so called prompts.
+Internally, the model translates the prompt first into tokens, integer codes representing certain words, symbols and syllabuls, and then into a lower dimensional vector representation of the text itself, which we call the input embedding.
+This input embedding is then passed through a neural network that yields a probability distribution over the next token in the sequence.
+To generate text, the model makes a draw from this distribution, and attached this generated token at the end of the prompt, repeating this process until it predicts the next token to be the end-of-text token (eos), i.e.
+until it predicts that the text ends.
+LLMs are pre-trained on large text corpora, such as the [Common Crawl](https://commoncrawl.org/) which contains over 250 billion pages of text.
+Their architecture is rooted in the attention mechanism.
+The attention mechanism models interactions in the text, such as negations, synonyms, or grammatical structures, across a long sequence of text.
+@vaswani2017attention implement this mechanism in the transformer block, which is a modular building block for the neural network underlying an LLM.
+The combination of these transformer blocks, giant pre-training data, and huge computing resources, lead to LLMs that have "emergent capabilities".
+These are capabilities, which were not explicitly trained for by the developers of the model, such as the ability of a model to translate between languages, pass the BAR exam, or write textual summaries of documents [see @weiEmergentAbilitiesLarge; @luAreEmergentAbilities2023].
 Unidirectional LLMs, such as GPT or LLaMa [@radford2018improving; @metaMetaLlama], are built for text-generation.
@@
 salutations in letters (long-range dependency).
+<!-- TODO: Simplify this paragraph -->
 
+- For generation of sequences, the LLM autoregressively predicts the next token, given the previous tokens.
+At each step in the generation, a selection of the next token needs to take place, for which there exist various procedures (CITE).
+As we are optimizing with respect to the joint likelihood of the target sequence, we also need to generate with a procedure that generates this most likely sequence.
+Finding this sequence from the generation tree is a combinatorical problem and computationally not feasible.
+Hence, we use the beam-search heuristic to make our generation.
+Rather than only tracking the most likely token at each generation step, beam search also tracks the $n$-most likely tokens.
+The more beams we search, the more likely we are to find a sequence with a higher likelihood, then we would have found otherwise [@huggingfaceGenerateText].
+
+Each document, in our case an advertising claim, $d$ consists of the tokens $t_1^{d}, \ldots, t_T^(d)$.
+
+The probability to generate the sequence $t_1, \ldots, t_T$ with the LLM when using the embedding $\boldsymbol{s}$ as an input is $p(t_1, \ldots, t_T \mid \boldsymbol{s})$.
+We can split this probability into conditional probabilities due to the autoregressive architecture of the LLM and because we observe the target sequence.
+This provides large computational gains, as we can calculate these parts in parallel.
+Hence, we define the log-likelihood of the summary embedding for the target sequence as
+$$
+\operatorname{\mathcal{L}} \left ( \boldsymbol{s} \mid t_1, \ldots, t_T \right) = \log p(t_1 | \boldsymbol{s}) + \log p(t_2 | t_1, \boldsymbol{s}) + \ldots + \log p(t_T | t_1, \ldots, t_{T-1}, \boldsymbol{s}) \text{, }
+$$
+
+and find the optimal summary embedding as
+
+$$
+\mathbf{s}^{*} = \argmax_{\boldsymbol{s}} \operatorname{\mathcal{L}} \left ( \boldsymbol{s} \mid t_1, \ldots, t_T \right) \text{. }
+$$
+
+We summarize the training process for the single summary embeddings in Algorithm 1.
+The summary embedding is a vector of length $E$, which is the same dimension as the dimension of the LLM's input embedding.
+
+\begin{algorithm}
+\begin{algorithmic}
+\label{alg:singletraining}
+\caption{Training of Single Summary Embeddings}
+\State $i \gets 0$
+\State $\epsilon \gets 0.01$
+\State $\boldsymbol{s} \gets \operatorname{Initialization} \left( \cdot \right)$
+\While{$True$}
+    \State $l^{(i)} \gets \operatorname{\mathcal{L}} \left ( \boldsymbol{s} \mid t_1, \ldots, t_T \right)$
+    \State $\nabla_{\boldsymbol{s}}^{(i)} l \gets \operatorname{ComputeGradient} \left ( l^{(i)}  \right )$
+    \State $\boldsymbol{s}^{(i + 1)} l \gets \operatorname{Optimizer} \left(\boldsymbol{s}^{(i)}, \; \nabla_{\boldsymbol{s}}^{(i)} l \right)$
+    \State $i \gets i + 1$    
+    \If{$l^{(i)} < \epsilon$}
+      \State \text{break}
+    \EndIf
+\EndWhile
+\end{algorithmic}
+\end{algorithm}
+
+As we are interested in finding a low-dimensional sub-space, in which we can represent the $D$ advertising claims, we use a factor model and estimate the summary embeddings for all advertising claims jointly, yielding the $D \times E$ matrix of summary embeddings $\mathbf{S}$.
+We restrict the degrees of freedom to $F$ hidden factors.
+To implement this factor model, we create an Autoencoder with one hidden layer, that is fully connected with linear activation functions [@goodfellowDeepLearning2016].
+The input data to this autoencoder is an identity matrix of dimension $D$.
+We illustrate this neural network representation of the factor model in @fig-factor.
+
+![Factor model in Neural Network form, example for document 2](factor_tikz.tex){#fig-factor}
+
+We denote the matrix of weights of the encoder and decoder layers by $\mathbf{W}^{Encoder}$ and $\mathbf{W}^{Decoder}$.
+These matrices have the dimensions $D \times F$ and $F \times E$ respectively.
+We also define the vectors of biases for the encoder and decoder by $\boldsymbol{b}^{Encoder}$ (length $F$) and $\boldsymbol{b}^{Decoder}$ (length $E$), and stack them into matrices by taking the Kronecker product with a length $D$ vector of ones, $\boldsymbol{1}_D$, to obtain the biases in matrix form as $\mathbf{B}^{Encoder} = \boldsymbol{1}_D \otimes \boldsymbol{b}^{Encoder}$ and $\mathbf{B}^{Decoder} = \boldsymbol{1}_F \otimes \boldsymbol{b}^{Decoder}$.
+For notational convenience, we denote this factor model by $\operatorname{Factor}_{\Omega_{Factor}} \left(\mathbb{I}_{D} \right): \lbrace 0, 1 \rbrace^{D \times D} \to \mathbb{R}^E$, with parameters $\Omega_{Factor}$, where $\Omega_{Factor} = \lbrace \mathbf{W}^{Encoder}, \mathbf{W}^{Decoder}, \mathbf{B}^{Encoder}, \mathbf{B}^{Decoder} \rbrace$.
+
+<!-- The Large Language Model provides a log-likelihood function $\operatorname{\mathcal{L}} \left ( \mathbf{S} \right): \mathbb{R}^{D \times E} \to \mathbb{R}^{\left (- \infty, 0 \right]}$, with parameters $\Omega_{LLM}$, which returns the log-likelihood for the summary embedding to generate the target documents.
+-->
+
+When we define the weights and biases in this form, we can write the factor representation of the input as
+
+$$
+\mathbf{F} = \mathbf{W}^{Encoder} + \mathbf{B}^{Encoder} \text{, }
+$$
+
+thereby the matrix of summary embeddings for all claims becomes
+
+$$
+\mathbf{S} = \left( \mathbf{W}^{Encoder} + \mathbf{B}^{Encoder} \right) \mathbf{W}^{Decoder} + \mathbf{B}^{Decoder} \text{. }
+$$
+
+When we estimate the matrix of summary embeddings $\mathbf{S}$, we use the joint log-likelihood to re-generate all documents in the dataset, $\operatorname{\mathcal{L}}_D = \sum_{d=1}^D \operatorname{\mathcal{L}}_d$, where $\operatorname{\mathcal{L}}_d$ is the log-likelihood to re-generate document $d$ with summary embedding $\mathbf{S}$.
+We estimate the summary embeddings by maximizing this joint log-likelihood with respect to the parameters of the factor model:
+
+<!-- TODO: Can I split this into separate optimizations? Why / why not? -->
+$$
+ \mathbf{S}^{*} = \argmax_{\mathbf{W}^{Encoder}, \; \mathbf{W}^{Decoder}, \; \mathbf{B}^{Encoder}, \; \mathbf{B}^{Decoder}} \operatorname{\mathcal{L}}_D \left ( \mathbf{S} \right) \text{. }
+$$
+
+<!-- 
+For one observation vector, $\boldsymbol{d}$, of length $D$, the resulting length $F$ vector in the factor layer, $\boldsymbol{f}$,  is $\boldsymbol{f} = \boldsymbol{i}^T \mathbb{W}^E + \boldsymbol{B}^E$.
+In turn, the corresponding vector in the output space, $\boldsymbol{e}$, is $\boldsymbol{e} = \boldsymbol{h} \mathbb{W}^D + \boldsymbol{B}^D$.
+Taking these two steps together we obtain the output vector as $\boldsymbol{e} = \left( \boldsymbol{i}^T \mathbb{W}^E + \boldsymbol{B}^E \right) \mathbb{W}^D + \boldsymbol{B}^D$.
+This shows how the elements of $\boldsymbol{e}$ are now linearly dependent and a function of the factors contained in $\boldsymbol{h}$.
+Taking into account that our input matrix is an identity matrix, one observation becomes a standard basis vector, i.e.
+a vector that only contains zeroes, except at one position, where it contains a one.
+Essentially, the standard basis vector with a one at position $j$, $\boldsymbol{i}_j^T$, selects the $j$-th column of a matrix when multiplying the two.
+Hence, we can represent observation $j$ in the hidden layer as $\boldsymbol{h}^{(j)} = \boldsymbol{W^E_{\cdot, \; j}} + \boldsymbol{B}^E$.
+ -->
+
+<!-- 
 - Notation
@@
 
-$$
-\boldsymbol{e^{*}} = \argmax_{\boldsymbol{e}} \operatorname{\mathcal{L}}_{\mathcal{M}} \left ( \boldsymbol{e}, \boldsymbol{t} \right) = \argmax_{\boldsymbol{e}} \sum_{i=1}^{L} \log p(t_1, \ldots, t_L | \boldsymbol{e})
-$$
+-->
 
+<!-- 
 We aim to find a summary embedding, $\boldsymbol{e^{*}}$, that maximizes the likelihood of generating a given target sequence, $\lbrace t_i \rbrace_{i=1}^{L}$, with a given LLM, $\mathcal{M}$.
@@
 In Algorithm 1, we summarize the training process.
+ -->
 
-<!-- TODO: Update this algorithm -->
+We summarize the training process for finding the sub-space representation in Algorithm 2.
+
+<!-- TODO: Update this -->
 
@@
 \label{alg:training}
-\caption{Training of Summary Embeddings}
-\State Initialize $\boldsymbol{s}$
+\caption{Training of Summary Embeddings based on factor model}
 \State $i \gets 0$
-\While{$Convergence == False$}
+\State $\epsilon \gets 0.01$
+\State $\mathbf{W}^{Encoder}_{(i)}, \; \mathbf{W}^{Decoder}_{(i)}, \; \mathbf{B}^{Encoder}_{(i)}, \; \mathbf{B}^{Decoder}_{(i)} \gets \operatorname{Initialization} \left( \cdot \right)$
+\While{$True$}
+    \State $\mathbf{S} \gets \left( \mathbf{W}^{Encoder}_{(i)} + \mathbf{B}^{Encoder}_{(i)} \right) \mathbf{W}^{Decoder}_{(i)} + \mathbf{B}^{Decoder}_{(i)}$
+    \State $l_{(i)} \gets \operatorname{\mathcal{L}}_D \left ( \mathbf{S} \right)$
+    \State $\nabla_{(i)} \gets \operatorname{ComputeGradient} \left ( l^{(i)}  \right )$
+    \State $\mathbf{W}^{Encoder}_{(i + 1)}, \; \mathbf{W}^{Decoder}_{(i + 1)}, \; \mathbf{B}^{Encoder}_{(i + 1)}, \; \mathbf{B}^{Decoder}_{(i + 1)} \gets \operatorname{Optimizer} \left(\mathbf{W}^{Encoder}_{(i)}, \; \mathbf{W}^{Decoder}_{(i)}, \; \mathbf{B}^{Encoder}_{(i)}, \; \mathbf{B}^{Decoder}_{(i)}, \; \nabla_{(i)} \right)$
     \State $i \gets i + 1$    
-    \State $\boldsymbol{o}^{(i)} \gets \mathcal{M}(\boldsymbol{s})$
-    \State $l^{(i)} \gets \operatorname{Likelihood}(\boldsymbol{o}^{(i)}, \lbrace t_i \rbrace_{1}^{L})$
-    \State $\boldsymbol{s}^{(i + 1)} \gets \operatorname{GradientDecent}(\boldsymbol{s}^{(i)}, l^{(i)})$
-    \State $Convergence \gets \operatorname{CheckConvergence}(\boldsymbol{s}^{(i + 1)}, \boldsymbol{s}^{(i)})$
+    \If{$l_{(i)} < \epsilon$}
+      \State \text{break}
+    \EndIf
 \EndWhile
@@
 <!-- TODO: Good idea or not? -->
-
-as the input-embedding that maximizes the likelihood of generating a given target sequence, $\lbrace t_i \rbrace_{i=1}^{L}$, where $L$ denotes the length of the target sequence and $t_i$ represents token $i$.
-We estimate these summary embeddings $\boldsymbol{e^{*}}$, by maximizing the conditional likelihood that this embedding generates the target sequence for a given LLM.
-We use a gradient-based optimization algorithm to maximize this likelihood.
-
-- Figure of generation tree
-  - For generation of sequences, the LLM autoregressively predicts the next token, given the previous tokens.
-At each step in the generation, a selection of the next token needs to take place, for which there exist various procedures (CITE).
-As we are optimizing with respect to the joint likelihood of the target sequence, we also need to generate with a procedure that generates this most likely sequence.
-Finding this sequence from the generation tree is a combinatorical problem and computationally not feasible.
-Hence, we use the beam-search heuristic to make our generation.
-Rather than only tracking the most likely token at each generation step, beam search also tracks the $n$-most likely tokens.
-The more beams we search, the more likely we are to find a sequence with a higher likelihood, then we would have found otherwise [@huggingfaceGenerateText].
-
-- LLM judge to rate claims
+<!-- TODO: Figure of generation tree -->
 
 # Data
+
+CITE: Prompt engineering
+
+:::{#fig-gpt}
+\noindent
+\begin{minipage}{\textwidth}
+\begin{tcolorbox}[colback=white, colframe=gray!75!black]
+\textbf{Prompt:} \textit{I work for a marketing agency.
+I want to market a haircare product and need some claims for this.
+\textbf{I want you to emphasize how shiny the haircare product consumers' hair makes}.
+I need claims that differ in style with respect to how tangible the claims are.
+Make these claims brief.
+Can you give me 10 claims that are rather tangible and 10 claims that are rather intangible?}
+\end{tcolorbox}
+\end{minipage}
+
+[Prompt to generate the benchmark advertising claims](https://chatgpt.com/share/6af9a0b9-f8ee-4c31-a1b8-00712384ba59)
+::::
+
+As as robustness check, we repeated this task for different attributes (instead of shiny, healthiness and colorfulness of hair), and for a different product category (surface cleaners).
+In total, we obtain 80 advertising claims, 20 for each of these settings.
+Half of these claims are supposed to be tangible and half intangible.
 
@@
 These data, stem from different variations of choice-based conjoint studies [@eggersChoiceBasedConjointAnalysis2022], which we cannot further disclose due to confidentiality agreements.
+The data includes choice-experiments aggregated at the level of the advertising claims, hence we do not have responses of individuals.
+These measurements of consumer-preferences are on multiple dimensions, such as relevance, uniqueness of a claim, brand fit, or an overarching rating of the claim.
+The advertising claims were designed for markets in different countries and are in english (US, UK) or translated into english (others).
+The date of the study is also included in the data.
+We also have information on the marketers motivation behind the design of an advertising claims, such as whether she designed the claims to pronounce the health benefits of a yoghurt, or the taste yoghurt, as well as the brand of the product.
 
-The data includes choice-experiments aggregated at the level of the advertising claims, hence we do not have responses of individuals.
-These measurements of consumer-preferences are on multiple dimensions, such as relevance, brand fit, or an overarching rating of the claim.
-
-The advertising claims were designed for markets in different countries and are in english (US, UK) or translated into english (others). The date of the study is also included in the data.
-
-We also have information on the marketers motivation behind the design of an advertising claims, such as whether she designed the claims to pronounce the health benefits of a yoghurt, or the taste yoghurt.
-
-The data also include the brand of a product.
+<!-- TODO: Some graphs / tables for descriptives -->
 
@@
 
+TODO: Have similarity measure vs euclidean distance in encoding space
+
+- Performed suite of tests of generated data
+  - haircare shiny
+  - haircare healthy
+  - haircare colorful
+
+[ ] Training of embeddings
+
+![Training of embeddings](figures/dashboard.png)
+
+- computation
+- dealing with spikes
+
+[ ] Analysis of embeddings; Plot correlation of claims and dims, PCA
+
+::: {#fig-correlation layout-ncol=2 layout-nrow=2}
+![RE](figures/correlation_re.pdf)
+
+![BoW](figures/correlation_bow.pdf)
+
+![W2V](figures/correlation_w2v.pdf)
+
+![BERT](figures/correlation_bert.pdf)
+
+Correlation matrices along the claims. 
+:::
+
+::: {#fig-correlation-surface layout-ncol=2 layout-nrow=2}
+![RE](figures/correlation_re_surface.pdf)
+
+![BoW](figures/correlation_bow_surface.pdf)
+
+![W2V](figures/correlation_w2v_surface.pdf)
+
+![BERT](figures/correlation_bert_surface.pdf)
+
+Correlation matrices along the claims. 
+:::
+
+::: {#fig-correlation-haircolor layout-ncol=2 layout-nrow=2}
+![RE](figures/correlation_re_color.pdf)
+
+![BoW](figures/correlation_bow_color.pdf)
+
+![W2V](figures/correlation_w2v_color.pdf)
+
+![BERT](figures/correlation_bert_color.pdf)
+
+Correlation matrices along the claims. 
+:::
+
+::: {#fig-correlation-health layout-ncol=2 layout-nrow=2}
+![RE](figures/correlation_re_health.pdf)
+
+![BoW](figures/correlation_bow_health.pdf)
+
+![W2V](figures/correlation_w2v_health.pdf)
+
+![BERT](figures/correlation_bert_health.pdf)
+
+Correlation matrices along the claims. 
+:::
+
+::: {#fig-correlationts layout-ncol=2 layout-nrow=2}
+![RE](figures/correlation_ts_re.pdf)
+
+![BoW](figures/correlation_ts_bow.pdf)
+
+![W2V](figures/correlation_ts_w2v.pdf)
+
+![BERT](figures/correlation_ts_bert.pdf)
+
+Correlation matrices along the embedding dimensions. 
+:::
+
+::: {#fig-pca layout-ncol=2 layout-nrow=2}
+![RE](figures/pca_re.pdf)
+
+![BoW](figures/pca_bow.pdf)
+
+![W2V](figures/pca_w2v.pdf)
+
+![BERT](figures/pca_bert.pdf)
+
+First two principal components of the embedding matrix, colored by class.
+:::
+
+::: {#fig-loo layout-ncol=2 layout-nrow=2}
+![RE](figures/loo_re.pdf)
+
+![BoW](figures/loo_bow.pdf)
+
+![W2V](figures/loo_w2v.pdf)
+
+![BERT](figures/loo_bert.pdf)
+
+Reuslts of leave-one-out regression, colored by class.
+:::
+
+  - GPT-2
+  - BERT
+  - Word2Vec
+  - RE
+
+[ ] Encoding dimension plot
+
+@fig-encoding, @tbl-claims
+
+::: {#fig-encoding layout-ncol=2}
+
+![All points](figures/encoding_space.pdf){#fig-encoding1}
+
+![Zoom-in](figures/encoding_space_zoomin.pdf){#fig-encoding2}
+
+Encoding space of the advertising claims.
+
+:::
+
+| Number | Claim |
+|:-:|:-----------:|
+0  (T) | Experience 50% more visible shine after just one use.
+1 (T) | Formulated with light-reflecting technology for a glossy finish.
+2 (T) | Transform dull strands into radiant, luminous locks.
+3 (T) | Infused with nourishing oils that enhance natural shine.
+4 (T) | See instant brilliance with our advanced shine-boosting formula.
+5 (T) | Locks in moisture to amplify hair's natural luster.
+6 (T) | Achieve salon-quality shine without leaving home.
+7 (T) | Visible reduction in dullness, replaced with stunning shine.
+8 (T) | Say goodbye to lackluster hair, hello to mirror-like shine.
+9 (T) | Clinically proven to enhance shine by up to 70%.
+10 (I) | Elevate your confidence with hair that gleams under any light.
+11 (I) | Embrace the allure of luminous hair that turns heads.
+12 (I) | Unleash the power of radiant hair that speaks volumes.
+13 (I) | Transform your look with hair that exudes brilliance.
+14 (I) | Feel the difference of hair that shines with vitality and health.
+15 (I) | Rediscover the joy of hair that beams with inner vibrancy.
+16 (I) | Indulge in the luxury of hair that shimmers with elegance.
+17 (I) | Step into the spotlight with hair that radiates beauty.
+18 (I) | Experience the magic of hair that dazzles with every movement.
+19 (I) | Unlock the secret to hair that shines from within, reflecting your inner glow.
+
+: Tangible (T) and Intangible (I) advertising claims for hair shampoo products. {#tbl-claims}
+
+- labels, accuracy
+- clustering
+- generation of new claims
+  - exploring the space
+  - measure of fit (area)
+
+[ ] generation of new claims [@Goodfellow2016]
+
+![Exploration of Encoding Space](figures/exploration_encoding_space.pdf)
+
+![Exploration of the likelihoods](figures/likelihoods_generation.pdf)
+
+<!-- TODO: Update these tables! -->
+
+|    | 3rd   | 2nd   | Choice        |   Prob. 3rd |   Prob. 2nd |   Prob. Choice |
+|---:|:------|:------|:--------------|------------:|------------:|---------------:|
+|  0 | Feel  | Un    | Step          | 0.0287145   | 0.0344409   |       0.887634 |
+|  1 | onto  | back  | into          | 0.000927151 | 0.00125316  |       0.992544 |
+|  2 | to    | class | the           | 5.98035e-05 | 6.91189e-05 |       0.99934  |
+|  3 | '     | sun   | spotlight     | 0.000623843 | 0.00123234  |       0.992407 |
+|  4 | into  | in    | with          | 0.00192485  | 0.00199383  |       0.992454 |
+|  5 | a     | hairs | hair          | 0.000946804 | 0.000969693 |       0.995565 |
+|  6 | above | .     | that          | 0.000743194 | 0.0014022   |       0.993072 |
+|  7 | X     | can   | radi          | 0.000538268 | 0.00169703  |       0.992411 |
+|  8 | ating | ats   | ates          | 0.000915091 | 0.00206985  |       0.996706 |
+|  9 | heat  | into  | beauty        | 0.00111988  | 0.00228065  |       0.987492 |
+| 10 | ).    | .)    | .             | 0.000529124 | 0.00114012  |       0.996219 |
+| 11 | \textit{2x linebreak}      | ]     | \textit{eos-token} | 0.000603324 | 0.00100365  |       0.995909 |
+| 12 | \textit{2x linebreak}    | \textit{linebreak}      | \textit{eos-token} | 7.61097e-06 | 0.00124311  |       0.99872  |
+| 13 | \textit{2x linebreak}    | \textit{linebreak}      | \textit{eos-token} | 2.16886e-05 | 0.000565556 |       0.999231 |
+| 14 | \textit{2x linebreak}    | \textit{linebreak}      | \textit{eos-token} | 3.80865e-05 | 0.00084979  |       0.998713 |
+| 15 | \textit{2x linebreak}    | \textit{linebreak}      | \textit{eos-token} | 5.16162e-05 | 0.00161664  |       0.997724 |
+| 16 | \textit{2x linebreak}    | \textit{linebreak}      | \textit{eos-token} | 0.000101122 | 0.0031923   |       0.995816 |
+: Example for a low 1st probability {#tbl-lowprob}
+
+|    | 3rd      | 2nd        | Choice        |   Prob. 3rd |   Prob. 2nd |   Prob. Choice |
+|---:|:---------|:-----------|:--------------|------------:|------------:|---------------:|
+|  0 | See      | Form       | Experience    | 0.00921188  | 0.013993    |       0.943786 |
+|  1 | 20       | 25         | 50            | 0.00203264  | 0.00445767  |       0.976235 |
+|  2 | %.       | .          | %             | 0.00142157  | 0.00205115  |       0.994173 |
+|  3 | better   | More       | more          | 0.000946993 | 0.00160339  |       0.996037 |
+|  4 | shine    | protective | visible       | 0.00178126  | 0.00218073  |       0.979003 |
+|  5 | glow     | light      | shine         | 0.000826363 | 0.00209097  |       0.983356 |
+|  6 | when     | by         | after         | 0.00171736  | 0.00367565  |       0.987    |
+|  7 | a        | well       | just          | 0.000570301 | 0.00172336  |       0.98959  |
+|  8 | two      | a          | one           | 0.00153683  | 0.00186334  |       0.989587 |
+|  9 | shine    | usage      | use           | 0.000427258 | 0.00324287  |       0.990624 |
+| 10 | ,        | !          | .             | 0.000354906 | 0.000630647 |       0.9981   |
+| 11 | .        |            | \textit{eos-token} | 0.000454232 | 0.00156762  |       0.994366 |
+| 12 | Transfer |            | \textit{eos-token} | 0.000179846 | 0.00032993  |       0.998234 |
+| 13 |          | More       | \textit{eos-token} | 0.000109437 | 0.000194042 |       0.998748 |
+| 14 | More     |            | \textit{eos-token} | 7.64376e-05 | 0.000174342 |       0.999236 |
+| 15 | Lab      |            | \textit{eos-token} | 5.29115e-05 | 0.000444251 |       0.998936 |
+| 16 | Lab      |            | \textit{eos-token} | 0.000276325 | 0.00155492  |       0.997141 |
+: Example for similar tokens {#tbl-similartoken}
+
+  - exploration of the likelihoods
+  - frequentist test of generating existing
+
+[ ] Repetition with actual marketing data
+
 ### Implementation
+
+- GPT-2; Far from state of the art, if it works here, it will work with other model
 
@@
 
+- Summary statistics of the data
+
 - Application to MR data
+
+<!-- TODO: Check whether this is okay to name -->
+
+- Two small sub-samples from the same market research study of product claims for a dairy product
+  - MR1: Major theme "Gut Health", country DE, language english
+  - MR2: Major theme "Scientific words", country ES, language english
+  - MR3: Major theme: "RTB vs. Benefit", minor theme: "Amino-Acid", "Energy", ..., country NL, fitness brand, language english
+  - MR4: Hair care, country BR, total sample, find difference between two brands: Both from same FMCG  corporation
+
+## MR1
+
+::: {#fig-correlation-mr1 layout-ncol=2}
+![RE](figures/correlation_re_mr1.pdf)
+
+![BERT](figures/correlation_bert_mr1.pdf)
+
+Correlation matrices along the claims. 
+:::
+
+![Encoding space of the advertising claims.](figures/encoding_space_mr1.pdf){#fig-encoding1-mr1}
+
+::: {#fig-loo_mr1 layout-ncol=2}
+
+![RE](figures/loo_re_mr1.pdf)
+
+![BERT](figures/loo_bert_mr1.pdf)
+
+Reuslts of leave-one-out regression, colored by class.
+:::
+
+## MR2
+
+::: {#fig-correlation-mr2 layout-ncol=2}
+![RE](figures/correlation_re_mr2.pdf)
+
+![BERT](figures/correlation_bert_mr2.pdf)
+
+Correlation matrices along the claims. 
+:::
+
+::: {#fig-encoding-mr2 layout-ncol=2}
+
+![All points](figures/encoding_space_mr2.pdf){#fig-encoding1-mr2}
+
+![Zoom-in](figures/encoding_space_zoomin_mr2.pdf){#fig-encoding2-mr2}
+
+Encoding space of the advertising claims.
+
+:::
+
+::: {#fig-loo_mr2 layout-ncol=2}
+
+![RE](figures/loo_re_mr2.pdf)
+
+![BERT](figures/loo_bert_mr2.pdf)
+
+Reuslts of leave-one-out regression, colored by class.
+:::
+
+## MR4
+
+::: {#fig-correlation-mr4 layout-ncol=2}
+![RE](figures/correlation_re_mr4.pdf)
+
+![BERT](figures/correlation_bert_mr4.pdf)
+
+Correlation matrices along the claims. 
+:::
+
+::: {#fig-encoding-mr4 layout-ncol=2}
+
+![All points](figures/encoding_space_mr4.pdf){#fig-encoding1-mr4}
+
+![Zoom-in](figures/encoding_space_zoomin_mr4.pdf){#fig-encoding2-mr4}
+
+Encoding space of the advertising claims.
+
+:::
+
+::: {#fig-loo_mr4 layout-ncol=2}
+
+![RE](figures/loo_re_mr4.pdf)
+
+![BERT](figures/loo_bert_mr4.pdf)
+
+Reuslts of leave-one-out regression, colored by class.
+:::
+
+## MR 5
+
+::: {#fig-anecdote-mr5 layout-ncol=2}
+
+![Occurrence of words taste, natural, and immune system.](figures/anecdote_immune_system_mr5.pdf){#fig-anecdote-mr5a}
+
+![Occurrence of words relating to morning theme](figures/anecdote_morning_system_mr5.pdf){#fig-anecdote-mr5b}
+
+Embedding of the yoghurt claims, colored by word features.
+:::
+
+- establish that the space is continuous, word features etc.
+
+![Uniqueness score of embedded yoghurt drink claims.](figures/uniqueness_mr5.pdf){#fig-ratings-mr5a}
+
+@fig-anecdote-mr5 shows the encoding space for the yoghurt drink advertising claims, and illustrates how the embedding picks up on language features of the documents.
+Namely, @fig-anecdote-mr5a colors the claims based on whether they contain the words "taste" or "sense", "natural", and/or "immune system".
+We color-coded these three by the basic colors (yellow, red, blue), and claims that contain multiple of these words by the mixtures of these colors.
+If a claim contains none of these words, we code it in grey.
+Claims with the same word features cluster together, while claims that contain none of these words, are pushed to the outside of the plot.
+Intuitively, we would expect that claims which contain multiple of these word features from a transition between the claims which only have one of the word features.
+For this, we only have claims 13 and 21 as examples, whereas the former does not form such a boundary, and the latter is at the edge of the "Natural" and "Immune System" class.
+
+<!-- TODO: Give example of transition between two points -->
+
+We expect advertising claims that are more unique to have a more isolated position in the encoding space, as they are less similar to other claims.
+In other words, the uniqueness score should be positively correlated with the euclidean distance of an encoded claim to its neighest neighbor.
+@fig-ratings-mr5a shows that this is not the case, the correlation of these two measures is substanial and negative ($-0.4375$).
+The opposite is true: The three most unique claims (4, 17, 18) cluster together closely, and the further claims are away from these three, the less unique they tend to be.
+Claims with low uniqueness scores are spread further apart and far away from the most unique claims.
+These three most unique claims have some things in common, all of these use a combination of the words "natural", "active", and talk about ingredients (with synonyms).
+On the other hand, claim 15 is the only claim in the dataset using the word "yoghurt drink" and is also the only claim about emotions.
+Thereby, it is unique from a language perspective (far away from other points in the encoding), but perhaps not unique to consumers, as there might be similar claims on the market already.
+Also, uniqueness is highly correlated ($0.8843$) with the overall rating of claims, some of this correlation could be due to consumers viewing uniqueness as a proxy for "satisfaction".
+Lastly, the researched advertising claims are self-selected, and perhaps more unique then the average claim that is on the market already.
+This might lead to a form of Simpson's paradox [@sprenger2021simpson], where in the population there is a positive correlation between uniqueness and unique language, but perhaps not in this special sub-sample of the data.
+
+![Overall rating of embedded yoghurt drink claims.](figures/rating_country_mr5.pdf){#fig-ratings-mr5b}
+
+@fig-ratings-mr5b again shows the encoding space, but this time colored by the overall rating of the claims.
+Claims of similar rating tend to cluster together, with higher ratings in the north-east and lower ratings towards the south-west of the graph.
+The clustering of similarly rated claims, could be due to these claims being similar in language.
+For marketers, there are two insights from this: One, getting the overall theme of a claim right, can ensure that customers perception of this claim is within a certain ballpark.
+Two, after identifying a fruitful theme for the advertising claim, it is still useful to explore this neighborhood in more fine-grained steps, as locally, there might be small alterations that have large effects on the perception: See e.g.
+claims 21 (rating in top 4% percentile ) and 14 (rating in top 12%), despite one of the shortest distances in the encoding space.
+@fig-ratings-mr5b also shows how ratings of uniqueness can differ by country, the reasons for this could lie in what consumers are used to from the domestic market and in cultural differences.
 
@@
 
-- Wording
+- Wording: Ballpark, finetuning idea
 
@@
 
+- Generator, predictor compatible embeddings; Generative AI to assist and automate design processes in marketing
+  - Existing approaches
+    - @hongWritingMoreCompelling2022 using a hierarchical attention network (HAN)
+  - @burnap2023product
+    - architecture of encoder, embedding space, predictor, generator
+    - images, VAE & GAN; Use semi-supervised learning of unlabelled and labeled images; These attributes can shape the generation
+    - augment the design process, rather than automate it
+    - Training requires large resources (2 weeks on multiple GPUs); However, use works on regular laptop
+    - Don't have pre-trained model available that can generate
+    - Performs better than research clinic, at a fraction of the cost, managers want to adapt this model in the organization
+    - Differences
+      - piggy-back of pre-trained LLM, can just learn the specifics of the domain rather than the language itself
+      - Deterministic generation
+      - When using bigger model, benefit of emergent capabilities; @brandUsingGPTMarket2023a; @goliFrontiersCanLarge2024 indicate that LLMs have ermergent capabilities wrt to consumer preferences (NOTE: Is this a fair point to make?
+We are getting rid of stochastic decoder in a sense)
+  - The approaches by @ludwigMachineLearningTool2023 and @burnap2023product model an encoding space for images, that can be used for generating new images as well as to create features that inform a supervised machine learning task
+  - These frameworks consist of three componenets: The encoder, which projects datapoints in an encoding space, the generator, which turns points from the encoding space into new datapoints, and the predictor, which predicts a certain label based on a point from the encoding space [compare terminology @burnap2023product].
+  - In our context, these three components perform the following tasks: The encoder represents an advertising claim as a vector, the generator generates and advertising claim based on such a vector, and the predictor predicts the marketeers motivation based on vector from the encoding space.
+  - Rather than inform product design, @ludwigMachineLearningTool2023, use such a framework to generate hypotheses for why a prisoner might be granted bail or not.
+For this, they predict judges decisions' from bail hearings, based on the mugshot photo of the defendant (predictor) and learn about the distribution of mugshots with a Generative Adversarial Network [CITE Goodfellow GAN].
+To generate new hypothesis, they perform gradient descent on the prediction of the predictor with respect to the encoding data point.
+They search for the alteration to the image, that changes the predicted outcome the most.
+By making equal steps along this gradient in opposing directions, they obtain to versions of the same image manipulated with respect to this feature.
+@ludwigMachineLearningTool2023 also find, that comparing these two images often yields people to name the specific change, creating new hypothesis on why an inmate might be granted bail or not.
+  - We call the input data, $x$, and its distribution $p(x)$. In our setting, $x$ is an advertising claim. <!-- TODO: How does it work for @burnap2023product ? -->
+  - We call labels, $y$ and its distribution $p(y)$.
+In our setting, a label is e.g.
+the motivation of the marketeer behind an advertising claim, such as whether they want the claim to be "tangible" or "intangible".
+The "label" can also be a continuous variable, such as a consumer rating, turning classification problems into regression problems.
+  - We assume, that we can predict $y$ based on the data $x$. The distribution of the label given the data is the likelihood $p(y | x)$.
+  - Generative Adversarial Networks (GANs) consist of two components: The generator and the discriminator.
+The discriminator takes a datapoint as its input and predicts whether this is a genuine or fake datapoint, while the generator creates new datapoints based on noise.
+The learning objective for the generator is to create samples that pass the scrutiny of the discriminator, while the objective for the discriminator is, to correctly classify its test cases.
+There exists an equilibrium for these two objectives, when the generator creates samples that are indistinguishable from real datapoints.
+In essence, the GAN learns the distribution $p(x)$ of the data.
+  - The GAN architecture was designed for continuous data (images), rather than discrete data (text) , and adaptations if this model are necessary [see @de2021survey]
+  - There is a potential to benefit of the pre-trained information in LLMs.
+Rather than GANs, which require more training data as these networks need to learn the properties of an image from the ground up, we can focus directly on the types of texts that are relevant to our domain.
+<!-- NOTE: Is this true?
+-->
+  - The training objective of LLMs is to predict the next token in a sequence of tokens
+  - Need to model $p(x)$ and $p(y | x)$
+  - For text data, we have options to model $p(y | x)$, e.g.
+with a supervised machine learning model.
+There are also potent ways to represent the input data $x$, sucha as by pooling BERT embeddings of a text [@devlinBERTPretrainingDeep2018; @shenBaselineNeedsMore2018].
+  - We can also make draws from $p(x)$, however, this only works through prompt engineering (CITE). And we do not obtain an encoding space, that can be shared with a predictor model.
+  - We solve this problem by finding an encoding space, that allows for informed generation and provides a representation useful for supervised learning tasks.
+  - We find this encoding space through an optimization.
+Namely, for every sequence in our dataset, we find an embedding vector, such that when we use this vector as an input to our LLM, we maximize the likelihood of generating this sequence.
+This we call the summary embedding.
+We perform this optimization for our full dataset simultaneously by using a factor model.
+  - By using this factor model, we create a bottleneck, forcing the model to learn helpful features; In other factor models, these features are often interpretable dimensions [@goodfellowDeepLearning2016].
+
 # Discussion
+
+- Further Applications
+  - Framework of generator and predictor
+  - optimization of prompts in continuous space
+  - automatic answer evaluation
+  - explainable AI for generative models
+
+## Problems
+
+- Issues with Generation
+  - How to evaluate the quality of the generated ad claims @tatsunorihashimotoUnifyingHumanStatistical2019
+  - Another explanaition could lie in the decoding strategie.
+@ariholtzmanCuriousCaseNeural2020 show how text generation based on a maximum likelihood objective leads to text that sounds bland and not human, propsing the use of different decoding strategies to make the generated text more human.
 
@@
 
-- Problems
+  - Need deeper architecture, the manifold on which advertising claims live is not a linear plane?
+
+  - does it make sense to have continuous representation of discrete thing?
+Then there would be other latent variable architectures, such as [Boltzman machines](https://www.cs.toronto.edu/~rgrosse/courses/csc321_2017/slides/lec20.pdf)
+
+- Implementation challenges
   - Computation time
-  - Benefits over existing approaches in practice?
-  - Conceptual issues of using AI in decision making @cathyoneilWeaponsMathDestruction2016; Gender biases @emilyshengWomanWorkedBabysitter2019
-  - How to evaluate the quality of the generated ad claims @tatsunorihashimotoUnifyingHumanStatistical2019
+  - [Tokenizer can cause problems?](https://www.promptingguide.ai/research/llm-tokenization) <!-- TODO: Is this true?
+Give an example?
+Won't be a problem forever ...
+But inherent to this type of model -->
+  - Not as flexible as general LLMs, as we need to have a target sequence and have a use case where we don't want a distribution
 
-- Expansion of this research
+## Expansion of this research
+
+<!-- TODO: Regularization of summary embeddings, have them absolute and not relative -->
+
+  - Incorporating the context of the product itself, embedding e.g. the brand, product features and competitors in the model, to move into part of the generation space that is "open" for the product
+  - Using fine-tuned LLMs
   - MAB around the fitted AE, to incorporate consumer feedback in online learning: Which areas of the space are valuable for which consumers?
@@
 
+# Conclusion
+
 {{< pagebreak >}}

### 2024-06-17 writing/paper.qmd
writing progress

@@
 There are various [blog posts](https://huggingface.co/docs/transformers/en/tasks/prompting#best-practices-of-llm-prompting) and books [e.g.
-@phoenix2024prompt] about how to write good prompts, in fact businesses are [hiring "Prompt Engineers"](https://indeed.com/q-prompt-engineer-jobs.html?vjk=98ac28acfd1328d7) even on dedicated [job boards](https://prompt-engineering-jobs.com/jobs/prompt-engineer-qwvmju/).
+@phoenix2024prompt] about how to write good prompts, in fact businesses are [hiring "Prompt Engineers"](https://indeed.com/q-prompt-engineer-jobs.html?vjk=98ac28acfd1328d7), even on dedicated [job boards](https://prompt-engineering-jobs.com/jobs/prompt-engineer-qwvmju/).
 Despite these, supposedly low-barrier, resources, end users without AI knowledge still struggle with prompt engineering [@zamfirescu2023johnny].
@@
 These issues pose risks for organizations, as the obscurity of manual prompt engineering can lead to loss of know-how, and the unpredictiability of the model can lead to lawsuits, damages to brands, or missed opportunities.
-
 Problems of reproduceability, transparency, and documentation, become critical for organizations with the advent of the [EU AI Act](https://digital-strategy.ec.europa.eu/en/policies/regulatory-framework-ai).
@@
 In this sense, these summary embeddings are optimal.
+To mitigate overfitting and to make these embeddings more interpretable, we also introduce a factor model to estimate these document summaries.
 We show, that these embeddings capture inherent information about the document, can be used for classification, and generation of new documents.
@@
 
-Beyond the methodological contribution, we show a marketing application of these summary embeddings and make a step towards using Generative AI in market research and design of the marketing-mix.
-We find that we can use these embeddings to capture design motivations of marketers behind advertising claims, and establish a measure for linguistic uniqueness of an advertising claim.
+Beyond the methodological contribution, we show a marketing application of these summary embeddings by applying them to market research data on product claims for yoghurt drinks.
+We find that we can use these embeddings to capture design elements of advertising claims, and establish a measure for linguistic uniqueness of an advertising claim.
 Suprisingly, we find that this linguistic uniqueness is negatively correlated with the perceived uniqueness of an advertising claim by consumers.
@@
 
-<!--
-
-- Consumer preferences in latent generation space & Ad-claims
-  - Market research is expensive. How can we leverage these insights better to improve its benefits?
-  - What is missing from preference learning? How is this different?
-    - Media products/things are different from products as attribute bundles
-      - We can directly represent the product itself?
-    - Brands and their role in advertising
-  - We can view products as attribute bundles and get utilities for these attributes 
-  - How do we do this for products that cannot be "discretized"?
-  - Generative AI for personalization
-    - Previously: Set pieces, e.g. recommendation systems
-    - True personalization: Generate solution for each individual consumer
-  - TODO: Add more product claim specific
-  - Advertisements can be effective, @daiFrontiersWhichFirms2023a observe an increase in purchase intentions for restaurants that adapt online advertising by 7%-19%.
-  - they are relevant @kongMadeUSAClaims2021
-  - costs
-  - need flexibility in online settings
-  - closely related problems, such as email targeting @sahniPersonalizationEmailMarketing2018
-  - advertising is one of the marketing-mix instruments which companies use to position themselves in the market and shapes the incentives of the system itself @moorthyTargetingAdvertisingSpending2023 @renAdvertisingContentCreation2023
-  - design question
-  - Ban of cookie-based advertising due to GDPR, adapting to observable behavior in live sessions or to context on the website itself becomes more important. GDPR
-  - Challenges to cookies in website morphing
-  - @beckerDoesItPay2019, point to the nuanced role that authenticity plays in advertising, especially with respect to its alignment with the brand and product. 
-  - @edelmanBrandingDigitalAge2010 Concept of the consumer decision journey, product claims could vary across multiple visits or encounters of the same customer
-
--->
+We proceed with a brief overview of the relevant literature and introduce our methodology.
+After testing our procedure on synthetic data, and exploring the low-dimensional factor space for the generation of new advertising claims, we apply our model to the market research data.
+We conclude with managerial implications, a discussion of the limitations of our approach and expansions for this research.
 
 ## Relevant Literature
+
+<!-- TODO: Merge these paragraphs -->
+
+LLMs predict the next word in a sequence of words.
+These models, such as [ChatGPT](https://chat.openai.com/) [@radford2018improving], generate text based on textual inputs, so called prompts.
+Internally, the model translates the prompt first into tokens, integer codes representing certain (sequences of) characters, and then into a lower dimensional vector representation of the text itself, which we call the input embedding.
+This input embedding is then passed through a neural network that yields a probability distribution over the next token in the sequence.
+To generate text, the model makes a draw from this distribution, and attached this generated token at the end of the prompt, repeating this process until it predicts the next token to be the end-of-text token (eos), i.e.
+until it predicts that the text ends.
+LLMs are pre-trained on large text corpora, such as the [Common Crawl](https://commoncrawl.org/) which contains over 250 billion pages of text.
+Their architecture is rooted in the attention mechanism.
+The attention mechanism models interactions in the text, such as negations, synonyms, or grammatical structures, across a long sequence of text.
+@vaswani2017attention implement this mechanism in the transformer block, which is a modular building block for the neural network underlying an LLM.
+The combination of these transformer blocks, giant pre-training data, and huge computing resources, lead to LLMs that have "emergent capabilities".
+These are capabilities, which were not explicitly trained for by the developers of the model, such as the ability of a model to translate between languages, pass the BAR exam, or write textual summaries of documents [see @weiEmergentAbilitiesLarge; @luAreEmergentAbilities2023].
 
@@
 The first objective is the so called "Masked Language Model" task.
-Here, the researchers replace a random selection of tokens in the text by a mask token, and the goal of the BERT model is to predict which token is underlying this mask.
-In order to learn about the information captures in a sentence, the researchers also pose the "Next Sentence Prediction" task to the model.
+Here, the researchers hide a random token in a sequence, and the goal of the BERT model is to predict which token is missing.
+This teaches the model the meaning of words in context.
+In order to learn about the information captured in a sentence, the researchers also pose the "Next Sentence Prediction" task to the model.
 For this training objective, the BERT model receives a pair of two sentences and needs to predict whether the second sentence follows the first one in a text.
@@
 BERT embeddings are versatile, but can also be fine-tuned to a specific task such as text summarization.
-However, such a fine-tuning requires the user to compile a dataset of text-summary and text pairs, that are representative of their application.
+Such a fine-tuning requires the user to compile a dataset of summary-text pairs, that are representative of their application.
 These BERT representations are only descriptive and cannot be used for the generation of new text [see @devlinBERTPretrainingDeep2018].
-There exist expansions to BERT, such as RoBERTa by @yinhanliuRoBERTaRobustlyOptimized2019, which improve the optimization procedure.
+There exist improvements to BERT, such as RoBERTa by @yinhanliuRoBERTaRobustlyOptimized2019, which improve the optimization procedure.
 
@@
 However, for many data science and automation tasks, we need a reliable numeric representation of the document, which is the focus of this paper.
+We could use these textual summaries for generation, in the sense that we could prompt the model to generate a new document based on the summary.
+A recent contribution, that can make prompt engineering more reliable is the framework proposed by @khattabDSPyCompilingDeclarative2023, called DSPy, which automatically learns how to combine different prompting and finetuning techniques, in order to improve the generated answer of the LLM with respect to a pre-defined metric.
+They show that their discrete optimization procedure leads to better generated output compared to few-shot prompting and expert designed prompts, even when applied to smaller LLMs.
+Their approach is text based and revolves around generating examples of the desired output, and then tuning the prompt and the LLM itself based on these examples.
+In spirit, their approach is similar to hyper-parameter tuning, in that it uses a type of (small) training data and a performance metric (e.g.
+whether the answer is an exact match to a certain label), and then adjusts the prompt and the LLM to maximize the scoring of its answer on this training set.
+The twist is, that the DSPy framework can perform such an hyper-parameter tuning in an automized fashion, by generating candidate prompts itself and selecting from them.
+Our proposed approach differs from DSPy, in that we optimize in a continuous space, and that our goal is to find a numeric summary for a document, rather than to generate a good response to natural language tasks, such as question answering.
 
@@
 | BERT `[CLS]` token                       | Next-sentence prediction task                  | @devlinBERTPretrainingDeep2018         | $\checkmark$   | $\times$       | Numeric    |
-| Pooled Word Embeddings              | Aggregation of token information               | @shenBaselineNeedsMore2018             | $\checkmark$   | $\times$       | Numeric    |
-| Prompt engineering                  | Emergent capability of LLM                     | CITE                                   | $\times$       | $\times$       | Textual    |
+| Pooled Word Embeddings              | Aggregation of token information               | @tomasmikolovDistributedRepresentationsWords2013 @shenBaselineNeedsMore2018             | $\checkmark$   | $\times$       | Numeric    |
+| (Optimized) Prompt engineering                  | Emergent capability of LLM                     | @khattabDSPyCompilingDeclarative2023; @huangenhanced                                   | $\times$       | $\checkmark$       | Textual    |
 | Reverse engineered document summaries | Maximize the likelihood to re-generate the focal document | *Proposed method*                        | $\checkmark$   | $\checkmark$   | Numeric |
@@
 <!-- TODO: Normalization layer and its role; Does it work without it? Why / not? -->
-
-Generative Large Language Models (LLM), such as [ChatGPT](https://chat.openai.com/) [@radford2018improving], generate text based on textual inputs, so called prompts.
-Internally, the model translates the prompt first into tokens, integer codes representing certain words, symbols and syllabuls, and then into a lower dimensional vector representation of the text itself, which we call the input embedding.
-This input embedding is then passed through a neural network that yields a probability distribution over the next token in the sequence.
-To generate text, the model makes a draw from this distribution, and attached this generated token at the end of the prompt, repeating this process until it predicts the next token to be the end-of-text token (eos), i.e.
-until it predicts that the text ends.
-LLMs are pre-trained on large text corpora, such as the [Common Crawl](https://commoncrawl.org/) which contains over 250 billion pages of text.
-Their architecture is rooted in the attention mechanism.
-The attention mechanism models interactions in the text, such as negations, synonyms, or grammatical structures, across a long sequence of text.
-@vaswani2017attention implement this mechanism in the transformer block, which is a modular building block for the neural network underlying an LLM.
-The combination of these transformer blocks, giant pre-training data, and huge computing resources, lead to LLMs that have "emergent capabilities".
-These are capabilities, which were not explicitly trained for by the developers of the model, such as the ability of a model to translate between languages, pass the BAR exam, or write textual summaries of documents [see @weiEmergentAbilitiesLarge; @luAreEmergentAbilities2023].
-Unidirectional LLMs, such as GPT or LLaMa [@radford2018improving; @metaMetaLlama], are built for text-generation.
-They predict the next token in a sequence of tokens autoregressively, therby, these models only work from left-to-right (or right-to-left).
-Bidirectional LLMs, such as BERT [@devlinBERTPretrainingDeep2018], are built for representing text.
-They predict a token given the context around this token, and can hence read the text both from left-to-right and right-to-left.
-Both types of models share two note-worthy components, that account for a large share of their power.
-One, developers pre-train both types of models on large bodies of text, to learn about the structure of language itself.
-Two, by using a transformer architecture [@vaswani2017attention], which enables these language models to learn dependencies between words, such as negations (short-range dependency), or document structures, e.g.
-salutations in letters (long-range dependency).
-<!-- TODO: Simplify this paragraph -->
+<!--
 
@@
 The more beams we search, the more likely we are to find a sequence with a higher likelihood, then we would have found otherwise [@huggingfaceGenerateText].
+  -->
 
@@
 The input data to this autoencoder is an identity matrix of dimension $D$.
+Before passing the summary embedding to the LLM, we pass these nodes into a Layer Normalization to stabilize the training process [@baLayerNormalization2016].
+Layer normalization normalizes across the hidden nodes, rather than across the observations in a batch [batch-normalization, see @ioffeBatchNormalizationAccelerating2015].
+Thereby, it can also be used with a single observation.
 We illustrate this neural network representation of the factor model in @fig-factor.
@@
 We also define the vectors of biases for the encoder and decoder by $\boldsymbol{b}^{Encoder}$ (length $F$) and $\boldsymbol{b}^{Decoder}$ (length $E$), and stack them into matrices by taking the Kronecker product with a length $D$ vector of ones, $\boldsymbol{1}_D$, to obtain the biases in matrix form as $\mathbf{B}^{Encoder} = \boldsymbol{1}_D \otimes \boldsymbol{b}^{Encoder}$ and $\mathbf{B}^{Decoder} = \boldsymbol{1}_F \otimes \boldsymbol{b}^{Decoder}$.
-For notational convenience, we denote this factor model by $\operatorname{Factor}_{\Omega_{Factor}} \left(\mathbb{I}_{D} \right): \lbrace 0, 1 \rbrace^{D \times D} \to \mathbb{R}^E$, with parameters $\Omega_{Factor}$, where $\Omega_{Factor} = \lbrace \mathbf{W}^{Encoder}, \mathbf{W}^{Decoder}, \mathbf{B}^{Encoder}, \mathbf{B}^{Decoder} \rbrace$.
+For notational convenience, we denote this factor model by $\operatorname{Factor}_{\Omega_{Factor}} \left(\mathbb{I}_{D} \right): \lbrace 0, 1 \rbrace^{D \times D} \to \mathbb{R}^{D \times E}$, with parameters $\Omega_{Factor}$, where $\Omega_{Factor} = \lbrace \mathbf{W}^{Encoder}, \mathbf{W}^{Decoder}, \mathbf{B}^{Encoder}, \mathbf{B}^{Decoder} \rbrace$.
 
@@
 
-<!-- TODO: Can I split this into separate optimizations? Why / why not? -->
+<!-- TODO: Is it S or S_{d, :}? Can I split this into separate optimizations? Why / why not? -->
 $$
@@
 
+<!-- TODO: Talk about Layer Normalization, explain difference to BatchNormalization -->
+<!-- TODO: Why does result depend on Layer Normalization? Is there a bug, why not? (Scrambled version works too) -->
+
+@baLayerNormalization2016
+
+- also used in @vaswani2017attention
+- can be used in test data
+- would like not to use it
+- majority of claims close to zero; However, some outliers which are pushed away from the others
+
+<!-- TODO: Note on optimization and behavior of likelihood, spike in gradient -->
 <!-- TODO: What does it really return? -->
@@
 
+## Diagnostics
+
+- Other methods as benchmarks
+  - Word2Vec
+  - BERT
+
+To analyze our results, we will use a suite of diagnostic tools, which we will explain here briefly.
+
+- Correlation matrix
+
+  - First, we de-mean each embedding dimension by subtracting the mean of this embedding dimension calculated across observations.
+The motivation for this is, to take out parts of the embedding that are due to the domain of the advertising claims.
+
+  - Correlation matrices across the claims and across the embedding dimensions: When analyzing the summary embeddings, we form correlations across the advertising claims and correlations across the embedding dimensions.
+The former is the same as a correlation matrix of a data matrix with as many rows as embedding dimensions and as many columns as the number of observations, while the latter is the correlation matrix of a data matrix with as many rows as observations and as many columns as embedding dimensions.
+
+- Leave-one-out regression (LOO)
+  - We also perform leave-one-out regression, here we regress a focal embedding onto the average embedding from the one class and the average embedding from the other class, and an intercept.
+We repeat this for all observations.
+Since all embeddings are de-meaned, we expect an average intercept of zero for all regressions.
+We also expect that the estimates for the focal claim's group are positive, while the estimates for the other group should be negative.
+
+- Principal Component Analysis (PCA)
+  - PCA is a dimensionality reduction technique, which is rooted in the singular value decomposition.
+Intuitively, it projects the data onto orthogonal axes in a way that each axis captures as much variance as possible.
+We select the first two principal components, as this allows us to visualize our data.
+  - Eigenvalue? Scree plot?
+
 # Data
 
-CITE: Prompt engineering
+As a first evaluation, we create a small synthetic dataset of 20 advertising claims for a haircare product by querying ChatGPT with the prompt in @fig-gpt.
+We want these claims to differ whether they are pronouncing tangible or intangible attributes of the product.
+These advertising claims all focus on how "shiny" the product ones hair makes.
+As as robustness check, we repeated this task for different attributes (instead of shinyness: Healthiness and colorfulness of hair), and for a different product category (surface cleaners).
+In total, we obtain 80 advertising claims, 20 for each of these settings.
+Half of these claims are supposed to be tangible and half intangible.
+Additionally, we check whether our results are robust to changes in the order of the claims, to rule out data leakage.
+We present these data in @tbl-claims.
 
@@
 ::::
-
-As as robustness check, we repeated this task for different attributes (instead of shiny, healthiness and colorfulness of hair), and for a different product category (surface cleaners).
-In total, we obtain 80 advertising claims, 20 for each of these settings.
-Half of these claims are supposed to be tangible and half intangible.
-
-We obtain data from a market research company, that includes advertising claims for different FMCG.
-In our application, we focus on dairy and hair shampoo products.
-These data, stem from different variations of choice-based conjoint studies [@eggersChoiceBasedConjointAnalysis2022], which we cannot further disclose due to confidentiality agreements.
-The data includes choice-experiments aggregated at the level of the advertising claims, hence we do not have responses of individuals.
-These measurements of consumer-preferences are on multiple dimensions, such as relevance, uniqueness of a claim, brand fit, or an overarching rating of the claim.
-The advertising claims were designed for markets in different countries and are in english (US, UK) or translated into english (others).
-The date of the study is also included in the data.
-We also have information on the marketers motivation behind the design of an advertising claims, such as whether she designed the claims to pronounce the health benefits of a yoghurt, or the taste yoghurt, as well as the brand of the product.
-
-<!-- TODO: Some graphs / tables for descriptives -->
-
-- Descriptives of the data
-
-- Prepocessing of the data
-
-<!-- TODO: Provide examples that are created with GPT, i.e. are fake -->
-
-# Results
-
-TODO: Have similarity measure vs euclidean distance in encoding space
-
-- Performed suite of tests of generated data
-  - haircare shiny
-  - haircare healthy
-  - haircare colorful
-
-[ ] Training of embeddings
-
-![Training of embeddings](figures/dashboard.png)
-
-- computation
-- dealing with spikes
-
-[ ] Analysis of embeddings; Plot correlation of claims and dims, PCA
-
-::: {#fig-correlation layout-ncol=2 layout-nrow=2}
-![RE](figures/correlation_re.pdf)
-
-![BoW](figures/correlation_bow.pdf)
-
-![W2V](figures/correlation_w2v.pdf)
-
-![BERT](figures/correlation_bert.pdf)
-
-Correlation matrices along the claims. 
-:::
-
-::: {#fig-correlation-surface layout-ncol=2 layout-nrow=2}
-![RE](figures/correlation_re_surface.pdf)
-
-![BoW](figures/correlation_bow_surface.pdf)
-
-![W2V](figures/correlation_w2v_surface.pdf)
-
-![BERT](figures/correlation_bert_surface.pdf)
-
-Correlation matrices along the claims. 
-:::
-
-::: {#fig-correlation-haircolor layout-ncol=2 layout-nrow=2}
-![RE](figures/correlation_re_color.pdf)
-
-![BoW](figures/correlation_bow_color.pdf)
-
-![W2V](figures/correlation_w2v_color.pdf)
-
-![BERT](figures/correlation_bert_color.pdf)
-
-Correlation matrices along the claims. 
-:::
-
-::: {#fig-correlation-health layout-ncol=2 layout-nrow=2}
-![RE](figures/correlation_re_health.pdf)
-
-![BoW](figures/correlation_bow_health.pdf)
-
-![W2V](figures/correlation_w2v_health.pdf)
-
-![BERT](figures/correlation_bert_health.pdf)
-
-Correlation matrices along the claims. 
-:::
-
-::: {#fig-correlationts layout-ncol=2 layout-nrow=2}
-![RE](figures/correlation_ts_re.pdf)
-
-![BoW](figures/correlation_ts_bow.pdf)
-
-![W2V](figures/correlation_ts_w2v.pdf)
-
-![BERT](figures/correlation_ts_bert.pdf)
-
-Correlation matrices along the embedding dimensions. 
-:::
-
-::: {#fig-pca layout-ncol=2 layout-nrow=2}
-![RE](figures/pca_re.pdf)
-
-![BoW](figures/pca_bow.pdf)
-
-![W2V](figures/pca_w2v.pdf)
-
-![BERT](figures/pca_bert.pdf)
-
-First two principal components of the embedding matrix, colored by class.
-:::
-
-::: {#fig-loo layout-ncol=2 layout-nrow=2}
-![RE](figures/loo_re.pdf)
-
-![BoW](figures/loo_bow.pdf)
-
-![W2V](figures/loo_w2v.pdf)
-
-![BERT](figures/loo_bert.pdf)
-
-Reuslts of leave-one-out regression, colored by class.
-:::
-
-  - GPT-2
-  - BERT
-  - Word2Vec
-  - RE
-
-[ ] Encoding dimension plot
-
-@fig-encoding, @tbl-claims
-
-::: {#fig-encoding layout-ncol=2}
-
-![All points](figures/encoding_space.pdf){#fig-encoding1}
-
-![Zoom-in](figures/encoding_space_zoomin.pdf){#fig-encoding2}
-
-Encoding space of the advertising claims.
-
-:::
 
@@
 
-- labels, accuracy
-- clustering
-- generation of new claims
-  - exploring the space
-  - measure of fit (area)
+For our empirical application, we obtain data from a market research company, that includes advertising claims for Yoghurt Drinks.
+These data, stem from different variations of choice-based conjoint studies [@eggersChoiceBasedConjointAnalysis2022], which we cannot further disclose due to confidentiality agreements.
+The data includes choice-experiments aggregated at the level of the advertising claims, hence we do not have responses of individuals.
+These measurements of consumer-preferences are on multiple dimensions, such as relevance, uniqueness of a claim, brand fit, or an overarching rating of the claim.
+The advertising claims were designed for markets in different countries and are in english (US, UK) or translated into english (others).
+The date of the study is also included in the data.
+All products are from the same brand.
+In total, we have 60 observations, 20 for each country.
+Many of the researched advertising claims occurr in multiple of these separate studies.
+In total, we have 32 unique advertising claims.
+Most advertising claims get a uniqueness measure around 5.8, with the whole distribution ranging from 4.8 to 7.4.
+When we look at the uniqueness measure with respect to the country in which the study was held, we see that the variance is the smallest for the German market and the largest for the spanish market.
+However, the means are not statistically different across the countries.
+The distribution of the rating measure is similar in shape, ranging from 43 to 80.
+The rating and uniqueness are positively correlated (correlation of $0.8843$)
 
-[ ] generation of new claims [@Goodfellow2016]
+<!-- We also have information on the marketers motivation behind the design of an advertising claims, such as whether she designed the claims to pronounce the health benefits of a yoghurt, or the taste yoghurt, as well as the brand of the product.
 
-![Exploration of Encoding Space](figures/exploration_encoding_space.pdf)
+## MR 5
 
-![Exploration of the likelihoods](figures/likelihoods_generation.pdf)
+- One brand, yoghurt drink
+- MaxDiff experiment, ranking of choices
+- Relevance, fit with brand, rank
+- 32 unique claims, 3 countries: Germany, Spain, UK
+
+ -->
+
+<!-- TODO: Some graphs / tables for descriptives -->
+
+The scale for the uniqueness measure is from 0 to 10 and for the rating measure from 0 to 100.
+
+::: {#fig-eda-mr4-uni layout-ncol=2}
+![Histogram across countries](figures/uniqueness_mr4_eda.pdf){#fig-eda-mr4-uni-a}
+
+![Boxplot per country](figures/uniqueness_box_mr4_eda.pdf){#fig-eda-mr4-uni-b}
+
+Distributions of measure for claim uniqueness and ratings across countries.
+:::
+
+::: {#fig-eda-mr4-rating layout-ncol=2}
+![Histogram across countries](figures/rating_mr4_eda.pdf)
+
+![Boxplot per country](figures/rating_box_mr4_eda.pdf)
+
+Distributions of measure for claim uniqueness and ratings across countries.
+:::
+
+![Scatterplot of Uniqueness and Rating measures, colored by country.](figures/scatterplot_mr4_eda.pdf)
+
+<!-- TODO: Provide examples that are created with GPT, i.e. are fake -->
+
+<!-- ## MR 6
+
+- 5 countries: Brazil, Mexico, Russia, France, Germany
+- 40 unique claims -->
+
+# Results
+
+## Single Embeddings
+
+- HairColor: 42 minutes
+- HairHealth: 14 minutes
+- Surface: 16 minutes
+- Surface Scramble: 9 minutes
+
+<!-- TODO: fill in the computation times -->
+<!-- TODO: Have model with e.g. 50 factors? Then for many claims? -->
+<!-- TODO: Have model without LayerNorm -->
+
+For this project, we obtained a [SURF NWO Small Compute Grant](https://www.surf.nl/en/small-compute-applications-nwo).
+We train these summary embeddings until we obtain a likelihood to generate the target claim of $0.99$.
+Performing this training on a Nvidia A100-GPU, takes about 30 seconds for all 20 claims.
+The training of the factor representation with 2 factors and for the same claims is more computationally intense and takes about Y minutes on the same computer.
+Here we traing for a joint likelihood of $0.99$.
+We always verified that the summary embeddings regenerate all focal claims correctly.
+
+@fig-correlation-single
+
+::: {#fig-correlation-single layout-ncol=3}
+![RE](figures/correlation_re_hair_single.pdf)
+
+![W2V](figures/correlation_w2v.pdf)
+
+![BERT](figures/correlation_bert.pdf)
+
+Correlation matrices along the claims. 
+:::
+
+- does not work to capture the two classes: Overfitting?
+- does regularization help?
+
+We can interpolate between to target strings by forming a weighted average of their summary embeddings.
+@tbl-interpolation shows the strings that we generated for a convex combination of the claims *"Unlock the secret to hair that shines from within, reflecting your inner glow."* (A) and *"Rediscover the joy of hair that beams with inner vibrancy."* (B).
+We combine them as $\boldsymbol{s}_{combine} = weight * \boldsymbol{s}_{A} + (1 - weight) * \boldsymbol{s}_{B}$, where $weight \in [0, 1]$.
+For weights that are close to either $0$ or $1$ we still generate the respective focal claim.
+When we start to step further into the middle between the two claims, first the end of the sequence changes (e.g.
+*"Unlock the secret to blackened nails that shine from the shine."* at for $weight = \frac{3}{20}$).
+For weights that are close to $\frac{1}{2}$, we stop generating eos-tokens and the generated strings become unintelligble.
+However, these strings still maintain some of the structure of the focal claims, such as having words related to beauty products and starting with the syllable "Un" or "Red".
+
+|    | Weight | Generated String                                                                                      |
+|---:|:-------|:------------------------------------------------------------------------------------------------------|
+|  0 | 0/20   | Unlock the secret to hair that shines from within, reflecting your inner glow.   \textit{eos-token}                     |
+|  1 | 1/20   | Unlock the secret to hair that shining from within, your bedroom. \textit{eos-token}                                     |
+|  2 | 2/20   | Unlock the secret to black metal's glow. \textit{multiple eos-token}                                                              |
+|  3 | 3/20   | Unlock the secret to blackened nails that shine from the shine.   \textit{eos-token}                                    |
+|  4 | 4/20   | Un to of for the team members upon                                                                    |
+|    |        | Moderation of the, that                                                                               |
+|  5 | 5/20   | Un to (93) of 15 + 18 + 36                                                                            |
+|    |        | Add to                                                                                               |
+|  6 | 6/20   | Un to (mit. of private eye and nirvana-ly-ly                                                          |
+|  7 | 7/20   | Uncle-healedered with a lifetime-changing blend of light and energy and                               |
+|  8 | 8/20   | Unclelyed by the Tone's Bright Side™ Lipstick.                                                        |
+|    |        | Bright                                                                                                |
+|  9 | 9/20   | Unclelying our clients and fory, we're sure about your smile.                                         |
+| 10 | 10/20  | Redmates on the Road to of Dreams.                                                                    |
+|    |        | Inspirating your soul with                                                                            |
+| 11 | 11/20  | Redmates at the Sunlight Your Hair.     \textit{multiple eos-token}                                                              |
+| 12 | 12/20  | Redhs founder, that a master the way she loves......                                                  |
+| 13 | 13/20  | Redhs founder, of where, at, and to and to, and can be                                                |
+| 14 | 14/20  | Rediscover the joy of purpose and detail with a strand of your hair. (and                             |
+| 15 | 15/20  | Rediscover the joy of success with long,istry's health and creativity. back to                        |
+| 16 | 16/20  | Rediscover the joy of hair that loves to making you.™ the way.br                                      |
+| 17 | 17/20  | Rediscover the joy of hair that's at the-corrects life with a healthy                                 |
+| 18 | 18/20  | Rediscover the joy of hair that beams with inner vibrancy. \textit{eos-token}                                           |
+| 19 | 19/20  | Rediscover the joy of hair that beams with inner vibrancy. \textit{eos-token}                                           |
+: Interpolation between two advertising claims. {#tbl-interpolation}
+
+## Factor Representation
+
+<!-- TODO: Unclear what area means in factor space -->
+<!-- TODO: Add freely estimated summary embeddings -->
+<!-- TODO: Give reason for need for regularization -->
+<!-- TODO: Give reason for using two factors -->
+<!-- TODO: Give the right methods / diagnostics -->
+<!-- TODO: Explain generation -->
+<!-- TODO: Give results without LayerNorm, why is it weird? -->
+
+For our factor model, we use two factors. The intuition for this is, that we want to distinguish two classes of tokens and want to be able to visualize our data in a two-dimensional space.
+
+<!-- TODO: Have similarity measure vs euclidean distance in encoding space -->
+
+- Performed suite of tests of generated data
+  - haircare shiny
+  - haircare healthy
+  - haircare colorful
+
+<!-- TODO: Just show likelihood graph -->
+
+![Training of embeddings](figures/dashboard.png)
+
+If our summary embeddings are able to capture relevant information in the advertising claims, then we expect to be able to separate between the tangible and intangible advertising claims.
+
+For this, we demean the embeddings across the embedding dimensions such that the only variation left should be due to the claims being tangible or intangible.
+
+When visualizing these correlations, we expect the first quartile and the third quartile to contain positive correlations, while the second and the fourth quartile should contain negative correlations.
+In @fig-correlation we present these correlation matrices for our Summary embeddings, Word2Vec embeddings, and BERT embeddings.
+We see this pattern for all three methods, indicating that all three methods are able to capture this part of information in the advertising claim.
+However, the correlations are large (in absolute value) for the summary embeddings.
+This is a first indication, that our factor structure captures this difference in the advertising claims.
+
+::: {#fig-correlation layout-ncol=4}
+
+![No Normalization](figures/correlation_re_hair20_nonorm.pdf)
+
+![With LayerNorm](figures/correlation_re.pdf)
+
+![W2V](figures/correlation_w2v.pdf)
+
+![BERT](figures/correlation_bert.pdf)
+
+Correlation matrices along the claims. 
+:::
+
+<!-- TODO: Smaller vocabulary than dims? Just compare RE and BERT -->
+
+@fig-correlationts shows correlation matrices across the embedding dimensions, i.e.
+each cell shows the correlation between two embedding dimensions calculated across the $20$ advertising claims.
+When comparing our summary embedding and the BERT based embeddings, we can see a stronger grid pattern with more correlations that are removed from zero.
+this is something we would anticipate for our factor model, as we restricted the degrees of freedom of the elements in the embedding vector to 2.
+Hence, many of the dimensions should be linearly dependent on each other.
+In contrast, for the BERT embedding most dimensions are barely correlated with each other.
+
+::: {#fig-correlationts layout-ncol=3}
+![RE](figures/correlation_ts_re.pdf)
+
+![W2V](figures/correlation_ts_w2v.pdf)
+
+![BERT](figures/correlation_ts_bert.pdf)
+
+Correlation matrices along the embedding dimensions. 
+:::
+
+We perform Principal-Component Analysis on the embedding matrices and visualize the first two principals components in @fig-pca.
+We color the data points by their respective class.
+We see that we can separate these two classes linearly for all three embeddings.
+
+::: {#fig-pca layout-ncol=3}
+![RE](figures/pca_re.pdf)
+
+![W2V](figures/pca_w2v.pdf)
+
+![BERT](figures/pca_bert.pdf)
+
+First two principal components of the embedding matrix, colored by class.
+:::
+
+@fig-loo shows the results of the leave-one-out regression. For the summary embeddings, the variance of the intercept is substantially larger than for the other embeddings.
+
+::: {#fig-loo layout-ncol=3}
+![RE](figures/loo_re.pdf)
+
+![W2V](figures/loo_w2v.pdf)
+
+![BERT](figures/loo_bert.pdf)
+
+Reuslts of leave-one-out regression, colored by class.
+:::
+
+@fig-encoding shows the low dimensional sub-space in which we located the advertising claims.
+It appears that we pick up the class with the two hidden factor, however, the class is not directly aligned with one of these axes.
+Many of the points are close to zero, with a few of the claims being far removed from the other points, e.g.
+claims 1, 9, and 3 for the tangible, and claims 12, 16, and 19 for the intangible claims.
+
+@tbl-claims
+
+::: {#fig-encoding layout-ncol=2}
+
+![All points](figures/encoding_space.pdf){#fig-encoding1}
+
+![Zoom-in](figures/encoding_space_zoomin.pdf){#fig-encoding2}
+
+Encoding space of the advertising claims.
+Numbered points are locations of claims that are part of the training data, colored dots are points where we generate one of the claims that are part of the training data, and red crosses are locations where we generate a claim that ends with the eos-token, but is not part of the training data.
+Whitespace marks coordinates where we generate a string that does not contain an eos token.
+:::
+
+<!-- 
+Another explanaition could lie in the decoding strategy.
+@ariholtzmanCuriousCaseNeural2020 show how text generation based on a maximum likelihood objective leads to text that sounds bland and not human, propsing the use of different decoding strategies to make the generated text more human.
+-->
+
+In @fig-space, we perform a grid search across the factor space.
+For each point in the grid search, we generate a text by passing the point through the decoder and the resulting embedding as an input into the LLM.
+In the plot, we have three types of different markers.
+A blue bubble with a number represents the location of the respective claim, while a colored dot represents a point where we generate one of the points that are part of the training data.
+Red crosses are points from which we generate a string, that comes to an end with the end of text token and is not part of the training data ("candidate points").
+Whitespace represents points where we do not generate a string that comes to an end within the number of tokens that we consider.
+Around each advertising claim, we find an island of points from which we can generate the same claim.
+The islands are surrounded by candidate points, areas which are further removed from any of the training claims tend to be whitespace.
+
+![Exploration of Encoding Space](figures/exploration_encoding_space.pdf){#fig-space}
+
+<!-- TODO: Turn this into bar-charts? -->
+
+We explore the generation based on our summary embedding deeper, by analysing the generation tree when generating with the summary embedding.
+We expect that at each generation step, the probability for the correct token approaches one, while the probability for all other tokens goes to zero.
+In @fig-like, we visualize the next token probabilities at each generation step for the tokens of the vocabulary.
+We see that indeed the probability for the correct token is close to one, while the probability for all other tokens is close to zero.
+In @tbl-lowprob, we show the generation probabilities for the three most likely tokens at each generation step.
+This example shows the claim, for which the probability of the correct token in the first position is the lowest out of all claims.
+The critical step in the generation appears to be the first token of the sequence, as the probability to generate the correct token in the second position jumps for $0.8876$ to $0.9925$.
+In @tbl-similartoken we show an example of a generation procedure, where also the second and third most likely tokens are insightful.
+For example, for the first token, where the true token is "Experience", which we correctly generate, the 2nd and 3rd most likely tokens are "Form" and "See".
+These are words, which are quite similar in meaning to the word "Experience".
+We can find similar examples for the second token (1st Choice: "50", 2n Choice: "25", 3rd Choice: "20") and the fifth position (1st Choice: "visible", 2nd Choice: "protective", 3rd Choice: "shine").
+The fact that these tokens get the second and third highest generation probabilties is likely due to the pre-training of the LLM, and reflects how these words tend to be used in similar situations.
+
+![Conditional generation probability for the most likely token at each step. All tokens match the target claims.](figures/likelihoods_generation.pdf){#fig-like}
 
@@
 | 10 | ).    | .)    | .             | 0.000529124 | 0.00114012  |       0.996219 |
-| 11 | \textit{2x linebreak}      | ]     | \textit{eos-token} | 0.000603324 | 0.00100365  |       0.995909 |
-| 12 | \textit{2x linebreak}    | \textit{linebreak}      | \textit{eos-token} | 7.61097e-06 | 0.00124311  |       0.99872  |
-| 13 | \textit{2x linebreak}    | \textit{linebreak}      | \textit{eos-token} | 2.16886e-05 | 0.000565556 |       0.999231 |
-| 14 | \textit{2x linebreak}    | \textit{linebreak}      | \textit{eos-token} | 3.80865e-05 | 0.00084979  |       0.998713 |
-| 15 | \textit{2x linebreak}    | \textit{linebreak}      | \textit{eos-token} | 5.16162e-05 | 0.00161664  |       0.997724 |
-| 16 | \textit{2x linebreak}    | \textit{linebreak}      | \textit{eos-token} | 0.000101122 | 0.0031923   |       0.995816 |
+| 11 |       | ]     | \textit{eos-token} | 0.000603324 | 0.00100365  |       0.995909 |
+| 12 |  \textit{linebreak}   |   \textit{linebreak}    | \textit{eos-token} | 7.61097e-06 | 0.00124311  |       0.99872  |
+| 13 |  \textit{linebreak}   |   \textit{linebreak}    | \textit{eos-token} | 2.16886e-05 | 0.000565556 |       0.999231 |
+| 14 |  \textit{linebreak}   |   \textit{linebreak}    | \textit{eos-token} | 3.80865e-05 | 0.00084979  |       0.998713 |
+| 15 |  \textit{linebreak}   |   \textit{linebreak}    | \textit{eos-token} | 5.16162e-05 | 0.00161664  |       0.997724 |
+| 16 |  \textit{linebreak}    |  \textit{linebreak}     | \textit{eos-token} | 0.000101122 | 0.0031923   |       0.995816 |
 : Example for a low 1st probability {#tbl-lowprob}
@@
 | 12 | Transfer |            | \textit{eos-token} | 0.000179846 | 0.00032993  |       0.998234 |
-| 13 |          | More       | \textit{eos-token} | 0.000109437 | 0.000194042 |       0.998748 |
-| 14 | More     |            | \textit{eos-token} | 7.64376e-05 | 0.000174342 |       0.999236 |
-| 15 | Lab      |            | \textit{eos-token} | 5.29115e-05 | 0.000444251 |       0.998936 |
-| 16 | Lab      |            | \textit{eos-token} | 0.000276325 | 0.00155492  |       0.997141 |
+| 13 |  \textit{linebreak}        | More       | \textit{eos-token} | 0.000109437 | 0.000194042 |       0.998748 |
+| 14 | More     |   \textit{linebreak}         | \textit{eos-token} | 7.64376e-05 | 0.000174342 |       0.999236 |
+| 15 | Lab      |   \textit{linebreak}         | \textit{eos-token} | 5.29115e-05 | 0.000444251 |       0.998936 |
+| 16 | Lab      |   \textit{linebreak}         | \textit{eos-token} | 0.000276325 | 0.00155492  |       0.997141 |
 : Example for similar tokens {#tbl-similartoken}
-
-  - exploration of the likelihoods
-  - frequentist test of generating existing
-
-[ ] Repetition with actual marketing data
-
-### Implementation
-
-- GPT-2; Far from state of the art, if it works here, it will work with other model
-
-- Autoencoder
-
-- Configuration of LLM
-
-- Connection to other measures, e.g. syntax or rating of consumers
-
-- Different generation strategies, use beam search to get closer to max LL
-
-- Optimization
-  - @diederikp.kingmaAdamMethodStochastic2014
-  - Computation time, scaling
-  - Issues with spikes; Figures where they come from
-
-- Encoding space
-  - example tangible, intangible
-  - generations from encoding space
-  - Wasteland, geometrical shape and where it comes from
-
-- Measures of fit
-  - areas in the encoding space (count out the pixels and compare to curvature of LL?)
-  - analysis of the trees
-  - Gap between most likely and second most likely token
-
-- Comparison to BERT embeddings?
-- Comparison to prompt engineering results?
 
@@
 
-- Summary statistics of the data
-
-- Application to MR data
-
-<!-- TODO: Check whether this is okay to name -->
-
+<!-- 
 - Two small sub-samples from the same market research study of product claims for a dairy product
@@
   - MR3: Major theme: "RTB vs. Benefit", minor theme: "Amino-Acid", "Energy", ..., country NL, fitness brand, language english
-  - MR4: Hair care, country BR, total sample, find difference between two brands: Both from same FMCG  corporation
+  - MR4: Hair care, country BR, total sample, find difference between two brands: Both from same FMCG  corporation -->
+
+<!--
 
@@
 
+-->
+
+Since we cannot disclose the actual advertising claims, we generate a set of similar claims with [ChatGPT](https://chatgpt.com/). However, we performed all computations on the actual data.
+
+Artificial examples for the yoghurt drink advertising claims:
+
+- *"A burst of fresh flavor to energize your morning"*
+- *"Refresh your day with a lively new taste"*
+- *"Dynamic flavor for an invigorating start"*
+- *"Begin your day with a crisp and revitalizing taste"*
+- *"Revitalize your senses with a pure, fresh flavor"*
+
 ::: {#fig-anecdote-mr5 layout-ncol=2}
@@
 ![Uniqueness score of embedded yoghurt drink claims.](figures/uniqueness_mr5.pdf){#fig-ratings-mr5a}
+
+We trained this model until we reached a joint likelihood of $0.99$, the training took around 8 minutes.
 
@@
 
-- Generative AI for personalization
+Managers can use these summary embeddings to augment the design process of advertising claims.
+The work by @burnap2023product and @ludwigMachineLearningTool2023 propose a framework for guided generation of images.
+The researchers model an embedding space for images, that fulfills two requirements: One, points in this embedding space can generate new images and two, they form useful features for a supervised machine learning task.
+For our setting of text data, we also estimate an embedding that can generate new text data and captures inherent information of the advertising claims, which could be used for classification.
+The framework which @burnap2023product and @ludwigMachineLearningTool2023 use consists of three components: The encoder, which projects datapoints into an encoding space, the generator, which turns points from the encoding space into new datapoints, and the predictor, which predicts a certain label based on a point from the encoding space [compare @burnap2023product].
+@burnap2023product propose a process to use this generative framework in the design process of cars, by optimizing for the predicted rating of a certain car-design.
+In a similar way, businesses could use such a setup to improve their advertising claims, e.g.
+by training the generator and predictor model based on existing market research data.
+Based on existing market research data which companies have on the performance of advertising claims that they workshopped in the past, we can train a predictor model which predicts a certain outcome measure based on the summary embedding of the advertising claim.
+We can combine such a predictor model with the low-dimensional factor space of the summary embeddings, to guide the exploration of this space.
+For this, we only need the decoder part of our factor model: For every point in the factor space, we can get the corresponding summary embedding and predict the respective outcome for this summary embedding with the predictor model.
+Through gradient descent, we can now find the point in the factor space that maximizes the predicted outcome.
+In a classification setting, we can use this technique, e.g.
+to find the closest point in the factor space which changes the predicted class of the advertising claim.
 
-- Brands
+A framework of generator and predictor model can also help in the design of market research itself.
+@ludwigMachineLearningTool2023 propose a workflow for generating research hypotheses based on such a model, by finding small steps in the encoding space which maximally change the prediction of the predictor model.
+They show that these steps yield interpretable and novel hypotheses for their research setting.
+In a similar vein, marketers might look for the minimal change to an advertising claim, such that this claim changes from being "tangible" to being "intangible".
+This could help to identify effects of certain design motivations more clearly, as it is closer to the "ceteris paribus" principle, which is difficult to uphold in research designs for modalities such as text.
 
-- Wording: Ballpark, finetuning idea
+We show that the distance of advertising claims in the factor space represents their linguistic similarity.
+Such a measure can be a useful tool when evaluating the market with respect to the positioning of different brands and products.
+With a visualization similar to @fig-ratings-mr5a,  managers can identify which products have similar sounding claims, and make predictions on consumers' ratings of competitors claims, even when they did not incorporate these into their own market research.
+Such a measure can also help in the identification of copy-cat product claims, which might steal market shares and damage the focal product's brand.
 
-- Interactive dashboard to explore the embedding space
+Managers can also explore the factor space to find new design motivations.
+@fig-anecdote-mr5 shows design motivations, which were not designated as themes in the market research study.
+By looking at the structure of the factor space, and comparing neighbouring claims for the similarities and differences, managers can identify linguistic features, such as the use of certain words of themes.
 
-- Costs of tailoring advertising claims
-  - When we take current prices for the [GPT-4o LLM; June 2024](https://openai.com/api/pricing/) (\$5 per 1 Mio tokens input, \$15 per 1 Mio tokens output), then training a dataset of 100 advertising claims, for 5,000 epochs will cost \$150.
-Freelancers on platforms such as [Upwork.com](https://www.upwork.com/) charge between \$20 and \$200 per hour for work on tasks such as SEO optimization, copy writing, and creation of advertising claims.
-    - Online services for conjoint studies, such as [Conjointly.com](https://conjointly.com/) and [Sawtooth](https://sawtoothsoftware.com/pricing) charge a few thousand dollars per year to use their services.
-Leveraging the resulting insights as much as possible, by using generative AI, could hence make market research more cost effective.
+<!-- - Paragraph on prompt engineering
+  - move away from prompt engineering
+  - certain tasks should be optimized, similar to automated grading
+  - validation of answers
+  - communication on social media
+  - automated customer service bots
+  - interpolation between different claims
+  - there are applications where distribution is desired -->
 
+<!-- - Rather than inform product design, @ludwigMachineLearningTool2023, use such a framework to generate hypotheses for why a prisoner might be granted bail or not.
+For this, they predict judges decisions' from bail hearings, based on the mugshot photo of the defendant (predictor) and learn about the distribution of mugshots with a Generative Adversarial Network [CITE Goodfellow GAN].
+To generate new hypothesis, they perform gradient descent on the prediction of the predictor with respect to the encoding data point.
+They search for the alteration to the image, that changes the predicted outcome the most.
+By making equal steps along this gradient in opposing directions, they obtain to versions of the same image manipulated with respect to this feature.
+@ludwigMachineLearningTool2023 also find, that comparing these two images often yields people to name the specific change, creating new hypothesis on why an inmate might be granted bail or not.
+-->
+
+<!-- 
 - Generator, predictor compatible embeddings; Generative AI to assist and automate design processes in marketing
@@
 @ludwigMachineLearningTool2023 also find, that comparing these two images often yields people to name the specific change, creating new hypothesis on why an inmate might be granted bail or not.
-  - We call the input data, $x$, and its distribution $p(x)$. In our setting, $x$ is an advertising claim. <!-- TODO: How does it work for @burnap2023product ? -->
+  - We call the input data, $x$, and its distribution $p(x)$. In our setting, $x$ is an advertising claim.
   - We call labels, $y$ and its distribution $p(y)$.
@@
 Rather than GANs, which require more training data as these networks need to learn the properties of an image from the ground up, we can focus directly on the types of texts that are relevant to our domain.
-<!-- NOTE: Is this true?
--->
+NOTE: Is this true?
   - The training objective of LLMs is to predict the next token in a sequence of tokens
@@
 We perform this optimization for our full dataset simultaneously by using a factor model.
-  - By using this factor model, we create a bottleneck, forcing the model to learn helpful features; In other factor models, these features are often interpretable dimensions [@goodfellowDeepLearning2016].
+ -->
+
+<!-- - Generative AI for personalization
+
+- Brands
+
+- Wording: Ballpark, finetuning idea
+
+- Interactive dashboard to explore the embedding space
+
+- Costs of tailoring advertising claims
+  - When we take current prices for the [GPT-4o LLM; June 2024](https://openai.com/api/pricing/) (\$5 per 1 Mio tokens input, \$15 per 1 Mio tokens output), then training a dataset of 100 advertising claims, for 5,000 epochs will cost \$150.
+Freelancers on platforms such as [Upwork.com](https://www.upwork.com/) charge between \$20 and \$200 per hour for work on tasks such as SEO optimization, copy writing, and creation of advertising claims.
+    - Online services for conjoint studies, such as [Conjointly.com](https://conjointly.com/) and [Sawtooth](https://sawtoothsoftware.com/pricing) charge a few thousand dollars per year to use their services.
+Leveraging the resulting insights as much as possible, by using generative AI, could hence make market research more cost effective.
+-->
 
@@
 
-- Issues with Generation
-  - How to evaluate the quality of the generated ad claims @tatsunorihashimotoUnifyingHumanStatistical2019
-  - Another explanaition could lie in the decoding strategie.
+Newly generated advertising claims are often unintelligible and there appears to be a lot of whitespace in the factor space.
+We find three reasons for this.
+One, we use the GPT-2 model in this paper, which is far from the state-of-the-art, e.g.
+on Huggingface's Open LLM Leaderboard, the best versions of GPT-2 achieve around half the score of the leadboard leader, which is a model based on [Llama-3](https://llama.meta.com/llama3/).
+To address this problem, we can use a more capable open-source LLM, such as Llama 3 instead, which comes at a larger computational costs.
+The second reason lies in the construction of our factor space.
+Accoding to @goodfellowDeepLearning2016, factor models can learn and represent features of the data well.
+However, they struggle with generation of new data points, which in practice, tend to be mixtures of the learned features rather than realistic data points.
+This is a pattern we see with out newly generated advertising claims, e.g.
+in @tbl-interpolation.
+We can address this problem by using a different architecture: Rather than using a factor model, we could use an autoencoder, which takes a different word embedding of the training data, e.g.
+BERT's `[CLS]` token [@devlinBERTPretrainingDeep2018], as an input and learns a mapping from these descriptive embeddings to our summary embeddings.
+The third reason also related to the architecture of our current model, but with respect to the number of layers and activation functions.
+Currently, we a searching for a representation of the advertising claims in 2 dimensions, and take linear combinations of the resulting factors to create our summary embeddings.
+Perhaps, this structure is too simplistic to represent this language domain well.
+Introducing additional layers and non-linear activation functions, such as the Swish function [@ramachandranSearchingActivationFunctions2017a], might yield to a representation of the space that is better suited for generation, as we learn the manifold on which these advertising claims live, better.
+
+- Improve generation by generating from 2nd and 3rd likeliest tokens in the generation tree. These are close variations on target sequences.
+
+- Can we identify higher levels of abstraction?
+
+There are some challenges, which we inherit from LLMs themselves, such as limited language modeling capabilities in non-english languages and safety and copyright issues when generating new text.
+However, these are problems that we can cirumvent, or at least mitigate, by using a more capable LLM with safeguards or an LLM which has been trained on a specific language if we are working with non-english texts.
+
+- computation issues
+  - single embeddings a lot faster than factor
+  - factor scales poorly with more data; Then need to use more factorss
+
+<!-- TODO: Do this with BERT and PCA? -->
+
+- problems with the market data application
+  - low sample; Even doubling of claims
+  - form labels ourselves
+  - these are not insights that we could have gotten with BERT & PCA?
+
+<!--
+
+- Implementation challenges
+  - Computation time
+  - [Tokenizer can cause problems?](https://www.promptingguide.ai/research/llm-tokenization)  TODO: Is this true? Give an example? Won't be a problem forever ... But inherent to this type of model
+
+- Another explanaition could lie in the decoding strategy.
 @ariholtzmanCuriousCaseNeural2020 show how text generation based on a maximum likelihood objective leads to text that sounds bland and not human, propsing the use of different decoding strategies to make the generated text more human.
-
-  - Using embedding rather than identity matrix as the input turns this into an autoencoder, rather than a factor model.
-According to @goodfellowDeepLearning2016, generation from factor models leads to a combination of the learned features in these factors, but not to the generation of sensible data points.
-Existing, autoencoder based, approaches that have succeeded in generation tasks, might be more suitable for this task (e.g.
-CITE: Variational Autoencoder).
 
@@
 Then there would be other latent variable architectures, such as [Boltzman machines](https://www.cs.toronto.edu/~rgrosse/courses/csc321_2017/slides/lec20.pdf)
-
-- Implementation challenges
-  - Computation time
-  - [Tokenizer can cause problems?](https://www.promptingguide.ai/research/llm-tokenization) <!-- TODO: Is this true?
-Give an example?
-Won't be a problem forever ...
-But inherent to this type of model -->
-  - Not as flexible as general LLMs, as we need to have a target sequence and have a use case where we don't want a distribution
+ -->
 
@@
 
+Another adjustments which could improve the quality of our text generation could be to introduce an adversary model, creating an adversarial structure.
+An adversarial structure consists of two models: A generator and an adversary.
+The generator creates new datapoints, while the adversary tries to predict whether a datapoint is genuine or has been generated [@goodfellowGenerativeAdversarialNets2014].
+This leads to a game between these two models, where the generator learns to represents the underlying data distribution, such that the adversary cannot distinguish between genuine and generated datapoints anymore.
+Training such an adversary can also be a way to evaluate the quality of the generated text itself, which is an ongoing research problem [@tatsunorihashimotoUnifyingHumanStatistical2019].
+Exploring how we can reconcile such an adversarial structure with our current training objective of maximizing the generation probability of a focal sequence, is an issue we leave for future research.
+
 <!-- TODO: Regularization of summary embeddings, have them absolute and not relative -->
 
-  - Incorporating the context of the product itself, embedding e.g. the brand, product features and competitors in the model, to move into part of the generation space that is "open" for the product
-  - Using fine-tuned LLMs
-  - MAB around the fitted AE, to incorporate consumer feedback in online learning: Which areas of the space are valuable for which consumers?
-  - For online learning, could wrap the decoder into an MAB and let the MAB explore the space by pulling arms <!-- TODO: Which kind of MAB? -->
-  - Could also learn preferences online with this setup
-  - Better LLM
-  - Image, video, website, audio data. Perhaps website a good application. Perhaps audio a rather simple way to make this multi-modal (content and sound)
+Can we tailor the training objective to a specific marketing goal?
+For example, rather than maximizing the likelihood to generate a known advertising claim, can we also maximize the likelihood to generate an advertising claim that is liked by a consumer?
+
+Can we adapt the LLM further to a specific domain? For example, the approach by 
+@khattabDSPyCompilingDeclarative2023 also allows for fine-tuning of the LLM with respect to a specifc performance metric.
+Rather than using an out of the box LLM, which has been trained on large text corpora, we could further train and adapt the LLM to the advertising claims, which are special in their use of words and short.
+Adapting the LLM to the language domain might prove to helpful, especially when we are interested in the factor space of a specific domain.
+Furthermore, such an adaptation might also improve the quality of new generated texts, as the LLM pronounces sequences of words that sounds like advertising claims more strongly.
+Besides fine-tuning, @ouyangTrainingLanguageModels2022 argue that especially Reinforcement Learning for Human Feedback (RL-HF), improves the capabilities of an LLM.
+Augmenting the training of summary embeddings by RL-HF, in order to regularize the fit of these embeddings by ensuring that they capture signal rather than noise, is another research avenue to purse in the future.
 
@@
 
+In this research, we propose a novel, optimal, type of document summary, which maximizes the likelihood of an LLM to generate the document that it summarizes.
+These document summaries are the only ones, which are numeric and can be used for generation of new text.
+In an application to synthetic advertising claims, we show that these document summaries capture relevant information about the documents, but need a form of regularization to prevent overfitting.
+We also propose a form of factor model to estimate these document summaries, which yields an interpretable factor space.
+In an application to market research data, we show that this factor space captures linguistic features and can provide insights on consumer ratings of uniqueness and overall appeal, as well as help with the discovery of design principles.
+We observe that linguistic uniqueness is not the same as perceived uniqueness by consumers, and that claims of similar appeal ratings cluster together.
+However, we also observe that small adjustments to the wording of claims can lead to relatively large differences in rating.
+From this we propose that market research on advertising claims, should consist of two stages, a first stage where the researchers explore the space of claims widely, trying to identify a type of claim with a high appeal ratings.
+In a second stage, the researchers should then zoom-in on the sub-region and research the effects of fine-adjustments to the wording of claims in this region.
+Open issues of our approach are to solve the problem of regularization in a more general way, to improve the quality of newly generated text, and to scale this approach to a larger corpus of documents.
+
 {{< pagebreak >}}
 
-# References
+# Appendix {.appendix}
+
+::: {#fig-correlation-surface layout-ncol=3}
+![RE](figures/correlation_re_surface.pdf)
+
+![W2V](figures/correlation_w2v_surface.pdf)
+
+![BERT](figures/correlation_bert_surface.pdf)
+
+Correlation matrices for the surface cleaner claims.
+:::
+
+::: {#fig-correlation-surfacescramble layout-ncol=3}
+![Scrambled order](figures/correlation_re_surface_scramble.pdf)
+
+![W2V](figures/correlation_w2v_surface_scramble.pdf)
+
+![BERT](figures/correlation_bert_surface_scramble.pdf)
+
+Correlation matrices for the scrambled surface cleaner claims.
+:::
+
+::: {#fig-correlation-haircolor layout-ncol=3}
+![RE](figures/correlation_re_color.pdf)
+
+![W2V](figures/correlation_w2v_color.pdf)
+
+![BERT](figures/correlation_bert_color.pdf)
+
+Correlation matrices for the hair claims with color attribute.
+:::
+
+::: {#fig-correlation-health layout-ncol=3}
+![RE](figures/correlation_re_health.pdf)
+
+![W2V](figures/correlation_w2v_health.pdf)
+
+![BERT](figures/correlation_bert_health.pdf)
+
+Correlation matrices for the hair color with health attribute.
+:::
+
+{{< pagebreak >}}
+
+# References {.references}

### 2024-06-21 writing/factor_tikz.tex
pre-defense sync

@@
     \tikzstyle{LLM}=[draw,thick,minimum width=2cm, minimum height=2cm]
+    \tikzstyle{Norm}=[draw,thick,minimum width=2cm, minimum height=2cm]
     \tikzstyle{NLL}=[draw,thick,minimum width=1cm, minimum height=1cm]
@@
 
-    \node[input neuron] (I-1) at (0, -1) {$i_1$};
-    \node[input neuron] (I-2) at (0, -2) {$i_2$};
+    \node[input neuron] (I-1) at (0, -1) {$0$};
+    \node[input neuron] (I-2) at (0, -2) {$1$};
     \node[input neuron] (I-3) at (0, -3) {$\vdots$};
-    \node[input neuron] (I-4) at (0, -4) {$i_I$};
-    \node[bias neuron] (I-5) at (0, -5) {$B^{E}$};
+    \node[input neuron] (I-4) at (0, -4) {$0$};
+    \node[bias neuron] (I-5) at (0, -5) {$B^{Encoder}$};
 
-    \node[hidden neuron] (H-1) at (\layersep, -1.5) {$h_1$};
+    \node[hidden neuron] (H-1) at (\layersep, -1.5) {$f_{2, 1}$};
     \node[bias neuron] (H-2) at (\layersep, -2.5) {$\vdots$};
-    \node[hidden neuron] (H-3) at (\layersep, -3.5) {$h_H$};
-    \node[bias neuron] (H-4) at   (\layersep, -4.5) {$B^{D}$};
+    \node[hidden neuron] (H-3) at (\layersep, -3.5) {$f_{2, H}$};
+    \node[bias neuron] (H-4) at   (\layersep, -4.5) {$B^{Decoder}$};
 
-    \node[output neuron] (O-1) at (\layersep * 2, -0.5) {$e_1$};
-    \node[output neuron] (O-2) at (\layersep * 2, -1.5) {$e_2$};
+    \node[output neuron] (O-1) at (\layersep * 2, -0.5) {$s_{1, 2}$};
+    \node[output neuron] (O-2) at (\layersep * 2, -1.5) {$s_{2, 2}$};
     \node[bias neuron] (O-3) at (\layersep * 2, -2.5) {$\vdots$};
-    \node[output neuron] (O-4) at (\layersep * 2, -3.5) {$e_{E-1}$};
-    \node[output neuron] (O-5) at (\layersep * 2, -4.5) {$e_E$};
+    \node[output neuron] (O-4) at (\layersep * 2, -3.5) {$s_{E-1, 2}$};
+    \node[output neuron] (O-5) at (\layersep * 2, -4.5) {$s_{E, 2}$};
 
-    \node[LLM] (M-1) at (\layersep * 3, -2.5) {LLM};
+    \node[Norm] (Norm-1) at (\layersep * 3, -2.5) {Layer-Norm.};
 
-    \node[NLL] (N-1) at (\layersep * 4, -2.5) {$- log \mathcal{L}$};
+    \node[LLM] (M-1) at (\layersep * 4, -2.5) {LLM};
+
+    \node[NLL] (N-1) at (\layersep * 5, -2.5) {$\operatorname{\mathcal{L}}_{2}$};
 
@@
     \foreach \source in {1,...,5}
-        \path (O-\source) edge (M-1);
+        \path (O-\source) edge (Norm-1);
 
-    \draw (I-1) -- (H-1) node [midway, fill=white, minimum size=2pt] {$W_{1,1}^{E}$};
-    \draw (H-1) -- (O-1) node [midway, fill=white, minimum size=2pt] {$W_{1,1}^{D}$};
+    \draw (I-1) -- (H-1) node [midway, fill=white, minimum size=2pt] {$W_{1,1}^{Encoder}$};
+    \draw (H-1) -- (O-1) node [midway, fill=white, minimum size=2pt] {$W_{1,1}^{Decoder}$};
 
+    \path (Norm-1) edge (M-1);
     \path (M-1) edge (N-1);
@@
     % Annotate the layers
-    \node[annot,above of=H-1, node distance=2cm] (hl) {Hidden ($H$)};
+    \node[annot,above of=H-1, node distance=2cm] (hl) {Factors ($F$)};
     \node[annot,right of=hl] {Summary-Embedding ($E$)};

### 2024-06-21 writing/paper.qmd
pre-defense sync

--- 2024-06-14 writing/paper.qmd
@@
 author: Finn-Ole Höner (657110)
-title: "working title"
+title: "Novel Generative Summaries: Safer Prompts & More insightful Market Research?"
 date: last-modified
@@
 cache: false
+links-as-notes: true
 ---
@@
 
-> *"Prompt engineering is the art of communicating eloquently to an AI."----Greg Brockman, President OpenAI*
+Many marketing problems in NLP require an accurate representation of the information contained in a text document.
+For example, we might want to analyze the design elements of advertising claims and predict their perception by consumers, or analyze the sentiment of a product review [@bergerUnitingTribesUsing2020a].
+A powerful advance in NLP, are Large Language Models (LLM), such as GPT [@radford2018improving] and the demand of businesses for such tools is [growing rapidly](https://www.statista.com/outlook/tmo/artificial-intelligence/natural-language-processing/worldwide#market-size).
+In marketing, summarizing texts with these LLMs is becoming an important application [@hartmannNaturalLanguageProcessing2023].
+However, using an LLM to summarize a document is difficult: The generated summaries depend on the wording of the used prompt and can be unreliable.
+Such textual summaries are also difficult to evaluate and are not practical to use in prevalent data science applications, such as prediction or clustering.
+In this work, we show how we can summarize documents with LLMs for this purpose.
 
@@
 <!-- TODO: Make this more about a summary, than about generation and prompts? Build the bridge of the two -->
-<!-- TODO: Talk about low-dim representation -->
-
-Generative Large Language Models (LLM) are powerful, but unreliable tools for generating text.
-[A study by the NGO "Algorithm Watch"](https://algorithmwatch.ch/en/wp-content/uploads/2023/12/AlgorithmWatch_Generative-AI-and-elections_An-analysis-of-Microsofts-Bing-Chat.pdf), which is partially funded by the European Union, finds that Microsoft's chatbot Bing Chat, delivers wrong information in a third of its replies to election related questions, posing a risk to exacerbate desinformation.
-
-Prompt engineering defines the task of finding the right prompt to generate a desired output.
-Typically, this is an interactive and iterative process between human and LLM.
-There are certain writing techniques that improve the quality of
-the LLM's response, such as providing examples for what the desired output should look like.
+<!-- TODO: Talk about low-dim representation
+
+A recent example of how prompts can lead to undesired behavior of the LLM is [a study by the European Union funded NGO "Algorithm Watch"](https://algorithmwatch.ch/en/wp-content/uploads/2023/12/AlgorithmWatch_Generative-AI-and-elections_An-analysis-of-Microsofts-Bing-Chat.pdf), which finds that Microsoft's chatbot Bing Chat delivers wrong information in a third of its replies to election related questions, exacerbating desinformation.
+-->
+
+A core distinction of our approach is, that we use an LLM to summarize a document without relying on prompt engineering.
+Prompt engineering defines the task of finding the right prompt to generate a desired output with a LLM.
+Typically, this is an interactive and iterative process between humans and LLM.
+Certain writing techniques improve the quality of the LLM's response, such as providing examples of what the desired output should look like.
 There are various [blog posts](https://huggingface.co/docs/transformers/en/tasks/prompting#best-practices-of-llm-prompting) and books [e.g.
-@phoenix2024prompt] about how to write good prompts, in fact businesses are [hiring "Prompt Engineers"](https://indeed.com/q-prompt-engineer-jobs.html?vjk=98ac28acfd1328d7) even on dedicated [job boards](https://prompt-engineering-jobs.com/jobs/prompt-engineer-qwvmju/).
+@phoenix2024prompt] about how to write good prompts.
+Businesses are [hiring "Prompt Engineers"](https://indeed.com/q-prompt-engineer-jobs.html?vjk=98ac28acfd1328d7) on dedicated [job boards](https://prompt-engineering-jobs.com/jobs/prompt-engineer-qwvmju/), indicating a demand for operating LLMs in a better way.
 Despite these, supposedly low-barrier, resources, end users without AI knowledge still struggle with prompt engineering [@zamfirescu2023johnny].
-Not only can prompt engineering pose a [security risk](https://www.promptingguide.ai/risks/adversarial), but the process is also fuzzy, hard to replicate, and might still lead to unpredictable behaviors of the model.
-These issues pose risks for organizations, as the obscurity of manual prompt engineering can lead to loss of know-how, and the unpredictiability of the model can lead to lawsuits, damages to brands, or missed opportunities.
-
-Problems of reproduceability, transparency, and documentation, become critical for organizations with the advent of the [EU AI Act](https://digital-strategy.ec.europa.eu/en/policies/regulatory-framework-ai).
-The act requires for "High risk applications", that the AI model features, e.g.
-"adequate mitigation systems", "traceability of results", and, "robustness".
-High risk applications are not common and important for society, they include applications e.g.
-in education, employment, and public services.
-
-We propose a new type of numeric document summary, which captures all the information contained in the document, as it can be used to recreate the document itself.
-This document summary comes in the form of a numeric vector, which we call the summary embedding.
-To obtain this summary embedding, we maximize the likelihood of generating the document with a Large Language Model (LLM).
-In this sense, these summary embeddings are optimal.
-We show, that these embeddings capture inherent information about the document, can be used for classification, and generation of new documents.
-There are only practical requirements for estimating these summary embeddings, namely that backpropagation of gradients is possible from the likelihood to the input of the LLM.
-
-Beyond the methodological contribution, we show a marketing application of these summary embeddings and make a step towards using Generative AI in market research and design of the marketing-mix.
-We find that we can use these embeddings to capture design motivations of marketers behind advertising claims, and establish a measure for linguistic uniqueness of an advertising claim.
-Suprisingly, we find that this linguistic uniqueness is negatively correlated with the perceived uniqueness of an advertising claim by consumers.
-We explore these linguistic features and find that claims with certain words and their synonyms, cluster together, e.g.
-claims that emphasize the taste of a yoghurt drink or present it as a breakfast.
-Furthermore, we show that claims that live in certain areas of a low dimensional representation of the summary embeddings, are more likely to be rated favorably by consumers, but that small differences in wording can lead to relatively large differences in evaluation.
+Not only can prompt engineering to pose a [security risk](https://www.promptingguide.ai/risks/adversarial), but the process is also fuzzy, hard to replicate, and might still lead to unpredictable behaviors of the LLM.
+These issues pose risks for organizations, as the obscurity of manual prompt engineering can lead to loss of know-how, and the unpredictability of the model can lead to lawsuits, damages to brands, and missed opportunities.
+Problems of reproducibility, transparency, and documentation become critical for organizations with the advent of the [EU AI Act](https://digital-strategy.ec.europa.eu/en/policies/regulatory-framework-ai).
+The act demands that the AI model features  "adequate mitigation systems", "traceability of results", and "robustness" when used for "High-risk applications".
+High-risk applications are common and important for society, as they include applications in education, employment, and public services.
+Hence, businesses and society alike have an interest in improving prompt engineering, in a way that makes the responses of LLMs better in quality and more robust.
+
+To make prompt engineering more safe, we need to make an LLM's output more reliable.
+In this study, we make a step towards more reliable text generation by proposing a novel type of document summary, which we call the "generative summary".
+The core idea of our method is to find an input to the LLM that captures all the information contained in a focal document, as it leads the LLM to recreate this document itself.
+To obtain this summary, we maximize the likelihood of generating the document with a Large Language Model (LLM) given a numeric vector.
+Hence, these generative summaries are optimal.
+To mitigate overfitting and to make these embeddings more interpretable, we also introduce a factor model to estimate these document summaries in a lower dimensional space.
+We show that these embeddings capture inherent information about the document, and can be used for the generation of new documents.
+We also find that we can form linear combinations of generative summaries that have meaning themselves, similar to the approach by @mikolovDistributedRepresentationsWords2013 with word embeddings.
+
+Furthermore, we illustrate the use of our generative summaries in a marketing setting and show how they can yield insights in market research.
+We apply our method to two datasets.
+First, an artificial dataset of advertising claims for hair shampoo and surface cleaners to validate that these generative summaries capture relevant information.
+Second, we show an application of these generative summaries in market research.
+We use these generative summaries to analyze market research data on product claims for a yogurt and a yogurt drink.
+We find that we can use these embeddings to capture the design elements of these claims, assess their linguistic uniqueness, and represent their design patterns.
+Surprisingly, we find that this linguistic uniqueness is negatively correlated with the perceived uniqueness of an advertising claim by consumers.
+We explore these linguistic features and find that claims with certain words and their synonyms cluster together, e.g.
+claims that emphasize the taste of a yogurt drink or present it as a breakfast.
+Furthermore, we show that claims that live in certain areas of a low dimensional representation of these summaries are more likely to be rated favorably by consumers, but that small differences in wording can lead to relatively large differences in evaluation.
 This suggests a two-step approach to designing a market research study for advertising claims, where the first stage should focus on a wide exploration of the design space, while the second stage should focus on a fine-grained evaluation of the most promising area in this design space.
-We end with a sketch on how these summary embeddings can be used in an AI augmented design process, similar to applications in image data by @burnap2023product and @ludwigMachineLearningTool2023.
-
-<!--
-
-- Consumer preferences in latent generation space & Ad-claims
-  - Market research is expensive. How can we leverage these insights better to improve its benefits?
-  - What is missing from preference learning? How is this different?
-    - Media products/things are different from products as attribute bundles
-      - We can directly represent the product itself?
-    - Brands and their role in advertising
-  - We can view products as attribute bundles and get utilities for these attributes 
-  - How do we do this for products that cannot be "discretized"?
-  - Generative AI for personalization
-    - Previously: Set pieces, e.g. recommendation systems
-    - True personalization: Generate solution for each individual consumer
-  - TODO: Add more product claim specific
-  - Advertisements can be effective, @daiFrontiersWhichFirms2023a observe an increase in purchase intentions for restaurants that adapt online advertising by 7%-19%.
-  - they are relevant @kongMadeUSAClaims2021
-  - costs
-  - need flexibility in online settings
-  - closely related problems, such as email targeting @sahniPersonalizationEmailMarketing2018
-  - advertising is one of the marketing-mix instruments which companies use to position themselves in the market and shapes the incentives of the system itself @moorthyTargetingAdvertisingSpending2023 @renAdvertisingContentCreation2023
-  - design question
-  - Ban of cookie-based advertising due to GDPR, adapting to observable behavior in live sessions or to context on the website itself becomes more important. GDPR
-  - Challenges to cookies in website morphing
-  - @beckerDoesItPay2019, point to the nuanced role that authenticity plays in advertising, especially with respect to its alignment with the brand and product. 
-  - @edelmanBrandingDigitalAge2010 Concept of the consumer decision journey, product claims could vary across multiple visits or encounters of the same customer
-
--->
+We end with a sketch of how managers can use these generative summaries to augment the design process and market research on product claims.
+
+We proceed with a brief overview of the relevant literature and introduce our methodology.
+After testing our procedure on synthetic data, and exploring the low-dimensional factor space for the generation of new advertising claims, we apply our model to the market research data.
+We conclude with managerial implications, a discussion of the limitations of our approach and expansions for this research.
 
@@
 
-In the following, we compare our summary embeddings with three established methods to summarize documents: The Bidirectional Encoder Representations from Transformers (BERT) model's `[CLS]` token [@devlinBERTPretrainingDeep2018], pooled word embeddings [@shenBaselineNeedsMore2018], and prompt engineering approaches [@@huangenhanced].
-
-The BERT model is a transformer [see @vaswani2017attention] based model to represent text.
+<!-- TODO: Merge these paragraphs -->
+
+In Natural Language Processing, words, symbols, and even syllables are represented by so-called tokens, integer codes representing a sequence of characters.
+A piece of text, represented by tokens, is what we call a document.
+To represent the meaning of words and phrases, rather than to just encode them by a token, we can use word embeddings.
+Word embeddings are numeric vectors that represent text.
+They are pre-trained on large amounts of text data and can capture the meaning of single words and phrases.
+Early approaches include the Word2Vec model by @mikolovDistributedRepresentationsWords2013.
+The researchers train these word embeddings through a model that predicts the text surrounding a focal word.
+Thereby they learn the context in which words occur, which is a way of representing their meaning.
+There are also approaches that create embeddings for sequences of words or even documents [e.g.
+@devlinBERTPretrainingDeep2018].
+In this work, we use Large Language Models (LLMs).
+LLMs, such as [ChatGPT](https://chat.openai.com/) [@radford2018improving], predict the next word in a sequence of words.
+These models generate text based on textual inputs ("prompts").
+Internally, the model translates these prompts into a sequence of vectors, the "input embedding".
+This input embedding is then passed through a neural network that predicts the next token in the sequence.
+To generate text, the model predicts the next token and attaches it to the prompt, repeating this process until it predicts the sequence to end (i.e.
+predicting the next token to be the "End-of-Sequence" token: "eos-token").
+LLMs are pre-trained on large text corpora, such as the [Common Crawl](https://commoncrawl.org/) which contains over 250 billion pages of text.
+Their architecture is rooted in the attention mechanism.
+The attention mechanism models interactions in the text, such as negations, synonyms, or grammatical structures, as well as topics, across a long sequence of text.
+@vaswani2017attention implements this mechanism in the transformer block, which is a modular building block for the LLM.
+The combination of these transformer blocks, giant pre-training data, and huge computing resources lead to LLMs that have "emergent capabilities".
+These are capabilities, that were not explicitly trained for by the developers of the model, such as the ability of a model to translate between languages, pass the BAR exam, or write textual summaries of documents [see @weiEmergentAbilitiesLarge; @luAreEmergentAbilities2023].
+
+In the following, we compare our *generative summaries* (*GS*) with three established methods to summarize documents:
+
+- The Bidirectional Encoder Representations from Transformers (*BERT*) model's classifier token (CLS) [@devlinBERTPretrainingDeep2018]
+- Pooled word embeddings (*PWE*)  [@shenBaselineNeedsMore2018] 
+- Prompt engineering approaches (*PE*) [@huangenhanced; @khattabDSPyCompilingDeclarative2023]
+
+<!-- TODO: Start with word embeddings themselves 
+Word embeddings are numeric vectors that represent text.
+-->
+
+__BERT__: An advanced approach to estimate a word embedding is the *BERT* model, which is a transformer-based model to represent text [see @vaswani2017attention].
 Its developers have pre-trained BERT on a large corpus of text, using two different self-supervised learning objectives.
-The first objective is the so called "Masked Language Model" task.
-Here, the researchers replace a random selection of tokens in the text by a mask token, and the goal of the BERT model is to predict which token is underlying this mask.
-In order to learn about the information captures in a sentence, the researchers also pose the "Next Sentence Prediction" task to the model.
+The first objective is the so-called "Masked Language Model" task.
+Here, the researchers hide a random token in a sequence, and the goal of the BERT model is to predict which token is missing.
+This teaches the model the meaning of words in context.
+To learn about the information captured in a sentence, the researchers also pose the "Next Sentence Prediction" task to the model.
 For this training objective, the BERT model receives a pair of two sentences and needs to predict whether the second sentence follows the first one in a text.
-This task trains the `[CLS]` token: As it is pre-pended before every such sentence pair, it learns a representation of the information contained in the focal sentence.
-In practice, the output for a `[CLS]` token serves as a powerful feature e.g.
+This task trains the classifier token (CLS): As it is pre-pended before every such sentence pair, it learns a representation of the information contained in the focal sentence.
+In practice, the output for a CLS token serves as a powerful feature e.g.
 for sentence classification.
 BERT embeddings are versatile, but can also be fine-tuned to a specific task such as text summarization.
-However, such a fine-tuning requires the user to compile a dataset of text-summary and text pairs, that are representative of their application.
+Such a fine-tuning requires the user to compile a dataset of summary-text pairs, that are representative of their application.
 These BERT representations are only descriptive and cannot be used for the generation of new text [see @devlinBERTPretrainingDeep2018].
-There exist expansions to BERT, such as RoBERTa by @yinhanliuRoBERTaRobustlyOptimized2019, which improve the optimization procedure.
-
-@shenBaselineNeedsMore2018 points to the value of using pooled word embeddings to represent documents.
-The idea behind a pooled word embedding is, that we can represent a sequence of tokens by aggregating the word-embeddings of these tokens.
-There are different methods we can use for this aggregation, some examples are the use of taking the average across the tokens (mean-pooling) or taking the maximum value across the tokens (max-pooling).
-There are various types of word embeddings that we could use for the pooling, often, these are non-attention word embeddings, such as Word2Vec [@tomasmikolovDistributedRepresentationsWords2013; @tomasmikolovEfficientEstimationWord2013] or Global Vectors (GLoVe) [@jeffreypenningtonGloveGlobalVectors2014].
-These pooled word embeddings, cannot be used for text generation and perform worse than BERT embeddings in language tasks, as they do not employ the attention mechanism and loose information in the aggregation step [see e.g.
+In contrast to generative LLMs, such as GPT [@radford2018improving], which consider context only unidirectional, BERT takes text from the left and the right into account when representing it as an embedding.
+There exist improvements to BERT, such as RoBERTa by @yinhanliuRoBERTaRobustlyOptimized2019, which improve the optimization procedure.
+
+__PWE__: @shenBaselineNeedsMore2018 argues that pooled word embeddings (*PWE*) are a simple, yet powerful way to represent documents.
+The idea behind a *PWE* is that we can represent a sequence of tokens by aggregating the word embeddings of these tokens.
+There are different methods we can use for this aggregation, such as taking the average across the tokens (mean-pooling) or taking the maximum value across the tokens (max-pooling).
+There are various types of word embeddings that we could use for the pooling, such as Word2Vec [@mikolovDistributedRepresentationsWords2013; @mikolovEfficientEstimationWord2013] or Global Vectors (GLoVe) [@jeffreypenningtonGloveGlobalVectors2014].
+These pooled word embeddings, cannot be used for text generation and perform worse than BERT embeddings in language tasks, as they do not employ the attention mechanism and lose information in the aggregation step [see e.g.
 @onanHierarchicalGraphbasedText2023].
 
-We can also summarize documents by passing these to an LLM and prompting the model to write a summary for us [e.g.
+__PE__: We can also summarize documents by passing these to an LLM and prompting the model to write a summary for us (Prompt engineering approaches, *PE*) [e.g.
 @huangenhanced; @chakrabortyAbstractiveSummarizationEvaluation2024].
 However, these summaries are textual, not deterministic, and their quality depends on the formulation of the prompt.
-For example, @liuLostMiddleHow2023 find that answers of LLMs are better, when we place relevant information at the beginning or end of the prompt, while inserting emotions into prompts also improves the quality of the LLMs response, in some cases even doubling its performance [see @liLargeLanguageModels2023].
+For example, LLMs deliver better answers when the prompt contains relevant information at the beginning or the end, or when the prompt is written emotionally [see @liuLostMiddleHow2023; @liLargeLanguageModels2023].
 Another issue is that these summaries can vary in length, and one needs to specify how detailed a summary should be.
@@
 represent the information in the focal document well, then they can be useful when we want to share a written summary of a document.
-However, for many data science and automation tasks, we need a reliable numeric representation of the document, which is the focus of this paper.
-
-In this work, we propose a method that optimizes the input prompt to an LLM, such that we obtain a desired outcome.
-We call these input prompts "document summaries", as the perfectly replicate a focal document.
-These input embeddings capture all information that is contained in a document, because we can re-generate the focal document just based on this summary embedding.
-These document summaries are deterministic, because we obtain them through maximum likelihood estimation and live in continuous space.
-We can evaluate the fit of such a document summary, as the likelihood indicates its fit directly.
+In this paper, we propose a solution that works for data science and automation tasks, as we create a numeric representation of the document.
+We could use these textual summaries for generation, in the sense that we could prompt the model to generate a new document based on the summary.
+
+A recent improvement to *PE* approaches is the framework proposed by @khattabDSPyCompilingDeclarative2023.
+This framework, called "Declarative Self-improving Language Programs, pythonically" (DSPy), learns how to combine different prompting and finetuning techniques to improve the generated answer.
+They show that their discrete optimization leads to better output compared to expert-designed prompts, even when applied to smaller LLMs.
+Their approach is text-based and revolves around generating examples of the desired output and then tuning the prompt and the LLM itself based on these examples.
+In spirit, their approach is similar to hyper-parameter tuning, in that it uses a type of (small) training data and a performance metric (e.g.
+whether the answer is an exact match to a certain label), and then adjusts the prompt and the LLM to maximize the scoring of its answer on this training set.
+The twist is, that the DSPy framework can perform such hyper-parameter tuning in an automized fashion, by generating candidate prompts itself and selecting from them.
+Our proposed approach differs from DSPy in that we optimize in a continuous space, and that our goal is to find a numeric summary for a document, rather than to generate a good response to language tasks, such as question answering.
+
+__GS__: In this work, we propose *generative summaries* (GS) which optimize the input prompt to an LLM, such that we obtain a desired outcome.
+These *generative summaries* should perfectly replicate the document they summarize, thereby capturing all information contained in it.
+These document summaries are deterministic because we obtain them through maximum likelihood estimation and they live in continuous numeric space.
+We can evaluate the quality of such a document summary directly through the obtained likelihood.
 
@@
 | Method                              | Origin                                          | Reference                              |  Deterministic? |  Generation?  |  Type?  |
-|:------------------------------------|:------------------------------------|:--------------------|:----------------------:|:----------------------:|:----------------------:|
-| BERT `[CLS]` token                       | Next-sentence prediction task                  | @devlinBERTPretrainingDeep2018         | $\checkmark$   | $\times$       | Numeric    |
-| Pooled Word Embeddings              | Aggregation of token information               | @shenBaselineNeedsMore2018             | $\checkmark$   | $\times$       | Numeric    |
-| Prompt engineering                  | Emergent capability of LLM                     | CITE                                   | $\times$       | $\times$       | Textual    |
-| Reverse engineered document summaries | Maximize the likelihood to re-generate the focal document | *Proposed method*                        | $\checkmark$   | $\checkmark$   | Numeric |
-: Overview of existing methods to summarize documents. {#tbl-docsum}
+|:---------|:------------------------------------|:--------------------|:---------------:|:---------------:|:--------------------:|
+| *BERT*                       | Next-sentence prediction task                  | \footnotesize @devlinBERTPretrainingDeep2018         | $\checkmark$   | $\times$       | Numeric    |
+| *PWE*              | Aggregation of token information               | \footnotesize @shenBaselineNeedsMore2018             | $\checkmark$   | $\times$       | Numeric    |
+| *PE*                  | Emergent capability of LLM                     | \footnotesize @khattabDSPyCompilingDeclarative2023                                   | $\times$       | $\checkmark$       | Textual    |
+|                   |                      |  |        |        |     |
+|                   |                      |  |        |        |     |
+| __*GS*__ | Maximize the likelihood to regenerate the focal document | __*Proposed method*__                        | $\checkmark$   | $\checkmark$   | Numeric |
+: Overview of methods to summarize documents. {#tbl-docsum}
+
+To validate whether our proposed approach is a viable alternative to these established methods, we will evaluate three aspects of our generative summaries on an artificial dataset, and explore their use in marketing applications.
+This yields these three *research objectives* for the validation and one explorative objective for the marketing application:
+
+1) Can we use these generative summaries for classification tasks?
+2) Can we regenerate the summarized document with our generative summaries?
+3) Can we generate new documents based on our generative summaries?
+4) *Explorative*: Can these generative summaries deliver useful insights in market research?
 
@@
 The more beams we search, the more likely we are to find a sequence with a higher likelihood, then we would have found otherwise [@huggingfaceGenerateText].
-
-Each document, in our case an advertising claim, $d$ consists of the tokens $t_1^{d}, \ldots, t_T^(d)$.
-
-The probability to generate the sequence $t_1, \ldots, t_T$ with the LLM when using the embedding $\boldsymbol{s}$ as an input is $p(t_1, \ldots, t_T \mid \boldsymbol{s})$.
+  -->
+
+<!-- Each document, in our case an advertising claim, $d$ consists of the tokens $t_1^{d}, \ldots, t_T^{(d)}$. -->
+
+The probability of generating the sequence $t_1, \ldots, t_T$ with the LLM when using the embedding $\boldsymbol{s}$ as an input is $p(t_1, \ldots, t_T \mid \boldsymbol{s})$.
 We can split this probability into conditional probabilities due to the autoregressive architecture of the LLM and because we observe the target sequence.
 This provides large computational gains, as we can calculate these parts in parallel.
-Hence, we define the log-likelihood of the summary embedding for the target sequence as
+Hence, we define the log-likelihood of the generative summary for the target sequence as
 $$
@@
 
-and find the optimal summary embedding as
+and find the optimal generative summary as
 
@@
 \mathbf{s}^{*} = \argmax_{\boldsymbol{s}} \operatorname{\mathcal{L}} \left ( \boldsymbol{s} \mid t_1, \ldots, t_T \right) \text{. }
-$$
-
-We summarize the training process for the single summary embeddings in Algorithm 1.
-The summary embedding is a vector of length $E$, which is the same dimension as the dimension of the LLM's input embedding.
-
-\begin{algorithm}
+$${#eq-single}
+
+We summarize the training process for these single generative summaries in Algorithm 1. The generative summary is a vector of length $E$, which is the same dimension as the LLM's input embedding.
+
+\begin{algorithm}[H]
 \begin{algorithmic}
@@
     \State $\nabla_{\boldsymbol{s}}^{(i)} l \gets \operatorname{ComputeGradient} \left ( l^{(i)}  \right )$
-    \State $\boldsymbol{s}^{(i + 1)} l \gets \operatorname{Optimizer} \left(\boldsymbol{s}^{(i)}, \; \nabla_{\boldsymbol{s}}^{(i)} l \right)$
+    \State $\boldsymbol{s}^{(i + 1)} \gets \operatorname{Optimizer} \left(\boldsymbol{s}^{(i)}, \; \nabla_{\boldsymbol{s}}^{(i)} l \right)$
     \State $i \gets i + 1$    
@@
 
-As we are interested in finding a low-dimensional sub-space, in which we can represent the $D$ advertising claims, we use a factor model and estimate the summary embeddings for all advertising claims jointly, yielding the $D \times E$ matrix of summary embeddings $\mathbf{S}$.
+To regularize these embeddings and make them more interpretable, we are interested in finding a low-dimensional sub-space in which we can represent the $D$ advertising claims.
+For this, we use a factor model and estimate the generative summaries for all advertising claims jointly, yielding the $D \times E$ matrix of generative summaries $\mathbf{S}$.
 We restrict the degrees of freedom to $F$ hidden factors.
-To implement this factor model, we create an Autoencoder with one hidden layer, that is fully connected with linear activation functions [@goodfellowDeepLearning2016].
-The input data to this autoencoder is an identity matrix of dimension $D$.
+To implement this factor model, we create an autoencoder with one hidden layer that is fully connected with linear activation functions [@goodfellowDeepLearning2016].
+The input data to this autoencoder is an identity matrix of dimension $D$, making it a factor model.
+Before passing the generative summary to the LLM, we normalize them with layer normalization [@baLayerNormalization2016].
+Layer normalization works across the hidden nodes, rather than across the observations in a batch [batch-normalization, see @ioffeBatchNormalizationAccelerating2015].
+Thereby, it can also be used with a single observation.
+In the optimization, we encounter sharp spikes in our gradients, that originate in the decoder of the model.
+We mitigate these by gradient clipping [@zhangWhyGradientClipping2020].
 We illustrate this neural network representation of the factor model in @fig-factor.
@@
 
-We denote the matrix of weights of the encoder and decoder layers by $\mathbf{W}^{Encoder}$ and $\mathbf{W}^{Decoder}$.
-These matrices have the dimensions $D \times F$ and $F \times E$ respectively.
-We also define the vectors of biases for the encoder and decoder by $\boldsymbol{b}^{Encoder}$ (length $F$) and $\boldsymbol{b}^{Decoder}$ (length $E$), and stack them into matrices by taking the Kronecker product with a length $D$ vector of ones, $\boldsymbol{1}_D$, to obtain the biases in matrix form as $\mathbf{B}^{Encoder} = \boldsymbol{1}_D \otimes \boldsymbol{b}^{Encoder}$ and $\mathbf{B}^{Decoder} = \boldsymbol{1}_F \otimes \boldsymbol{b}^{Decoder}$.
-For notational convenience, we denote this factor model by $\operatorname{Factor}_{\Omega_{Factor}} \left(\mathbb{I}_{D} \right): \lbrace 0, 1 \rbrace^{D \times D} \to \mathbb{R}^E$, with parameters $\Omega_{Factor}$, where $\Omega_{Factor} = \lbrace \mathbf{W}^{Encoder}, \mathbf{W}^{Decoder}, \mathbf{B}^{Encoder}, \mathbf{B}^{Decoder} \rbrace$.
-
-<!-- The Large Language Model provides a log-likelihood function $\operatorname{\mathcal{L}} \left ( \mathbf{S} \right): \mathbb{R}^{D \times E} \to \mathbb{R}^{\left (- \infty, 0 \right]}$, with parameters $\Omega_{LLM}$, which returns the log-likelihood for the summary embedding to generate the target documents.
+We denote the matrix of encoder and decoder weights by $\mathbf{W}^{Encoder}$ and $\mathbf{W}^{Decoder}$.
+These matrices have the dimensions $D \times F$ and $F \times E$.
+We also define the encoder and decoder biases as $\boldsymbol{b}^{Encoder}$ (length $F$) and $\boldsymbol{b}^{Decoder}$ (length $E$) and stack them into the matrices $\mathbf{B}^{Encoder}$ and $\mathbf{B}^{Decoder}$^[We do this by taking the Kronecker product with a length $D$ vector of ones, $\boldsymbol{1}_D$, to obtain the biases in matrix form as $\mathbf{B}^{Encoder} = \boldsymbol{1}_D \otimes \boldsymbol{b}^{Encoder}$ and $\mathbf{B}^{Decoder} = \boldsymbol{1}_F \otimes \boldsymbol{b}^{Decoder}$.].
+We denote the resulting factor model by $\operatorname{Factor}_{\Omega_{Factor}} \left(\mathbb{I}_{D} \right): \mathbb{I}_{D} \to \mathbb{R}^{D \times E}$ and collect the weights and biases in $\Omega_{Factor}$.
+
+<!-- The Large Language Model provides a log-likelihood function $\operatorname{\mathcal{L}} \left ( \mathbf{S} \right): \mathbb{R}^{D \times E} \to \mathbb{R}^{\left (- \infty, 0 \right]}$, with parameters $\Omega_{LLM}$, which returns the log-likelihood for the generative summary to generate the target documents.
 -->
@@
 
-thereby the matrix of summary embeddings for all claims becomes
+thereby the matrix of generative summaries for all claims becomes
 
@@
 
-When we estimate the matrix of summary embeddings $\mathbf{S}$, we use the joint log-likelihood to re-generate all documents in the dataset, $\operatorname{\mathcal{L}}_D = \sum_{d=1}^D \operatorname{\mathcal{L}}_d$, where $\operatorname{\mathcal{L}}_d$ is the log-likelihood to re-generate document $d$ with summary embedding $\mathbf{S}$.
-We estimate the summary embeddings by maximizing this joint log-likelihood with respect to the parameters of the factor model:
-
-<!-- TODO: Can I split this into separate optimizations? Why / why not? -->
+When we estimate the matrix of generative summaries $\mathbf{S}$, we use the joint log-likelihood to re-generate all documents in the dataset, $\operatorname{\mathcal{L}}_D$.
+We estimate the generative summaries by maximizing this joint log-likelihood with respect to the parameters of the factor model:
+
+<!-- TODO: Is it S or S_{d, :}? Can I split this into separate optimizations? Why / why not? -->
 $$
  \mathbf{S}^{*} = \argmax_{\mathbf{W}^{Encoder}, \; \mathbf{W}^{Decoder}, \; \mathbf{B}^{Encoder}, \; \mathbf{B}^{Decoder}} \operatorname{\mathcal{L}}_D \left ( \mathbf{S} \right) \text{. }
-$$
+$${#eq-factor}
+
+<!-- TODO: Write out what happens in Layer Norm! -->
+<!-- TODO: Update the method section wrt to the factor form! -->
 
@@
 <!-- 
-We aim to find a summary embedding, $\boldsymbol{e^{*}}$, that maximizes the likelihood of generating a given target sequence, $\lbrace t_i \rbrace_{i=1}^{L}$, with a given LLM, $\mathcal{M}$.
+We aim to find a generative summary, $\boldsymbol{e^{*}}$, that maximizes the likelihood of generating a given target sequence, $\lbrace t_i \rbrace_{i=1}^{L}$, with a given LLM, $\mathcal{M}$.
 We use a gradient-based optimization algorithm to maximize this likelihood.
-We initialize the summary embedding as the element-wise average of the embedding of the target sequence.
+We initialize the generative summary as the element-wise average of the embedding of the target sequence.
 We then generate a sequence of length $L$, with the LLM and the current input embedding.
@@
 
-<!-- TODO: Update this -->
-
-\begin{algorithm}
+\begin{algorithm}[H]
 \begin{algorithmic}
@@
 
-<!-- TODO: What does it really return? -->
-<!-- TODO: Good idea or not? -->
-<!-- TODO: Figure of generation tree -->
+For this project, we obtained a [SURF NWO Small Compute Grant](https://www.surf.nl/en/small-compute-applications-nwo) and perform all computations on the [Snellius High-Performance Cluster](https://www.surf.nl/en/services/snellius-the-national-supercomputer).
+We train these generative summaries until we obtain a likelihood to generate the target claim of $0.99$.
+We use the Adam optimizer [@diederikp.kingmaAdamMethodStochastic2014] with a learning rate of $0.1$, no weight decay, and the values of $0.9$ and $0.999$ for the exponential decay rates of the first and second moment respectively.
+To initialize our generative summary and factor model, we use the standard initialization of [PyTorch](https://github.com/pytorch/pytorch).
+We train on a Nvidia A100-GPU, which takes about 30 seconds for 20 claims.
+The training of the factor representation with 2 factors and for the same claims is more computationally intense and takes about 60 minutes on the same computer.
+Here we train for a joint likelihood of $0.99$.
+We always verified that the generative summaries regenerate all focal claims correctly.
+
+<!-- TODO: Note on optimization and behavior of likelihood, spike in gradient -->
+<!-- TODO: Figure of generation tree? -->
+
+## Diagnostics
+
+As benchmarks for our model, we use the BERT and PWE methods.
+Each of these methods yields one embedding vector per document.
+For the PWE method, this embedding is of length $100$ and for the BERT method, it is of length $768$.
+We obtain the PWE embedding, by applying Word2Vec embedding to each token in the document and averaging out over the tokens [@mikolovDistributedRepresentationsWords2013; @shenBaselineNeedsMore2018].
+For the BERT method, we pass the document through the pre-trained BERT model and extract the prediction for the CLS token, which serves as the document summary.
+
+To analyze our results, we will use a suite of diagnostic tools, which we will explain here briefly.
+
+First, we form two correlation matrices.
+One between the claims $\mathbf{C}^{Claims}$ and one between the embedding dimensions $\mathbf{C}^{Embeddings}$.
+For the former, the correlation at element $i, j$ of the correlation matrix is the correlation of the i-th and j-th claim:
+
+$$
+\mathbf{C}_{i,j}^{Claims} = \frac{\sum_{k=1}^E \left ( \mathbf{S}_{i, k} - \frac{1}{E} \sum_{k=1}^E \mathbf{S}_{i, k} \right) \left ( \mathbf{S}_{j, k} - \frac{1}{E} \sum_{k=1}^E \mathbf{S}_{j, k} \right)}{\sqrt{\sum_{k=1}^E \left( \mathbf{S}_{i, k} - \frac{1}{E} \sum_{k=1}^E \mathbf{S}_{i, k} \right)^2 \sum_{k=1}^E \left( \mathbf{S}_{j, k} - \frac{1}{E} \sum_{k=1}^E \mathbf{S}_{j, k} \right)^2 } } \text{.
+}
+$$
+
+For the latter, the correlation at element $i, j$ of the correlation matrix is the correlation of the i-th and j-th embedding dimension:
+$$
+\mathbf{C}_{i,j}^{Embeddings} = \frac{\sum_{k=1}^D \left ( \mathbf{S}_{k, i} - \frac{1}{D} \sum_{k=1}^D \mathbf{S}_{k, i} \right) \left ( \mathbf{S}_{k, j} - \frac{1}{D} \sum_{k=1}^D \mathbf{S}_{k, j} \right)}{\sqrt{\sum_{k=1}^D \left( \mathbf{S}_{k, i} - \frac{1}{D} \sum_{k=1}^D \mathbf{S}_{k, i} \right)^2 \sum_{k=1}^D \left( \mathbf{S}_{k, j} - \frac{1}{D} \sum_{k=1}^D \mathbf{S}_{k, j} \right)^2 } } \text{.
+}
+$$
+
+Besides these correlation matrices, we also use Principal Component Analysis (PCA).
+For a data matrix $\mathbf{X}$, PCA extracts $k$ principal components by taking the eigenvectors that correspond to the $k$ largest eigenvalues of the matrix $\mathbf{X}^T \mathbf{X}$.
+These $k$ eigenvectors contain the coordinates of the data points in the $k$-dimensional space spanned by these principal components [@pearsonLIIILinesPlanes1901].
 
@@
 
-CITE: Prompt engineering
+As a first evaluation, we create a small synthetic dataset of 20 advertising claims for hair shampoo by querying [ChatGPT](https://chat.openai.com/) with the prompt in @fig-gpt.
+We want half of these claims to advertise tangible aspects of the product and the other half to advertise intangible aspects of the product.
+These advertising claims all focus on how "shiny" the product ones hair makes.
+As as robustness check, we repeated this task for different attributes (healthiness and colorfulness), and a different product category (surface cleaners).
+In total, we obtain 80 advertising claims, 20 for each of these settings.
+We show the shiny hair shampoo claims in @tbl-claims and moved the other analyses to the appendix.
 
@@
 
-: Tangible (T) and Intangible (I) advertising claims for hair shampoo products. {#tbl-claims}
-
-- labels, accuracy
-- clustering
-- generation of new claims
-  - exploring the space
-  - measure of fit (area)
-
-[ ] generation of new claims [@Goodfellow2016]
-
-![Exploration of Encoding Space](figures/exploration_encoding_space.pdf)
-
-![Exploration of the likelihoods](figures/likelihoods_generation.pdf)
+: Tangible (T) and Intangible (I) advertising claims for hair shampoo products. (Created by ChatGPT) {#tbl-claims}
+
+For our empirical application, we obtain two datasets from a market research company.
+The first one contains advertising claims for yogurt drinks and their evaluation by consumers.
+The second dataset contains advertising claims for yogurts and the design motivations behind these advertising claims, such as whether a claim talks about the natural ingredients of the yogurt or its local origin.
+These data stem from different variations of choice-based conjoint studies [@eggersChoiceBasedConjointAnalysis2022], which we cannot further disclose due to confidentiality agreements.
+The data includes choice experiments aggregated at the level of the advertising claims, so we do not observe individual responses.
+We focus on measures of a claim's uniqueness and appeal, as perceived by the consumers.
+The advertising claims were designed for markets in different countries and are (translated into) English.
+Each dataset contains only one brand.
+
+For the yogurt drinks, we have 60 aggregate claim evaluations, 20 for each country (Germany, UK, Spain).
+Many of the researched advertising claims occur in multiple of these separate studies.
+In total, we have 26 unique advertising claims for yogurt drinks.
+Most of these advertising claims get a uniqueness measure of around 5.8, with the whole distribution ranging from 4.8 to 7.4 on 10 point scale.
+When we look at the uniqueness measure concerning the country in which the study was conducted, we see that the variance is the smallest for the German market and the largest for the Spanish market.
+However, the means are not statistically different across the countries.
+The distribution of the rating measure is similar in shape, ranging from 43 to 80 on a 100 point scale.
+The rating and uniqueness are positively correlated (correlation of $0.8843$).
+The yogurt drink advertising claims are similar to these synthetic claims:
+
+- *"A burst of fresh flavor to energize your morning"*
+- *"Refresh your day with a lively new taste"*
+- *"Dynamic flavor for an invigorating start"*
+- *"Begin your day with a crisp and revitalizing taste"*
+- *"Revitalize your senses with a pure, fresh flavor"*
+
+For the second set of claims, the claims for yogurts, we have $215$ observations of $82$ unique advertising claims.
+The data stems from studies in the UK, France, Germany, Spain, and Sweden.
+This data also contains a design motivation for the advertising claims.
+In total, there are $44$ different classes for these design themes.
+We focus on the five largest classes: "Packaging", "Local & Responsible", "Sourcing", "Sustainability", and "Naturality", which account for 70% of the observations.
+Below, we show a short description of what these themes look like:
+
+- __Packaging__: The claim emphasizes e.g. that the packaging is recyclable or made from eco-friendly materials
+- __Local & Responsible__: The claim talks about the local origin of the yogurt or how the yogurt company supports local farmers
+- __Sourcing__: The claim emphasizes the origin of the ingredients
+- __Sustainability__: These claims state that the yogurt is e.g. climate-friendly
+- __Naturality__: These claims state that the yogurt is e.g. made from natural ingredients or has been processed as little as possible
+
+<!-- We also have information on the marketers motivation behind the design of an advertising claims, such as whether she designed the claims to pronounce the health benefits of a yogurt, or the taste yogurt, as well as the brand of the product.
+
+## MR 5
+
+- One brand, yogurt drink
+- MaxDiff experiment, ranking of choices
+- Relevance, fit with brand, rank
+- 32 unique claims, 3 countries: Germany, Spain, UK
+
+ -->
+
+<!-- TODO: Some graphs / tables for descriptives -->
+
+::: {#fig-eda-mr4-uni layout-ncol=2}
+![Distribution of the uniqueness measure across all countries](figures/uniqueness_mr4_eda.pdf){#fig-eda-mr4-uni-a}
+
+![Distribution of the uniqueness measure per country. The black line indicates the mean.](figures/uniqueness_box_mr4_eda.pdf){#fig-eda-mr4-uni-b}
+
+Distributions for the claim uniqueness measure across the whole sample and as a boxplot per country.
+:::
+
+::: {#fig-eda-mr4-rating layout-ncol=2}
+![Distribution of the uniqueness measure across all countries](figures/rating_mr4_eda.pdf)
+
+![Distribution of the appeal rating per country. The black line indicates the mean.](figures/rating_box_mr4_eda.pdf)
+
+Distributions for the appeal rating across the whole sample and as a boxplot per country.
+:::
+
+![Scatterplot of uniqueness and rating , colored by country.](figures/scatterplot_mr4_eda.pdf)
+
+<!-- TODO: Provide examples that are created with GPT, i.e. are fake -->
+
+<!-- ## MR 6
+MR7: F4224
+- 5 countries: Brazil, Mexico, Russia, France, Germany
+- 40 unique claims -->
+
+# Results
+
+<!-- - HairColor: 42 minutes
+- HairHealth: 14 minutes
+- Surface: 16 minutes
+- Surface Scramble: 9 minutes
+- For our factor model, we use two factors. The intuition for this is, that we want to distinguish two classes of tokens and want to be able to visualize our data in a two-dimensional space. 
+- 
+- -->
+
+<!-- TODO: discuss computation times -->
+<!-- TODO: Have model with e.g. 50 factors? Then for many claims? -->
+
+We start by applying our method to the synthetic dataset of hair shampoo claims, that differ by whether they are formulated in a tangible or intangible way.
+We estimate these generative summaries in 3 versions: The single estimations, which estimate the generative summaries for each claim separately and without the factor structure (compare @eq-single), and two-factor models with two factors each (compare @eq-factor).
+One of these uses the Layer Normalization, while the other one does not.
+We present these three versions, to illustrate the effects of the factor structure and the layer normalization.
+In the appendix, we show the same diagnostics for other datasets, such as advertising claims for hair shampoo that emphasize how healthy/colorful the hair becomes (instead of shinyness), as well as for a surface cleaner product.
+These results are robust to changing the order of the claims in the dataset.
+@fig-optimization shows the training process for the generative summaries estimated with the 2-factor model with layer normalization.
+The y-axis shows the negative log-likelihood (a value of 0.01 corresponds to a likelihood of $e^{-0.01} = 0.99$).
+Despite our efforts to mitigate spikes, we still see some smaller spikes in the tail of the optimization.
+
+![Training of embeddings](figures/dashboard.png){#fig-optimization}
+
+## Objective 1: Can we use these generative summaries for classification tasks?
+
+If our generative summaries capture relevant information in the advertising claims, then we can separate between the tangible and intangible advertising claims.
+To be left with only the difference due to being tangible or intangible we demean the embeddings across the embedding dimensions [compare @mikolovDistributedRepresentationsWords2013].
+We then calculate the correlation matrix across the claims and the embedding dimensions.
+When visualizing these correlations across the claims, we expect the upper-left square and the lower-right square of the matrix to contain positive correlations, since these are the correlations for the claims of the same class (i.e.
+tangible-tangible or intangible-intangible).
+On the contrary, the upper-right and the lower-left quartile should contain negative correlations, since these are correlations between the two classes (tangible-intangible).
+
+In @fig-correlation we present these correlation matrices for our GS method and the PWE and BERT approaches.
+These correlation matrices show how the embedding values of two claims vary with each other.
+Identifying these two classes works well for the generative summaries which we estimate with 2 factors and layer normalization.
+When we compare this version to the PWE and BERT approaches, we see this pattern for all methods, indicating that all four methods can capture this part of the information in the advertising claim.
+We observe, that the correlations are the largest (in absolute value) for the generative summaries and only a few correlations are close to zero.
+In the appendix, we show the claim correlation matrices for our other generated datasets.
+The results mirror the results for the hair shampoo claims that emphasize the shinyness of the hair, with one exception: The hair shampoo claims are about how colorful it makes the hair.
+Here, we find that 2 factors are not enough to separate between tangible and intangible claims.
+We can recover these two classes by using a 15-factor model instead.
+Here we also observe, that the correlations are then weaker, as these document embeddings have more degrees of freedom in how their values come to be.
+
+<!--
+HF seems to compute loss across all labels
+https://discuss.huggingface.co/t/source-and-target-vs-input-and-labels-for-causal-autoregressive-language-models/19915
+
+- TODO: Check whether we need [attention mask in generate](https://discuss.huggingface.co/t/batch-generation-with-gpt2/1517/2)
+
+- TODO: Analyze the attention given to the summary embedding?
+
+ -->
+
+::: {#fig-correlation layout="[[1, 1, 1], [1, 1]]"}
+
+![GS: Single estimation](figures/correlation_re_hair_single.pdf)
+
+![GS: 2 Factors](figures/correlation_re_hair20_nonorm.pdf)
+
+![GS: 2 Factors with LayerNorm.](figures/correlation_re.pdf)
+
+![PWE](figures/correlation_w2v.pdf)
+
+![BERT](figures/correlation_bert.pdf)
+
+Correlation matrices along the claims. 
+:::
+
+@fig-correlationts shows correlation matrices across the embedding dimensions and illustrates the effects of using a factor structure on the generative summaries.
+Each cell shows the correlation between two embedding dimensions calculated across the $20$ advertising claims.
+When comparing our generative summaries with the two benchmarks, we can see a stronger grid pattern with more correlations that are removed from zero.
+This stems from our factor structure, as this pattern is not present in the single estimation version.
+Making many of the embedding dimensions linearly dependent strengthens the correlations between them.
+In contrast, for the BERT embedding most dimensions are barely correlated with each other.
+The Word2Vec embedding, which is used in PWE, only has 100 dimensions, as opposed to the 768 dimensions of our generative generative summaries and the BERT embedding.
+Its correlations appear to be stronger but do not exhibit the same grid pattern as for the generative summaries.
+
+::: {#fig-correlationts layout="[[1, 1, 1], [1, 1]]"}
+
+![GS: Single estimation](figures/correlation_ts_re_hair_single.pdf)
+
+![GS: 2 Factors](figures/correlation_ts_re_hair20_nonorm.pdf)
+
+![GS: 2 Factors with LayerNorm.](figures/correlation_ts_re.pdf)
+
+![PWE](figures/correlation_ts_w2v.pdf)
+
+![BERT](figures/correlation_ts_bert.pdf)
+
+Correlation matrices along the embedding dimensions. 
+:::
+
+After establishing that our generative summaries capture similar information to established document summary techniques, we want to explore the low-dimensional factor space of our model.
+@fig-encoding shows the low dimensional sub-space in which we located the advertising claims, which we color in orange (tangible claims) and green (intangible claims).
+The assigned numbers correspond to the number of the claims in @tbl-claims.
+It appears that we can perfectly separate the two classes with the two hidden factors.
+Many of the points are close to zero, with a few of the claims being far removed from the other points, e.g.
+claims 1, 3, and 9 for the tangible, and claims 12, 16, and 19 for the intangible claims.
+When comparing distances in the encoding space, we can find some patterns that relate to the types of words that are used in the advertising claims.
+For example, the triangle of claims 0, 4 and 7 uses words related to vision ("...
+visible ...", "See ...", and "Visible ..."), and the two claims 2 and 13 both start with the word "Transform." The tangible claims far removed from the center, claims 1, 3, and 9, have in common that they talk about a form of "technology", whereas claims 3 and 9 both use the phrase "enhance [...] shine".
+For the intangible claims, the two claims 12 and 19 both start with the close words "Unleash" and "Unlock".
+However, there are also structures in this representation which are not intuitive.
+We would anticipate that claim 4 would have been close to these 3 tangible claims in the south-west of the plot, as it is about an "advanced shine-boosting formula", or the outlier claim 16 ("Indulge in the luxury of hair that shimmers with elegance"), would be part of the cluster at the center because it is similarly constructed as the 7 intangible claims there (pattern: *Feeling/Experience of the consumer*---"of hair that"---*feature of the hair*).
+
+::: {#fig-encoding layout-ncol=2}
+
+![All claims estimated  by the factor model](figures/encoding_space.pdf){#fig-encoding1}
+
+![Zoom-in on the center claims estimated by the factor model](figures/encoding_space_zoomin.pdf){#fig-encoding2}
+
+Two-dimensional factor space of the advertising claims. Numbers correspond to the numbers in @tbl-claims, orange points are tangible claims, and green points are intangible claims.
+:::
+
+## Objective 2: Can we use these generative summaries to re-generate the focal document?
+
+While our generative summaries appear to capture the information contained in a document, one of their main advantages is that we can use them for generation.
+In the following, we analyze the generation with the generative summary.
+If our generative summaries indeed maximize the likelihood of generating the target text, then the most likely token at each position in the generated sequence should be the corresponding token from the summarized document.
+We expect that at each generation step, the probability for the correct token approaches one, while the probability for all other tokens goes to zero.
+In @fig-like, we visualize the next token probabilities at each generation step for the correct token.
+We see that the probability for the correct token is close to one, with the first token in a sequence having the lowest probability of being generated correctly.
+This holds across all claims.
+An explanation for this is, that at the first position, the generation relies on the generative summary alone.
+At later stages the joint information of generative summary and previously generated tokens stabilizes the generation, driving the generation probabilities closer to 1.
+
+![Conditional generation probability for the most likely token at each step for the factor model. All tokens match the target claims.](figures/likelihoods_generation.pdf){#fig-like}
+
+Next, we explore also the generation probability for the 2nd and 3rd most likely token at each position, to get a better understanding of the distribution at each step.
+We present two examples.
+One for the generation of the claim with the lowest probability to correctly generate the first token, in @tbl-lowprob, and another example that illustrates how we benefit from the LLMs pre-training, in @tbl-similartoken.
+Both tables show these generation probabilities for the three most likely tokens at each generation step.
+@tbl-lowprob again shows how the critical step in the generation is the first token of the sequence, as the probability to generate the correct token in the second position jumps for $0.8876$ to $0.9925$ and never drops lower than $0.9875$ again.
+In @tbl-similartoken, we see that the most likely alternatives to the first token, "Experience", are "Form" and "See".
+We find similar examples for the second token (1st Choice: "50", 2nd Choice: "25", 3rd Choice: "20") and the fifth position (1st Choice: "visible", 2nd Choice: "protective", 3rd Choice: "shine").
+The fact that these tokens get the second and third-highest generation probabilities is likely due to the pre-training of the LLM, and reflects how these words tend to be used in similar situations, such as in shampoo product claims.
 
@@
 | 10 | ).    | .)    | .             | 0.000529124 | 0.00114012  |       0.996219 |
-| 11 | \textit{2x linebreak}      | ]     | \textit{eos-token} | 0.000603324 | 0.00100365  |       0.995909 |
-| 12 | \textit{2x linebreak}    | \textit{linebreak}      | \textit{eos-token} | 7.61097e-06 | 0.00124311  |       0.99872  |
-| 13 | \textit{2x linebreak}    | \textit{linebreak}      | \textit{eos-token} | 2.16886e-05 | 0.000565556 |       0.999231 |
-| 14 | \textit{2x linebreak}    | \textit{linebreak}      | \textit{eos-token} | 3.80865e-05 | 0.00084979  |       0.998713 |
-| 15 | \textit{2x linebreak}    | \textit{linebreak}      | \textit{eos-token} | 5.16162e-05 | 0.00161664  |       0.997724 |
-| 16 | \textit{2x linebreak}    | \textit{linebreak}      | \textit{eos-token} | 0.000101122 | 0.0031923   |       0.995816 |
-: Example for a low 1st probability {#tbl-lowprob}
+| 11 |       | ]     | \textit{eos-token} | 0.000603324 | 0.00100365  |       0.995909 |
+| 12 |  \textit{linebreak}   |   \textit{linebreak}    | \textit{eos-token} | 7.61097e-06 | 0.00124311  |       0.99872  |
+| 13 |  \textit{linebreak}   |   \textit{linebreak}    | \textit{eos-token} | 2.16886e-05 | 0.000565556 |       0.999231 |
+| 14 |  \textit{linebreak}   |   \textit{linebreak}    | \textit{eos-token} | 3.80865e-05 | 0.00084979  |       0.998713 |
+| 15 |  \textit{linebreak}   |   \textit{linebreak}    | \textit{eos-token} | 5.16162e-05 | 0.00161664  |       0.997724 |
+| 16 |  \textit{linebreak}    |  \textit{linebreak}     | \textit{eos-token} | 0.000101122 | 0.0031923   |       0.995816 |
+: Example for a low 1st probability based on factor model {#tbl-lowprob}
 
@@
 | 12 | Transfer |            | \textit{eos-token} | 0.000179846 | 0.00032993  |       0.998234 |
-| 13 |          | More       | \textit{eos-token} | 0.000109437 | 0.000194042 |       0.998748 |
-| 14 | More     |            | \textit{eos-token} | 7.64376e-05 | 0.000174342 |       0.999236 |
-| 15 | Lab      |            | \textit{eos-token} | 5.29115e-05 | 0.000444251 |       0.998936 |
-| 16 | Lab      |            | \textit{eos-token} | 0.000276325 | 0.00155492  |       0.997141 |
-: Example for similar tokens {#tbl-similartoken}
-
-  - exploration of the likelihoods
-  - frequentist test of generating existing
-
-[ ] Repetition with actual marketing data
-
-### Implementation
-
-- GPT-2; Far from state of the art, if it works here, it will work with other model
-
-- Autoencoder
-
-- Configuration of LLM
-
-- Connection to other measures, e.g. syntax or rating of consumers
-
-- Different generation strategies, use beam search to get closer to max LL
-
-- Optimization
-  - @diederikp.kingmaAdamMethodStochastic2014
-  - Computation time, scaling
-  - Issues with spikes; Figures where they come from
-
-- Encoding space
-  - example tangible, intangible
-  - generations from encoding space
-  - Wasteland, geometrical shape and where it comes from
-
-- Measures of fit
-  - areas in the encoding space (count out the pixels and compare to curvature of LL?)
-  - analysis of the trees
-  - Gap between most likely and second most likely token
-
-- Comparison to BERT embeddings?
-- Comparison to prompt engineering results?
-
-# Empirical Application
-
-- Summary statistics of the data
-
-- Application to MR data
-
-<!-- TODO: Check whether this is okay to name -->
-
+| 13 |  \textit{linebreak}        | More       | \textit{eos-token} | 0.000109437 | 0.000194042 |       0.998748 |
+| 14 | More     |   \textit{linebreak}         | \textit{eos-token} | 7.64376e-05 | 0.000174342 |       0.999236 |
+| 15 | Lab      |   \textit{linebreak}         | \textit{eos-token} | 5.29115e-05 | 0.000444251 |       0.998936 |
+| 16 | Lab      |   \textit{linebreak}         | \textit{eos-token} | 0.000276325 | 0.00155492  |       0.997141 |
+: Example for similar tokens based on factor model {#tbl-similartoken}
+
+## Objective 3: Can we use these generative summaries to generate new documents?
+
+To evaluate whether we can use our generative summaries to generate new documents, we want to explore the space in between two generative summaries.
+Below, we explore combinations of generative summaries and points from the factor space to see what kinds of documents they generate.
+We start by interpolating between two generative summaries which we have estimated by themselves, i.e.
+without a factor model connecting them.
+The second analysis will be on the factor space of the hair shampoo 2-factor model.
+
+@tbl-interpolation shows the documents that we generate for a convex combination of the single generative summaries for the claims *"Unlock the secret to hair that shines from within, reflecting your inner glow."* (A) and *"Rediscover the joy of hair that beams with inner vibrancy."* (B), based on their generative summaries, which we estimated directly.
+We combine them as $\boldsymbol{s}_{combine} = weight * \boldsymbol{s}_{A} + (1 - weight) * \boldsymbol{s}_{B}$, where $weight \in [0, 1]$.
+For weights that are close to either $0$ or $1$ we still generate one of the two claims.
+When we step further into the middle between these two generative summaries the end of the sequence changes first (e.g.
+*"Unlock the secret to blackened nails that shine from the shine."* at for $weight = \frac{3}{20}$).
+For weights that are close to $\frac{1}{2}$, we stop generating eos-tokens and the generated strings become unintelligible.
+However, these strings still maintain some of the structure of the focal claims such as having words related to beauty products and starting with the syllable "Un" or "Red".
+
+|    | Weight | Generated String                                                                                      |
+|---:|:-------|:------------------------------------------------------------------------------------------------------|
+|  0 | 0/20   | Unlock the secret to hair that shines from within, reflecting your inner glow.   \textit{eos-token}                     |
+|  1 | 1/20   | Unlock the secret to hair that shining from within, your bedroom. \textit{eos-token}                                     |
+|  2 | 2/20   | Unlock the secret to black metal's glow. \textit{multiple eos-token}                                                              |
+|  3 | 3/20   | Unlock the secret to blackened nails that shine from the shine.   \textit{eos-token}                                    |
+|  4 | 4/20   | Un to of for the team members upon                                                                    |
+|    |        | Moderation of the, that                                                                               |
+|  5 | 5/20   | Un to (93) of 15 + 18 + 36                                                                            |
+|    |        | Add to                                                                                               |
+|  6 | 6/20   | Un to (mit. of private eye and nirvana-ly-ly                                                          |
+|  7 | 7/20   | Uncle-healedered with a lifetime-changing blend of light and energy and                               |
+|  8 | 8/20   | Unclelyed by the Tone's Bright Side™ Lipstick.                                                        |
+|    |        | Bright                                                                                                |
+|  9 | 9/20   | Unclelying our clients and fory, we're sure about your smile.                                         |
+| 10 | 10/20  | Redmates on the Road to of Dreams.                                                                    |
+|    |        | Inspirating your soul with                                                                            |
+| 11 | 11/20  | Redmates at the Sunlight Your Hair.     \textit{multiple eos-token}                                                              |
+| 12 | 12/20  | Redhs founder, that a master the way she loves......                                                  |
+| 13 | 13/20  | Redhs founder, of where, at, and to and to, and can be                                                |
+| 14 | 14/20  | Rediscover the joy of purpose and detail with a strand of your hair. (and                             |
+| 15 | 15/20  | Rediscover the joy of success with long,istry's health and creativity. back to                        |
+| 16 | 16/20  | Rediscover the joy of hair that loves to making you.™ the way.br                                      |
+| 17 | 17/20  | Rediscover the joy of hair that's at the-corrects life with a healthy                                 |
+| 18 | 18/20  | Rediscover the joy of hair that beams with inner vibrancy. \textit{eos-token}                                           |
+| 19 | 19/20  | Rediscover the joy of hair that beams with inner vibrancy. \textit{eos-token}                                           |
+: Interpolation between two advertising claims based on non-factor model. {#tbl-interpolation}
+
+Next, we want to explore the encoding space of a factor model and generate new texts from points in this encoding space.
+In @fig-space, we perform a grid search across the factor space of our 2-factor model for shiny hair shampoo claims.
+For each point in the grid search, we generate a text by passing this point through the decoder to create a generative summary.
+We then pass this generative summary into the LLM to generate a text for this point from the encoding space.
+In the plot, we have three types of different markers.
+A blue bubble with a number represents the location of the respective claim, while a colored dot represents a point where we generate one of the points that are part of the training data.
+Red crosses are points from which we generate a string, that comes to an end with the end of text token and is not part of the training data ("candidate points").
+Whitespace represents points where we do not generate a string that comes to an end within the number of tokens that we consider.
+Around each advertising claim, we find an island of points from which we can generate the same claim.
+The islands are surrounded by candidate points, and areas that are further removed from any of the training claims tend to be whitespace.
+
+![Exploration of Encoding Space of the factor model.
+Numbered bubbles are the locations of the respective claim, red crosses represent candidate points.
+These are points from which we generate a text that ends with the eos-token, but is not part of the training data.
+Points that do not generate an eos-token are marked by whitespace.](figures/exploration_encoding_space.pdf){#fig-space}
+
+Some examples of candidate points are:
+
+- "Experience 50% more visible shine after-changing lighting.<|endoftext|>"
+- "Visible reduction in dullness, but with a touch of menace.<|endoftext|>"
+- "Step into the spotlight with hair that to website in.).<|endoftext|>"
+- "Elevate your confidence with hair that luxeaks.<|endoftext|>"
+- "Experience the magic of-life.<|endoftext|>"
+
+Many of these candidate points generate the same texts (if the points are close to each other).
+Most of these generated claims do not seem to be intelligible.
+However, they still tend to use words from the training data and have a sentence structure that is similar to product claims.
+
+<!-- TODO: Unclear what area means in factor space -->
+<!-- TODO: Add freely estimated generative summaries -->
+<!-- TODO: Give reason for need for regularization -->
+<!-- TODO: Give reason for using two factors -->
+<!-- TODO: Give the right methods / diagnostics -->
+<!-- TODO: Explain generation -->
+<!-- TODO: Give results without LayerNorm, why is it weird? -->
+<!-- TODO: Have similarity measure vs euclidean distance in encoding space -->
+<!-- TODO: Just show likelihood graph -->
+<!-- TODO: Smaller vocabulary than dims? Just compare RE and BERT -->
+
+<!--
+
+We perform Principal-Component Analysis on the embedding matrices and visualize the first two principals components in @fig-pca.
+We color the data points by their respective class.
+We see that we can separate these two classes linearly for all three embeddings.
+
+::: {#fig-pca layout-ncol=3}
+![RE](figures/pca_re.pdf)
+
+![W2V](figures/pca_w2v.pdf)
+
+![BERT](figures/pca_bert.pdf)
+
+First two principal components of the embedding matrix, colored by class.
+:::
+
+@fig-loo shows the results of the leave-one-out regression. For the generative summaries, the variance of the intercept is substantially larger than for the other embeddings.
+
+::: {#fig-loo layout-ncol=3}
+![RE](figures/loo_re.pdf)
+
+![W2V](figures/loo_w2v.pdf)
+
+![BERT](figures/loo_bert.pdf)
+
+Reuslts of leave-one-out regression, colored by class.
+:::
+
+Another explanaition could lie in the decoding strategy.
+@ariholtzmanCuriousCaseNeural2020 show how text generation based on a maximum likelihood objective leads to text that sounds bland and not human, propsing the use of different decoding strategies to make the generated text more human.
+-->
+
+<!-- TODO: Turn this into bar-charts? -->
+
+# Objective 4: Empirical Application
+
+<!-- 
 - Two small sub-samples from the same market research study of product claims for a dairy product
@@
 
-::: {#fig-anecdote-mr5 layout-ncol=2}
-
-![Occurrence of words taste, natural, and immune system.](figures/anecdote_immune_system_mr5.pdf){#fig-anecdote-mr5a}
-
-![Occurrence of words relating to morning theme](figures/anecdote_morning_system_mr5.pdf){#fig-anecdote-mr5b}
-
-Embedding of the yoghurt claims, colored by word features.
-:::
-
-- establish that the space is continuous, word features etc.
-
-![Uniqueness score of embedded yoghurt drink claims.](figures/uniqueness_mr5.pdf){#fig-ratings-mr5a}
-
-@fig-anecdote-mr5 shows the encoding space for the yoghurt drink advertising claims, and illustrates how the embedding picks up on language features of the documents.
+-->
+
+After validating the properties of our method on a synthetic dataset, we now apply it to market research data, to illustrate how it can be used in practice.
+In the following, we look at two empirical applications of our generative summaries.
+The first application is on product claims for yogurt drinks and their appeal to consumers from three different countries.
+In the second application, we analyze product claims for yogurt.
+For this dataset, we also have design motivations behind these product claims, which we will use to validate whether our algorithm captures these design motivations.
+Since we cannot disclose the actual advertising claims, we generate a set of similar claims with [ChatGPT](https://chatgpt.com/) or use "stand-ins" for advertising claims, which are supposed to represent these and their properties without actually disclosing them.
+However, we performed all computations on the actual data.
+
+## Yoghurt-drink claims, word features, and consumer ratings (Application 1)
+
+Besides the advertising claims themselves, we also have measures on the uniqueness of a claim (as perceived by consumers) and an overarching appeal rating of the claim by the consumer.
+To aid the market research process of finding the best advertising claims, we investigate the encoding space of the advertising claims, as estimated with a 2-factor generative summary.
+First, we explore this space, to identify linguistic features of the advertising claims.
+These are not features that were reported in the data but rather features that we interpret ourselves by looking at claims that are close to each other in the encoding space.
+We look at both word features (i.e.
+a certain word occurs in the claim) and a higher level theme of the claim, e.g.
+whether the claims frame the yogurt drink as a "breakfast" or "start in the day" ("morning" theme).
+Next, we are interested in whether the uniqueness measure is positively correlated to the distance of claims to each other in the encoding space.
+Namely, whether more unique claims are more separated from other claims in the encoding space.
+Finally, we explore whether certain regions of the encoding space are associated with higher appeal ratings.
+We trained this 2-factor document summary until we reached a joint likelihood of $0.99$, which training took around 8 minutes on the Nvidia A100 GPU.
+We also embed these advertising claims by summarizing them with BERT and extract the first two principal components from the resulting embedding matrix [see @pearsonLIIILinesPlanes1901].
+We want to caution, that these findings are based on a very small sample and are supposed to serve as a proof of concept on how these generative summaries, or here their encoding space, can be useful in market research.
+
+::: {#fig-anecdote-mr5a layout-ncol=2}
+
+![Generative Summaries (factor model)](figures/anecdote_immune_system_mr5.pdf){#fig-anecdote-mr5a}
+
+![BERT CLS + PCA](figures/anecdote_immune_system_bert_mr5.pdf){#fig-anecdote-mr5a}
+
+Embedding of the yogurt claims, colored by word features.
+:::
+
+::: {#fig-anecdote-mr5b layout-ncol=2}
+
+![Generative Summaries (factor model)](figures/anecdote_morning_system_mr5.pdf)
+
+![BERT CLS + PCA](figures/anecdote_morning_system_bert_mr5.pdf)
+
+Occurrence of words relating to the morning theme in the low-dimensional spaces.
+:::
+
+::: {#fig-uniqueness layout-ncol=2}
+
+![Generative Summaries (factor model)](figures/uniqueness_mr5.pdf)
+
+![BERT CLS + PCA](figures/uniqueness_bert_mr5.pdf)
+
+Uniqueness score of embedded yogurt drink claims.
+:::
+
+@fig-anecdote-mr5a shows the encoding space for the yogurt drink advertising claims and illustrates how the embedding picks up on language features of the documents.
 Namely, @fig-anecdote-mr5a colors the claims based on whether they contain the words "taste" or "sense", "natural", and/or "immune system".
@@
 If a claim contains none of these words, we code it in grey.
-Claims with the same word features cluster together, while claims that contain none of these words, are pushed to the outside of the plot.
-Intuitively, we would expect that claims which contain multiple of these word features from a transition between the claims which only have one of the word features.
-For this, we only have claims 13 and 21 as examples, whereas the former does not form such a boundary, and the latter is at the edge of the "Natural" and "Immune System" class.
+These "flag-words" come from looking at the factor space and comparing close advertising claims with each other, through this exploration we discovered these flag-words.
+Hence, this is not an external label.
+
+Claims with the same word features live in similar regions of the encoding space, while claims that contain none of these words are pushed to the outside of the plot.
+Intuitively, we would expect that claims that contain multiple of these word features form a transition between the claims that only have one of the word features.
+For this, we only have claims 13 and 21 as examples, whereas the former does not form such a boundary, and the latter is at the edge of the "Natural" and "Immune System" classes.
+Despite not being an external label, these word features tend to cluster to certain regions in the BERT visualization as well.
+
+Similarly, @fig-anecdote-mr5b colors the advertising claims blue if they are written with a "morning" theme.
+An example of such a claim would be *"Begin your day with a crisp and revitalizing taste"* (synthetic example).
+Like the word features in @fig-anecdote-mr5a, we hand-coded these themes.
+The split of the two groups is not crisp, but morning-themed advertising claims tend to occur in the south-west of the space, while the other claims are in the north-east.
+Similarly, for the BERT visualization, morning-themed claims occur mostly in the southeast of the plot.
 
@@
 We expect advertising claims that are more unique to have a more isolated position in the encoding space, as they are less similar to other claims.
-In other words, the uniqueness score should be positively correlated with the euclidean distance of an encoded claim to its neighest neighbor.
-@fig-ratings-mr5a shows that this is not the case, the correlation of these two measures is substanial and negative ($-0.4375$).
-The opposite is true: The three most unique claims (4, 17, 18) cluster together closely, and the further claims are away from these three, the less unique they tend to be.
+In other words, the uniqueness score should be positively correlated with the Euclidean distance of an encoded claim to its nearest neighbor.
+This is the case for the BERT-based space, however, not for our factor space.
+A potential reason for this could be that two factors are not enough to pick up on higher levels of abstraction, such as uniqueness, while the principal components of a 768-dimensional embedding could.
+
+For the 2-factor generative summary space, @fig-uniqueness shows that the uniqueness measure and nearest neighbor distance are negatively correlated ($r=-0.44$).
+Here the three most unique claims (4, 17, 18) cluster together, and the further claims are away from these three, the less unique they tend to be.
 Claims with low uniqueness scores are spread further apart and far away from the most unique claims.
 These three most unique claims have some things in common, all of these use a combination of the words "natural", "active", and talk about ingredients (with synonyms).
-On the other hand, claim 15 is the only claim in the dataset using the word "yoghurt drink" and is also the only claim about emotions.
+On the other hand, claim 15 is the only claim in the dataset using the word "yogurt drink" and is also the only claim about emotions.
 Thereby, it is unique from a language perspective (far away from other points in the encoding), but perhaps not unique to consumers, as there might be similar claims on the market already.
-Also, uniqueness is highly correlated ($0.8843$) with the overall rating of claims, some of this correlation could be due to consumers viewing uniqueness as a proxy for "satisfaction".
+Also, uniqueness is highly correlated ($r=0.88$) with the overall rating of claims, some of this correlation could be due to consumers viewing uniqueness as a proxy for "satisfaction".
 Lastly, the researched advertising claims are self-selected, and perhaps more unique then the average claim that is on the market already.
-This might lead to a form of Simpson's paradox [@sprenger2021simpson], where in the population there is a positive correlation between uniqueness and unique language, but perhaps not in this special sub-sample of the data.
-
-![Overall rating of embedded yoghurt drink claims.](figures/rating_country_mr5.pdf){#fig-ratings-mr5b}
-
-@fig-ratings-mr5b again shows the encoding space, but this time colored by the overall rating of the claims.
-Claims of similar rating tend to cluster together, with higher ratings in the north-east and lower ratings towards the south-west of the graph.
-The clustering of similarly rated claims, could be due to these claims being similar in language.
-For marketers, there are two insights from this: One, getting the overall theme of a claim right, can ensure that customers perception of this claim is within a certain ballpark.
+This might lead to a form of Simpson's paradox [@sprenger2021simpson], wherein the population there is a positive correlation between uniqueness and unique language, but perhaps not in this special sub-sample of the data.
+
+::: {#fig-rating layout-nrow=2}
+
+![Generative Summaries (Factor Model)](figures/rating_country_mr5.pdf)
+
+![BERT](figures/rating_country_bert_mr5.pdf)
+
+Embedding of the yogurt claims, colored by appeal rating.
+:::
+
+@fig-rating again shows the encoding space, but this time colored by the overall rating of the claims and split by the country for which the market was researched.
+The same claims can get different ratings by country, the reasons for this could lie in what consumers are used to from the domestic market and in cultural differences.
+For both our proposed method and the BERT-based visualization, we find that claims with similar appeal ratings cluster together.
+For our document summaries, higher ratings occurr in the northeast and lower ratings towards the southwest of the graph.
+The clustering of similarly rated claims could be due to these claims being similar in language.
+For marketers, there are two insights from this: One, getting the overall theme of a claim right, can ensure that customers' perception of this claim is within a certain ballpark.
 Two, after identifying a fruitful theme for the advertising claim, it is still useful to explore this neighborhood in more fine-grained steps, as locally, there might be small alterations that have large effects on the perception: See e.g.
-claims 21 (rating in top 4% percentile ) and 14 (rating in top 12%), despite one of the shortest distances in the encoding space.
-@fig-ratings-mr5b also shows how ratings of uniqueness can differ by country, the reasons for this could lie in what consumers are used to from the domestic market and in cultural differences.
+claims 21 (rating in top 4% percentile ) and 14 (rating in top 12%), even though these claims are very close to each other.
+When we consider these observations together with the identified word features and themes from above, it appears that advertising claims for yogurt drinks, which mention the immune system and do not play into a morning theme are appealing to consumers, however, these claims are not perceived as the most unique.
+Claims that mention the word "natural" are among the most unique claims, again not mentioning a morning theme.
+
+## Yoghurt-claims and design motivations (Application 2)
+
+For the yogurt claims, we estimate a 20-factor model and train it until we reach a likelihood of $0.99$.
+We train this model on all unique advertising claims in this dataset.
+Since the encoding space is of dimension 20, we cannot use it for visualizations directly.
+Instead, we apply Principal Component Analysis to these 20 factors and extract the first two principal components.
+The intuition behind this is, that if we capture relevant information with these 20 factors, then we might also pick up on this information with the principal components of these factors.
+As a comparison, we again apply BERT's CLS token to the same advertising claims and perform PCA on the resulting embedding.
+
+As a first step, we are interested in the location of claims that mention a specific country in the principal component space.
+An example of such a claim would be: "Yogurt fermented in Germany".
+These types of claims should have a similar effect on consumers in the domestic market, and hence a good representation of these advertising claims should cluster the different "country-versions" of the same advertising claims together.
+For this, we flag claims if they mention a specific country and then group these claims into "Claim Types".
+Some claims don't mention a specific country but rather a local origin.
+We collect these claims under the flag "Local".
+
+We find three claim types: "Support", "Ingredients", "Origin", and "Made in X".
+These four claim types are identical in wording and only differ by the mentioned country.
+The first one is about support for "local" producers, the second one is about domestic ingredients, the third one is about the origin of the product and the fourth one uses the phrase "made in (country)".
+
+In @fig-mr7-country we show the advertising claims in the principal component space and color claims by their mentioned country and adjust the marker type by claim type.
+For the BERT embedding, we see that the different claim types cluster together across all groups.
+For the "Ingredients" and "Made in X" types, the claim from the "Local" location is a bit removed from the cluster.
+This is plausible, as the wording of these claims deviates grammatically from that of the respective claims for the nations.
+We find a similar result for the PCA visualization of our 20-factor model.
+However, some of the clusters are not as clearly separated from each other (e.g.
+"Support" & "Origin").
+It appears that both types of document summaries capture the information in the texts in such a way, that similar claims cluster together in a two-dimensional principal component space.
+
+We expand our analysis by investigating whether we can identify the design motivations of the advertising claims again in a two-dimensional principal component space.
+In @fig-mr7, we show the first two principal components of the generative summary and the BERT summary for all advertising claims from the five major categories "Packaging", "Local & Responsible", "Sourcing", "Naturality", and "Sustainability", which are labels provided in the data.
+In the PCA visualization of the yogurt claims based on the BERT embeddings, we see that claims with the "Packaging" theme are separated from the other types of claims in the northwest corner of the plot.
+The claims with the themes of "Local & Responsible", "Sourcing", "Sustainability", and "Naturality" do not seem to exhibit a specific pattern and are meshed together in the southwest and east areas of the plot.
+
+The visualization is different for the generative summaries.
+Here, advertising claims with the theme "Naturality" follow the second principal component, while the "Local & Responsible", "Sustainability", and the "Packaging" themes follow the first principal component.
+It also appears that there is a transition between the "Local & Responsible" and "Packaging" themes along this axis.
+
+The theme "Sourcing" appears to be situated at the intersection of "Local & Responsible" and "Naturality." The three "Sourcing" claims, with a first principal component between -4 and -2, and a second principal component between 0 and 2, all have the same phrasing which emphasizes the local origin of the yogurt.
+The only difference between these three claims is that they use a different "country" to describe the origin (e.g.
+*"Fermented with just French ingredients"*).
+We observe a similar pattern for the four southwest "Sourcing" claims, which only differ by an adjective that specifies the origin of the farmers from which the product is sourced.
+The two most south "Naturaliy" claims are claims that mention the local and organic ingredients of the yogurt.
+This ties in with the surrounding claims, which are part of either the "Sourcing" or "Local & Responsible" themes.
+The two "Packaging" claims, which have a first principal component of smaller than one, emphasize how the packaging is not wasteful.
+The most "northern" packaging claim, i.e.
+the one with the highest second principal component, points to "plant-based" (i.e.
+"natural") packaging materials.
+Since most of these "Packaging" claims revolve around e.g.
+recyclable or eco-friendly packaging, it fits that most of the "Sustainability" claims are also located among the "Packaging" claims.
+
+In the word embedding literature, a prominent example to explore the information content of these embeddings is to compare linear combinations of word embeddings to other word embeddings.
+@mikolovDistributedRepresentationsWords2013 present the following example: The word embedding of the word "Queen" is the closest word embedding to the combination of the word embedding for "King", minus the word embedding for "Man", plus the word embedding for "Woman" (see illustration @eq-word).
+Here, we try something similar with the advertising claims: We form linear combinations of the generative summaries and generate a text from the resulting vector.
+For example, we can alter the country of origin for an advertising claim, as follows: The generative summary for "Yoghurt produced in Sweden" - "Swedish ingredients" + "UK ingredients" generates the text "Yoghurt produced in the UK", which is another advertising claim from the training data (see illustration @eq-generation).
+Similarly, in an attempt to generate the claim "Yoghurt produced by local farmer" (which is not part of the training data), we form the combination "Yoghurt produced in Germany" - "German ingredients" + "Local Product", which yields: "Producing farmer Produced by the local farmer [repetition]".
+While the generated text is not grammatically correct, does not end with an eos-token and contains a repetition, it does seem to capture that we want an advertising claim that emphasizes the local producer of the product, and contains no mention of the ingredients themselves.
+Finally, we find indications that generated claims make use of associations that the LLM has learned, which go beyond grammatical structures.
+The combination "Vegetarian ingredients" + "Local production" - "Overpackaging" generates the text "Locally sourced in the United States of America.
+We are in the process of moving to the United States of America.
+[repetiton]".
+While this is not a valid advertising claim, it contains mentions of local production.
+Perhaps, subtracting the generative summary for "Overpackaging" leads to the generation of "United States of America", which is a country that is never mentioned in the training data, but might be associated with packaging waste in the pre-training data of the used LLM.
+
+$$
+\boldsymbol{\text{King}} - \boldsymbol{\text{Man}} + \boldsymbol{\text{Woman}} \approx \boldsymbol{\text{Queen}}
+$${#eq-word}
+
+$$
+\operatorname{LLM} \left( \mathbf{S}_{\text{Yoghurt produced in Sweden}} - \mathbf{S}_{\text{Swedish ingredients}} + \mathbf{S}_{\text{UK ingredients}} \right) = \text{"Yoghurt produced in the UK"}
+$${#eq-generation}
+
+::: {#fig-mr7-country layout-ncol=2}
+
+![First two principal components of the 20-factor model](figures/PCA_mr7_country_mr7.pdf)
+
+![First two principal components of the BERT model](figures/PCA_mr7_country_bert_mr7.pdf)
+
+Coloring of claims by whether they mention a certain origin country.
+:::
+
+::: {#fig-mr7 layout-ncol=2}
+
+![First two principal components of the 20-factor model](figures/PCA_mr7_mr7.pdf)
+
+![First two principal components of the BERT model](figures/PCA_mr7_bert_mr7.pdf)
+
+Embedding of the yogurt claims, colored by word features.
+:::
 
@@
 
-- Generative AI for personalization
-
-- Brands
-
-- Wording: Ballpark, finetuning idea
-
-- Interactive dashboard to explore the embedding space
-
-- Costs of tailoring advertising claims
-  - When we take current prices for the [GPT-4o LLM; June 2024](https://openai.com/api/pricing/) (\$5 per 1 Mio tokens input, \$15 per 1 Mio tokens output), then training a dataset of 100 advertising claims, for 5,000 epochs will cost \$150.
-Freelancers on platforms such as [Upwork.com](https://www.upwork.com/) charge between \$20 and \$200 per hour for work on tasks such as SEO optimization, copy writing, and creation of advertising claims.
-    - Online services for conjoint studies, such as [Conjointly.com](https://conjointly.com/) and [Sawtooth](https://sawtoothsoftware.com/pricing) charge a few thousand dollars per year to use their services.
-Leveraging the resulting insights as much as possible, by using generative AI, could hence make market research more cost effective.
-
+Managers can use these generative summaries to augment the design process of advertising claims.
+The work by @burnap2023product and @ludwigMachineLearningTool2023 propose a framework for guided generation of images.
+The researchers model an embedding space for images, that fulfills two requirements: One, points in this embedding space can generate new images and two, they form useful features for a supervised machine learning task.
+For our setting of text data, we also estimate an embedding that can generate new text data and captures inherent information of the advertising claims, which could be used for classification.
+The frameworks of @burnap2023product and @ludwigMachineLearningTool2023 use three components: The encoder, which projects data points into an encoding space (our factor model), the generator (our LLM), which turns points from the encoding space into new data points, and the predictor, which predicts a certain label based on a point from the encoding space (not implemented).
+In a similar way to @burnap2023product, businesses could improve their advertising claims, e.g.
+by training the generator and predictor model based on their market research data, e.g.
+predicting overall appeal for a claim.
+We can combine such a predictor model with the low-dimensional factor space of the generative summaries, to guide the exploration of new claims.
+For this, we only need the decoder part of our factor model: For every point in the factor space, we can get the corresponding generative summary and predict the respective outcome for this generative summary with the predictor model.
+Through gradient descent, we can now find the points in the factor space that maximizes this predicted outcome.
+In a classification setting, we can use this technique, e.g.
+to find the closest point in the factor space which changes the predicted class of the advertising claim.
+
+A framework of generator and predictor model can also help in the design of market research itself.
+@ludwigMachineLearningTool2023 propose a workflow for generating research hypotheses based on such a model, by finding small steps in the encoding space that maximally change the prediction of the predictor model.
+They show that these steps yield interpretable and novel hypotheses for their research setting.
+In a similar vein, marketers might look for a minimal change to an advertising claim, such that this claim changes from being "tangible" to being "intangible".
+This could help to identify the effects of certain design motivations more clearly, as it operationalizes the "ceteris paribus" principle, by making the smallest possible alteration for this change in class.
+Typically, such alteration is difficult to obtain in research designs for flexible modalities such as text.
+
+Managers can also explore the factor space to find new design motivations.
+@fig-anecdote-mr5a shows design motivations, which were not designated as themes in the market research study.
+By looking at the structure of the factor space, and comparing neighbouring claims for the similarities and differences, managers can identify linguistic features, such as the use of certain words of themes.
+
+<!-- - Paragraph on prompt engineering
+  - move away from prompt engineering
+  - certain tasks should be optimized, similar to automated grading
+  - validation of answers
+  - communication on social media
+  - automated customer service bots
+  - interpolation between different claims
+  - there are applications where distribution is desired -->
+
+<!-- - Rather than inform product design, @ludwigMachineLearningTool2023, use such a framework to generate hypotheses for why a prisoner might be granted bail or not.
+For this, they predict judges decisions' from bail hearings, based on the mugshot photo of the defendant (predictor) and learn about the distribution of mugshots with a Generative Adversarial Network [CITE Goodfellow GAN].
+To generate new hypothesis, they perform gradient descent on the prediction of the predictor with respect to the encoding data point.
+They search for the alteration to the image, that changes the predicted outcome the most.
+By making equal steps along this gradient in opposing directions, they obtain to versions of the same image manipulated with respect to this feature.
+@ludwigMachineLearningTool2023 also find, that comparing these two images often yields people to name the specific change, creating new hypothesis on why an inmate might be granted bail or not.
+-->
+
+<!-- 
 - Generator, predictor compatible embeddings; Generative AI to assist and automate design processes in marketing
@@
 Namely, for every sequence in our dataset, we find an embedding vector, such that when we use this vector as an input to our LLM, we maximize the likelihood of generating this sequence.
-This we call the summary embedding.
+This we call the generative summary.
 We perform this optimization for our full dataset simultaneously by using a factor model.
-  - By using this factor model, we create a bottleneck, forcing the model to learn helpful features; In other factor models, these features are often interpretable dimensions [@goodfellowDeepLearning2016].
+ -->
+
+<!-- - Generative AI for personalization
+
+- Brands
+
+- Wording: Ballpark, finetuning idea
+
+- Interactive dashboard to explore the embedding space
+
+- Costs of tailoring advertising claims
+  - When we take current prices for the [GPT-4o LLM; June 2024](https://openai.com/api/pricing/) (\$5 per 1 Mio tokens input, \$15 per 1 Mio tokens output), then training a dataset of 100 advertising claims, for 5,000 epochs will cost \$150.
+Freelancers on platforms such as [Upwork.com](https://www.upwork.com/) charge between \$20 and \$200 per hour for work on tasks such as SEO optimization, copy writing, and creation of advertising claims.
+    - Online services for conjoint studies, such as [Conjointly.com](https://conjointly.com/) and [Sawtooth](https://sawtoothsoftware.com/pricing) charge a few thousand dollars per year to use their services.
+Leveraging the resulting insights as much as possible, by using generative AI, could hence make market research more cost effective.
+-->
 
@@
 
-- Further Applications
-  - Framework of generator and predictor
-  - optimization of prompts in continuous space
-  - automatic answer evaluation
-  - explainable AI for generative models
-
-## Problems
-
-- Issues with Generation
-  - How to evaluate the quality of the generated ad claims @tatsunorihashimotoUnifyingHumanStatistical2019
-  - Another explanaition could lie in the decoding strategie.
+In this research, we propose a novel type of document summary, which maximizies the probability to re-generate the focal document.
+We show that it requires a form of regularization in practice to capture useful features.
+We also show that these generative summaries let the generation distribution collapse at the desired target sequence.
+We can interpolate between two document summaries, and a low-dimensional factor representation captures linguistic similarity of documents.
+Since this is a new method, further validation is required.
+Below, we discuss some limitations and challenges of our document summaries, issues with our applications, and end with next steps in this research.
+
+Here, we summarize the results for our four *research objectives*:
+
+1) Yes, we capture meaningful features with the generative summaries, that we can use for classification.
+2) Yes, we can re-generate the focal document and it appears that the generative distribution collapses at the target sequence.
+3) We can interpolate between two document summaries, and the low-dimensional factor representation appears to make sensible generations, as e.g.
+linear combinations of generative summaries have meaning.
+However, the quality of the generations is limited, as many of the generated advertising claims are not intelligible.
+4) Yes, we can find marketing insights with our generative summaries, which are relevant and can lead to managerial implications.
+However, further validation and experiments are required, especially to validate the use of their text generation abilities.
+
+## Limitations
+
+Newly generated advertising claims are often unintelligible and there appears to be a lot of whitespace in the factor space.
+We find three reasons for this.
+One, we use the GPT-2 model in this paper, which is far from state-of-the-art, e.g.
+on Huggingface's Open LLM Leaderboard, the best versions of GPT-2 achieve around half the score of the leaderboard leader, which is a model based on [Llama-3](https://llama.meta.com/llama3/).
+To address this problem, we can use a more capable open-source LLM, such as Llama 3 instead, which comes at a larger computational cost.
+The second reason lies in the construction of our factor space.
+According to @goodfellowDeepLearning2016, factor models can learn and represent features of the data well.
+However, they struggle with the generation of new data points, which in practice, tend to be mixtures of the learned features rather than realistic data points.
+This is a pattern we see without newly generated advertising claims, e.g.
+in @tbl-interpolation.
+The third reason is also related to the architecture of our current model but with respect to the number of layers and activation functions.
+Currently, we a searching for a representation of the advertising claims in 2 dimensions, and take linear combinations of the resulting factors to create our generative summaries.
+Perhaps, this structure is too simplistic to represent this language domain well.
+Introducing additional layers and non-linear activation functions, such as the Swish function [@ramachandranSearchingActivationFunctions2017a], might yield a representation of the space that is better suited for generation, as we learn the manifold on which these advertising claims live, better.
+We also observe that more factors yield a more powerful generative summary, however that an estimation of the generative summaries without a factor structure makes them overfit.
+Currently, there is no data-driven method to regularize the generative summaries, which leaves the choice of the number of factors to the researcher.
+
+<!-- - Linear combinations of generation work, use many more factors than before! Still see similar patterns ... -->
+
+There are some challenges, which we inherit from LLMs themselves, such as limited language modeling capabilities in non-English languages and safety and copyright issues when generating new text.
+However, these are problems that we can circumvent, or at least mitigate, by using a more capable LLM with safeguards, or an LLM that has been trained in a specific language if we are working with non-English texts.
+
+We also want to discuss some observations in the computation of these document summaries.
+The computation time of our factor model depends on two aspects.
+The first aspect is the length of the supplied documents and the second is the number of factors.
+Models that have fewer factors get more and more difficult to compute, especially if we are using more and more claims.
+The intuition for this is, that it gets difficult to "squeeze" many different documents through only a few factors, while still regenerating each document.
+One way how this affects the optimization is, that the log-likelihood plateaus and only very slowly increases (if at all).
+While this might be less of a problem, and more of a statistical fact (we cannot represent certain complexities with a low-dimensional factor model with an arbitrarily high likelihood), it can be an issue in certain applications, e.g.
+when the researcher needs a specific (low-dimensional) factor representation for a large number of claims.
+
+<!-- TODO: Do this with BERT and PCA? -->
+
+Our empirical application has low power and does not allow for generalizations, as it is based on a sample of at most 215 observations, with even fewer unique claims.
+To validate whether our factor model indeed has regions where claims exist that have e.g.
+a higher appeal rating, we would need to generate new claims from this region and investigate them in an experiment together with the existing claims, as our interpretations are purely correlational.
+
+<!--
+
+- Implementation challenges
+  - Computation time
+  - [Tokenizer can cause problems?](https://www.promptingguide.ai/research/llm-tokenization)  TODO: Is this true? Give an example? Won't be a problem forever ... But inherent to this type of model
+
+- Another explanaition could lie in the decoding strategy.
 @ariholtzmanCuriousCaseNeural2020 show how text generation based on a maximum likelihood objective leads to text that sounds bland and not human, propsing the use of different decoding strategies to make the generated text more human.
-
-  - Using embedding rather than identity matrix as the input turns this into an autoencoder, rather than a factor model.
-According to @goodfellowDeepLearning2016, generation from factor models leads to a combination of the learned features in these factors, but not to the generation of sensible data points.
-Existing, autoencoder based, approaches that have succeeded in generation tasks, might be more suitable for this task (e.g.
-CITE: Variational Autoencoder).
 
@@
 
-<!-- TODO: Regularization of summary embeddings, have them absolute and not relative -->
-
-  - Incorporating the context of the product itself, embedding e.g. the brand, product features and competitors in the model, to move into part of the generation space that is "open" for the product
-  - Using fine-tuned LLMs
-  - MAB around the fitted AE, to incorporate consumer feedback in online learning: Which areas of the space are valuable for which consumers?
-  - For online learning, could wrap the decoder into an MAB and let the MAB explore the space by pulling arms <!-- TODO: Which kind of MAB? -->
-  - Could also learn preferences online with this setup
-  - Better LLM
-  - Image, video, website, audio data. Perhaps website a good application. Perhaps audio a rather simple way to make this multi-modal (content and sound)
+We identify three avenues for future research based on our generative summaries.
+
+The first avenue is with respect to the summary's generative capabilities.
+Here, we want to research two extensions, that could improve the quality of its generated text.
+The first one is to replace the factor mode by a non-linear autoencoder which takes a different document summary, e.g.
+BERT's CLS token [@devlinBERTPretrainingDeep2018], as an input and learns a mapping from these descriptive embeddings to our generative summaries.
+The motivation for this is, that non-linear autoencoders appear to be better suited for generation than factor models [see @goodfellowDeepLearning2016].
+A second extension to improve the quality of our text generation could be to introduce an adversary model, creating an adversarial structure.
+An adversarial structure consists of two models: A generator and an adversary.
+The generator creates new data points, while the adversary tries to predict whether a data point is genuine or has been generated [@goodfellowGenerativeAdversarialNetworks2020].
+This leads to a game between these two models, where the generator learns to represent the underlying data distribution, such that the adversary cannot distinguish between genuine and generated datapoints anymore.
+Training such an adversary can also be a way to evaluate the quality of the generated text itself, which is an ongoing research problem [@tatsunorihashimotoUnifyingHumanStatistical2019].
+Exploring how we can reconcile such an adversarial structure with our current training objective of maximizing the generation probability of a focal sequence, is an issue we leave for future research.
+
+The second avenue is the selection of an adequate number of factors.
+The number of factors is a hyper-parameter that the researcher must set when using this model.
+For the number of factors, there are two interesting research questions.
+The first one is, to explore how different numbers of factors capture different levels of abstraction in the text data.
+The second one is, to develop a data-driven method to choose the number of factors.
+While there can be theoretical motivations for the number of factors (e.g.
+having two classes as in the test case of intangible and tangible claims), developing a data-driven method would be helpful.
+Such a method could work by penalizing the use of more factors while rewarding a high likelihood.
+Perhaps, adapting an information criterion that is based on the likelihood, such as the Bayesian Information Criterion (BIC), could be the starting point for such a development.
+
+<!-- TODO: Regularization of generative summaries, have them absolute and not relative -->
+
+<!-- Can we tailor the training objective to a specific marketing goal?
+For example, rather than maximizing the likelihood to generate a known advertising claim, can we also maximize the likelihood to generate an advertising claim that is liked by a consumer?
+-->
+
+<!-- The third avenue, are applications of the generative summaries to improve prompt engineering. -->
+
+Can we adapt the LLM further to a specific domain? For example, the approach by 
+@khattabDSPyCompilingDeclarative2023 also allows for fine-tuning of the LLM with respect to a specific performance metric.
+Rather than using an out of the box LLM, which has been trained on large text corpora, we could further train and adapt the LLM to the advertising claims, which are special in their use of words and short.
+Adapting the LLM to the language domain might prove to helpful, especially when we are interested in the factor space of a specific domain.
+Furthermore, such an adaptation might also improve the quality of new generated texts, as the LLM pronounces sequences of words that sounds like advertising claims more strongly.
+Besides fine-tuning, @ouyangTrainingLanguageModels2022 argue that especially Reinforcement Learning for Human Feedback (RL-HF), improves the capabilities of an LLM.
+Augmenting the training of generative summaries by RL-HF, in order to regularize the fit of these embeddings by ensuring that they capture signal rather than noise, is another research avenue to purse in the future.
 
@@
 
+In this research, we propose a novel, optimal, type of document summary, which maximizes the likelihood of an LLM to generate the document that it summarizes.
+These document summaries are the only ones, that are numeric and can be used for the generation of new text.
+In an application to synthetic advertising claims, we show that these document summaries capture relevant information about the documents, but need a factor structure prevent overfitting.
+We also propose a form of factor model to estimate these document summaries, which yields an interpretable factor space.
+In an application to market research data, we show that this factor space captures linguistic features and can provide insights on consumer ratings of uniqueness and overall appeal, as well as help with the discovery of design principles.
+We observe that linguistic uniqueness is not the same as perceived uniqueness by consumers and that claims of similar appeal ratings cluster together.
+However, we also observe that small adjustments to the wording of claims can lead to relatively large differences in rating.
+We also show that linear combinations of our generative summaries can be meaningful.
+We sketch a framework that can help business to adapt their text-based marketing measures, without having to rely on prompt engineering.
+Open issues of our approach are to solve the problem of selecting the number of factors in a more general way and to improve the quality of newly generated text.
+
 {{< pagebreak >}}
 
-# References
+# Appendix {.appendix}
+
+| Number | Claim |
+|:-:|:-----------:|
+0  (T) | Transforms dull hair into vibrant, eye-catching color in just one use.
+1 (T) | Infuses hair with rich, lasting color that doesn't fade after multiple washes.
+2 (T) | Provides 3X more vibrant color compared to leading brands.
+3 (T) | Boosts color intensity for up to 8 weeks with regular use.
+4 (T) | Enhances natural hair color with luminous, multi-dimensional shades.
+5 (T) | Leaves hair visibly shinier and smoother with every application.
+6 (T) | Protects color-treated hair from fading and dullness.
+7 (T) | Locks in color and moisture for a soft, silky finish.
+8 (T) | Revives faded hair color in just 5 minutes.
+9 (T) | Strengthens hair while delivering vivid, long-lasting color.
+10 (I) | Unleash your most vibrant self with our haircare product.
+11 (I) | Experience the magic of bold, beautiful hair color.
+12 (I) | Let your hair reflect your inner radiance.
+13 (I) | Feel the confidence of truly vibrant hair.
+14 (I) | Embrace your colorful side with every wash.
+15 (I) | Discover the joy of brilliantly colored hair.
+16 (I) | Transform your hair into a work of art.
+17 (I) | Reveal the true potential of your hair's color.
+18 (I) | Express your personality through stunning hair color.
+19 (I) | Let your hair be a reflection of your vibrant spirit.
+: Tangible (T) and Intangible (I) advertising claims for hair shampoo products. (Created by ChatGPT) {#tbl-claims-color}
+
+| Number | Claim |
+|:-:|:-----------:|
+0  (T) | Reduces hair breakage by up to 75% in just 4 weeks.
+1 (T) | Increases hair strength by 50% after 3 uses.
+2 (T) | Visibly smoother hair after one application.
+3 (T) | Improves hair thickness by 30% within a month.
+4 (T) | Reduces split ends by 60% with regular use.
+5 (T) | Enhances hair shine by 40% in 2 weeks.
+6 (T) | Reduces frizz by 80% in high humidity conditions.
+7 (T) | Promotes 2x faster hair growth.
+8 (T) | Decreases scalp dryness by 50% after first use.
+9 (T) | Improves hair elasticity by 45% with continuous use.
+10 (I) | Transforms your hair’s overall vitality.
+11 (I) | Revitalizes dull, lifeless hair.
+12 (I) | Restores natural hair vibrancy.
+13 (I) | Nourishes your hair from root to tip.
+14 (I) | Gives your hair a healthy, radiant glow.
+15 (I) | Revives the natural beauty of your hair.
+16 (I) | Infuses your hair with deep moisture.
+17 (I) | Enhances the natural texture of your hair.
+18 (I) | Makes hair feel luxuriously soft and silky.
+19 (I) | Elevates your hair’s natural brilliance.
+: Tangible (T) and Intangible (I) advertising claims for hair shampoo products. (Created by ChatGPT) {#tbl-claims-health}
+
+| Number | Claim |
+|:-:|:-----------:|
+0  (T) | Removes 99% of all dirt and grime in one wipe.
+1 (T) | Restores surfaces to their original shine in seconds.
+2 (T) | Leaves no streaks, just a sparkling clean finish.
+3 (T) | Eliminates tough stains with ease.
+4 (T) | Provides a long-lasting protective shine.
+5 (T) | Fast-acting formula, see results immediately.
+6 (T) | Safe for all surfaces including glass, metal, and wood.
+7 (T) | Dries quickly, leaving surfaces spotless and gleaming.
+8 (T) | Contains no harsh chemicals, gentle yet effective.
+9 (T) | Reduces cleaning time by half with superior efficiency.
+10 (I) | Experience a new level of clean.
+11 (I) | Transform your home with a radiant shine.
+12 (I) | Let your surfaces shine like never before.
+13 (I) | Make every room sparkle with brilliance.
+14 (I) | Enjoy the confidence of a spotless home.
+15 (I) | Feel the freshness in every corner.
+16 (I) | Unlock the secret to dazzling surfaces.
+17 (I) | Embrace the power of effortless shine.
+18 (I) | Rediscover the beauty of your surfaces.
+19 (I) | Breathe easy in a home that gleams.
+: Tangible (T) and Intangible (I) advertising claims for surface cleaner. (Created by ChatGPT) {#tbl-claims-surface}
+
+::: {#fig-correlation-surface layout-ncol=3}
+![RE](figures/correlation_re_surface.pdf)
+
+![W2V](figures/correlation_w2v_surface.pdf)
+
+![BERT](figures/correlation_bert_surface.pdf)
+
+Correlation matrices for the surface cleaner claims.
+:::
+
+::: {#fig-correlation-surfacescramble layout-ncol=3}
+![Scrambled order](figures/correlation_re_surface_scramble.pdf)
+
+![W2V](figures/correlation_w2v_surface_scramble.pdf)
+
+![BERT](figures/correlation_bert_surface_scramble.pdf)
+
+Correlation matrices for the scrambled surface cleaner claims.
+:::
+
+::: {#fig-correlation-haircolor layout-ncol=4}
+![2-factors](figures/correlation_re_color.pdf)
+
+![15-factors](figures/correlation_re_hair20color.pdf)
+
+![W2V](figures/correlation_w2v_color.pdf)
+
+![BERT](figures/correlation_bert_color.pdf)
+
+Correlation matrices for the hair claims with color attribute.
+:::
+
+::: {#fig-correlation-health layout-ncol=3}
+![RE](figures/correlation_re_health.pdf)
+
+![W2V](figures/correlation_w2v_health.pdf)
+
+![BERT](figures/correlation_bert_health.pdf)
+
+Correlation matrices for the hair color with health attribute.
+:::
+
+{{< pagebreak >}}
+
+# References {.references}

### 2024-06-21 writing/slides_pre_defense.qmd
pre-defense sync

---
title: "TITLE"
subtitle: "Pre-Defense MPhil assessed by dr. Meike Morren & dr. Jonne Guyt"
author: "Finn-Ole Höner; Supervisors: prof. Dennis Fok & prof. Bas Donkers"
date: 2024-06-28
date-format: long
format: beamer
aspectratio: 169
incremental: false
include-in-header: ../beamer_preamble.tex
bibliography: ../references.bib
fig-align: center
fig-cap-location: bottom
fig-format: pdf
links-as-notes: true
number-depth: 1
eval: false
echo: true
fig-width: 12
fig-height: 8
pdf-engine: pdflatex
---

## One-Slide Summary

- New method to calculate document summaries

---

\tiny

| Method                              | Origin                                          | Reference                              |  Deterministic? |  Generation?  |  Type?  |
|:------------------------------------|:------------------------------------|:--------------------|:--------------:|:--------------:|:--------------:|
| BERT `[CLS]` token                       | Next-sentence prediction task                  | @devlinBERTPretrainingDeep2018         | $\checkmark$   | $\times$       | Numeric    |
| Pooled Word Embeddings              | Aggregation of token information               | @tomasmikolovDistributedRepresentationsWords2013 @shenBaselineNeedsMore2018             | $\checkmark$   | $\times$       | Numeric    |
| (Optimized) Prompt engineering                  | Emergent capability of LLM                     | @khattabDSPyCompilingDeclarative2023; @huangenhanced                                   | $\times$       | $\checkmark$       | Textual    |
|  |  |                         |    |    |  |
| Reverse engineered document summaries | Maximize the likelihood to re-generate the focal document | *Proposed method*                        | $\checkmark$   | $\checkmark$   | Numeric |
: Overview of existing methods to summarize documents. {#tbl-docsum}

---

![Training of embeddings](figures/dashboard.png){#fig-optimization}

## two

::: {#fig-correlation layout-ncol=2}

![SE: 2 Factors with LayerNorm.](figures/correlation_re.pdf)

![BERT](figures/correlation_bert.pdf)

Correlation matrices along the claims. 
:::

---

::: {#fig-encoding layout-ncol=2}

![All points](figures/encoding_space.pdf){#fig-encoding1}

![Zoom-in](figures/encoding_space_zoomin.pdf){#fig-encoding2}

Encoding space of the advertising claims. Numbered points are locations of claims that are part of the training data, colored dots are points where we generate one of the claims that are part of the training data, and red crosses are locations where we generate a claim that ends with the eos-token, but is not part of the training data. Whitespace marks coordinates where we generate a string that does not contain an eos token.
:::

---

![Exploration of Encoding Space](figures/exploration_encoding_space.pdf){#fig-space}

---

\small

|    | 3rd      | 2nd        | Choice        |   Prob. 3rd |   Prob. 2nd |   Prob. Choice |
|---:|:---------|:-----------|:--------------|------------:|------------:|---------------:|
|  0 | See      | Form       | Experience    |       0.009 |       0.014 |          0.944 |
|  1 | 20       | 25         | 50            |       0.002 |       0.004 |          0.976 |
|  2 | %.       | .          | %             |       0.001 |       0.002 |          0.994 |
|  3 | better   | More       | more          |       0.001 |       0.002 |          0.996 |
|  4 | shine    | protective | visible       |       0.002 |       0.002 |          0.979 |
|  5 | glow     | light      | shine         |       0.001 |       0.002 |          0.983 |
|  6 | when     | by         | after         |       0.002 |       0.004 |          0.987 |
|  7 | a        | well       | just          |       0.001 |       0.002 |          0.990 |
|  8 | two      | a          | one           |       0.002 |       0.002 |          0.990 |
|  9 | shine    | usage      | use           |       0.000 |       0.003 |          0.991 |
| 10 | ,        | !          | .             |       0.000 |       0.001 |          0.998 |
| 11 | .        | \textit{space}  | \textit{eos-token} |       0.000 |       0.002 |          0.994 |
: Example for similar tokens {#tbl-similartoken}

---

\small

|    | Weight | Generated String                                                                                      |
|---:|:-------|:------------------------------------------------------------------------------------------------------|
|  0 | 0/20   | Unlock the secret to hair that shines from within, reflecting your inner glow.   \textit{eos-token}   |
|  1 | 1/20   | Unlock the secret to hair that shining from within, your bedroom. \textit{eos-token}                  |
|  2 | 2/20   | Unlock the secret to black metal's glow. \textit{multiple eos-token}                                  |
|  3 | 3/20   | Unlock the secret to blackened nails that shine from the shine.   \textit{eos-token}                  |
|  4 | 4/20   | Un to of for the team members upon \textit{linebreak} Moderation of the, that|
|  5 | ...    | ...                                                                                                   |
| 17 | 17/20  | Rediscover the joy of hair that's at the-corrects life with a healthy                                 |
| 18 | 18/20  | Rediscover the joy of hair that beams with inner vibrancy. \textit{eos-token}                         |
| 19 | 19/20  | Rediscover the joy of hair that beams with inner vibrancy. \textit{eos-token}                         |
: Example for similar tokens {#tbl-weight}

---

\small

\begin{algorithm}[H]
\begin{algorithmic}
\label{alg:singletraining}
\caption{Training of Single Summary Embeddings}
\State $i \gets 0$
\State $\epsilon \gets 0.01$
\State $\boldsymbol{s} \gets \operatorname{Initialization} \left( \cdot \right)$
\While{$True$}
    \State $l^{(i)} \gets \operatorname{\mathcal{L}} \left ( \boldsymbol{s} \mid t_1, \ldots, t_T \right)$
    \State $\nabla_{\boldsymbol{s}}^{(i)} l \gets \operatorname{ComputeGradient} \left ( l^{(i)}  \right )$
    \State $\boldsymbol{s}^{(i + 1)} l \gets \operatorname{Optimizer} \left(\boldsymbol{s}^{(i)}, \; \nabla_{\boldsymbol{s}}^{(i)} l \right)$
    \State $i \gets i + 1$    
    \If{$l^{(i)} < \epsilon$}
      \State \text{break}
    \EndIf
\EndWhile
\end{algorithmic}
\end{algorithm}

---

\small

\begin{algorithm}[H]
\begin{algorithmic}
\label{alg:training}
\caption{Training of Summary Embeddings based on factor model}
\State $i \gets 0$
\State $\epsilon \gets 0.01$
\State $\mathbf{W}^{Encoder}_{(i)}, \; \mathbf{W}^{Decoder}_{(i)}, \; \mathbf{B}^{Encoder}_{(i)}, \; \mathbf{B}^{Decoder}_{(i)} \gets \operatorname{Initialization} \left( \cdot \right)$
\While{$True$}
    \State $\mathbf{S} \gets \left( \mathbf{W}^{Encoder}_{(i)} + \mathbf{B}^{Encoder}_{(i)} \right) \mathbf{W}^{Decoder}_{(i)} + \mathbf{B}^{Decoder}_{(i)}$
    \State $l_{(i)} \gets \operatorname{\mathcal{L}}_D \left ( \mathbf{S} \right)$
    \State $\nabla_{(i)} \gets \operatorname{ComputeGradient} \left ( l^{(i)}  \right )$
    \State $\mathbf{W}^{Encoder}_{(i + 1)}, \; \mathbf{W}^{Decoder}_{(i + 1)}, \; \mathbf{B}^{Encoder}_{(i + 1)}, \; \mathbf{B}^{Decoder}_{(i + 1)} \gets \operatorname{Optimizer} \left(\mathbf{W}^{Enc., Dec.}_{(i + 1)}, \; \mathbf{B}^{Enc., Dec.}_{(i + 1)}, \; \nabla_{(i)} \right)$
    \State $i \gets i + 1$    
    \If{$l_{(i)} < \epsilon$}
      \State \text{break}
    \EndIf
\EndWhile
\end{algorithmic}
\end{algorithm}

---

![Document Summaries](figures/anecdote_morning_system_mr5.pdf){width=75%}

---

![Document Summaries](figures/uniqueness_mr5.pdf){width=75%}

## Appendix

![Document Summaries](figures/anecdote_morning_system_bert_mr5.pdf){width=75%}

---

![Document Summaries](figures/uniqueness_bert_mr5.pdf){width=75%}

## References {.allowframebreaks}

### 2024-06-21 src/analysis/toolbox.py

@@ -37,6 +37,17 @@ def TokenizeClaims(lClaims, tokenizer):
 def create_input_data(iClaims):
     return torch.eye(iClaims)
 
+def load_data(sDataPath, bUnique = False):
+    with open(sDataPath, 'r') as file:
+        lClaims = file.readlines()
+    lClaims = [claim.replace('\n', '') for claim in lClaims]
+    
+    if bUnique:
+    
+        return list(set(lClaims))
+    else:
+        return lClaims
+
 # TODO: Consider removing this function
 def PadScores(tOut, tLengths, iClaims, iVocab: int=50257):
 

### 2024-06-21 src/analysis/visualize_embeddings.py

@@ -20,6 +20,14 @@ sPathEmbeddings = "data/processed/summary_embeddings"
 sNameEmbeddingA = "tangible_hair.pkl"
 sNameEmbeddingB = "intangible_hair.pkl"
 
+# plt.rcParams.update({
+#     "font.family": "serif",
+#     "text.usetex": True,
+#     'text.latex.preamble': r'\usepackage{cm-super}' + r'\usepackage{kpfonts}' + r"\usepackage{amsmath}" + r'\usepackage{amssymb}' + r"\usepackage{siunitx}"
+# })
+
+# csfont = {'fontname':'kpfonts'}
+
 # FUNCTIONS --------------------------
 # https://stackoverflow.com/questions/56369030/runtimeerror-attempting-to-deserialize-object-on-a-cuda-device
 class CPU_Unpickler(pickle.Unpickler):
@@ -52,16 +60,17 @@ class EmbeddingVisualizer:
         self.lEmbeddingB = lEmbeddingB
         
         
-def visualize_correlations(similarity_matrix, sName, bEmbedding=False):
+def visualize_correlations(similarity_matrix, sName, sFilePath, bEmbedding=False):
 
     plt.imshow(similarity_matrix, cmap='coolwarm', vmin=-1, vmax=1)
     plt.colorbar()
     plt.title(f'Similarity Matrix ({sName})')
-    plt.xticks(range(similarity_matrix.shape[0]), range(1, similarity_matrix.shape[0] + 1))
-    plt.yticks(range(similarity_matrix.shape[0]), range(1, similarity_matrix.shape[0] + 1))
+    #plt.xticks([])
+    #plt.yticks([])
+    plt.savefig(sFilePath)
     plt.show()
     
-def visualize_pca(embeddings, sName, vColorAssign=None):
+def visualize_pca(embeddings, sName, sFilePath, vColorAssign=None):
     
     if vColorAssign is None:
         vColorAssign = np.zeros(embeddings.shape[0])
@@ -75,6 +84,11 @@ def visualize_pca(embeddings, sName, vColorAssign=None):
     for i, (x, y) in enumerate(zip(pca_embeddings[:, 0], pca_embeddings[:, 1])):
         plt.text(x, y, str(i+1), color='black', fontsize=10)
     plt.title(f'PCA ({sName})')
+    plt.xlabel('PC1')
+    plt.ylabel('PC2')
[…]

### 2024-06-21 src/cluster/run.py

@@ -1,0 +1,197 @@
+# SETUP --------------------------
+from src.toolbox import *
+import numpy as np
+import matplotlib.pyplot as plt
+
+import torch
+import torch.optim as optim
+from transformers import AutoTokenizer, GPT2LMHeadModel
+
+import time
+import yaml
+
+import mlflow
+
+# GPU
+device = (
+    "cuda"
+    if torch.cuda.is_available()
+    else "mps"
+    if torch.backends.mps.is_available()
+    else "cpu"
+)
+
+print(torch.cuda.get_device_name(torch.cuda.current_device()))
+torch.set_default_device(device)
+print(torch.cuda.get_device_name(torch.cuda.current_device()))
+
+print(f"Using {device} device")
+print(torch.tensor([1, 2, 3]).device)
+
+
+# load config
+# Read YAML file
+with open("config/experiment_test.yaml", 'r') as stream:
+    ddConfig = yaml.safe_load(stream)
+
+print(ddConfig)
+
+# MAGICKS --------------------------
+s_model = ddConfig["s_model"]
+iSeed = 4523522
+sPadding = "left"
+sNameAEModel = ddConfig["sNameAEModel"]
+
+lTargetStrings = load_data(ddConfig["sDataPath"])
+iClaims = len(lTargetStrings)
+iEncodingDim = ddConfig["iEncodingDim"]
+iOutputSize = ddConfig["iOutputSize"]
+
+torch.manual_seed(iSeed)
+model = GPT2LMHeadModel.from_pretrained(s_model, device_map=device, resume_download=True)
+#model = model.to(memory_format=torch.channels_last)
+tokenizer = AutoTokenizer.from_pretrained(s_model, resume_download=True)
+
+if tokenizer.pad_token is None:
+            tokenizer.pad_token = tokenizer.eos_token
+
+tOHETarget = create_input_data(iClaims)
+tTargetStrings, lLengths = TokenizeClaims(lTargetStrings, tokenizer)
+iMaxTokens = max(lLengths)
+
+tInputEmbeddings = model.get_input_embeddings()(tTargetStrings)
+# TODO: Look into `past_key_values`, can speed things up by caching the hidden states
+tTarget = torch.cat([torch.tensor([-100] * iClaims).unsqueeze(-1), tTargetStrings], dim = 1)
+
+
+# activation, hidden units, bias, other
[…]

### 2024-06-21 src/cluster/src/toolbox.py

@@ -1,0 +1,256 @@
+# Read in the sentences from lClaims
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+
+import torch
+import torch.nn as nn
+import torch.optim as optim
+
+from transformers import AutoTokenizer, GPT2LMHeadModel, PhrasalConstraint
+
+import time
+import pickle
+import io
+import os
+
+import mlflow
+
+def TokenizeClaims(lClaims, tokenizer):
+    # Tokenize the sentences
+    tokenized_sentences = [tokenizer.encode(sentence, add_special_tokens=True) for sentence in lClaims]
+
+    # Find the maximum length of the tokenized sentences
+    lLengths = [len(sentence) for sentence in tokenized_sentences]
+    max_length = max(len(sentence) for sentence in tokenized_sentences) + 1 # add on to account for the eos token
+
+    # Pad the tokenized sentences on the left to have the same length
+    padded_sentences = [sentence + [tokenizer.eos_token_id] + [tokenizer.pad_token_id] * (max_length - len(sentence) - 1) for sentence in tokenized_sentences]
+    
+    # Convert the padded sentences to tensors
+    padded_sentences = torch.tensor(padded_sentences)
+
+    return padded_sentences, torch.tensor(lLengths)
+
+def create_input_data(iClaims):
+    return torch.eye(iClaims)
+
+def load_data(sDataPath):
+    with open(sDataPath, 'r') as file:
+        lClaims = file.readlines()
+    lClaims = [claim.replace('\n', '') for claim in lClaims]
+    return lClaims
+
+# Defining Autoencoder model
+class Autoencoder(nn.Module):
+   def __init__(self, input_size, encoding_dim, output_size):
+       super(Autoencoder, self).__init__()
+       self.encoder = nn.Sequential(
+           nn.Linear(input_size, encoding_dim, bias=True)
+       )
+       self.decoder = nn.Sequential(
+           nn.Linear(encoding_dim, output_size, bias=True)#,
+           #nn.Dropout(p = 10/768),
+           #nn.LayerNorm(output_size)
+       )
+
+   def forward(self, x):
+       x = self.encoder(x)
+       x = self.decoder(x)
+       return x
+   
+   
[…]

### 2024-06-21 src/new_imply.py

@@ -1,0 +1,473 @@
+# toolbox.py^ --------------------------
+
+# Read in the sentences from lClaims
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+
+import torch
+import torch.nn as nn
+import torch.optim as optim
+
+from transformers import AutoTokenizer, GPT2LMHeadModel, PhrasalConstraint
+
+import time
+import pickle
+import io
+import os
+
+import mlflow
+
+def TokenizeClaims(lClaims, tokenizer):
+    # Tokenize the sentences
+    tokenized_sentences = [tokenizer.encode(sentence, add_special_tokens=True) for sentence in lClaims]
+
+    # Find the maximum length of the tokenized sentences
+    lLengths = [len(sentence) for sentence in tokenized_sentences]
+    max_length = max(len(sentence) for sentence in tokenized_sentences) + 1 # add on to account for the eos token
+
+    # Pad the tokenized sentences on the left to have the same length
+    padded_sentences = [sentence + [tokenizer.eos_token_id] + [tokenizer.pad_token_id] * (max_length - len(sentence) - 1) for sentence in tokenized_sentences]
+    
+    # Convert the padded sentences to tensors
+    padded_sentences = torch.tensor(padded_sentences)
+
+    return padded_sentences, torch.tensor(lLengths)
+
+def create_input_data(iClaims):
+    return torch.eye(iClaims)
+
+def load_data(sDataPath):
+    with open(sDataPath, 'r') as file:
+        lClaims = file.readlines()
+    lClaims = [claim.replace('\n', '') for claim in lClaims]
+    return lClaims
+
+# Defining Autoencoder model
+class Autoencoder(nn.Module):
+   def __init__(self, input_size, encoding_dim, output_size):
+       super(Autoencoder, self).__init__()
+       self.encoder = nn.Sequential(
+           nn.Linear(input_size, encoding_dim, bias=True)
+       )
+       self.decoder = nn.Sequential(
+           nn.Linear(encoding_dim, output_size, bias=True)#,
+           #nn.Dropout(p = 10/768),
+           #nn.LayerNorm(output_size)
+       )
+
+   def forward(self, x):
+       x = self.encoder(x)
+       x = self.decoder(x)
[…]

### 2024-06-21 remove_file_tracking.sh
sync

@@ -1,0 +1,34 @@
+#!/bin/bash
+
+# Check if the file path argument is provided
+if [ -z "$1" ]; then
+    echo "Usage: $0 path/to/your/file"
+    exit 1
+fi
+
+FILE_PATH=$1
+
+# Check if git-filter-repo is installed
+if ! command -v git-filter-repo &> /dev/null; then
+    echo "git-filter-repo is not installed. Please install it first."
+    exit 1
+fi
+
+# Remove the file from the repository history
+git filter-repo --path "$FILE_PATH" --invert-paths
+
+# Add the file to .gitignore
+echo "$FILE_PATH" >> .gitignore
+
+# Remove the file from the index
+git rm --cached "$FILE_PATH"
+
+# Commit the changes
+git add .gitignore
+git commit -m "Stop tracking $FILE_PATH and add to .gitignore"
+
+# Force push to the remote repository
+git push origin --force --all
+git push origin --force --tags
+
+echo "Successfully removed $FILE_PATH from Git history and stopped tracking it."
